### 2019-12-30 data-raw/phenotypes_2019.Rmd

---
title: "Phenotypes_Planting_2019"
author: "Alice MacQueen"
date: "12/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(XLConnect)
library(rlang)
library(switchgrassGWAS)
```

## Load all phenotypic data from 2019

```{r}
wb_pheno <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
                                   paste0("GWAS DOE Switchgrass, ",
                                          "Consolidated Data"),
                                   paste0("DOE_GWAS_2019_Consolidated ",
                                          "Pheno Data_12-25-19.xlsx")))
lst_pheno <- readWorksheet(wb_pheno, sheet = getSheets(wb_pheno))
phenos_planting1 <- lst_pheno$`GWAS 2019 Greenup Data`
phenos_planting2 <- lst_pheno$`GWAS 2019 Core Phenotype Data`
phenos_key <- lst_pheno$`Column Key and Notes`

wb_KING2019 <- loadWorkbook(file.path("C:", "Users", "alice", "Dropbox",
                                      paste0("GWAS DOE Switchgrass, ",
                                             "Consolidated Data"),
                                      "Individual Site Files",
                                      "KING_GWAS_2019_Chlorosis Data.xlsx"))
lst_KING2019 <- readWorksheet(wb_KING2019, sheet = getSheets(wb_KING2019))
KING2019 <- as_tibble(lst_KING2019$`KING GWAS 2019 Chlorosis`)
key_KING19 <- as_tibble(lst_KING2019$`COLUMN KEY`)

all_plants_phenos <- phenos_planting1 %>%
  left_join(PKLE_2018, by = c("SITE", "PLOT_GL", "PLOT_LC", "ACC", "INDV", "PLANT_ID")) %>%
  left_join(KBSM2018, by = c("PLOT_GL")) %>%
  left_join(TMPL2018, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  mutate(DAM = DAM/10) %>%
  left_join(KING2018, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  left_join(CLMB2018, by = "PLOT_GL") %>%
  left_join(KING2019, by = "PLOT_GL") %>%
  left_join(LOST2018, by = c("SITE", "PLOT_GL", "PLOT_LC", "PLANT_ID",
                             "PLANT_.ID_GL")) %>%
  mutate(PLANT_ID = ifelse(ACC == "AP13",
                           "AP13",
                           PLANT_ID),
         DEAD_2018 = ifelse(is.na(DEAD_2018), 0, DEAD_2018))
```

# Look at 2019 data

Remove 2019 phenotype scores for plants that didn't survive the winter of 2018.
```{r}
all_plants_phenos %>%
  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  group_by(SRV, DEAD_2018) %>% tally()

all_plants_phenos %>%
  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  filter(SRV == "W" & DEAD_2018 == 1)

all_plants_phenos %>%
  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  filter(SRV == "Y" & DEAD_2018 == 1)

```
12 called as "N" but I called previously as survived (DEAD_2018 = 0) - I'm willing to call these dead. Little data was collected.

13 called as "W" but I called previously as dead (and were noted (in 'NOTES') as "Dead" or "dead".) 8 have phenotype data collected in 2019.
CLMB	C6920	6920	J216	A	J216.A	J216.A.CL	
FRMI	F1002	1002	J462	A	J462.A	J462.A.FR	
FRMI	F1714	1714	J458	B	J458.B	J458.B.FR	
FRMI	F2716	2716	J498	C	J498.C	J498.C.FR	
FRMI	F2724	2724	J235	A	J235.A	J235.A.FR	
KING	K1909	1909	J465	C	J465.C	J465.C.KG	
KING	K2008	2008	J597	A	J597.A	J597.A.KG	
PKLE	P5601	5601	J036	A	J036.A	J036.A.PK	
TMPL	T1303	1303	J597	A	J597.A	J597.A.TP	
TMPL	T1609	1609	J538	A	J538.A	J538.A.TP
TMPL	T2308	2308	J504	A	J504.A	J504.A.TP	
TMPL	T2721	2721	J477	B	J477.B	J477.B.TP	
TMPL	T2803	2803	J592	C	J592.C	J592.C.TP

6 with issues on whether or not the plant died:
For these, SRV was called as "Y", but were noted (in 'NOTES') as "Dead". All 6 have some phenotypic data collected in 2019.

FRMI	F2519	2519	J482	A	J482.A	J482.A.FR	
FRMI	F2523	2523	J577	B	J577.B	J577.B.FR	
FRMI	F2720	2720	J465	A	J465.A	J465.A.FR	
FRMI	F3302	3302	J340	A	J340.A	J340.A.FR	
TMPL	T2106	2106	J019	A	J019.A	J019.A.TP	
OVTN	V1319	1319	J580	A	J580.A	J580.A.VN

Ask Jason about 19 of these. Drop all 31 for now.

```{r}
phenotypes_2019 <- all_plants_phenos %>%
  left_join(phenos_planting2, by = c("SITE", "PLOT_GL", "PLOT_LC")) %>%
  filter(!(SRV == "W" & DEAD_2018 == 1) & !(SRV == "Y" & DEAD_2018 == 1) & !(SRV == "N")) %>%
  dplyr::select(SITE:PLANT_ID, GR1:GR100, EM1:NOTE_HARVEST)
```

# 2019 Phenotype Cleanup

Ensure all of the columns have the correct data type - mostly, numeric.

```{r}
phenotypes_2019 %>%
  mutate(GR1 = as.numeric(GR1),
         GR50 = as.numeric(GR50),
         GR100 = as.numeric(GR100),
         EM1 = as.numeric(EM1),
         EM50 = as.numeric(EM50),
         FL1 = as.numeric(FL1),
         FL50 = as.numeric(FL50),
         TC_EOS = as.numeric(TC_EOS),
         HT_PAN_EOS = as.numeric(HT_PAN_EOS),
         LD_EOS = as.numeric(LD_EOS),
         REGR = as.numeric(REGR),
         MOISTURE = as.numeric(MOISTURE),
         BIOMASS = as.numeric(BIOMASS))

```

## Look for outlier values by site

and maybe by genetic subgroup, or ACC, or PLANT_ID