### 2020-01-06 analysis/1_bigsnpr_gwas_prep.Rmd
PAG talk draft from Tom's slides

---
title: "1_bigsnpr_gwas_prep"
author: "Alice MacQueen"
date: "1/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(XLConnect)
library(rlang)
library(bigsnpr)
library(switchgrassGWAS)
library(AnnotationDbi)
library(cowplot)
library(viridis)
library(bigsnpr)
library(scales)
library(lubridate)

source("../../Functions_ggplot-theme-adjustments_2018-01-03.R")
theme_set(theme_poster)

load("../data/sites.rda")
load("../data/metadata.rda")
load("../data/Taxa.rda")
load("../data/phenotypes_2019.rda")

usa <- map_data("state")
```

## prepare phenotypes for GWAS

```{r}
colremove <- names(which(colSums(!is.na(phenotypes_wide_2019)) < 200))

phe_gwas_4big <- phenotypes_wide_2019 %>%
  dplyr::select(-colremove, -ends_with("1"), -ends_with("100"), 
                -starts_with("BRKG"), -starts_with("FRMI"), 
                -starts_with("LINC"), -starts_with("OVTN"), 
                -starts_with("TMPL"), -starts_with("KING"))

```

# gwas with 9 PCs for 509 individuals

```{r}

# Load files used in creating the annotation tables.

txdb <- loadDb(file = file.path("/", "home", "alice", "pvdiv", "switchgrassGWAS", 
                                "data",
                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))

anno_info <-
  read_delim(file = file.path("/", "home", "alice",  "pvdiv", "switchgrassGWAS", 
                              "data",
                              "Pvirgatum_516_v5.1.annotation_info.txt"),
             col_names = TRUE, delim = "\t")
gene_anno_info <- tbl_df(anno_info) %>%
  distinct(locusName, .keep_all = TRUE)

svd <- readRDS(file.path("/", "home", "alice", "Github", "pvdiv-phenology-gxe",
                         "large-files", "SVD_upland_lowland_continuous_k20.rds"))

snp <- snp_attach(file.path("~", "pvdiv", "VCF_and_Numeric", "Pvirgatum_4x_784g_imp_phased_up_low_continuous_subset_maf0.05.rds"))

  G <- snp$genotypes
  CHR <- snp$map$chromosome
  POS <- snp$map$physical.pos
  plants <- snp$fam$sample.ID
  bonferroni <- -log10(0.05/length(POS))
  NCORES <- nb_cores()

  CHRN <- enframe(CHR, name = NULL) %>%
    dplyr::rename(CHR = .data$value) %>%
    mutate(CHRN = case_when(.data$CHR == "Chr01K" ~ 1,
                            .data$CHR == "Chr01N" ~ 2,
                            .data$CHR == "Chr02K" ~ 3,
                            .data$CHR == "Chr02N" ~ 4,
                            .data$CHR == "Chr03K" ~ 5,
                            .data$CHR == "Chr03N" ~ 6,
                            .data$CHR == "Chr04K" ~ 7,
                            .data$CHR == "Chr04N" ~ 8,
                            .data$CHR == "Chr05K" ~ 9,
                            .data$CHR == "Chr05N" ~ 10,
                            .data$CHR == "Chr06K" ~ 11,
                            .data$CHR == "Chr06N" ~ 12,
                            .data$CHR == "Chr07K" ~ 13,
                            .data$CHR == "Chr07N" ~ 14,
                            .data$CHR == "Chr08K" ~ 15,
                            .data$CHR == "Chr08N" ~ 16,
                            .data$CHR == "Chr09K" ~ 17,
                            .data$CHR == "Chr09N" ~ 18,
                            TRUE ~ 19
    ))
  markers <- tibble(CHR = CHR, POS = POS)

  plants <- plants %>% enframe(value="PLANT_ID") %>% dplyr::select(-name)
  plants %>%
    left_join(phe_gwas_4big) %>% arrange(PLANT_ID)

  df_phe <- plants %>%
    left_join(phe_gwas_4big)

  gwas <- pvdiv_gwas(df = df_phe, type = "linear", snp = snp, ncores = NCORES, 
                     covar = svd, npcs = 9)

```

# Mash

STIL, KBSM, PKLE, CLMB

GR50
EM50
FL50
TC_EOS
HT_PAN_EOS
T_GR_EM
T_GR_FL
T_EM_FL

```{r}
markers2sub <- attr(svd, "subset")
markers2 <- tibble(CHR = CHR[markers2sub], POS = POS[markers2sub])

gwas_rds <- pvdiv_results_in_folder(path = ".", pattern = "GWAS_object_") # 8 phenotypes from 4 sites - 32 phenotypes total with > 200 individuals phenotyped and sequenced.

phenotype_names <- str_sub(gwas_rds, start = 13, end = -5)

numSNPs <- ceiling(1500000/length(phenotype_names)^2)
numSNPs <- 1500 # the above gave 1465, so just rounded this to 1500.

mash_input <- pvdiv_bigsnp2mashr(path = ".", gwas_rds = gwas_rds, 
                           phenotypes = phenotype_names, numSNPs = numSNPs, 
                           markers = markers, markers2 = markers2, G = G, 
                           model = "linear", saveoutput = TRUE)

mash_output <- mash_standard_run(path = ".", list_input = mash_input, 
                                  numSNPs = numSNPs, saveoutput = TRUE)
```

## Visualize bigsnpr GWAS

```{r}
for(i in 2:ncol(phe_gwas_4big)){

  df1 <- phe_gwas_4big %>%
    dplyr::select(PLANT_ID, i)
  gwas <- pvdiv_gwas(df = df1, type = "linear", snp = snp, ncores = NCORES, 
                     covar = svd, npcs = 9)

  ggsave(snp_manhattan(gwas, infos.chr = CHRN$CHRN,
                       infos.pos = snp$map$physical.pos, npoints = 2.1E06) +
           geom_hline(yintercept = bonferroni, lty = 2),
         filename = paste0("Manhattan_logistic_winter_survival", 
                           "_subset_9_PCs.png"), width = 10, height = 6,
         path = file.path("~", "pvdiv", "climateGWAS", 
                          "results-upland-vs-lowland", 
                          "Logistic_winter_survival"))
  anno_tables <- pvdiv_table_topsnps(df = gwas, type = "bigsnp", 
                                     n = c(10,100,1000), FDRalpha = 0.1, 
                                     rangevector = c(0, 20000, 100000),
                                     markers = markers, 
                                     anno_info = gene_anno_info,
                                     txdb = txdb)
  saveRDS(anno_tables, file.path("~", "pvdiv", "climateGWAS", 
                                 "results-upland-vs-lowland", 
                                 "Logistic_winter_survival",
                                 paste0("Annotation_tables_", 
                                        "logistic_winter_survival", 
                                        "_subset_9_PCs.rds")))
  ggsave(snp_qq(gwas[which(!is.na(gwas$score)),]),
         filename = paste0("QQplot_logistic_winter_survival_", 
                           "subset_9_PCs.png"), width = 4,
         height = 4, path = file.path("~", "pvdiv", "climateGWAS", 
                                      "results-upland-vs-lowland", 
                                      "Logistic_winter_survival"))
}

```

# Local score post processing of bigsnpr GWAS

```{r}
log10p <- predict(gwas)
pval <- 10^log10p

data1 <- data.table(chr = CHR, pos = POS, estim = gwas$estim, 
                    std.err = gwas$std.err, bigsnpscore = gwas$score, 
                    pval = pval)

setkey(data1, chr)
data1$pval[data1$pval==0]=1e-16#min(data1$pval[-which(data1$pval ==0)])
Nchr=length(data1[,unique(chr)])

### Computation of absolute position in the genome. 
#This is useful for doing genomewide plots. 
chrInfo=data1[,.(L=.N,cor=autocor(pval)),chr]
setkey(chrInfo,chr)
data.table(chr=data1[,unique(chr),], S=cumsum(c(0,chrInfo$L[-Nchr])))

# The score mean must be negative, ksi must be chosen between mean and max of -log10(p-value)

xi=2
data1[,score:= -log10(pval)-xi]
data1[,lindley:=lindley(score),chr]

### Choice of $\xi$ (1,2,3 or 4)

mean(data1$score)

# verify mean is negative and max is positive
mean(-log10(data1$pval));max(-log10(data1$pval))

#hist(data1$score)
max(data1$lindley)

# Compute significance threshold for each chromosome
## Uniform distribution of p-values
chrInfo[,th:=thresUnif(L, cor, xi),chr]
(data1=data1[chrInfo])

(sigZones=data1[chrInfo,sig_sl(lindley, pos, unique(th)),chr])

saveRDS(data1, "Logistic_winter_tolerance_gwas_lindley_scores.rds")
saveRDS(sigZones, "Logistic_winter_tolerance_Lindley_sigZones.rds")
```

# Visualize local score GWAS

```{r}
phenos <- "winter_tolerance"
i = 1

  ggmanobject <- ggplot(data = data1, aes(x = pos, y = lindley)) +
    geom_point(aes(color = chr, fill = chr)) +
    geom_hline(yintercept = data1$th[1], linetype = 2) + 
    facet_wrap(~ chr, nrow = 1, scales = "free_x", 
               strip.position = "bottom") +
    scale_color_viridis(option = "B", end = 0.8, discrete = TRUE) + 
    scale_fill_viridis(option = "B", end = 0.8, discrete = TRUE) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          panel.background = element_rect(fill=NA)) +
    labs(x = "Chromosome", y = "Lindley") +
    scale_x_continuous(expand = c(0.22, 0.22)) #+
    #ylim(0, 40)
    #scale_shape_manual(values = rep(c(21,22),9), guide = FALSE)

    save_plot(paste0("Manhattan_", phenos[i], "_ksi_2_ylim_", 
                     str_replace_all(Sys.time(), ":", "."), ".png"), 
              plot = ggmanobject, base_aspect_ratio = 2.5, base_height = 5)
```