### 2020-01-22 analysis/common-garden-allele-balancing/1_VCF_file_filtering_by_common_garden_site.Rmd
Site by year subsets for common garden allele balancing

---
title: "VCF file filtering by common garden site"
author: "Alice MacQueen"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
load("~/Github/pvdiv-phenology-gxe/data/phenotypes_2019.rda")
load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
```

## Goals & Outline

    1. Use the 33.9 million SNP file, located at:
    `/home/SharedUser/PV/NatPop/All_chr_combined_636g_80pct_maf0.005.vcf.bz2`
      Convert this file to vcf.gz first.
    2. Use or make a data frame of PLANT_ID's that were present at each site in 2019.
    3. Subset the 33.9 million SNP file to 5% MAF for each site.
    4. Use vcftools to find allele frequencies for each SNP at each site.

Below are the steps necessary to make a subset of a VCF file for use for GWAS.

## Subset function

```{r}
pvdiv_pop_subset_bigsnp <- function(subset, selector){
  subset_name <- paste0("Pvirgatum_4x_784g_imp_phased_", subset, "_subset")
  snpfile_subset <- subset(snp, ind.row = selector)
  snp_subset <- snp_attach(snpfile_subset)
  snp_writeBed(snp_subset, bedfile = paste0(subset_name, ".bed"))
  mac <- round(25/length(selector), digits = 2) # what should the MAF be?
  if(mac > 0.05){
    mac <- 0.05
  }
  plink <- download_plink()
  bedfile <- snp_plinkQC(plink, prefix.in = subset_name,
                         maf = mac, geno = 0.1, hwe = 1e-50, 
                         prefix.out = paste0(subset_name, "_maf", mac),
                         extra.options = "--allow-no-sex --allow-extra-chr --chr Chr01K Chr01N Chr02K Chr02N Chr03K Chr03N Chr04K Chr04N Chr05K Chr05N Chr06K Chr06N Chr07K Chr07N Chr08K Chr08N Chr09K Chr09N")
  rdsfile <- snp_readBed(bedfile)
}

# Load data - if you get errors, make sure to run R from the GWAS/ folder in pubjuenger
snp <- snp_attach("/home/pubjuenger/GWAS/Pvirgatum_4x_784g_imp_phased_maf0.005.rds")
switchgrassGWAS::phenotypes

# Use phenotypes to set up selection criteria for subsets. These need to be vectors of the row numbers in the Taxa dataframe that you have phenotypes for.
gwasct_sel <- which(!is.na(phenotypes$GWAS_CT)) # gwas_ct GWAS

## Run function to generate subset SNP files for the subset
pvdiv_pop_subset_bigsnp(subset = "gwasct", selector = gwasct_sel)

```

## 2. Use or make a data frame of PLANT_ID's that were present at each site in 2019.

Plants alive EOS 2019 (higher of HT_PAN_EOS and TC_EOS):
BRKG: 254
CLMB: 496
FRMI: 185
KBSM: 516
KING: 286
LINC: 260
OVTN: 331
PKLE: 614
STIL: 451
TMPL: 373

For each site: # SNPs above a MAF of 5%
% SNPs above a MAF of 5% (at all 10 sites), # SNPs > 5% MAF at all 10 sites, distribution of these SNPs across the chromosomes
Number of plants alive from each subpopulation (sort of did this already)

Average, minimum, maximum, standard deviation, coefficient of variation of MAF across all 10 sites
histograms of these values across all SNPs

Compare these values to their values in 2018 to see how much frequencies shifted in the last year?

```{r}
colSums(!is.na(phenotypes_wide_2019))

for(i in 1:10){
site <- sites$Site[i]

site_2019 <- phenotypes_wide_2019 %>%
  pivot_longer(cols = BRKG_EM1:TMPL_TC_EOS, names_to = "SITE_PHE", values_to = "Measurement") %>%
  separate(col = SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  filter(SITE == site & !is.na(Measurement)) %>%
  group_by(PLANT_ID) %>%
  summarise(count = n())

writeLines(site_2019$PLANT_ID, con = paste0(site, "_2019"))
}

for(i in 1:10){
site <- sites$Site[i]
site_2018 <- phenotypes_2018 %>%
  pivot_longer(cols = BRKG_DEAD_2018:TMPL_TC_EOS_2018, names_to = "SITE_PHE", values_to = "Measurement") %>%
  separate(col = SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  filter(SITE == site & !is.na(Measurement)) %>%
  group_by(PLANT_ID) %>%
  summarise(count = n()) 
writeLines(site_2018$PLANT_ID, con = paste0(site, "_2018"))
}

```

Plants alive EOS 2018, then EOS 2019:
BRKG: 412 down to 254
CLMB: 591 down to 496
FRMI: 273 down to 185
KBSM: 544 down to 516
KING: 305 down to 286
LINC: 437 down to 260
OVTN: 346 down to 331
PKLE: 623 down to 614
STIL: 469 down to 451
TMPL: 388 down to 373

# VCFtools lines
```{}
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2018j &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2018 --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2018 &
```