### 2020-01-24 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
MAF Analysis and Plots for Common Garden Allele Balancing

---
title: "Allele Frequencies Common Gardens"
author: "Alice MacQueen"
date: "1/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(viridis)
library(cowplot)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r}
load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
```

```{r}
year <- "2019"
i=1
k= i+1
nmax = 33890000

site_year <- read_delim(paste0(sites$Site[i],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[i],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[i],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[i],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[i],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[i],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[i],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[i],"_", year)), sep = ":", 
           convert = TRUE)
site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
freq_sitebyyear <- site_year %>%
  full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))

for(k in 3:10){
  site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
  freq_sitebyyear <- freq_sitebyyear %>%
    full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))
}
year <- "2018"
for(k in 1:10){
  site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
  freq_sitebyyear <- freq_sitebyyear %>%
    full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))
}
saveRDS(freq_sitebyyear, file = paste0("Allele_frequencies_site_by_year_", nmax, "SNPs.rds"))
```

```{r}
freq_sitebyyear <- readRDS("Allele_frequencies_site_by_year_100000SNPs.rds")
freq_sitebyyear
```

```{r}
sites
freq_sitebyyear %>%
  group_by(between(MINF_PKLE_2018, 0.05, 0.95)) %>% tally()
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & 
           between(MINF_PKLE_2018, 0.05, 0.95) &
           between(MINF_FRMI_2018, 0.05, 0.95) & 
           between(MINF_KING_2018, 0.05, 0.95) & 
           between(MINF_OVTN_2018, 0.05, 0.95) & 
           between(MINF_KBSM_2018, 0.05, 0.95) & 
           between(MINF_LINC_2018, 0.05, 0.95) & 
           between(MINF_STIL_2018, 0.05, 0.95) & 
           between(MINF_TMPL_2018, 0.05, 0.95) & 
           between(MINF_CLMB_2018, 0.05, 0.95)) 
freq_sitebyyear %>%
  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
           between(MINF_PKLE_2018, 0.05, 0.95),
           between(MINF_FRMI_2018, 0.05, 0.95), 
           between(MINF_KING_2018, 0.05, 0.95), 
           between(MINF_OVTN_2018, 0.05, 0.95),  
           between(MINF_KBSM_2018, 0.05, 0.95),  
           between(MINF_LINC_2018, 0.05, 0.95),  
           between(MINF_STIL_2018, 0.05, 0.95),  
           between(MINF_TMPL_2018, 0.05, 0.95),  
           between(MINF_CLMB_2018, 0.05, 0.95),
           between(MINF_BRKG_2019, 0.05, 0.95), 
           between(MINF_PKLE_2019, 0.05, 0.95),
           between(MINF_FRMI_2019, 0.05, 0.95), 
           between(MINF_KING_2019, 0.05, 0.95), 
           between(MINF_OVTN_2019, 0.05, 0.95),  
           between(MINF_KBSM_2019, 0.05, 0.95),  
           between(MINF_LINC_2019, 0.05, 0.95),  
           between(MINF_STIL_2019, 0.05, 0.95),  
           between(MINF_TMPL_2019, 0.05, 0.95),  
           between(MINF_CLMB_2019, 0.05, 0.95)) %>% 
  tally() %>% arrange(desc(n)) %>%
  mutate(frac = n/100000)# %>%
  #write_csv(., path = "MAF_allows_gwas_10_sites_2_years.csv")
freq_sitebyyear %>%
  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
           between(MINF_PKLE_2018, 0.05, 0.95),
           between(MINF_FRMI_2018, 0.05, 0.95), 
           between(MINF_KING_2018, 0.05, 0.95), 
           between(MINF_OVTN_2018, 0.05, 0.95),  
           between(MINF_KBSM_2018, 0.05, 0.95),  
           between(MINF_LINC_2018, 0.05, 0.95),  
           between(MINF_STIL_2018, 0.05, 0.95),  
           between(MINF_TMPL_2018, 0.05, 0.95),  
           between(MINF_CLMB_2018, 0.05, 0.95)) %>% 
  tally() %>% arrange(desc(n)) %>%
  mutate(approx = n/100000*33880000) %>%
  write_csv(., path = "MAF_allows_gwas_10_sites_2018.csv")

freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")

36007-28544
28544/36007
36007/100000*33880000
28544/100000*33880000
```

# 36007 out of 100k sites had MAF > 0.05 at Pickle in 2018, our best site.
12.2M SNPs total.
# 28544 out of 100k sites have MAF > 0.05 at all 10 locations.
That's 7463 SNPs less, or 79.27% of the SNPs we would have had.
9.67M SNPs total. Lose 20% of the SNPs by constraining to 5% MAF at all sites.

BRKG 35452

```{r}
freq_sitebyyear %>%
  filter(between(MINF_TMPL_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_TMPL_2018, y = MINF_TMPL_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_LINC_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_LINC_2018, y = MINF_LINC_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_CLMB_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_CLMB_2018, y = MINF_CLMB_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  ggplot(aes(x = MINF_BRKG_2019, y = MINF_KING_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
```

```{r}
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & between(MINF_KING_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_KING_2018)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2019, 0.05, 0.95) & between(MINF_KING_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2019, y = MINF_KING_2019)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
freq_sitebyyear %>%
  filter(between(MINF_LINC_2019, 0.05, 0.95) & between(MINF_PKLE_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_LINC_2019, y = MINF_PKLE_2019)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
```

```{r}
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Brookings, 2018") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_KING_2019, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_KING_2019, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Kingsville, 2019") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2019, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2019, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2019") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2018, 0.05, 0.95) & between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2018, y = MINF_BRKG_2018)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2018") +
  ylab("MAF in Brookings, 2018")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2018, 0.05, 0.95) & between(MINF_LINC_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2018, y = MINF_LINC_2018)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2018") +
  ylab("MAF in Lincoln, 2018")
```

# Plots

```{r}
maf_long <- freq_sitebyyear %>%
  dplyr::select(-starts_with("MAJF"), -starts_with("N_CHR_")) %>%
  pivot_longer(cols = starts_with("MINF"), names_to = "SITE_YEAR", values_to = "MINF") %>%
  separate(col = "SITE_YEAR", into = c("MINF_", "SITE", "YEAR"), sep = "_")

```