### 2020-02-06 analysis/common-garden-allele-balancing/2_allele_frequencies_common_gardens.Rmd
Analyze genomic distribution of winter survival selected SNPs

---
title: "Allele Frequencies Common Gardens"
author: "Alice MacQueen"
date: "1/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(viridis)
library(cowplot)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Join data on server

```{r}
load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
year <- "2019"
i=1
k= i+1
nmax = 33890000

site_year <- read_delim(paste0(sites$Site[i],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[i],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[i],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[i],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[i],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[i],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[i],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[i],"_", year)), sep = ":", 
           convert = TRUE)
site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
freq_sitebyyear <- site_year %>%
  full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))

for(k in 3:10){
  site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
  freq_sitebyyear <- freq_sitebyyear %>%
    full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))
}
year <- "2018"
for(k in 1:10){
  site_year2 <- read_delim(paste0(sites$Site[k],"_", year, ".frq"), delim = "\t", 
                        n_max = nmax, col_types = "cdddcc", skip = 1, 
                        col_names=c("CHR", "POS", "N_ALLELES", 
                                    paste0("N_CHR_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ1_", sites$Site[k],"_", year),
                                    paste0("ALLELEFREQ2_", sites$Site[k],"_", year))) %>%
  separate(col = paste0("ALLELEFREQ1_", sites$Site[k],"_", year), 
           into = c("MAJOR",
                    paste0("MAJF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE) %>%
  separate(col = paste0("ALLELEFREQ2_", sites$Site[k],"_", year), 
           into = c("MINOR",
                    paste0("MINF_", sites$Site[k],"_", year)), sep = ":", 
           convert = TRUE)
  freq_sitebyyear <- freq_sitebyyear %>%
    full_join(site_year2, by = c("CHR", "POS", "N_ALLELES", "MAJOR", "MINOR"))
}
saveRDS(freq_sitebyyear, file = paste0("Allele_frequencies_site_by_year_", nmax, "SNPs.rds"))
```

## Load data

```{r}
freq_sitebyyear <- readRDS("Allele_frequencies_site_by_year_100000SNPs.rds")
freq_sitebyyear
```

## Tally SNPs by site

```{r}
sites
freq_sitebyyear %>%
  group_by(between(MINF_PKLE_2018, 0.05, 0.95)) %>% tally()
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & 
           between(MINF_PKLE_2018, 0.05, 0.95) &
           between(MINF_FRMI_2018, 0.05, 0.95) & 
           between(MINF_KING_2018, 0.05, 0.95) & 
           between(MINF_OVTN_2018, 0.05, 0.95) & 
           between(MINF_KBSM_2018, 0.05, 0.95) & 
           between(MINF_LINC_2018, 0.05, 0.95) & 
           between(MINF_STIL_2018, 0.05, 0.95) & 
           between(MINF_TMPL_2018, 0.05, 0.95) & 
           between(MINF_CLMB_2018, 0.05, 0.95)) 
freq_sitebyyear %>%
  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
           between(MINF_PKLE_2018, 0.05, 0.95),
           between(MINF_FRMI_2018, 0.05, 0.95), 
           between(MINF_KING_2018, 0.05, 0.95), 
           between(MINF_OVTN_2018, 0.05, 0.95),  
           between(MINF_KBSM_2018, 0.05, 0.95),  
           between(MINF_LINC_2018, 0.05, 0.95),  
           between(MINF_STIL_2018, 0.05, 0.95),  
           between(MINF_TMPL_2018, 0.05, 0.95),  
           between(MINF_CLMB_2018, 0.05, 0.95),
           between(MINF_BRKG_2019, 0.05, 0.95), 
           between(MINF_PKLE_2019, 0.05, 0.95),
           between(MINF_FRMI_2019, 0.05, 0.95), 
           between(MINF_KING_2019, 0.05, 0.95), 
           between(MINF_OVTN_2019, 0.05, 0.95),  
           between(MINF_KBSM_2019, 0.05, 0.95),  
           between(MINF_LINC_2019, 0.05, 0.95),  
           between(MINF_STIL_2019, 0.05, 0.95),  
           between(MINF_TMPL_2019, 0.05, 0.95),  
           between(MINF_CLMB_2019, 0.05, 0.95)) %>% 
  tally() %>% arrange(desc(n)) %>%
  mutate(frac = n/100000)# %>%
  #write_csv(., path = "MAF_allows_gwas_10_sites_2_years.csv")
freq_sitebyyear %>%
  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
           between(MINF_PKLE_2018, 0.05, 0.95),
           between(MINF_FRMI_2018, 0.05, 0.95), 
           between(MINF_KING_2018, 0.05, 0.95), 
           between(MINF_OVTN_2018, 0.05, 0.95),  
           between(MINF_KBSM_2018, 0.05, 0.95),  
           between(MINF_LINC_2018, 0.05, 0.95),  
           between(MINF_STIL_2018, 0.05, 0.95),  
           between(MINF_TMPL_2018, 0.05, 0.95),  
           between(MINF_CLMB_2018, 0.05, 0.95)) %>% 
  tally() %>% arrange(desc(n)) %>%
  mutate(approx = n/100000*33880000) %>%
  write_csv(., path = "MAF_allows_gwas_10_sites_2018.csv")

freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")

36007-28544
28544/36007
36007/100000*33880000
28544/100000*33880000
```

    *36007 out of 100k sites had MAF > 0.05 at Pickle in 2018, our best site.
    *12.2M SNPs total.
    *28544 out of 100k sites have MAF > 0.05 at all 10 locations.

That's 7463 SNPs less, or 79.27% of the SNPs we would have had.
9.67M SNPs total. Lose 20% of the SNPs by constraining to 5% MAF at all sites.

BRKG 35452

# ------------------
# Start of Plots

```{r}
freq_sitebyyear %>%
  filter(between(MINF_TMPL_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_TMPL_2018, y = MINF_TMPL_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_LINC_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_LINC_2018, y = MINF_LINC_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_CLMB_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_CLMB_2018, y = MINF_CLMB_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
freq_sitebyyear %>%
  ggplot(aes(x = MINF_BRKG_2019, y = MINF_KING_2019)) + 
  geom_point(alpha = 0.05) + geom_abline(slope = 1, intercept = 0, color = "red")
```

```{r}
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & between(MINF_KING_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_KING_2018)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2019, 0.05, 0.95) & between(MINF_KING_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2019, y = MINF_KING_2019)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
freq_sitebyyear %>%
  filter(between(MINF_LINC_2019, 0.05, 0.95) & between(MINF_PKLE_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_LINC_2019, y = MINF_PKLE_2019)) + 
  geom_hex() + geom_abline(slope = 1, intercept = 0, color = "red") + 
  scale_fill_viridis(option = "B", trans = "log2") 
```

## Labeled hex plots

```{r}
freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Brookings, 2018") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_KING_2019, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_KING_2019, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Kingsville, 2019") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2019, 0.05, 0.95) & between(MINF_BRKG_2019, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2019, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2019") +
  ylab("MAF in Brookings, 2019")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2018, 0.05, 0.95) & between(MINF_BRKG_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2018, y = MINF_BRKG_2018)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2018") +
  ylab("MAF in Brookings, 2018")

freq_sitebyyear %>%
  filter(between(MINF_OVTN_2018, 0.05, 0.95) & between(MINF_LINC_2018, 0.05, 0.95)) %>%
  ggplot(aes(x = MINF_OVTN_2018, y = MINF_LINC_2018)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.43)) + 
  xlab("MAF in Overton, 2018") +
  ylab("MAF in Lincoln, 2018")
```

## Grid of plots?

```{r}
maf_long <- freq_sitebyyear %>%
  dplyr::select(-starts_with("MAJF"), -starts_with("N_CHR_")) %>%
  pivot_longer(cols = starts_with("MINF"), names_to = "SITE_YEAR", values_to = "MINF") %>%
  separate(col = "SITE_YEAR", into = c("MINF_", "SITE", "YEAR"), sep = "_")

```

# ---------------------------
## Full SNP file analyses

```{r}

freq_sitebyyear %>%
  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
           between(MINF_PKLE_2018, 0.05, 0.95),
           between(MINF_FRMI_2018, 0.05, 0.95), 
           between(MINF_KING_2018, 0.05, 0.95), 
           between(MINF_OVTN_2018, 0.05, 0.95),  
           between(MINF_KBSM_2018, 0.05, 0.95),  
           between(MINF_LINC_2018, 0.05, 0.95),  
           between(MINF_STIL_2018, 0.05, 0.95),  
           between(MINF_TMPL_2018, 0.05, 0.95),  
           between(MINF_CLMB_2018, 0.05, 0.95),
           between(MINF_BRKG_2019, 0.05, 0.95), 
           between(MINF_PKLE_2019, 0.05, 0.95),
           between(MINF_FRMI_2019, 0.05, 0.95), 
           between(MINF_KING_2019, 0.05, 0.95), 
           between(MINF_OVTN_2019, 0.05, 0.95),  
           between(MINF_KBSM_2019, 0.05, 0.95),  
           between(MINF_LINC_2019, 0.05, 0.95),  
           between(MINF_STIL_2019, 0.05, 0.95),  
           between(MINF_TMPL_2019, 0.05, 0.95),  
           between(MINF_CLMB_2019, 0.05, 0.95)) %>% 
  tally() %>% arrange(desc(n)) %>%
  write_csv(., path = "SNPs_where_MAF_allows_gwas_10_sites_2_years.csv")
freq_sitebyyear %>%
  group_by(between(MINF_BRKG_2018, 0.05, 0.95), 
           between(MINF_PKLE_2018, 0.05, 0.95),
           between(MINF_FRMI_2018, 0.05, 0.95), 
           between(MINF_KING_2018, 0.05, 0.95), 
           between(MINF_OVTN_2018, 0.05, 0.95),  
           between(MINF_KBSM_2018, 0.05, 0.95),  
           between(MINF_LINC_2018, 0.05, 0.95),  
           between(MINF_STIL_2018, 0.05, 0.95),  
           between(MINF_TMPL_2018, 0.05, 0.95),  
           between(MINF_CLMB_2018, 0.05, 0.95)) %>% 
  tally() %>% arrange(desc(n)) %>%
  write_csv(., path = "SNPs_where_MAF_allows_gwas_10_sites_2018.csv")
```

```{r}
freq_maf_2018 <- freq_sitebyyear %>%
  filter(between(MINF_BRKG_2018, 0.05, 0.95) & 
           between(MINF_PKLE_2018, 0.05, 0.95) &
           between(MINF_FRMI_2018, 0.05, 0.95) & 
           between(MINF_KING_2018, 0.05, 0.95) & 
           between(MINF_OVTN_2018, 0.05, 0.95) & 
           between(MINF_KBSM_2018, 0.05, 0.95) & 
           between(MINF_LINC_2018, 0.05, 0.95) & 
           between(MINF_STIL_2018, 0.05, 0.95) & 
           between(MINF_TMPL_2018, 0.05, 0.95) & 
           between(MINF_CLMB_2018, 0.05, 0.95)) 
saveRDS(freq_maf_2018, "~/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/Allele_frequencies_with_maf_above_5percent.rds")
```

## Labeled hex plots

```{r}
brkg2yr <- freq_maf_2018 %>%
  ggplot(aes(x = MINF_BRKG_2018, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.475)) + 
  xlab("MAF in Brookings, 2018") +
  ylab("MAF in Brookings, 2019")
save_plot(filename = "BRKG_2_years_MAF_change.png", plot = brkg2yr)

brkg_king9 <- freq_maf_2018 %>%
  ggplot(aes(x = MINF_KING_2019, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.475)) + 
  xlab("MAF in Kingsville, 2019") +
  ylab("MAF in Brookings, 2019")
save_plot(filename = "BRKG_2019_vs_KING_2019_MAF.png", plot = brkg_king9)

brkg_ovtn9 <- freq_maf_2018 %>%
  ggplot(aes(x = MINF_OVTN_2019, y = MINF_BRKG_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.475)) + 
  xlab("MAF in Overton, 2019") +
  ylab("MAF in Brookings, 2019")
save_plot(filename = "BRKG_2019_vs_OVTN_2019_MAF.png", plot = brkg_ovtn9)

brkg_ovtn8 <- freq_maf_2018 %>%
  ggplot(aes(x = MINF_OVTN_2018, y = MINF_BRKG_2018)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.475)) + 
  xlab("MAF in Overton, 2018") +
  ylab("MAF in Brookings, 2018")
save_plot(filename = "BRKG_2018_vs_OVTN_2018_MAF.png", plot = brkg_ovtn8)

linc_ovtn8 <- freq_maf_2018 %>%
  ggplot(aes(x = MINF_OVTN_2018, y = MINF_LINC_2018)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.475)) + 
  xlab("MAF in Overton, 2018") +
  ylab("MAF in Lincoln, 2018")
save_plot(filename = "LINC_2018_vs_OVTN_2018_MAF.png", plot = linc_ovtn8)
###
linc_ovtn9 <- freq_maf_2018 %>%
  ggplot(aes(x = MINF_OVTN_2019, y = MINF_LINC_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.475)) + 
  xlab("MAF in Overton, 2019") +
  ylab("MAF in Lincoln, 2019")
save_plot(filename = "LINC_2019_vs_OVTN_2019_MAF.png", plot = linc_ovtn9)

frmi_king9 <- freq_maf_2018 %>%
  ggplot(aes(x = MINF_FRMI_2019, y = MINF_KING_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.475)) + 
  xlab("MAF in Fermilab, IL, 2019") +
  ylab("MAF in Kingsville, TX, 2019")
save_plot(filename = "FRMI_2019_vs_KING_2019_MAF.png", plot = frmi_king9)
##

frmi_2yr <- freq_maf_2018 %>%
  ggplot(aes(x = MINF_FRMI_2018, y = MINF_FRMI_2019)) + 
  geom_hline(yintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_vline(xintercept = c(0.25, 0.5, 0.75), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.95, 0.475)) + 
  xlab("MAF in Fermilab, IL, 2018") +
  ylab("MAF in Fermilab, IL, 2019")
save_plot(filename = "FRMI_2018_vs_2019_MAF.png", plot = frmi_2yr)
##

```

# Minor allele count instead of MAF

Minor allele needs to be present at least 15 times (for a MAF of 5% in 2019 at FRMI). 18 for mean

But for BRKG/LINC/KING 21 - 24 present is better. Let's do 25, every location has to have 25 or more of an allele to keep that SNP.

Or (if FRMI leads to dropping a ton of SNPs) possibly keep some SNPs that are above 15 for FRMI, and above 25 everywhere else.

```{r}
freq
mean(freq$N_CHR_FRMI_2019)*.05 # 286 # 416 BRKG 428 LINC 478 KING
(286)*.05
mean(freq$N_CHR_BRKG_2019)*.05
mean(freq$N_CHR_LINC_2019)*.05
mean(freq$N_CHR_KING_2019)*.05
```
STIL 762

```{r}
mac_sitebyyear <- freq_sitebyyear %>%
  mutate(MAC_PKLE_2019 = ifelse(round(N_CHR_PKLE_2019*MINF_PKLE_2019)>=N_CHR_PKLE_2019/2, round(N_CHR_PKLE_2019*MAJF_PKLE_2019), round(N_CHR_PKLE_2019*MINF_PKLE_2019)),
         MAC_FRMI_2019 = ifelse(round(N_CHR_FRMI_2019*MINF_FRMI_2019)>=N_CHR_FRMI_2019/2, round(N_CHR_FRMI_2019*MAJF_FRMI_2019), round(N_CHR_FRMI_2019*MINF_FRMI_2019)),
         MAC_LINC_2019 = ifelse(round(N_CHR_LINC_2019*MINF_LINC_2019)>=N_CHR_LINC_2019/2, round(N_CHR_LINC_2019*MAJF_LINC_2019), round(N_CHR_LINC_2019*MINF_LINC_2019)),
         MAC_BRKG_2019 = ifelse(round(N_CHR_BRKG_2019*MINF_BRKG_2019)>=N_CHR_BRKG_2019/2, round(N_CHR_BRKG_2019*MAJF_BRKG_2019), round(N_CHR_BRKG_2019*MINF_BRKG_2019)),
         MAC_CLMB_2019 = ifelse(round(N_CHR_CLMB_2019*MINF_CLMB_2019)>=N_CHR_CLMB_2019/2, round(N_CHR_CLMB_2019*MAJF_CLMB_2019), round(N_CHR_CLMB_2019*MINF_CLMB_2019)),
         MAC_OVTN_2019 = ifelse(round(N_CHR_OVTN_2019*MINF_OVTN_2019)>=N_CHR_OVTN_2019/2, round(N_CHR_OVTN_2019*MAJF_OVTN_2019), round(N_CHR_OVTN_2019*MINF_OVTN_2019)),
         MAC_STIL_2019 = ifelse(round(N_CHR_STIL_2019*MINF_STIL_2019)>=N_CHR_STIL_2019/2, round(N_CHR_STIL_2019*MAJF_STIL_2019), round(N_CHR_STIL_2019*MINF_STIL_2019)),
         MAC_TMPL_2019 = ifelse(round(N_CHR_TMPL_2019*MINF_TMPL_2019)>=N_CHR_TMPL_2019/2, round(N_CHR_TMPL_2019*MAJF_TMPL_2019), round(N_CHR_TMPL_2019*MINF_TMPL_2019)),
         MAC_KBSM_2019 = ifelse(round(N_CHR_KBSM_2019*MINF_KBSM_2019)>=N_CHR_KBSM_2019/2, round(N_CHR_KBSM_2019*MAJF_KBSM_2019), round(N_CHR_KBSM_2019*MINF_KBSM_2019)),
         MAC_KING_2019 = ifelse(round(N_CHR_KING_2019*MINF_KING_2019)>=N_CHR_KING_2019/2, round(N_CHR_KING_2019*MAJF_KING_2019), round(N_CHR_KING_2019*MINF_KING_2019)),
         MAC_KING_2018 = ifelse(round(N_CHR_KING_2018*MINF_KING_2018)>=N_CHR_KING_2018/2, round(N_CHR_KING_2018*MAJF_KING_2018), round(N_CHR_KING_2018*MINF_KING_2018)),
         MAC_KBSM_2018 = ifelse(round(N_CHR_KBSM_2018*MINF_KBSM_2018)>=N_CHR_KBSM_2018/2, round(N_CHR_KBSM_2018*MAJF_KBSM_2018), round(N_CHR_KBSM_2018*MINF_KBSM_2018)),
         MAC_PKLE_2018 = ifelse(round(N_CHR_PKLE_2018*MINF_PKLE_2018)>=N_CHR_PKLE_2018/2, round(N_CHR_PKLE_2018*MAJF_PKLE_2018), round(N_CHR_PKLE_2018*MINF_PKLE_2018)),
         MAC_OVTN_2018 = ifelse(round(N_CHR_OVTN_2018*MINF_OVTN_2018)>=N_CHR_OVTN_2018/2, round(N_CHR_OVTN_2018*MAJF_OVTN_2018), round(N_CHR_OVTN_2018*MINF_OVTN_2018)),
         MAC_BRKG_2018 = ifelse(round(N_CHR_BRKG_2018*MINF_BRKG_2018)>=N_CHR_BRKG_2018/2, round(N_CHR_BRKG_2018*MAJF_BRKG_2018), round(N_CHR_BRKG_2018*MINF_BRKG_2018)),
         MAC_CLMB_2018 = ifelse(round(N_CHR_CLMB_2018*MINF_CLMB_2018)>=N_CHR_CLMB_2018/2, round(N_CHR_CLMB_2018*MAJF_CLMB_2018), round(N_CHR_CLMB_2018*MINF_CLMB_2018)),
         MAC_TMPL_2018 = ifelse(round(N_CHR_TMPL_2018*MINF_TMPL_2018)>=N_CHR_TMPL_2018/2, round(N_CHR_TMPL_2018*MAJF_TMPL_2018), round(N_CHR_TMPL_2018*MINF_TMPL_2018)),
         MAC_LINC_2018 = ifelse(round(N_CHR_LINC_2018*MINF_LINC_2018)>=N_CHR_LINC_2018/2, round(N_CHR_LINC_2018*MAJF_LINC_2018), round(N_CHR_LINC_2018*MINF_LINC_2018)),
         MAC_STIL_2018 = ifelse(round(N_CHR_STIL_2018*MINF_STIL_2018)>=N_CHR_STIL_2018/2, round(N_CHR_STIL_2018*MAJF_STIL_2018), round(N_CHR_STIL_2018*MINF_STIL_2018)),
         MAC_FRMI_2018 = ifelse(round(N_CHR_FRMI_2018*MINF_FRMI_2018)>=N_CHR_FRMI_2018/2, round(N_CHR_FRMI_2018*MAJF_FRMI_2018), round(N_CHR_FRMI_2018*MINF_FRMI_2018))) %>%
  dplyr::select(CHR, POS, MAC_PKLE_2019:MAC_FRMI_2018)

saveRDS(mac_sitebyyear, file =  "SNPs_where_minor_allele_count_allows_gwas_10_sites_2_years.rds")

mac_sitebyyear %>%
    group_by(MAC_PKLE_2019 > 25, MAC_PKLE_2018 > 25,
           MAC_FRMI_2019 > 15, MAC_FRMI_2018 > 15,
           MAC_OVTN_2019 > 25, MAC_OVTN_2018 > 25,
           MAC_LINC_2019 > 25, MAC_LINC_2018 > 25,
           MAC_STIL_2019 > 25, MAC_STIL_2018 > 25,
           MAC_TMPL_2019 > 25, MAC_TMPL_2018 > 25,
           MAC_CLMB_2019 > 25, MAC_CLMB_2018 > 25,
           MAC_BRKG_2019 > 25, MAC_BRKG_2018 > 25,
           MAC_KBSM_2019 > 25, MAC_KBSM_2018 > 25,
           MAC_KING_2019 > 25, MAC_KING_2018 > 25) %>%
  tally() %>%
  write_csv(., path = "Minor_allele_count_that_allows_gwas_10_sites_2_years_relaxed_FRMI.csv")

mac_sitebyyear %>%
    group_by(MAC_PKLE_2019 > 25, MAC_PKLE_2018 > 25,
           MAC_FRMI_2019 > 25, MAC_FRMI_2018 > 25,
           MAC_OVTN_2019 > 25, MAC_OVTN_2018 > 25,
           MAC_LINC_2019 > 25, MAC_LINC_2018 > 25,
           MAC_STIL_2019 > 25, MAC_STIL_2018 > 25,
           MAC_TMPL_2019 > 25, MAC_TMPL_2018 > 25,
           MAC_CLMB_2019 > 25, MAC_CLMB_2018 > 25,
           MAC_BRKG_2019 > 25, MAC_BRKG_2018 > 25,
           MAC_KBSM_2019 > 25, MAC_KBSM_2018 > 25,
           MAC_KING_2019 > 25, MAC_KING_2018 > 25) %>%
  tally() %>%
  write_csv(., path = "Minor_allele_count_that_allows_gwas_10_sites_2_years.csv")
```

```{r}
mac %>%
  group_by(MAC_PKLE_2019 > 25, MAC_PKLE_2018 > 25,
           MAC_FRMI_2019 > 15, MAC_FRMI_2018 > 15,
           MAC_OVTN_2019 > 25, MAC_OVTN_2018 > 25,
           MAC_LINC_2019 > 25, MAC_LINC_2018 > 25,
           MAC_STIL_2019 > 25, MAC_STIL_2018 > 25,
           MAC_TMPL_2019 > 25, MAC_TMPL_2018 > 25,
           MAC_CLMB_2019 > 25, MAC_CLMB_2018 > 25,
           MAC_BRKG_2019 > 25, MAC_BRKG_2018 > 25,
           MAC_KBSM_2019 > 25, MAC_KBSM_2018 > 25,
           MAC_KING_2019 > 25, MAC_KING_2018 > 25) %>%
  tally()
```

```{r}
mac_sitebyyear <- readRDS(file =  "SNPs_where_minor_allele_count_allows_gwas_10_sites_2_years.rds")

mac_2018_gt25 <- mac_sitebyyear %>%
  filter(MAC_PKLE_2018 >= 25,
         MAC_FRMI_2018 >= 15,
         MAC_OVTN_2018 >= 25,
         MAC_LINC_2018 >= 25,
         MAC_STIL_2018 >= 25,
         MAC_TMPL_2018 >= 25,
         MAC_CLMB_2018 >= 25,
         MAC_BRKG_2018 >= 25,
         MAC_KBSM_2018 >= 25,
         MAC_KING_2018 >= 25)

brkg_2yr <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_BRKG_2018, y = MAC_BRKG_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Brookings, 2018") +
  ylab("Minor Allele Count in Brookings, 2019")
save_plot(filename = "BRKG_2018_v_2019_minor_allele_count_change.png", plot = brkg_2yr)

brkg_king9 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_KING_2019, y = MAC_BRKG_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Kingsville, 2019") +
  ylab("Minor Allele Count in Brookings, 2019")
save_plot(filename = "BRKG_2019_vs_KING_2019_minor_allele_count.png", plot = brkg_king9)

brkg_ovtn9 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_OVTN_2019, y = MAC_BRKG_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + + 
  xlab("Minor Allele Count in Overton, 2019") +
  ylab("Minor Allele Count in Brookings, 2019")
save_plot(filename = "BRKG_2019_vs_OVTN_2019_minor_allele_count.png", plot = brkg_ovtn9)

brkg_ovtn8 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_OVTN_2018, y = MAC_BRKG_2018)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Overton, TX, 2018") +
  ylab("Minor Allele Count in Brookings, SD, 2018")
save_plot(filename = "BRKG_2018_vs_OVTN_2018_minor_allele_count.png", plot = brkg_ovtn8)

linc_ovtn8 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_OVTN_2018, y = MAC_LINC_2018)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) +  
  xlab("Minor Allele Count in Overton, 2018") +
  ylab("Minor Allele Count in Lincoln, 2018")
save_plot(filename = "LINC_2018_vs_OVTN_2018_minor_allele_count.png", plot = linc_ovtn8)
###

linc_ovtn9 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_OVTN_2019, y = MAC_LINC_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Overton, TX, 2019") +
  ylab("Minor Allele Count in Lincoln, NE, 2019")
save_plot(filename = "LINC_2019_vs_OVTN_2019_minor_allele_count.png", plot = linc_ovtn9)

frmi_king9 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_FRMI_2019, y = MAC_KING_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Fermilab, IL, 2019") +
  ylab("Minor Allele Count in Kingsville, TX, 2019")
save_plot(filename = "FRMI_2019_vs_KING_2019_minor_allele_count.png", plot = frmi_king9)
##

frmi_2yr <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_FRMI_2018, y = MAC_FRMI_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Fermilab, IL, 2018") +
  ylab("Minor Allele Count in Fermilab, IL, 2019")
save_plot(filename = "FRMI_2018_vs_2019_minor_allele_count.png", plot = frmi_2yr)
##
linc_2yr <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_LINC_2018, y = MAC_LINC_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Lincoln, NE, 2018") +
  ylab("Minor Allele Count in Lincoln, NE, 2019")
save_plot(filename = "LINC_2018_vs_2019_minor_allele_count.png", plot = linc_2yr)

clmb_2yr <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_CLMB_2018, y = MAC_CLMB_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Columbia, MO, 2018") +
  ylab("Minor Allele Count in Columbia, MO, 2019")
save_plot(filename = "CLMB_2018_vs_2019_minor_allele_count.png", plot = clmb_2yr)

kbsm_2yr <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_KBSM_2018, y = MAC_KBSM_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in KBS, MI, 2018") +
  ylab("Minor Allele Count in KBS, MI, 2019")
save_plot(filename = "KBSM_2018_vs_2019_minor_allele_count.png", plot = kbsm_2yr)
###
frmi_king8 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_FRMI_2018, y = MAC_KING_2018)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Fermilab, IL, 2018") +
  ylab("Minor Allele Count in Kingsville, TX, 2018")
save_plot(filename = "FRMI_2018_vs_KING_2018_minor_allele_count.png", plot = frmi_king9)
brkg_king8 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_KING_2018, y = MAC_BRKG_2018)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Kingsville, 2018") +
  ylab("Minor Allele Count in Brookings, 2018")
save_plot(filename = "BRKG_2018_vs_KING_2018_minor_allele_count.png", plot = brkg_king9)
#
clmb_pkle9 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_CLMB_2019, y = MAC_PKLE_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Columbia, MO, 2019") +
  ylab("Minor Allele Count in Pickle, TX, 2019")
save_plot(filename = "CLMB_2019_vs_PKLE_2019_minor_allele_count.png", plot = clmb_pkle9)

kbsm_pkle9 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_KBSM_2019, y = MAC_PKLE_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in KBS, MI, 2019") +
  ylab("Minor Allele Count in Pickle, TX, 2019")
save_plot(filename = "KBSM_2019_vs_PKLE_2019_minor_allele_count.png", plot = kbsm_pkle9)

brkg_pkle9 <- mac_2018_gt25 %>%
  ggplot(aes(x = MAC_BRKG_2019, y = MAC_PKLE_2019)) + 
  geom_hline(yintercept = c(0, 100, 200), color = "lightgrey") +
  geom_vline(xintercept = c(0, 100, 200), color = "lightgrey") +
  geom_hex() + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_fill_viridis(option = "B", trans = "log2") + 
  theme(legend.position = c(0.2, 0.7)) + 
  xlab("Minor Allele Count in Brookings, SD, 2019") +
  ylab("Minor Allele Count in Pickle, TX, 2019")
save_plot(filename = "BRKG_2019_vs_PKLE_2019_minor_allele_count.png", plot = brkg_pkle9)
```

5 plots of 2018 vs 2019: any place where more than 20 plants died. That's basically just the north.

Some 2019 north vs south plots.

# -----------------

# Impute on 11M SNPs

```{r}
snps <- readRDS("SNPs_where_minor_allele_count_allows_gwas_10_sites_2_years_.rds")
```

Here is a breakdown of the number of individuals we have for GWAS at each site, for phenotypes from 2018 and from 2019. The first two data columns have individuals with high quality SNP data and phenotypes. The next two have tetraploid plants with phenotypes – there is a lot of overlap, and only 5-6 plants that could have their libraries resequenced to help the GWAS data.

	Plants with SNP data & phenotypes	Plants with phenotypes	Plants Dead
	2018	2019	2018	2019	2018 to 2019
BRKG	408	254	412	254	154
KBSM	539	517	544	517	22
FRMI	272	185	273	185	87
LINC	433	262	437	262	171
CLMB	585	496	591	496	89
STIL	465	451	469	451	14
OVTN	343	332	346	332	11
TMPL	384	373	388	373	11
PKLE	617	615	623	615	2
KING	303	294	305	294	9

Given that information, I then subdivided the SNPs from John’s new 33.9M SNP dataset. These are SNPs missing in 20% or less of individuals from each subpopulation. Instead of a minor allele frequency, I used a minor allele count of 25 or more (except for FRMI where I used 15). This was the same as a MAF of 5% at BRKG, FRMI, and LINC, and a slightly lower MAF for the other sites with more individuals. Here’s the breakdown of the site*year combinations where SNPs can and can’t be used for GWAS:

Number of SNPs	Category
12536071
	Can't be used for GWAS – these SNPs have less than 25 copies of the minor allele in the 631 individuals planted at the common gardens.
10452584	Can be used for GWAS at all 10 sites and both 2018 and 2019 data.
1160326	Can’t be used in LINC & BRKG in 2019
1155906	Only at high enough frequency to be used at the big 3 sites
546381	Can only be used at PKLE & CLMB
439918	Can only be used at PKLE
346189	Can only be used at big 4 sites – KBSM CLMB PKLE STIL
341773	Can’t be used at KING OVTN FRMI
331082	Can’t be used at LINC in 2019 
327364	Can’t be used at KING
315522	Can only be used at big 4 and LINC 2018
305573	Missing 2019 at 4 freezing sites – LINC FRMI CLMB BRKG
285838	Can only be used at PKLE x2yr & CLMB 2018
246340	Missing 2019 at 3 freezing sites, not big site (BRKG)
231650	Missing KING x2yr and FRMI 2019
192819	Missing KING x2yr and FRMI x2yr
……	4.67M SNPs have some other combination of Site x Year that is useable.

10_site_2_yr 10452584 all MAC > 25 except FRMI 2019 > 15.
big_three   10452584 + 1155906 PKLE KBSM CLMB > 25
fourway_seven   PKLE TMPL BRKG KBSM LINC CLMB STIL > 25
 10452584 + 341773 + 327364 + 231650 + 192819 + ...
miss_freeze 305573 + 246340 + 1160326 + 331082 4 sites 2018 > 25 and 1-4 sites 2019 < 25

## 10site_2yr
all MAC >= 25 except FRMI 2019 > 15.

10,183,856 SNPs
```{r}
tensite_twoyear <- snps %>%
  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
           MAC_FRMI_2019 >= 15, MAC_FRMI_2018 >= 25,
           MAC_OVTN_2019 >= 25, MAC_OVTN_2018 >= 25,
           MAC_LINC_2019 >= 25, MAC_LINC_2018 >= 25,
           MAC_STIL_2019 >= 25, MAC_STIL_2018 >= 25,
           MAC_TMPL_2019 >= 25, MAC_TMPL_2018 >= 25,
           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
           MAC_BRKG_2019 >= 25, MAC_BRKG_2018 >= 25,
           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25,
           MAC_KING_2019 >= 25, MAC_KING_2018 >= 25)
saveRDS(tensite_twoyear, file = "SNPs_above_MAC_25_in_10sites_2years.rds")
```

## big_three

18,704,908 SNPs

PKLE KBSM CLMB >= 25
```{r}
big_three <- snps %>%
  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25)
saveRDS(big_three, "SNPs_above_MAC_25_in_big_three_sites.rds")
10183856+6027776+2493276
```

## fourway_seven

12,677,132 SNPs

PKLE TMPL BRKG KBSM LINC CLMB STIL >= 25
```{r}
fourway_seven <- snps %>%
  filter(MAC_PKLE_2019 >= 25, MAC_PKLE_2018 >= 25,
           MAC_LINC_2019 >= 25, MAC_LINC_2018 >= 25,
           MAC_STIL_2019 >= 25, MAC_STIL_2018 >= 25,
           MAC_TMPL_2019 >= 25, MAC_TMPL_2018 >= 25,
           MAC_CLMB_2019 >= 25, MAC_CLMB_2018 >= 25,
           MAC_BRKG_2019 >= 25, MAC_BRKG_2018 >= 25,
           MAC_KBSM_2019 >= 25, MAC_KBSM_2018 >= 25)
saveRDS(fourway_seven, file = "SNPs_above_MAC_25_in_fourway_7_sites.rds")
10183856+2493276
```

## microbiome

Just got the results from PKLE thus far.

21,525,180 SNPs

```{r}
microbiome <- snps %>%
  filter(MAC_PKLE_2019 >= 25)
10183856+6027776+2493276+2820272
```

## missing_frz

4 sites 2018 >= 25 and 1-4 sites 2019 < 25

2,288,528 SNPs
```{r}
missing_frz <- snps %>%
  filter(MAC_FRMI_2018 >= 25 & MAC_LINC_2018 >= 25 & MAC_CLMB_2018 >= 25 &
           MAC_BRKG_2018 >= 25, MAC_FRMI_2019 < 15 | MAC_LINC_2019 < 25 | 
           MAC_CLMB_2019 < 25 | MAC_BRKG_2019 < 25) %>%
  mutate(avediff = ((MAC_FRMI_2018 - MAC_FRMI_2019) + (MAC_LINC_2018 - MAC_LINC_2019) + (MAC_CLMB_2018 - MAC_CLMB_2019) + (MAC_BRKG_2018 - MAC_BRKG_2019))/4)

saveRDS(missing_frz, file = "Missing_SNPs_2019_4_freezing_sites.rds")

25/(615*2)
25/(496*2)
```

```{r}
snp_subsets <- snps %>%
  mutate(SNP_tensite_twoyear = ifelse(MAC_PKLE_2019 >= 25 & 
                                        MAC_PKLE_2018 >= 25 & 
                                        MAC_FRMI_2019 >= 15 & 
                                        MAC_FRMI_2018 >= 25 & 
                                        MAC_OVTN_2019 >= 25 & 
                                        MAC_OVTN_2018 >= 25 & 
                                        MAC_LINC_2019 >= 25 & 
                                        MAC_LINC_2018 >= 25 & 
                                        MAC_STIL_2019 >= 25 & 
                                        MAC_STIL_2018 >= 25 & 
                                        MAC_TMPL_2019 >= 25 & 
                                        MAC_TMPL_2018 >= 25 & 
                                        MAC_CLMB_2019 >= 25 & 
                                        MAC_CLMB_2018 >= 25 & 
                                        MAC_BRKG_2019 >= 25 & 
                                        MAC_BRKG_2018 >= 25 & 
                                        MAC_KBSM_2019 >= 25 & 
                                        MAC_KBSM_2018 >= 25 & 
                                        MAC_KING_2019 >= 25 & 
                                        MAC_KING_2018 >= 25, "Y", "N"),
         SNP_big_three = ifelse(MAC_PKLE_2019 >= 25 & 
                                        MAC_PKLE_2018 >= 25 & 
                                        MAC_CLMB_2019 >= 25 & 
                                        MAC_CLMB_2018 >= 25 & 
                                        MAC_KBSM_2019 >= 25 & 
                                        MAC_KBSM_2018 >= 25, "Y", "N"),
         SNP_fourway_seven = ifelse(MAC_PKLE_2019 >= 25 & 
                                        MAC_PKLE_2018 >= 25 & 
                                        MAC_LINC_2019 >= 25 & 
                                        MAC_LINC_2018 >= 25 & 
                                        MAC_STIL_2019 >= 25 & 
                                        MAC_STIL_2018 >= 25 & 
                                        MAC_TMPL_2019 >= 25 & 
                                        MAC_TMPL_2018 >= 25 & 
                                        MAC_CLMB_2019 >= 25 & 
                                        MAC_CLMB_2018 >= 25 & 
                                        MAC_BRKG_2019 >= 25 & 
                                        MAC_BRKG_2018 >= 25 & 
                                        MAC_KBSM_2019 >= 25 & 
                                        MAC_KBSM_2018 >= 25, "Y", "N"),
         SNP_microbiome = ifelse(MAC_PKLE_2019 >= 25 & 
                                        MAC_PKLE_2018 >= 25, "Y", "N"),
         SNP_missing_frz = ifelse((MAC_FRMI_2018 >= 25 & MAC_BRKG_2018 >= 25 &
                                     MAC_LINC_2018 >= 25 & 
                                     MAC_CLMB_2018 >= 25) &
                                     (MAC_LINC_2019 < 25 | 
                                        MAC_CLMB_2019 < 25 | 
                                        MAC_BRKG_2019 < 25 |
                                    MAC_FRMI_2019 < 15), "Y", "N")
         )

snp_subsets %>%
  dplyr::select(CHR, POS, starts_with("SNP_")) %>%
  saveRDS(., file = "SNP_subsets_33M_markers.rds")

missing_frz <- snp_subsets %>%
  filter(SNP_missing_frz == "Y")

snp_subsets %>%
  filter(SNP_missing_frz == "Y") %>%
  saveRDS(., file = "Missing_SNPs_2019_4_freezing_sites.rds")

snp_subsets %>%
  group_by(SNP_missing_frz) %>% tally()
snp_subsets %>%
  group_by(SNP_tensite_twoyear, SNP_big_three, SNP_fourway_seven, SNP_microbiome) %>% tally()
```

```{r}
missing_frz <- readRDS(file = "Missing_SNPs_2019_4_freezing_sites.rds")
```

```{r}
theme_set(theme_poster)
missing_frz %>%
  ggplot(aes(x = POS, y = avediff)) + 
  geom_point() + 
  facet_wrap(~CHR)
save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites.png", plot = last_plot())

missing_frz %>%
  ggplot(aes(x = POS, y = avediff)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss, 4 sites") +
  theme(legend.position = "right")
save_plot(filename = "Distribution_of_SNPs_lost_in_freezing_four_sites_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_FRMI_2018 >= 25 & MAC_LINC_2018 >= 25 & MAC_CLMB_2018 >= 25 &
           MAC_BRKG_2018 >= 25 & MAC_FRMI_2019 < 15 & MAC_LINC_2019 < 25 & 
           MAC_CLMB_2019 < 25 & MAC_BRKG_2019 < 25) %>%
  ggplot(aes(x = POS, y = avediff)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss, 4 sites") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_all_four_freezing_sites_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_BRKG_2018 >= 25 & MAC_BRKG_2019 < 15) %>%
  mutate(Allele_loss = MAC_BRKG_2018 - MAC_BRKG_2019) %>%
  ggplot(aes(x = POS, y = Allele_loss)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_BRKG_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_LINC_2018 >= 25 & MAC_LINC_2019 < 15) %>%
  mutate(Allele_loss = MAC_LINC_2018 - MAC_LINC_2019) %>%
  ggplot(aes(x = POS, y = Allele_loss)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_LINC_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_FRMI_2018 >= 25 & MAC_FRMI_2019 < 15) %>%
  mutate(Allele_loss = MAC_FRMI_2018 - MAC_FRMI_2019) %>%
  ggplot(aes(x = POS, y = Allele_loss)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_FRMI_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz %>%
  filter(MAC_CLMB_2018 >= 25 & MAC_CLMB_2019 < 15) %>%
  mutate(Allele_loss = MAC_CLMB_2018 - MAC_CLMB_2019) %>%
  ggplot(aes(x = POS, y = Allele_loss)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_lost_in_CLMB_log10.png", plot = last_plot(), base_asp = 4, base_height = 4)

missing_frz <- missing_frz %>%
  mutate(Loss_LINC = MAC_LINC_2018 - MAC_LINC_2019,
         Loss_BRKG = MAC_BRKG_2018 - MAC_BRKG_2019,
         Loss_FRMI = MAC_FRMI_2018 - MAC_FRMI_2019,
         Loss_CLMB = MAC_CLMB_2018 - MAC_CLMB_2019)

missing_frz %>%
  group_by(Loss_BRKG > 75, Loss_LINC > 78, Loss_FRMI > 54, Loss_CLMB > 51) %>%
  tally() %>%
  write_csv(., path = "SNPs_loss_overlaps_1percent_tail_2019_four_freezing_sites.csv")

plot1 <- missing_frz %>%
  ggplot(aes(x = Loss_CLMB)) + geom_histogram(binwidth = 2)
save_plot(filename = "Histogram_SNPs_lost_CLMB.png", plot = plot1)
plot1 <- missing_frz %>%
  ggplot(aes(x = Loss_FRMI)) + geom_histogram(binwidth = 2)
save_plot(filename = "Histogram_SNPs_lost_FRMI.png", plot = plot1)

plot1 <- missing_frz %>%
  ggplot(aes(x = Loss_BRKG)) + geom_histogram(binwidth = 2)
save_plot(filename = "Histogram_SNPs_lost_BRKG.png", plot = plot1)

plot1 <- missing_frz %>%
  ggplot(aes(x = Loss_LINC)) + geom_histogram(binwidth = 2)
save_plot(filename = "Histogram_SNPs_lost_LINC.png", plot = plot1)

missing_frz %>%
  summarise(top_LINC = quantile(Loss_LINC, 0.99),
            top_BRKG = quantile(Loss_BRKG, 0.99),
            top_FRMI = quantile(Loss_FRMI, 0.99),
            top_CLMB = quantile(Loss_CLMB, 0.99),
            )

# LINC & BRKG are the most similar

all_frz <- snp_subsets %>%
  mutate(Loss_LINC = (MAC_LINC_2018 - MAC_LINC_2019)/MAC_LINC_2018,
         Loss_BRKG = (MAC_BRKG_2018 - MAC_BRKG_2019)/MAC_BRKG_2018,
         Loss_FRMI = (MAC_FRMI_2018 - MAC_FRMI_2019)/MAC_FRMI_2018,
         Loss_CLMB = (MAC_CLMB_2018 - MAC_CLMB_2019)/MAC_CLMB_2018)
all_frz %>%
  summarise(top_LINC = quantile(Loss_LINC, 0.99, na.rm = TRUE),
            top_BRKG = quantile(Loss_BRKG, 0.99, na.rm = TRUE),
            top_FRMI = quantile(Loss_FRMI, 0.99, na.rm = TRUE),
            top_CLMB = quantile(Loss_CLMB, 0.99, na.rm = TRUE)
            )
all_frz %>%
  group_by(Loss_BRKG > 159, Loss_LINC > 179, Loss_FRMI > 84, Loss_CLMB > 94) %>%
  tally() %>%
  write_csv(., path = "All_SNPs_lost_alleles_1percent_tail_2019_four_freezing_sites.csv")

missing_frz %>%
  filter(Loss_BRKG > 75 | Loss_LINC > 78 | Loss_FRMI > 54 | Loss_CLMB > 51) %>% saveRDS(., file = "Lost_SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
all_frz %>%
  filter(Loss_BRKG > 159 | Loss_LINC > 179 | Loss_FRMI > 84 | Loss_CLMB > 94) %>% saveRDS(., file = "SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
```

  top_LINC top_BRKG top_FRMI top_CLMB
     <dbl>    <dbl>    <dbl>    <dbl>
1       78       75       54       51

# Frequencies lost 
```{r}
frz_freq <- readRDS("SNPs_with_1percent_tail_frequency_change_freezing_sites.rds") 
loss_freq <- readRDS("Lost_SNPs_with_1percent_tail_frequency_change_freezing_sites.rds")
```

```{r}
loss_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_LINC/MAC_LINC_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_LINC.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_LINC/MAC_LINC_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_LINC.png", plot = last_plot(), base_asp = 4, base_height = 4)
```

```{r}
loss_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_CLMB/MAC_CLMB_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_CLMB.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_CLMB/MAC_CLMB_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_CLMB.png", plot = last_plot(), base_asp = 4, base_height = 4)

loss_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_FRMI/MAC_FRMI_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_FRMI.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_FRMI/MAC_FRMI_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_FRMI.png", plot = last_plot(), base_asp = 4, base_height = 4)

loss_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_BRKG/MAC_BRKG_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_lost_in_BRKG.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
ggplot(aes(x = POS/1000000, y = Loss_BRKG/MAC_BRKG_2018)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_BRKG.png", plot = last_plot(), base_asp = 4, base_height = 4)
```

```{r}
loss_freq %>%
    filter(Loss_BRKG > 75 | Loss_LINC > 78 | Loss_FRMI > 54 | 
             Loss_CLMB > 51) %>% 
  mutate(Loss_freq = (Loss_BRKG/MAC_BRKG_2018 + Loss_FRMI/MAC_FRMI_2018 +
                            Loss_LINC/MAC_LINC_2018 +
                            Loss_CLMB/MAC_CLMB_2018)/4) %>% 
  ggplot(aes(x = POS/1000000, y = Loss_freq)) + 
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_one_lost_all_four_sites.png", plot = last_plot(), base_asp = 4, base_height = 4)
frz_freq %>%
  filter(Loss_BRKG > 159 | Loss_LINC > 179 | Loss_FRMI > 84 | 
           Loss_CLMB > 94) %>%
  mutate(Loss_freq = (Loss_BRKG/MAC_BRKG_2018 + Loss_FRMI/MAC_FRMI_2018 +
                            Loss_LINC/MAC_LINC_2018 +
                            Loss_CLMB/MAC_CLMB_2018)/4) %>% 
  ggplot(aes(x = POS/1000000, y = Loss_freq)) +
  geom_hex() + 
  facet_wrap(~CHR, scales = "free_x", nrow = 1) + 
  scale_fill_viridis(trans = "log10") + 
  xlab("Position (Mb)") + ylab("Minor Allele Loss") +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "Distribution_of_SNPs_1per_delta_all_four_sites.png", plot = last_plot(), base_asp = 4, base_height = 4)
```