### 2020-03-06 R/Fast_QQ_plot.Rmd

---
title: "Fast QQ plots"
author: "Alice MacQueen"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(AnnotationDbi)
library(multtest)
library(viridis)
library(cowplot)
library(data.table)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")

NCORES <- nb_cores()
```

##

```{r}
gwas_data2 <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/all-phenotypes-two-years/GWAS_datatable_KBSM_GR50_3_PCs_.rds")

```

# Faster QQ plotting

```{r}
round2 <- function(x, at) ceiling(x / at) * at
round_xy <- function(x, y, cl = NA, cu = NA, roundby = 0.001){
  expected <- round2(x, at = roundby)
  observed <- round2(y, at = roundby)
  if(!is.na(cl[1]) & !is.na(cu[1])){
    clower <- round2(cl, at = roundby)
    cupper <- round2(cu, at = roundby)
    tp <- cbind(expected, observed, clower, cupper)
    return(tp[!duplicated(tp),])
  } else {
    tp <- cbind(expected, observed)
    return(tp[!duplicated(tp),])
  }
}

  # Save a QQ plot
  ggsave(snp_qq(gwas[which(!is.na(gwas$score)),]),
         filename = paste0("QQplot_", names(df1)[2], "_", PCdf1$NumPCs,
                           "_PCs.png"), width = 4,
         height = 4, path = file.path("/", "home", "alice", "Github",
                                      "pvdiv-phenology-gxe", "analysis",
                                      "all-phenotypes-two-years"))
```

```{r}
#' Create a quantile-quantile plot with ggplot2.
#'
#' Assumptions:
#'   - Expected P values are uniformly distributed.
#'   - Confidence intervals assume independence between tests.
#'     We expect deviations past the confidence intervals if the tests are
#'     not independent.
#'     For example, in a genome-wide association study, the genotype at any
#'     position is correlated to nearby positions. Tests of nearby genotypes
#'     will result in similar test statistics.
#'
#' @param ps Vector of p-values.
#' @param ci Size of the confidence interval, 95% by default.
#' @param lambdaGC Logical. Add the Genomic Control coefficient as subtitle to the plot?
#' @return A ggplot2 plot.
#' @examples
#' library(ggplot2)
#' gg_qqplot(runif(1e2)) + theme_grey(base_size = 24)
gg_qqplot <- function(ps, ci = 0.95, lambdaGC = FALSE) {
  n  <- length(ps)
  df <- data.frame(
    observed = -log10(sort(ps)),
    expected = -log10(ppoints(n)),
    clower   = -log10(qbeta(p = (1 - ci) / 2, shape1 = 1:n, shape2 = n:1)),
    cupper   = -log10(qbeta(p = (1 + ci) / 2, shape1 = 1:n, shape2 = n:1))
  )
  df_round <- round_xy(df$expected, df$observed, cl = df$clower, cu = df$cupper)
  log10Pe <- expression(paste("Expected -log"[10], plain("("), italic(p-value),
                              plain(")")))
  log10Po <- expression(paste("Observed -log"[10], plain("("), italic(p-value),
                              plain(")")))
  p1 <- ggplot(as_tibble(df_round)) +
    geom_point(aes(expected, observed), shape = 1, size = 1) +
    geom_abline(intercept = 0, slope = 1, size = 1.5, color = "red") +
    geom_line(aes(expected, cupper), linetype = 2) +
    geom_line(aes(expected, clower), linetype = 2) +
    xlab(log10Pe) +
    ylab(log10Po)

  if (lambdaGC) {
      xtr <- log10(ps)
      MEDIAN <- log10(0.5)
      f.opt <- function(x) (x - MEDIAN)
      xtr_p <- stats::median(xtr) / stats::uniroot(f.opt, interval = range(xtr),
                                                   check.conv = TRUE, 
                                                   tol = tol)$root
      lamGC <- signif(xtr_p)
      expr <- substitute(expression(lambda[GC] == l), list(l = lamGC))
      p1 + labs(subtitle = eval(expr))
    } else {
      p1
    }
}

  xtr <- -stats::na.omit(gwas$log10p)
  MEDIAN <- log10(0.5)
  f.opt <- function(x) (x - MEDIAN)
  stats::median(xtr) / stats::uniroot(f.opt, 
                                              interval = range(xtr),
                                      check.conv = TRUE, tol = tol)$root

  # Save a QQ plot
  ggsave(gg1,
         filename = paste0("QQplot_", "test2", ".png"), width = 4,
         height = 4, path = file.path("/", "home", "alice", "Github",
                                      "pvdiv-phenology-gxe", "analysis",
                                      "all-phenotypes-two-years"))
```