### 2020-03-24 R/Make_kinship_van_Raden.Rmd
Make van Raden Kinship matrices for switchgrass

---
title: "Make Kinship Van Raden"
author: "Alice MacQueen"
date: "3/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
```

##

In the GAPIT numerical object, columns are SNPs and rows are individuals. It's the same for bigsnpr objects, luckily.

# Making a kinship matrix using the van Raden method.

If you don't need to subset by individuals or SNPs, you can use this function as-is. Otherwise see the code in the next section for options for subsetting.

```{r}
snp <- snp_attach("Pvirgatum_V5_GWAS_630g_33M_tensite_twoyear.rds")

pvdiv_kinship_van_raden <- function(snp){
  G <- snp$genotypes
  nInd <- snp$genotypes$nrow
  # Centered (mean of rows subtracted) transposed crossproduct of snp file.
  K <- big_tcrossprodSelf(G, fun.scaling = big_scale(center = TRUE, 
                                                                 scale = FALSE))

  #Extract diagonals
  i = 1:nInd
  j=(i-1)*nInd
  index=i+j
  d=K[index]
  DL=min(d)
  DU=max(d)
  floor=min(K[i,i])

  K=(K[i,i]-floor)/(DL-floor)
  MD=(DU-floor)/(DL-floor)

  rownames(K) <- snp$fam$sample.ID
  colnames(K) <- snp$fam$sample.ID

  if(MD>2){
    K[index]=K[index]/(MD-1)+1
  }
  return(K)
}
```

# Subset to just single ecotypes, and to just single genetic subpopulations.

This method could be used to remove any individuals that don't meet specific criteria.

```{r}
metadata <- read_rds("~/Github/pvdiv-phenotypes/data/metadata.rds")

upeco <- metadata[which(metadata$ECOTYPE_NNET == "Upland"),]$PLANT_ID
loeco <- metadata[which(metadata$ECOTYPE_NNET == "Lowland"),]$PLANT_ID
coeco <- metadata[which(metadata$ECOTYPE_NNET == "Coastal"),]$PLANT_ID

mgeno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Midwest"),]$PLANT_ID
ageno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Atlantic"),]$PLANT_ID
ggeno <- metadata[which(metadata$GENETIC_SUBPOP_50PER == "Gulf"),]$PLANT_ID

upland_eco <- which(snp$fam$sample.ID %in% upeco)
lowland_eco <- which(snp$fam$sample.ID %in% loeco)
coastal_eco <- which(snp$fam$sample.ID %in% coeco)
midwest_geno <- which(snp$fam$sample.ID %in% mgeno)
atlantic_geno <- which(snp$fam$sample.ID %in% ageno)
gulf_geno <- which(snp$fam$sample.ID %in% ggeno)

subsets <- list(upland_eco = upland_eco, lowland_eco = lowland_eco, coastal_eco = coastal_eco, midwest_geno = midwest_geno, atlantic_geno = atlantic_geno, gulf_geno = gulf_geno)

for(q in seq_along(subsets)){
    ind.row <- subsets[[q]]
    G <- snp$genotypes
    nInd <- length(ind.row)
    # Centered (mean of rows subtracted) transposed crossproduct of snp file.
    K <- big_tcrossprodSelf(G, ind.row = ind.row, 
                            fun.scaling = big_scale(center = TRUE, 
                                                    scale = FALSE))

    #Extract diagonals
    i = 1:nInd
    j=(i-1)*nInd
    index=i+j
    d=K[index]
    DL=min(d)
    DU=max(d)
    floor=min(K[i,i])

    K2=(K[i,i]-floor)/(DL-floor)
    MD=(DU-floor)/(DL-floor)
    # Name rows and columns
    rownames(K2) <- snp$fam$sample.ID[ind.row]
    colnames(K2) <- snp$fam$sample.ID[ind.row]

    if(MD>2){
      K2[index]=K2[index]/(MD-1)+1
    }

  write_rds(K2, paste0("Kinship_van_Raden_method_", nInd, "_individuals_", names(subsets)[q], "type.rds")) 
}
```

Other bigstatsr commands that I didn't end up using.
`snpSum <- big_colstats(G)# GAPIT: apply(snps,2,mean) Take the mean over columns.`
`snpMean <- snpSum$sum / nInd # bigsnpr has a function that can center like this.`

# GAPIT's method, for comparison purposes
```{r}
`GAPIT.kinship.VanRaden` <-
function(snps,hasInbred=TRUE) {
# Object: To calculate the kinship matrix using the method of VanRaden (2009, J. Dairy Sci. 91:4414???C4423)
# Authors: Zhwiu Zhang
# Last update: August 15, 2011 
##############################################################################################

print("Calculating kinship with VanRaden method...")

#snps=hm$GD 
nSNP=ncol(snps)
nInd=nrow(snps)
n=nInd

snpMean= apply(snps,2,mean)
snps=snps-snpMean
print("Getting X'X...")
K=tcrossprod((snps), (snps))

print("Adjusting...")
#Extract diagonals
i =1:n
j=(i-1)*n
index=i+j
d=K[index]
DL=min(d)
DU=max(d)
floor=min(K)

K=(K-floor)/(DL-floor)
MD=(DU-floor)/(DL-floor)

#Handler of diagonals over 2
#print("MD")
#print(MD)
#print(K[1:5,1:5])

if(is.na(K[1,1])) stop ("GAPIT says: Missing data is not allowed for numerical genotype data")
if(MD>2)K[index]=K[index]/(MD-1)+1

#Handler of inbred
if(MD<2 & hasInbred) K=2*K/((DU-floor)/(DL-floor))

print("Calculating kinship with VanRaden method: done")

return(K)
}
```