### 2020-04-11 R/D2F_annotations.Rmd
Analysis for three most compelling flowering time candidate genes

---
title: "Annos_D2F"
author: "Alice MacQueen"
date: "4/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
```

##

```{r}
anno_all <- pvdiv_results_in_folder(path = "../analysis/gwas/", pattern = "Annotation")
anno_d2f <- read_delim("../analysis/gwas/D2Fannos.txt", delim = " ", col_names = "filename")

annos_fl <- anno_all %>% 
  enframe(name = NULL, value = "filename") %>%
  filter(!grepl("GR50", filename) & !grepl("SEN50", filename))
```
# Top 100 or so overlaps, not filtering to fourway hits
```{r}
inputdir <- "~/Github/pvdiv-phenology-gxe/analysis/gwas/"
i=1
  annos <- read_rds(paste0(inputdir, annos_fl$filename[i]))
  anno_overlap100 <- annos$top500SNPs_within50000bp

  anno_overlap100$Anno_descr <- paste0(str_sub(annos_fl$filename[i], start = 19, 
                                          end = -12))

for(i in 2:nrow(annos_fl)){
  annos <- read_rds(paste0(inputdir, annos_fl$filename[i]))
  anno_overlap1001 <- annos$top500SNPs_within50000bp

    anno_overlap1001$Anno_descr <- str_sub(annos_fl$filename[i], start = 19, end = -12)
  anno_overlap100 <- anno_overlap100 %>%
    dplyr::select(-TXID) %>%
    full_join(anno_overlap1001)

}
anno_overlap100 <- anno_overlap100 %>%
    separate(Anno_descr, into = c("SITE", "PHE", "type", "genos", "nSNP", 
                                  "SNP"), sep = "_") %>%
    mutate(genos = str_sub(genos, start = 6, end = -2),
           subpop = case_when(nSNP == "11.5M" ~ "Atlantic_Gulf",
                              nSNP == "8.8M" ~ "Midwest",
                              nSNP == "12.3M" ~ "Atlantic_Midwest",
                              nSNP == "12.1M" ~ "Gulf_Midwest",
                              nSNP == "10.3M" ~ "Gulf",
                              nSNP == "10.6M" ~ "Atlantic",
                              TRUE ~ "Other"),
           Atlantic = case_when(grepl("Atlantic", subpop) ~ 1,
                                TRUE ~ 0),
           Midwest = case_when(grepl("Midwest", subpop) ~ 1,
                                TRUE ~ 0),
           Gulf = case_when(grepl("Gulf", subpop) ~ 1,
                                TRUE ~ 0)
           )

anno_overlap100 %>% 
  dplyr::select(SITE, PHE, subpop:Gulf, everything()) %>%
  mutate(group = ceiling(POS/1000000)) %>%
  group_by(CHR, group, SITE) %>%
  tally() %>% tally() %>% arrange(desc(n), CHR, group)

anno_overlap100 %>% 
  dplyr::select(SITE, PHE, subpop:Gulf, everything()) %>%
  filter(Atlantic == 0 & Midwest == 1) %>%
  mutate(group = round(POS/1000000, digits = 1)) %>%
  group_by(CHR, group, SITE) %>%
  tally() %>% tally() %>% filter(n >= 5) %>% arrange(CHR, group)

anno_overlap100 %>%
  dplyr::select(PHE, SITE, subpop, POS, everything()) %>%
  mutate(group = round(POS/1000000, digits = 1),
         log10p = -log10(`p value`)) %>%
  filter((CHR %in% c("Chr07K") & group %in% c(10.4)) | (CHR %in% c("Chr07N") & group %in% c(45.7))) %>%
  dplyr::select(CHR, group, log10p, `SNP Effect`, PHE, SITE, `Gene ID`, everything()) %>%
  arrange(CHR, group)
```
Chr07K	10.4	5	 	Pavir.7KG042400	 PKLE LINC CLMB FRMI KBSM KING   	Encodes a cold-activated plasma membrane protein cold-responsive protein kinase that phosphorylates 14-3-3 proteins. The phosphorylated 14-3-3 proteins shuttle from the cytosol to the nucleus, where they interact with and destabilize the key cold-responsive C-repeat-binding factor (CBF) proteins, modulate CBF stability and the response to cold stress. ~9-11d
Chr07N	45.7	5	 LINC STIL TMPL KBSM CLMB ~6-8d 
Pavir.7NG394500 (expressed protein in rice)
Pavir.7NG399825

##----------------------------
KBSM & CLMB Chr07N 45.7 ~15d KBSM ~8d CLMB Gulf_Midwest	Pavir.7NG394600	AT1G30620
HIGH SUGAR RESPONSE8, HSR8, MUR4, MURUS 4, UDP-D-XYLOSE 4-EPIMERASE 1, UXE1
Description (?) 	  	encodes a type-II membrane protein that catalyzes 4-epimerization of UDP-D-Xylose to UDP-L-Arabinose in vitro, the nucleotide sugar used by glycosyltransferases in the arabinosylation of cell wall polysaccharides and wall-resident proteoglycans.
Highly expressed in axis of inflorescence, flower, carpel, sensecent leaf

# FL50 & D2F annos 2020-04-10

As FL50 and D2F QTL regions are basically the same, just a few more FL50 QTL regions, and as the FL50 & D2F GWAS look quite similar in most cases, I will just compare annotations for all the flowering results here.
```{r}
inputdir <- "~/Github/pvdiv-phenology-gxe/analysis/gwas/"
i=1
  annos <- read_rds(paste0(inputdir, annos_fl$filename[i]))
  anno_overlap <- annos$top500SNPs_within50000bp %>%
  filter((CHR == "Chr02K" & between(POS, 63687000, 65490000)) | 
           (CHR == "Chr02N" & between(POS, 58030000, 61280000)) | 
           (CHR == "Chr03K" & between(POS, 17360000, 26180000)) | 
           (CHR == "Chr04K" & between(POS, 1950000, 5500000)) | 
           (CHR == "Chr05N" & between(POS, 20000, 4330000)) | 
           (CHR == "Chr05N" & between(POS, 58240000, 66000000)) | 
           (CHR == "Chr08N" & between(POS, 550000, 34420000)) | 
           (CHR == "Chr09K" & between(POS, 13650000, 18180000)))

  anno_overlap$Anno_descr <- paste0(str_sub(annos_fl$filename[i], start = 19, 
                                          end = -12))

for(i in 2:nrow(annos_fl)){
  annos <- read_rds(paste0(inputdir, annos_fl$filename[i]))
  anno_overlap1 <- annos$top500SNPs_within50000bp %>%
    dplyr::select(-TXID) %>%
  filter((CHR == "Chr02K" & between(POS, 63687000, 65490000)) | 
           (CHR == "Chr02N" & between(POS, 58030000, 61280000)) | 
           (CHR == "Chr03K" & between(POS, 17360000, 26180000)) | 
           (CHR == "Chr04K" & between(POS, 1950000, 5500000)) | 
           (CHR == "Chr05N" & between(POS, 20000, 4330000)) | 
           (CHR == "Chr05N" & between(POS, 58240000, 66000000)) | 
           (CHR == "Chr08N" & between(POS, 550000, 34420000)) | 
           (CHR == "Chr09K" & between(POS, 13650000, 18180000)))

    anno_overlap1$Anno_descr <- str_sub(annos_fl$filename[i], start = 19, end = -12)
  anno_overlap <- anno_overlap %>%
    full_join(anno_overlap1)

}
anno_overlap <- anno_overlap %>%
    separate(Anno_descr, into = c("SITE", "PHE", "type", "genos", "nSNP", 
                                  "SNP"), sep = "_") %>%
    mutate(genos = str_sub(genos, start = 6, end = -2),
           subpop = case_when(nSNP == "11.5M" ~ "Atlantic_Gulf",
                              nSNP == "8.8M" ~ "Midwest",
                              nSNP == "12.3M" ~ "Atlantic_Midwest",
                              nSNP == "12.1M" ~ "Gulf_Midwest",
                              nSNP == "10.3M" ~ "Gulf",
                              nSNP == "10.6M" ~ "Atlantic",
                              TRUE ~ "Other"),
           Atlantic = case_when(grepl("Atlantic", subpop) ~ 1,
                                TRUE ~ 0),
           Midwest = case_when(grepl("Midwest", subpop) ~ 1,
                                TRUE ~ 0),
           Gulf = case_when(grepl("Gulf", subpop) ~ 1,
                                TRUE ~ 0)
           )

anno_overlap %>% 
  dplyr::select(SITE, PHE, subpop:Gulf, everything()) %>%
  mutate(group = ceiling(POS/1000000)) %>%
  group_by(CHR, group, SITE) %>%
  tally() %>% tally() %>% arrange(desc(n), CHR, group)

anno_overlap %>% 
  dplyr::select(SITE, PHE, subpop:Gulf, everything()) %>%
  filter(Atlantic == 0 & Midwest == 1) %>%
  mutate(group = round(POS/1000000, digits = 1)) %>%
  group_by(CHR, group, SITE) %>%
  tally() %>% tally() %>% filter(n >= 3) %>% arrange(CHR, group)

anno_overlap %>% 
  mutate(group = round(POS/1000000, digits = 1)) %>%
  filter(CHR == "Chr08N" & group %in% c(0.8, 15.6, 27.7)) %>%
  dplyr::select(CHR, POS, subpop, `SNP Effect`, PHE, SITE, `Gene ID`, everything()) %>%
  arrange(CHR, POS)
```
Chr08N 0.8Mb in Gulf Midwest LINC TMPL STIL BRKG (colder) ~13 - 14d
Pavir.8NG355401 AT3G05550 Hypoxia responsive family protein only sort of highly expressed in early flower
Pavir.8NG355404 AT2G30330 Protein - protein interaction with BLOS2 and also with SNX1.Located in endomembrane system and hypothesized to be involved in endomembrane transport.
Pavir.8NG355405	AT5G05340 Pavir.8NG355404 Homolog to a peroxidase superfamily protein highly expressed only in senescent leaf petioles.
Pavir.8NG002300	AT5G03610 highly expressed in axis of inflorescence  	GDSL-motif esterase/acyltransferase/lipase. Enzyme group with broad substrate specificity that may catalyze acyltransfer or hydrolase reactions with lipid and non-lipid substrates.

I gotta say, all of these hits I've been looking at, I've been shocked by the number of hits that have very high gene expression in the axis of inflorescence (in Arabidopsis). This is almost certainly a good sign, given that I've been looking at flowering hits!

Chr02N	59	8	CLMB BRKG TMPL PKLE STIL ... CLMB ... KBSM ... LINC FRMI KBSM
Pavir.2NG496800 58.19  AT4G25230 RIN2 RPM1 interacting protein. Positive regulator of RPM1 and PRS2 mediated hypersensitive response.Functions as ubiquitin ligase and binds to RPM1. Could certainly hasten flowering if it increases a basal defense response (MacQueen et al 2016).

KBSM STIL?
Pavir.2NG491000 58.61Mb AT2G35660 CTF2A: Encodes a member of a novel gene family with homology to known proteins involved in hydroxylation and oxidation of an aromatic ring. Highly expressed in sepals, pedicel, leaves, and siliques.

Chr02N	59	8	
Chr03K	24	8	
Chr03K	26	9	
Chr04K	3	8		
Chr04K	4	8	
Chr05N	2	9	
Chr05N	3	8	
Chr05N	4	9	
Chr05N	60	8	
Chr05N	64	9		
Chr05N	65	9		
Chr05N	66	9		
Chr08N	10	9		
Chr08N	13	9

Without Atlantic group:
Chr05N	65	9		
Chr02N	59	7		
Chr05N	4	7		
Chr05N	66	7

Chr08N	0.8	4		
Chr08N	15.6	4		
Chr08N	27.7	4

Without Midwest group:
Chr03K	26	8		
Chr05N	2	8		
Chr05N	4	8		
Chr08N	10	8		
Chr04K	3	7		
Chr05N	3	7		
Chr05N	65	7

Without Gulf group:
Chr08N	3	6

With Gulf group:
Chr03K	26	8		
Chr04K	4	8		
Chr05N	2	8		
Chr05N	3	7		
Chr05N	4	9		
Chr05N	62	7		
Chr05N	64	7		
Chr05N	65	9		
Chr05N	66	7		
Chr08N	10	8	
Chr08N	13	7

With Midwest group:
Chr02N	59	7		
Chr03K	24	7		
Chr03K	26	7		
Chr05N	4	7		
Chr05N	5	7		
Chr05N	64	7		
Chr05N	65	8		
Chr05N	66	9		
Chr08N	13	7

Anything by 100Kb interval
Chr03K	24.9	5		
Chr04K	3.7	5		
Chr05N	61.4	5		
Chr05N	65.9	5		
Chr08N	15.6	5		
Chr08N	17.7	5		
Chr08N	33.1	5

Midwest & Gulf
Chr05N	4.3	3		
Chr08N	0.8	4		
Chr08N	15.6	3		
Chr08N	27.7	3

Gulf, no Atlantic
Chr08N	0.8	4		
Chr08N	15.6	4		
Chr08N	27.7	4

Midwest, no Atlantic
Chr05N	4.3	3		
Chr05N	59.1	3		
Chr08N	0.8	4		
Chr08N	15.6	3		
Chr08N	27.7	3

Other intervals to look in: 
Ch05N 4.3
Chr05N 59.1
Chr05N 61.4
Chr08N 0.8
Chr08N 15.6
Chr08N 27.7

```{r}
anno_overlap %>% 
  dplyr::select(PHE, SITE, subpop, POS, everything()) %>%
  mutate(group = round(POS/1000000, digits = 1),
         log10p = -log10(`p value`)) %>%
  filter((CHR %in% c("Chr05N") & group %in% c(4.3, 59.1, 61.4)) | (CHR %in% c("Chr08N") & group %in% c(0.8, 15.6, 27.7))) %>%
  dplyr::select(CHR, group, log10p, `SNP Effect`, PHE, SITE, `Gene ID`, everything()) %>%
  arrange(CHR, group)

anno_overlap %>% 
  dplyr::select(PHE, SITE, subpop, POS, everything()) %>%
  mutate(group = round(POS/1000000, digits = 1),
         log10p = -log10(`p value`)) %>%
  filter((CHR %in% c("Chr03K") & group %in% c(24.9)) | (CHR %in% c("Chr04K") & group %in% c(3.7))) %>%
  dplyr::select(CHR, group, log10p, `SNP Effect`, PHE, SITE, `Gene ID`, everything()) %>%
  arrange(CHR, group)

```
Chr03K 24.9 CLMB FRMI KBSM PKLE STIL probably Atlantic
	AT1G34340 mature & senescent leaf; senescent flower
Chr04K 3.7 BRKG CLMB PKLE STIL TMPL Atlantic or Gulf ~5d effect Pavir.4KG031100 homolog to AT1G71696 which is highly expressed in root, root apical meristem, axis of inflorescence, carpel, and senescing flowers. sol1 partially suppresses the short root phenotype caused by CLE19 overexpression.

5N 4.3 Mb: TMPL, STIL, KING, CLMB. In populations with Midwest plants.
Pavir.5N036700 (oxidation related zinc finger), to Pavir.5N037500
Best candidate:
Pavir.5NG037100. in A. thaliana, highly expressed in the axis of inflorescence and in dividing cells.
Plant Aurora kinases play a role in the maintenance of primary meristems and control of endoreduplication (Petrovska et al 2011 New Phytologist)

Chr05N 61.4 Mb KING & BRKG (LINC & PKLE & TMPL close) Pavir.5NG557850 is a
homolog to AT3G17450 which is a hAT dimerization domain-containing protein that is highly expressed in the axis of inflorescence, the carpel of developing and mature flowers, and early seeds.
AT4G10360 (less likely though is expressed in pollen germination and tube growth).

Chr08N 0.8Mb LINC STIL TMPL BRKG Gulf & Midwest. ~13d effect at all sites but BRKG. Pavir.8NG355404 Homolog to a peroxidase superfamily protein highly expressed only in senescent leaf petioles.

Chr08N 15.6Mb PKLE LINC KING KBSM CLMB 14-18 d effect. Mostly in Atlantic subpop.
Pavir.8NG252400 (Gulf effect probably)

Chr08N 27.7Mb CLMB LINC STIL TMPL. Gulf effect probably.

# D2F annos 2020-04-08
```{r}
inputdir <- "~/Github/pvdiv-phenology-gxe/analysis/gwas/"
i=1
  annos <- read_rds(paste0(inputdir, anno_d2f$filename[i]))
  anno_overlap <- annos$top500SNPs_within0bp %>%
  filter((CHR == c("Chr02K") & between(POS, 63687000, 64420000)) | (CHR == "Chr02N" & between(POS, 60250000, 61280000)) | (CHR == "Chr04K" & between(POS, 1950000, 5450000)) | (CHR == "Chr05N" & between(POS, 20000, 4330000)) | (CHR == "Chr05N" & between(POS, 58910000, 65260000)))
  anno_overlap$SITE_PHE <- str_sub(anno_d2f$filename[i], start = 19, end = 26)

for(i in 2:nrow(anno_d2f)){
  annos <- read_rds(paste0(inputdir, anno_d2f$filename[i]))
  anno_overlap1 <- annos$top500SNPs_within0bp %>%
  filter((CHR == c("Chr02K") & between(POS, 63687000, 64420000)) | (CHR == "Chr02N" & between(POS, 60250000, 61280000)) | (CHR == "Chr04K" & between(POS, 1950000, 5450000)) | (CHR == "Chr05N" & between(POS, 20000, 4330000)) | (CHR == "Chr05N" & between(POS, 58910000, 65260000)))
    anno_overlap1$SITE_PHE <- str_sub(anno_d2f$filename[i], start = 19, end = 26)
  anno_overlap <- anno_overlap %>%
    full_join(anno_overlap1)

}
  anno_overlap %>% 
    mutate(group = ceiling(POS/1000000)) %>%
    group_by(CHR, group, SITE_PHE) %>%
    tally() %>% arrange(CHR, group)
anno_overlap %>% 
  filter(CHR == "Chr05N" & between(POS, 63800000, 65200000)) %>% arrange(POS)
```
Pavir.5NG107400 is a gene found in the major QTL region in the fourway cross on Chr05N. It is also found in GWAS of Gulf and Midwestern individuals in the diversity panel, and has SNPs that fall within the gene at PKLE (28d effect) & CLMB (11d effect). Pavir.5NG107400 is located on Chr05N at 64.39Mb, and is homologous to ATCPL2, a carboxyl-terminal domain (CTD) phosphatase that regulates plant growth, stress, and auxin responses. In A. thaliana, mutants in CPL2 have early flowering, low fertility, and increased salt sensitivity. In this 100kb region of Chr05N, we also see associations in the top 500 SNPs by p-value for TMPL (13d effect size, 90kb away from this gene), LINC (-13d, 3kb), KING (34d, 61kb).
Overall there are many overlaps on Chr05N between the QTL and GWAS analyses for days to flowering (FL50 - GR50), and overlaps at single sites for Chr02K (TMPL) and Chr04K (KBSM).

2K tmpl
4k KBSM
5N: KBSM, STIL, PKLE, CLMB, KING (early) tmpl LINC (500)
5N (LATE) PKLE CLMB KBSM KING STIL (TOP 500: CLMB KING STIL KBSM TMPL PKLE)
```{r}
fourways
```
Chr02K_63.687871	Chr02K_64.419301    no
Chr02N_60.251689	Chr02N_61.274395    Chr02N	60985754	60985754	
Chr04K_1.959391	Chr04K_5.499253       no
Chr05N_2.1318	Chr05N_4.328965         Chr05N	4568275	4568275
Chr05N_58.916323	Chr05N_65.255869    Chr05N	58930463	58930463 and 11 SNPs to 
Chr05N	64393503	64393503

# D2F
```{r}
annos$FDR0.1_within0bp %>%
  filter(CHR %in% c("Chr02K", "Chr02N", "Chr04K", "Chr05N"))

# Most permissive set of FL50 & D2F hits - seeing as they all basically overlap #
annos$FDR0.1_within0bp %>%
  filter((CHR == c("Chr02K") & between(POS, 63687000, 65490000)) | 
           (CHR == "Chr02N" & between(POS, 58030000, 61280000)) | 
           (CHR == "Chr03K" & between(POS, 17360000, 26180000)) | 
           (CHR == "Chr04K" & between(POS, 1950000, 5500000)) | 
           (CHR == "Chr05N" & between(POS, 20000, 4330000)) | 
           (CHR == "Chr05N" & between(POS, 58240000, 66000000)) | 
           (CHR == "Chr08N" & between(POS, 550000, 34420000)) | 
           (CHR == "Chr09K" & between(POS, 13650000, 18180000)))

annos$FDR0.1_within50000bp %>%
  filter(CHR %in% c("Chr05N") & between(POS, 58830460, 64493505))
```
Chr02N: No gene within 50kb
```{r}
-log10(2.03e-06)
```

```{r}
phe <- Phenotypes_D2F_fourway_project

phe %>%
  dplyr::select(PLANT_ID, ends_with("D2F"), ends_with("FL50"),
                ends_with("GR50")) %>%
  write_rds("../data/Phenotypes_D2F_fourway_project.rds")
```