### 2020-04-24 R/D2F_breakdown_by_site.Rmd
D2F environmental index analysis for two indices and CV of phenotypes

---
title: "Untitled"
author: "Alice MacQueen"
date: "4/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
```

#

```{r}
phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
```

```{r}
lodToPval <-  function(x){
  pchisq(x * (2 * log(10)), df = 1, lower.tail = FALSE) / 2
}
```

```{r}
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(!(SITE %in% c("FRMI_D2F", "TMPL_D2F", "BRKG_D2F"))) %>%
  ggplot(aes(x = D2F)) + 
  geom_histogram() + 
  facet_wrap(~SITE)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("D2F")) %>%
  pivot_longer(cols = ends_with("FL50"), names_to = "SITE", values_to = "FL50") %>%
  #filter(!(SITE %in% c("FRMI_FL50", "TMPL_FL50", "BRKG_FL50"))) %>%
  ggplot(aes(x = FL50)) + 
  geom_histogram() + 
  facet_wrap(~SITE)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F < 30)
```

```{r}
phe[1:15,] %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID))

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE)) %>%
  ggplot(aes(x = SITE, y = mean_D2F)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP))

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(GENETIC_SUBPOP_KINSHIP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = cv_D2F)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 0.02, angle = 45) + ylim(c(0,0.6)) +
  scale_color_viridis(discrete = TRUE, end = 0.8)

?sd()
?var() # - variance (square of sd)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  ggplot(aes(x = D2F)) + 
  geom_histogram(aes(fill = GENETIC_SUBPOP_KINSHIP)) + facet_grid(~SITE)
```
Very similar sites: BRKG, CLMB, LINC, (FRMI)
STIL
KBSM, PKLE
KING
...
TMPL

Let's consider the idea that the Midwest and Gulf populations are receiving extremely different environmental signals at each of the planting locations. What can I use to support this idea? Well, the days to flowering

```{r}
sites
```
From Houle 1992:

Crow 1958: "opportunity for selection": I = V_P / X_bar^2
X_bar represents the population mean before selection.
V_P phenotypic trait variance. `var()`

To calculate CV (coefficient of variation) you take the standard deviation of the data and divide it by the mean of the data.  sqrt(V_P) / X_bar

I is related to the phenotypic coefficient of variation, CV_P = 100*sqrt(V_P) / X_bar
(this is just CV * 100 to be a percentage).

Relative evolvability of fitness:

I_A = V_A / X_bar^2 = (CV_A / 100)^2
I_A is also an appropriate indicator of the variability of fitness as large values must be the result of powerful forces maintaining genetic variance.

A variance to mean ratio is the most appropriate measure of the evolvability of a trait.

Compared three quantities:

Residual coefficient of variation CV_R = 100 * sqrt(V_P - V_A) / X_bar

h^2 narrow sense heritability = V_A / V_P

CV_A = 100*sqrt(V_A) / X_bar

```{r}
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(GENETIC_SUBPOP_KINSHIP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = cv_D2F)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 0.02, angle = 45) + ylim(c(0,0.6)) +
  scale_color_viridis(discrete = TRUE, end = 0.8)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(GENETIC_SUBPOP_KINSHIP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = mean_D2F)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 50, angle = 45) +
  scale_color_viridis(discrete = TRUE, end = 0.8) +
  geom_errorbar(aes(x = Latitude, ymin = mean_D2F - sd_D2F, ymax = mean_D2F + sd_D2F, group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP))

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, GENETIC_SUBPOP_KINSHIP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(GENETIC_SUBPOP_KINSHIP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = sd_D2F^2)) + 
  geom_point(aes(color = GENETIC_SUBPOP_KINSHIP)) +
  geom_line(aes(group = GENETIC_SUBPOP_KINSHIP, color = GENETIC_SUBPOP_KINSHIP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 0.02, angle = 45) +
  scale_color_viridis(discrete = TRUE, end = 0.8)
```

# Fourway site plots
```{r}
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, SUBPOP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(SUBPOP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(!(SITE %in% c("FRMI", "TMPL"))) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = cv_D2F)) + 
  geom_point(aes(color = SUBPOP)) +
  geom_line(aes(group = SUBPOP, color = SUBPOP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 0.02, angle = 45) + ylim(c(0,0.6)) +
  scale_color_viridis(discrete = TRUE)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, SUBPOP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(SUBPOP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(!(SITE %in% c("FRMI", "TMPL"))) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = mean_D2F)) + 
  geom_line(aes(group = SUBPOP, color = SUBPOP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 50, angle = 45) +
  scale_color_viridis(discrete = TRUE) +
  geom_errorbar(aes(x = Latitude, ymin = mean_D2F - sd_D2F, ymax = mean_D2F + sd_D2F, group = SUBPOP, color = SUBPOP)) +
  geom_point(aes(color = SUBPOP))

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  #filter(GENETIC_SUBPOP_KINSHIP %in% c("Midwest", "Gulf")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10) %>%
  group_by(SITE, SUBPOP) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE), 
            median_D2F = median(D2F, na.rm = TRUE),
            sd_D2F = sd(D2F, na.rm = TRUE),
            cv_D2F = sd_D2F/mean_D2F) %>%
  arrange(SUBPOP, cv_D2F) %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(!(SITE %in% c("FRMI", "TMPL"))) %>%
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = sd_D2F^2)) + 
  geom_point(aes(color = SUBPOP)) +
  geom_line(aes(group = SUBPOP, color = SUBPOP)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(x = Latitude, label = SITE), y = 2000, angle = 45) + ylim(c(0, 4200)) +
  scale_color_viridis(discrete = TRUE, end = 1)

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Gulf") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("KING", "STIL")) %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none")
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("KING", "STIL")) %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none")

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Gulf") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("PKLE", "STIL")) %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none")
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("PKLE", "STIL")) %>%
  ggplot(aes(x = SITE, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none")

phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Gulf") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("STIL", "CLMB", "KING", "LINC", "KBSM")) %>%   # "PKLE", "BRKG"
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 130, angle = 45)
phe %>%
  dplyr::select(-ends_with("GR50"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  filter(SITE %in% c("STIL","KING","KBSM")) %>%   # "BRKG", "CLMB", "LINC", 
  left_join(sites) %>%
  ggplot(aes(x = Latitude, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = PLANT_ID)) + 
  theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 130, angle = 45)
```

Instead of D2F, PTT2F? photothermal time to flowering, from greenup.

```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

gdd_df <- ptt_df %>%
  mutate(GDD_0C_SM = ifelse((TMAX_SM + TMIN_SM)/2 >= 0,
                      (TMAX_SM + TMIN_SM)/2 - 0,
                      0)) %>%
  dplyr::select(DOY, SITE, Day_of_year, GDD_13C_SM:GDD_8C_SM, GDD_0C_SM)

GR_start <- phe %>% 
  dplyr::select(-ends_with("D2F"), -ends_with("FL50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("GR50"), names_to = "SITE", values_to = "GR50") %>%
  #filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  group_by(SITE, PLANT_ID) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- phe %>% 
  dplyr::select(-ends_with("D2F"), -ends_with("GR50")) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  pivot_longer(cols = ends_with("FL50"), names_to = "SITE", values_to = "FL50") %>%
  #filter(D2F > 10 & SUBPOP == "Midwest") %>%
  ungroup() %>%
  mutate(SITE = str_sub(SITE, end = 4)) %>%
  group_by(SITE, PLANT_ID) %>%
  summarise(FL_end = ceiling(max(FL50)))

GDD_PT <- NULL

tmp3 <- gdd_df %>%
  dplyr::select(DOY, SITE, Day_of_year, GDD_13C_SM:GDD_0C_SM) %>%
  left_join(GR_start, by = "SITE") %>%
  left_join(FL_end, by = c("SITE", "PLANT_ID")) %>%
  group_by(SITE, Day_of_year, PLANT_ID) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup()

tmp1 <- ptt_df %>%
  dplyr::select(DOY, SITE, Day_of_year, PTT_8C:PTT_13C) %>%
  left_join(GR_start, by = "SITE") %>%
  left_join(FL_end, by = c("SITE", "PLANT_ID")) %>%
  group_by(SITE, Day_of_year, PLANT_ID) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup()

tmp1 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CPTT_8C = sum(PTT_8C),
            CPTT_9C = sum(PTT_9C),
            CPTT_10C = sum(PTT_10C),
            CPTT_11C = sum(PTT_11C),
            CPTT_12C = sum(PTT_12C),
            CPTT_13C = sum(PTT_13C)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  filter(!(SITE %in% c("TMPL", "BRKG", "FRMI"))) %>%
  filter(SUBPOP %in% c("Gulf", "Atlantic", "Midwest")) %>%
  ggplot(aes(x = Latitude, y = CPTT_8C)) + 
  geom_line(aes(group = PLANT_ID, color = SUBPOP)) + 
  #theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 58000, angle = 45)
```

```{r}
pid_all_sites <- phe %>%
  dplyr::select(PLANT_ID, ends_with("D2F")) %>%
  pivot_longer(cols = ends_with("D2F"), names_to = "SITE", values_to = "D2F") %>%
  group_by(PLANT_ID) %>%
  filter(!is.na(D2F) & !(SITE %in% c("FRMI_D2F", "BRKG_D2F", "TMPL_D2F"))) %>%
  count() %>% 
  arrange(desc(n)) %>%
  filter(n >= 4)

tmp3 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_0C = sum(GDD_0C_SM),
            CGDD_8C = sum(GDD_8C_SM),
            CGDD_13C = sum(GDD_13C_SM)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  #filter(!(SITE %in% c("TMPL"))) %>%
  filter(PLANT_ID %in% pid_all_sites$PLANT_ID & SUBPOP %in% c("Midwest", "Gulf")) %>%
  ggplot(aes(x = Latitude, y = D2F)) + 
  geom_line(aes(group = PLANT_ID, color = SUBPOP)) + 
  #theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 10, angle = 45) +
  geom_smooth(aes(group = SUBPOP), method = "lm")  +
  #geom_smooth(aes(group = SUBPOP), color = "green") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1) +
  ylim(c(0, 275))
save_plot(filename = "../analysis/environmental-index/Reaction_norm_days_to_flowering_Gulf_and_Midwest_grown_4_or_more_4way_sites.png", plot = last_plot(), base_height = 6)

tmp3 %>%
  group_by(SITE, PLANT_ID, GR_start, FL_end) %>%
  summarise(CGDD_0C = sum(GDD_0C_SM),
            CGDD_8C = sum(GDD_8C_SM),
            CGDD_13C = sum(GDD_13C_SM)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  filter(D2F > 10) %>%
  left_join(metadata) %>%
  left_join(sites) %>%
  #filter(!(SITE %in% c("TMPL"))) %>%
  filter(PLANT_ID %in% pid_all_sites$PLANT_ID & SUBPOP %in% c("Midwest", "Gulf")) %>%
  ggplot(aes(x = Latitude, y = CGDD_13C)) + 
  geom_line(aes(group = PLANT_ID, color = SUBPOP)) + 
  #theme(legend.position = "none") +
  geom_text(aes(x = Latitude, label = SITE), y = 100, angle = 45) +
  geom_smooth(aes(group = SUBPOP), method = "lm")  +
  #geom_smooth(aes(group = SUBPOP), color = "green") +
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1)  + 
  ylim(c(0,3200))
save_plot(filename = "../analysis/environmental-index/Reaction_norm_cumulative_GDD_to_flowering_Gulf_and_Midwest_grown_4_or_more_4way_sites.png", plot = last_plot(), base_height = 6)
```
 61 PLANT ID have D2F at all 6 fourway sites.  209 at 5 fourway sites.
PTTGR2FL