### 2020-04-29 R/h2_var_p_and_cov_p.Rmd
rrBLUP and h^2 analysis for D2F

---
title: "h2 and cov_P"
author: "Alice MacQueen"
date: "4/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
```

#

```{r}
phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenotypes_D2F_fourway_project.rds")
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_))
```