### 2020-06-11 data-raw/Flowering_GWAS_Phenotypes.Rmd

---
title: "Flowering_GWAS_Phenotypes"
author: "Alice MacQueen"
date: "6/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
phenotypes <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_cleaned.rds")
```

# Make phenotypes for flowering time paper from all the cleaned phenotypes.

Specifically using GR50 & FL50 Julian dates from 2019. And using them to make cGDD between greenup and flowering for another GWAS phenotype.
```{r}

```