### 2020-06-17 R/Analyse_univarate_gwas.Rmd
GWAS with GDD results overviews for writeup

---
title: "Analyze Univariate GWAS"
author: "Alice MacQueen"
date: "6/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#
```{r}
summary <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary.csv",
                    col_names = "filename")

summary %>%
  separate(col = "filename", into = c("path", "site", "phe", NA, NA, "individuals", "SNPs", NA, "PCs", NA, NA, "date_run", "time_run"), sep = "_") %>%
  separate(individuals, into = c("individuals", NA), sep = -1, convert = TRUE) %>%
  mutate(subpop = case_when(SNPs == "12.1M" ~ "Midwest_Gulf",
                            SNPs == "10.6M" ~ "Atlantic",
                            SNPs == "8.8M" ~ "Midwest",
                            SNPs == "10.3M" ~ "Gulf",
                            SNPs == "12.3M" ~ "Midwest_Atlantic",
                            SNPs == "11.5M" ~ "Atlantic_Gulf",
                       TRUE ~ NA_character_)) %>%
  write_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary.csv", 
            col_names = TRUE)

```

```{r}
summary1 <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/GWAS_summary_Bonferroni.csv", col_names = TRUE)
summary2 <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary.csv", col_names = TRUE)

summary2 %>%
  rename(Manh_time = time_run) %>%
  select(-path) %>%
  left_join(summary1) %>%
  mutate(time_run = coalesce(time_run, Manh_time),
         time_run = str_sub(time_run, end = -5)) %>%
  select(-Qqplot_order, -path, -Manh_time) %>%
  write_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary.csv", 
            col_names = TRUE, na = "")
```

```{r}
summary <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/gwas/GDD/Bonferroni/GWAS_summary_Bonferroni_mash_decisions.csv", col_names = TRUE)

summary %>%
  filter(grepl("GR", phe)) %>%
  group_by(subpop, site, qqplot) %>%
  add_tally(sort = FALSE) %>%
  select(indv:notes, n, everything()) %>%
  arrange(subpop, site, phe)

summary %>%
  filter(grepl("GR", phe), qqplot != "drop", qqplot != "iffy") %>%
  mutate(bonferroni = as.numeric(bonferroni)) %>%
  arrange(subpop, site, phe) %>%
  filter(qqplot == "signal" & bonferroni > 0) %>%
  group_by(subpop) %>% 
  tally()

```

```{r}
summary %>%
  filter(grepl("FL50$", phe), qqplot != "drop", qqplot != "iffy") %>%
  mutate(bonferroni = as.numeric(bonferroni)) %>%
  arrange(subpop, site, phe) %>%
  filter(qqplot == "signal" & bonferroni > 0) %>%
  group_by(subpop) %>% 
  tally()
summary %>%
  filter(grepl("FL50$", phe))
summary %>%
  filter(grepl("FL50$", phe), qqplot != "drop", qqplot != "iffy") %>%
  mutate(bonferroni = as.numeric(bonferroni)) %>%
  arrange(subpop, site, phe) %>%
  filter(qqplot == "signal" & bonferroni > 0)

summary %>%
  filter(qqplot != "drop", qqplot != "iffy", notes != "structure") %>%
  mutate(bonferroni = as.numeric(bonferroni)) %>%
  group_by(subpop, phe) %>%
  summarise(sum = sum(bonferroni, na.rm = TRUE))
```
CGDD: 37 total. 14 have signal in QQplot & hits above Bonferroni
Atlantic	1			
Atlantic_Gulf	4			
Gulf	1			
Midwest	2			
Midwest_Atlantic	2			
Midwest_Gulf	5	
PTT: 37 total. 13 have signal & hits above BOnferroni.
Atlantic	1			
Atlantic_Gulf	3			
Midwest	1			
Midwest_Atlantic	3			
Midwest_Gulf	5	
CRAIN: 17 of 37.
Atlantic	2			
Atlantic_Gulf	4			
Gulf	3			
Midwest	1			
Midwest_Atlantic	3			
Midwest_Gulf	4	
CRAIN.5d: 16 of 37. 
Atlantic	3			
Atlantic_Gulf	3			
Gulf	2			
Midwest	1			
Midwest_Atlantic	4			
Midwest_Gulf	3

Very few GR have signal in QQplot & hits above Bonferroni. 
GR1 13 of 40. GR50 12 of 40. GR100 5 of 38.
Atlantic	3			
Atlantic_Gulf	5			
Gulf	7			
Midwest 0 
Midwest_Atlantic	7			
Midwest_Gulf	8

EM has more: 33 out of 70. EM1 16 of 35.  EM50 17 of 35.
Atlantic	6			
Atlantic_Gulf	6			
Gulf	5			
Midwest	1			
Midwest_Atlantic	7			
Midwest_Gulf	8

FL also has more: 33 out of 76. FL1 17 of 39. FL50 16 of 37.
Atlantic	4			
Atlantic_Gulf	10			
Gulf	4			
Midwest	2			
Midwest_Atlantic	7			
Midwest_Gulf	6