### 2020-06-19 R/Draft_models_and_stats_2020-06-19.Rmd
Start writing fourway GDD section

---
title: "Draft_models_and_stats_2020-06-19"
author: "Alice MacQueen"
date: "6/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(cowplot)
library(corpcor)
library(viridis)
```

# ASReml-R analyses on server

```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
library(corpcor)
```

# Load data

```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "manuscript", "plots")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe <- read_rds(file.path(workingdir, "data",
                          "Phenology_CGDD_PTT_and_rainfall_phenotypes.rds"))
```

Unfortunately, no parents flowered exclusively during shortening days at the three Texas sites; WBC flowered during shortening days in the majority years in the Texas common gardens.
```{r}
phe %>%
  filter(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "WBC", "DAC", "VS16", "VxW", "AxD")) %>%
  filter(PHE == "FL50" & SITE %in% c("PKLE", "TMPL", "KING", "OVTN", "STIL")) %>%
  group_by(PLANT_ID, SITE, MEAS >= 173) %>%
  tally() 
```
AxD KING 0 / 41
PKLE 0 / 39
TMPL 0/39

VxW 
PKLE 2 / 43
TMPL: 21 / 41
KING: 0 / 43

AP13:
TMPL: 56 / 66 >= 173
PKLE: 52 / 67 
KING: 0 / 40

WBC: 
TMPL: 40/40
PKLE: 39 / 46
KING: 10 / 18

DAC: 0% of 15
VS16: 3.6% of 56

```{r}
56/66
52/67
0
(56+52)/(66+67+40)

100
39/46
10/18
(40+39+10)/(40+46+18)
2/56
23/(43+41+43)
41+39+39
```