### 2020-07-08 R/Make_mash_U_hyp.Rmd
Make U_hyp object for gulf subpop and use in mash

---
title: "Make_mash_U_hyp"
author: "Alice MacQueen"
date: 2020-07-07
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(geosphere)
library(spdep)
library(rlist)
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
sitedf <- as.data.frame(sites)
coordinates(sitedf) <- c("Longitude", "Latitude")
```

# Covariance matrices for mash

Want to make these based on the inverse of the distance between sites, and also 
on the covariances

```{r}
sitenb <- tri2nb(sitedf, row.names = sitedf$manu_site)
summary(sitenb)
wm <- nb2mat(sitenb, style = 'W')
ww <- nb2listw(sitenb, style = 'W')
moran(sitedf$Elevation, ww, n=length(ww$neighbours), S0=Szero(ww), NAOK = TRUE)
moran.test(sitedf$Elevation, ww, randomisation=FALSE, na.action = na.omit)
moran.mc(sitedf$Elevation, ww, nsim=499)

wmw <- nb2mat(sitenb, style = 'W')
colnames(wmw) <- rownames(wmw)

U1 <- as_tibble(wmw, .name_repair = "unique") %>%
    mutate(rowU = sitedf$manu_site) %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                 values_to = "covar") %>%
    filter(!is.na(.data$covar))

  U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    CDBNgenomics::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "#27AD81FF", "#5DC863FF", "#FDE725FF")) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "right",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("")
```

## Inverse distances between sites, in 100km
```{r}
distsites <- distm(sitedf)
colnames(distsites) <- sites$SITE
rownames(distsites) <- sites$SITE
W <- 1 / (distsites/100000) # 100km distance measure
W[is.infinite(W)] <- NA
diag(W) <- max(W, na.rm = TRUE) + min(W, na.rm = TRUE)
U_dist <- W
U1 <- as_tibble(W, .name_repair = "unique") %>%
    mutate(rowU = rownames(W)) %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                 values_to = "covar") %>%
    filter(!is.na(.data$covar))

  U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    CDBNgenomics::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "#27AD81FF", "#5DC863FF", "#FDE725FF")) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "right",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("")
```

## Weather covariances in the time period between greenup and flowering, plus in the 5d before flowering
```{r}
phe_wea <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")

weadf_cov <- phe_wea %>%
  pivot_longer(cols = GR50:CRAIN_5d, names_to = "phe", values_to = "MEAS") %>%
  unite(col = "site_phe", SITE, phe, sep = ".") %>%
  pivot_wider(id_cols = PLANT_ID, names_from = site_phe, values_from = MEAS, values_fn = mean)

weadf_cov2 <- cov(weadf_cov[-1], use = "pairwise")
colnames(weadf_cov2)

pattern <- c("GR50", "FL50", "CGDD_12C$", "CGDD_12C_5d", "PTT_12C$", 
             "PTT_12C_5d", "CRAIN$", "CRAIN_5d" )
Ureturn <- list()

for(i in seq_along(pattern)){
  U_wea <- weadf_cov2 %>%
    as_tibble(rownames = "rowU") %>%
    dplyr::select(rowU, matches(pattern[i])) %>%
    filter(grepl(pattern[i], rowU)) %>%
    separate(rowU, into = c("site", "phe"), sep = 4) %>%
    dplyr::select(-phe) %>%
    column_to_rownames(var = "site") %>%
    as.data.frame()
  colnames(U_wea) <- str_sub(colnames(U_wea), end = 4)

  Ureturn <- list.append(Ureturn, U_wea = U_wea)
  names(Ureturn)[i] <- paste0("U_", str_extract(pattern[i], "[^$]+")) # matches one or more characters other than $
}
names(Ureturn)
Ureturn <- list.append(Ureturn, U_dist = U_dist)
```

# Use these U_hyp in mash_standard_run....
# mash runs

Midwest: CLMB, KBSM, PKLE
Gulf: CLMB KBSM KING PKLE TMPL
Midwest_Gulf: BRKG CLMB KBSM KING LINC PKLE STIL TMPL
Atlantic: CLMB KBSM LINC PKLE TMPL
always in alphabetical order, so that's good.
Also, each location is repeated twice. Once for CGDD and once for FL50. So covar matrices should be duplicated also? Ehhhh might also consider running these separately first! Honestly. And join them in a mash analysis later if I want to compare them. But just duplicate them now to test??

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")

Ugulf <- list()
i = 1 
for(i in seq_along(names(Ureturn))){
  Utest <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% gulf_loc),
                 which(colnames(Ureturn[[i]]) %in% gulf_loc)]

  Urow <- rep(Utest, each = 2)
  Utr <- matrix(unlist(Urow), nrow = nrow(Utest), ncol = ncol(Utest)*2)
  Ut <- rep(Utr, each = 2)
  Udup <- matrix(unlist(Ut), nrow = nrow(Utr)*2, ncol = ncol(Utr))
  colnames(Udup) <- rep(gulf_loc, each = 2)
  rownames(Udup) <- rep(gulf_loc, each = 2)
  Ugulf <- list.append(Ugulf, Udup = Udup)
  names(Ugulf)[i] <- names(Ureturn)[i]
}

write_rds(Ugulf, path = file.path(path, "U_hyp_gulf.rds"))
```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")
U_gulf <- read_rds(file.path(path, "U_hyp_gulf.rds"))
i = 2
  m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_hyp = U_gulf,
                             suffix = paste0(phe_suffix[i], "_U_hyp_v1"),
                             saveoutput = TRUE)
```

## Look at old mash_standard_runs

Need to rerun these without using ref = 'mean'. Then I'll be able to look at the covariance matrices.
```{r}
library(switchgrassGWAS)

path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c( "atlantic", "gulf", "midwest", "midwest_gulf")
numSNPs <- c(10000, 10000, 28000, 4000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")
mash_rds <- pvdiv_results_in_folder(path = path, pattern = "^Strong_Effects")
mash_suffix <- str_sub(mash_rds, start = 24, end = -5) %>%
  str_replace(., "^_", "")

i = 2
for(i in seq_along(mash_rds)){
  m <- readRDS(file.path(path, mash_rds[i]))
  suffix <- mash_suffix[i]
  p1 <- mash_plot_manhattan_by_condition(m, saveoutput = TRUE, suffix = suffix)
  p2 <- mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
  p3 <- mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
  p4 <- mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
  p5 <- mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
  p6 <- mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
  p7 <- mash_plot_Ulist(m, range = get_U_by_mass(m), saveoutput = TRUE, 
                  suffix = suffix)
}
```