### 2020-07-09 R/Make_mash_U_hyp.Rmd
Test smaller Uhyp matrices for midwest and gulf subpops

---
title: "Make_mash_U_hyp"
author: "Alice MacQueen"
date: 2020-07-07
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(geosphere)
library(spdep)
library(rlist)
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
sitedf <- as.data.frame(sites)
coordinates(sitedf) <- c("Longitude", "Latitude")
```

# Covariance matrices for mash

Want to make these based on the inverse of the distance between sites, and also 
on the covariances

```{r}
sitenb <- tri2nb(sitedf, row.names = sitedf$manu_site)
summary(sitenb)
wm <- nb2mat(sitenb, style = 'W')
ww <- nb2listw(sitenb, style = 'W')
moran(sitedf$Elevation, ww, n=length(ww$neighbours), S0=Szero(ww), NAOK = TRUE)
moran.test(sitedf$Elevation, ww, randomisation=FALSE, na.action = na.omit)
moran.mc(sitedf$Elevation, ww, nsim=499)

wmw <- nb2mat(sitenb, style = 'W')
colnames(wmw) <- rownames(wmw)

U1 <- as_tibble(wmw, .name_repair = "unique") %>%
    mutate(rowU = sitedf$manu_site) %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                 values_to = "covar") %>%
    filter(!is.na(.data$covar))

  U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    CDBNgenomics::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "#27AD81FF", "#5DC863FF", "#FDE725FF")) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "right",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("")
```

## Inverse distances between sites, in 100km
```{r}
distsites <- distm(sitedf)
colnames(distsites) <- sites$SITE
rownames(distsites) <- sites$SITE
W <- 1 / (distsites/100000) # 100km distance measure
W[is.infinite(W)] <- NA
diag(W) <- max(W, na.rm = TRUE) + min(W, na.rm = TRUE)
U_dist <- W
U1 <- as_tibble(W, .name_repair = "unique") %>%
    mutate(rowU = rownames(W)) %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                 values_to = "covar") %>%
    filter(!is.na(.data$covar))

  U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    CDBNgenomics::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "#27AD81FF", "#5DC863FF", "#FDE725FF")) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "right",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("")
```

## Weather covariances in the time period between greenup and flowering, plus in the 5d before flowering
```{r}
phe_wea <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")

weadf_cov <- phe_wea %>%
  pivot_longer(cols = GR50:CRAIN_5d, names_to = "phe", values_to = "MEAS") %>%
  unite(col = "site_phe", SITE, phe, sep = ".") %>%
  pivot_wider(id_cols = PLANT_ID, names_from = site_phe, values_from = MEAS, values_fn = mean)

weadf_cov2 <- cov(weadf_cov[-1], use = "pairwise")
colnames(weadf_cov2)

pattern <- c("GR50", "FL50", "CGDD_12C$", "CGDD_12C_5d", "PTT_12C$", 
             "PTT_12C_5d", "CRAIN$", "CRAIN_5d" )
Ureturn <- list()

for(i in seq_along(pattern)){
  U_wea <- weadf_cov2 %>%
    as_tibble(rownames = "rowU") %>%
    dplyr::select(rowU, matches(pattern[i])) %>%
    filter(grepl(pattern[i], rowU)) %>%
    separate(rowU, into = c("site", "phe"), sep = 4) %>%
    dplyr::select(-phe) %>%
    column_to_rownames(var = "site") %>%
    as.data.frame()
  colnames(U_wea) <- str_sub(colnames(U_wea), end = 4)

  Ureturn <- list.append(Ureturn, U_wea = U_wea)
  names(Ureturn)[i] <- paste0("U_", str_extract(pattern[i], "[^$]+")) # matches one or more characters other than $
}
names(Ureturn)

Ureturn$U_dist <- Ureturn$U_dist %>% as.data.frame()
Ureturn <- list.append(Ureturn, U_dist = U_dist)

str(Ureturn)
```

# Use these U_hyp in mash_standard_run....
# mash runs

Midwest: CLMB, KBSM, PKLE
Gulf: CLMB KBSM KING PKLE TMPL
Midwest_Gulf: BRKG CLMB KBSM KING LINC PKLE STIL TMPL
Atlantic: CLMB KBSM LINC PKLE TMPL
always in alphabetical order, so that's good.
Also, each location is repeated twice. Once for CGDD and once for FL50. So covar matrices should be duplicated also? Ehhhh might also consider running these separately first! Honestly. And join them in a mash analysis later if I want to compare them. But just duplicate them now to test??

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")

Ugulf <- list()
i = 1 
for(i in seq_along(names(Ureturn))){
  Utest <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% gulf_loc),
                 which(colnames(Ureturn[[i]]) %in% gulf_loc)]

  Urow <- rep(Utest, each = 2)
  Utr <- matrix(unlist(Urow), nrow = nrow(Utest), ncol = ncol(Utest)*2)
  Ut <- rep(Utr, each = 2)
  Udup <- matrix(unlist(Ut), nrow = nrow(Utr)*2, ncol = ncol(Utr))
  colnames(Udup) <- rep(gulf_loc, each = 2)
  rownames(Udup) <- rep(gulf_loc, each = 2)
  Ugulf <- list.append(Ugulf, Udup = Udup)
  names(Ugulf)[i] <- names(Ureturn)[i]
}
str(Ugulf)
write_rds(Ugulf, path = file.path(path, "U_hyp_gulf.rds"))

Ugulf$U_GR50
U_gulf_sym <- list()
isSymmetric(Ugulf$U_dist)
i = 2
for(i in seq_along(names(Ugulf))){
  Usym <- Ugulf[[i]]
  Usym1 <- Usym
  Usym2 <- Usym
  for(j in nrow(Usym1)){
    for(k in ncol(Usym1)){
      if(j < k){
        Usym1[j, k] <- Usym[j, k]
        Usym2[j, k] <- Usym[k, j]
      } else {
        Usym1[j, k] <- Usym[k, j]
        Usym2[j, k] <- Usym[j, k]
      }
    }
  }
  U_gulf_sym <- list.append(U_gulf_sym, U1 = Usym1, U2 = Usym2)
  length(names(U_gulf_sym))
  names(U_gulf_sym)[length(names(U_gulf_sym))-1] <- paste0(names(Ugulf)[i], "_1")
  names(U_gulf_sym)[length(names(U_gulf_sym))] <- paste0(names(Ugulf)[i], "_2")
}

write_rds(U_gulf_sym, path = file.path(path, "U_hyp_gulf_symmetric.rds"))

```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")
U_gulf <- read_rds(file.path(path, "U_hyp_gulf.rds"))
i = 2
U_ed <- file.path(path, "Mash_data_driven_covariances_10000_gulf.rds")
  m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_ed = U_ed,
                             U_hyp = U_gulf, suffix = phe_suffix[i],
                             saveoutput = TRUE)
```

## Look at old mash_standard_runs

Need to rerun these without using ref = 'mean'. Then I'll be able to look at the covariance matrices.
```{r}
library(switchgrassGWAS)

path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c( "atlantic", "gulf", "midwest", "midwest_gulf")
numSNPs <- c(10000, 10000, 28000, 4000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")
mash_rds <- pvdiv_results_in_folder(path = path, pattern = "^Strong_Effects")
mash_suffix <- str_sub(mash_rds, start = 24, end = -5) %>%
  str_replace(., "^_", "")

i = 2
for(i in seq_along(mash_rds)){
  m <- readRDS(file.path(path, mash_rds[i]))
  suffix <- mash_suffix[i]
  p1 <- mash_plot_manhattan_by_condition(m, saveoutput = TRUE, suffix = suffix)
  p2 <- mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
  p3 <- mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
  p4 <- mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
  p5 <- mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
  p6 <- mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
  p7 <- mash_plot_Ulist(m, range = get_U_by_mass(m), saveoutput = TRUE, 
                  suffix = suffix)
}
load_all("~/Github/switchgrassGWAS")
```

# Test mash with U_hyp

I got lots of errors the first time I ran mash to completion with U_hyp, and I am not yet certain why. I'd like to run mash on just FL50 in the gulf, and GDD in the midwest maybe, to see if I can diagnose what's going on or what the covariance matrices should look like.

## Rerun bigsnp2mashr

```{r}
trim_name <- function(x){
    for(i in seq_along(x)){
  if((str_sub(x[i], start = -1) %in% c("_")))
  { x[i] <- str_sub(x[i], end = -2) }
    }
  return(x)
}

pathin <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD")
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")

inputfiles <- read_delim(file.path("~", "Github", "pvdiv-phenology-gxe", 
                                   "analysis", "gwas", "inputkinship.txt"), 
                         delim = " ", col_names = "SNPfiles") 
# Midwest & Gulf, Midwest, Gulf, Atlantic, Midwest & Atlantic, Gulf & Atlantic

# Midwest
snp <- snp_attach(inputfiles$SNPfiles[2])
gwas_rds <- pvdiv_results_in_folder(path = pathin, 
                                    pattern = paste0("GWAS_datatable_",
                                                     "(KBSM|PKLE|CLMB)", ".+",
                                                     "(CGDD_12C)", 
                                                     ".+",
                                                     "\\d+g_8.8M"))

site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -41))
numSNPs <- 20000 #ceiling((1000000 / length(site_phe)^2)/1000)*1000

midwest_mashin <- 
  pvdiv_bigsnp2mashr(path = pathin, snp = snp,
                    gwas_rds = gwas_rds, phenotypes = site_phe, 
                    numSNPs = numSNPs, scale = FALSE, 
                    suffix = "midwest_uhyp1", saveoutput = TRUE)
# Gulf [3]
snp <- snp_attach(inputfiles$SNPfiles[3])
gwas_rds <- pvdiv_results_in_folder(path = pathin, 
                                    pattern = paste0("GWAS_datatable_", ".+",
                                                     "(FL50)", 
                                                     ".+",
                                                     "\\d+g_10.3M"))

site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -42))
numSNPs <- 30000 #ceiling((1000000 / length(site_phe)^2)/1000)*1000

gulf_mashin <- 
  pvdiv_bigsnp2mashr(path = pathin, snp = snp,
                    gwas_rds = gwas_rds, phenotypes = site_phe, 
                    numSNPs = numSNPs, scale = FALSE, suffix = "gulf_uhyp1", 
                    saveoutput = TRUE)
```

# Uhyp for midwest and gulf

```{r}
gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")

Ugulfhyp <- list()
Umidwhyp <- list()

for(i in seq_along(names(Ureturn))){
  Ugulf1 <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% gulf_loc),
                 which(colnames(Ureturn[[i]]) %in% gulf_loc)]
  Umidw1 <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% midw_loc),
                 which(colnames(Ureturn[[i]]) %in% midw_loc)]

  Ugulfhyp <- list.append(Ugulfhyp, Ugulf1 = Ugulf1)
  Umidwhyp <- list.append(Umidwhyp, Umidw1 = Umidw1)
  names(Ugulfhyp)[i] <- names(Ureturn)[i]
  names(Umidwhyp)[i] <- names(Ureturn)[i]
}
str(Umidwhyp)
write_rds(Ugulfhyp, path = file.path(path, "U_hyp_gulf_1phe.rds"))
write_rds(Umidwhyp, path = file.path(path, "U_hyp_midw_1phe.rds"))
```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
numSNPs <- 20000
suffix = "midwest_uhyp1"
Umidwhyp <- read_rds(path = file.path(path, "U_hyp_midw_1phe.rds"))

m <- mash_standard_run(path = path, numSNPs = numSNPs, U_hyp = Umidwhyp, 
                       suffix = suffix, saveoutput = TRUE)
```