### 2020-07-09 R/Make_mash_U_hyp.Rmd
Try rescaling all FL50 gulf effects to fall on same after-solstice scale

---
title: "Make_mash_U_hyp"
author: "Alice MacQueen"
date: 2020-07-07
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(geosphere)
library(spdep)
library(rlist)
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
sitedf <- as.data.frame(sites)
coordinates(sitedf) <- c("Longitude", "Latitude")
```

# Covariance matrices for mash

Want to make these based on the inverse of the distance between sites, and also 
on the covariances

```{r}
sitenb <- tri2nb(sitedf, row.names = sitedf$manu_site)
summary(sitenb)
wm <- nb2mat(sitenb, style = 'W')
ww <- nb2listw(sitenb, style = 'W')
moran(sitedf$Elevation, ww, n=length(ww$neighbours), S0=Szero(ww), NAOK = TRUE)
moran.test(sitedf$Elevation, ww, randomisation=FALSE, na.action = na.omit)
moran.mc(sitedf$Elevation, ww, nsim=499)

wmw <- nb2mat(sitenb, style = 'W')
colnames(wmw) <- rownames(wmw)

U1 <- as_tibble(wmw, .name_repair = "unique") %>%
    mutate(rowU = sitedf$manu_site) %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                 values_to = "covar") %>%
    filter(!is.na(.data$covar))

  U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    CDBNgenomics::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "#27AD81FF", "#5DC863FF", "#FDE725FF")) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "right",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("")
```

## Inverse distances between sites, in 100km
```{r}
distsites <- distm(sitedf)
colnames(distsites) <- sites$SITE
rownames(distsites) <- sites$SITE
W <- 1 / (distsites/100000) # 100km distance measure
W[is.infinite(W)] <- NA
diag(W) <- max(W, na.rm = TRUE) + min(W, na.rm = TRUE)
U_dist <- W
U1 <- as_tibble(W, .name_repair = "unique") %>%
    mutate(rowU = rownames(W)) %>%
    pivot_longer(cols = -.data$rowU, names_to = "colU",
                 values_to = "covar") %>%
    filter(!is.na(.data$covar))

  U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    CDBNgenomics::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "#27AD81FF", "#5DC863FF", "#FDE725FF")) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = "right",
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("")
```

## Weather covariances in the time period between greenup and flowering, plus in the 5d before flowering
```{r}
phe_wea <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")

weadf_cov <- phe_wea %>%
  pivot_longer(cols = GR50:CRAIN_5d, names_to = "phe", values_to = "MEAS") %>%
  unite(col = "site_phe", SITE, phe, sep = ".") %>%
  pivot_wider(id_cols = PLANT_ID, names_from = site_phe, values_from = MEAS, values_fn = mean)

weadf_cov2 <- cov(weadf_cov[-1], use = "pairwise")
colnames(weadf_cov2)

pattern <- c("GR50", "FL50", "CGDD_12C$", "CGDD_12C_5d", "PTT_12C$", 
             "PTT_12C_5d", "CRAIN$", "CRAIN_5d" )
Ureturn <- list()

for(i in seq_along(pattern)){
  U_wea <- weadf_cov2 %>%
    as_tibble(rownames = "rowU") %>%
    dplyr::select(rowU, matches(pattern[i])) %>%
    filter(grepl(pattern[i], rowU)) %>%
    separate(rowU, into = c("site", "phe"), sep = 4) %>%
    dplyr::select(-phe) %>%
    column_to_rownames(var = "site") %>%
    as.data.frame()
  colnames(U_wea) <- str_sub(colnames(U_wea), end = 4)

  Ureturn <- list.append(Ureturn, U_wea = U_wea)
  names(Ureturn)[i] <- paste0("U_", str_extract(pattern[i], "[^$]+")) # matches one or more characters other than $
}
names(Ureturn)

Ureturn$U_dist <- Ureturn$U_dist %>% as.data.frame()
Ureturn <- list.append(Ureturn, U_dist = U_dist)

str(Ureturn)
```

# Use these U_hyp in mash_standard_run....
# mash runs

Midwest: CLMB, KBSM, PKLE
Gulf: CLMB KBSM KING PKLE TMPL
Midwest_Gulf: BRKG CLMB KBSM KING LINC PKLE STIL TMPL
Atlantic: CLMB KBSM LINC PKLE TMPL
always in alphabetical order, so that's good.
Also, each location is repeated twice. Once for CGDD and once for FL50. So covar matrices should be duplicated also? Ehhhh might also consider running these separately first! Honestly. And join them in a mash analysis later if I want to compare them. But just duplicate them now to test??

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")

Ugulf <- list()
i = 1 
for(i in seq_along(names(Ureturn))){
  Utest <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% gulf_loc),
                 which(colnames(Ureturn[[i]]) %in% gulf_loc)]

  Urow <- rep(Utest, each = 2)
  Utr <- matrix(unlist(Urow), nrow = nrow(Utest), ncol = ncol(Utest)*2)
  Ut <- rep(Utr, each = 2)
  Udup <- matrix(unlist(Ut), nrow = nrow(Utr)*2, ncol = ncol(Utr))
  colnames(Udup) <- rep(gulf_loc, each = 2)
  rownames(Udup) <- rep(gulf_loc, each = 2)
  Ugulf <- list.append(Ugulf, Udup = Udup)
  names(Ugulf)[i] <- names(Ureturn)[i]
}
str(Ugulf)
write_rds(Ugulf, path = file.path(path, "U_hyp_gulf.rds"))

Ugulf$U_GR50
U_gulf_sym <- list()
isSymmetric(Ugulf$U_dist)
i = 2
for(i in seq_along(names(Ugulf))){
  Usym <- Ugulf[[i]]
  Usym1 <- Usym
  Usym2 <- Usym
  for(j in nrow(Usym1)){
    for(k in ncol(Usym1)){
      if(j < k){
        Usym1[j, k] <- Usym[j, k]
        Usym2[j, k] <- Usym[k, j]
      } else {
        Usym1[j, k] <- Usym[k, j]
        Usym2[j, k] <- Usym[j, k]
      }
    }
  }
  U_gulf_sym <- list.append(U_gulf_sym, U1 = Usym1, U2 = Usym2)
  length(names(U_gulf_sym))
  names(U_gulf_sym)[length(names(U_gulf_sym))-1] <- paste0(names(Ugulf)[i], "_1")
  names(U_gulf_sym)[length(names(U_gulf_sym))] <- paste0(names(Ugulf)[i], "_2")
}

write_rds(U_gulf_sym, path = file.path(path, "U_hyp_gulf_symmetric.rds"))

```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c("midwest", "gulf", "midwest_gulf", "atlantic")
numSNPs <- c(28000, 10000, 4000, 10000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")
U_gulf <- read_rds(file.path(path, "U_hyp_gulf.rds"))
i = 2
U_ed <- file.path(path, "Mash_data_driven_covariances_10000_gulf.rds")
  m_out <- mash_standard_run(path = path, numSNPs = numSNPs[i], U_ed = U_ed,
                             U_hyp = U_gulf, suffix = phe_suffix[i],
                             saveoutput = TRUE)
```

## Look at old mash_standard_runs

Need to rerun these without using ref = 'mean'. Then I'll be able to look at the covariance matrices.
```{r}
library(switchgrassGWAS)

path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                  "gwas", "GDD")
phe_suffix <- c( "atlantic", "gulf", "midwest", "midwest_gulf")
numSNPs <- c(10000, 10000, 28000, 4000)

gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")
atl_loc <- c("CLMB", "KBSM", "LINC", "PKLE", "TMPL")
midwgulf_loc <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL",
                  "TMPL")
mash_rds <- pvdiv_results_in_folder(path = path, pattern = "^Strong_Effects")
mash_suffix <- str_sub(mash_rds, start = 24, end = -5) %>%
  str_replace(., "^_", "")

i = 2
for(i in seq_along(mash_rds)){
  m <- readRDS(file.path(path, mash_rds[i]))
  suffix <- mash_suffix[i]
  p1 <- mash_plot_manhattan_by_condition(m, saveoutput = TRUE, suffix = suffix)
  p2 <- mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
  p3 <- mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
  p4 <- mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
  p5 <- mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
  p6 <- mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
  p7 <- mash_plot_Ulist(m, range = get_U_by_mass(m), saveoutput = TRUE, 
                  suffix = suffix)
}
load_all("~/Github/switchgrassGWAS")
```

# Test mash with U_hyp

I got lots of errors the first time I ran mash to completion with U_hyp, and I am not yet certain why. I'd like to run mash on just FL50 in the gulf, and GDD in the midwest maybe, to see if I can diagnose what's going on or what the covariance matrices should look like.

## Rerun bigsnp2mashr

```{r}
trim_name <- function(x){
    for(i in seq_along(x)){
  if((str_sub(x[i], start = -1) %in% c("_")))
  { x[i] <- str_sub(x[i], end = -2) }
    }
  return(x)
}

pathin <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD")
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")

inputfiles <- read_delim(file.path("~", "Github", "pvdiv-phenology-gxe", 
                                   "analysis", "gwas", "inputkinship.txt"), 
                         delim = " ", col_names = "SNPfiles") 
# Midwest & Gulf, Midwest, Gulf, Atlantic, Midwest & Atlantic, Gulf & Atlantic

# Midwest
snp <- snp_attach(inputfiles$SNPfiles[2])
gwas_rds <- pvdiv_results_in_folder(path = pathin, 
                                    pattern = paste0("GWAS_datatable_",
                                                     "(KBSM|PKLE|CLMB)", ".+",
                                                     "(CGDD_12C)", 
                                                     ".+",
                                                     "\\d+g_8.8M"))

site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -41))
numSNPs <- 20000 #ceiling((1000000 / length(site_phe)^2)/1000)*1000

midwest_mashin <- 
  pvdiv_bigsnp2mashr(path = pathin, snp = snp,
                    gwas_rds = gwas_rds, phenotypes = site_phe, 
                    numSNPs = numSNPs, scale = FALSE, 
                    suffix = "midwest_uhyp1", saveoutput = TRUE)
# Gulf [3]
snp <- snp_attach(inputfiles$SNPfiles[3])
gwas_rds <- pvdiv_results_in_folder(path = pathin, 
                                    pattern = paste0("GWAS_datatable_", ".+",
                                                     "(FL50)", 
                                                     ".+",
                                                     "\\d+g_10.3M"))

site_phe <- trim_name(str_sub(gwas_rds, start = 16, end = -42))
numSNPs <- 30000 #ceiling((1000000 / length(site_phe)^2)/1000)*1000

gulf_mashin <- 
  pvdiv_bigsnp2mashr(path = pathin, snp = snp,
                    gwas_rds = gwas_rds, phenotypes = site_phe, 
                    numSNPs = numSNPs, scale = FALSE, suffix = "gulf_uhyp1", 
                    saveoutput = TRUE)
```

# Uhyp for midwest and gulf

```{r}
gulf_loc <- c("CLMB", "KBSM", "KING", "PKLE", "TMPL")
midw_loc <- c("CLMB", "KBSM", "PKLE")

Ugulfhyp <- list()
Umidwhyp <- list()

for(i in seq_along(names(Ureturn))){
  Ugulf1 <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% gulf_loc),
                 which(colnames(Ureturn[[i]]) %in% gulf_loc)]
  Umidw1 <- Ureturn[[i]][which(rownames(Ureturn[[i]]) %in% midw_loc),
                 which(colnames(Ureturn[[i]]) %in% midw_loc)]

  Ugulfhyp <- list.append(Ugulfhyp, Ugulf1 = as.matrix(Ugulf1))
  Umidwhyp <- list.append(Umidwhyp, Umidw1 = as.matrix(Umidw1))
  names(Ugulfhyp)[i] <- names(Ureturn)[i]
  names(Umidwhyp)[i] <- names(Ureturn)[i]
}
str(Umidwhyp)
is.matrix(Umidwhyp$U_GR50)
isSymmetric(Umidwhyp$U_GR50)
write_rds(Ugulfhyp, path = file.path(path, "U_hyp_gulf_1phe.rds"))
write_rds(Umidwhyp, path = file.path(path, "U_hyp_midw_1phe.rds"))
```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
numSNPs <- 20000
suffix = "midwest_uhyp1"
Umidwhyp <- read_rds(path = file.path(path, "U_hyp_midw_1phe.rds"))
Udist <- Umidwhyp$U_dist

m <- mash_standard_run(path = path, numSNPs = numSNPs, U_hyp = Udist, 
                       suffix = suffix, saveoutput = TRUE)

m <- mash_standard_run(path = path, numSNPs = numSNPs, U_hyp = Umidwhyp, 
                       suffix = suffix, saveoutput = TRUE)
```
Error in mashr::mash_set_data(Bhat_random, Shat_random) :
  Shat cannot contain NaN/Inf values
  problem for midwest...

U_hyp must be numeric matrices. They must be symmetric, square, in the same order as Bhat etc, and with the same # rows and cols as Bhat etc.

They can't contain NaN, NA, Inf values. And it might help if they're positive definite.

That's all the info in the mash package that I can find.

I suppose that I want them to be on the same scale as the mash input data, not the same scale as whatever weather or distance data they came from. This is a sense in which correlation matrices would be better. Take a look at the U_ed. Rescale to a similar scale as ED_PCA_1. But then...

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
numSNPs <- 30000
suffix = "gulf_uhyp1"
Ugulfhyp <- read_rds(path = file.path(path, "U_hyp_gulf_1phe.rds"))
Udistg <- Ugulfhyp$U_dist
U_h1 <- list(Udist = as.matrix(Udistg))
isSymmetric(U_h1$Udist)
typeof(Udistg)

U_h2 <- list(Udist = Ugulfhyp$U_dist, UGDD = Ugulfhyp$U_CGDD_12C)

U_ed <- file.path(path, "Mash_data_driven_covariances_30000_gulf_uhyp1.rds")

m <- mash_standard_run(path = path, numSNPs = numSNPs, U_ed = U_ed,
                       U_hyp = U_h1, suffix = suffix, saveoutput = TRUE)
m <- mash_standard_run(path = path, numSNPs = numSNPs, U_ed = U_ed,
                       U_hyp = U_h2, suffix = suffix, saveoutput = TRUE)
```

```{r}
path <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", 
                       "gwas", "GDD", "Uhyp")
m <- readRDS(file.path(path, "Strong_Effects30000SNPs_gulf_uhyp1.rds"))
suffix = "gulf_uhyp1"
mash_plot_covar(m, saveoutput = TRUE, suffix = suffix)
mash_plot_Ulist(m, range = get_U_by_mass(m))
mash_plot_sig_by_condition(m, saveoutput = TRUE, suffix = suffix)
mash_plot_pairwise_sharing(m, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 1, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 2, saveoutput = TRUE, suffix = suffix)
mash_plot_effects(m, n = 8)

mash_plot_manhattan_by_condition(m, saveoutput = TRUE, suffix = suffix)
```

```{r}
load_all("~/Github/switchgrassGWAS")
fitted_g <- list(Ulist = u_ed)
m_mock <- list(fitted_g = fitted_g)
mash_plot_Ulist(m_mock, limits = FALSE)

bhat %>% as_tibble() %>%
  colMeans()
summary(bhat)

# CLMB_FL50 KBSM_FL50 KING_FL50 PKLE_FL50 TMPL_FL50 
240.7987 - 173  
261.9135 - 173   
209.8265 - 173  
210.2195 - 173  
240.0100 - 173

B_hat_strong_df_30000topSNPs_gulf_solstice <- bhat %>%
  mutate(Bhat_CLMB_FL50 = Bhat_CLMB_FL50 + (240.7987 - 173),
         Bhat_KING_FL50 = Bhat_KING_FL50 + (209.8265 - 173 ),
         Bhat_KBSM_FL50 = Bhat_KBSM_FL50 + (261.9135 - 173),
         Bhat_PKLE_FL50 = Bhat_PKLE_FL50 + (210.2195 - 173),
         Bhat_TMPL_FL50 = Bhat_TMPL_FL50 + (240.0100 - 173),
         ) #%>%
  #summary()
write_rds(B_hat_strong_df_30000topSNPs_gulf_solstice, file.path(path, "B_hat_strong_df_30000topSNPs_gulf_solstice.rds"))

B_hat_random_df_30000topSNPs_gulf_solstice <- B_hat_random_df_30000topSNPs_gulf_uhyp1 %>%
  mutate(Bhat_CLMB_FL50 = Bhat_CLMB_FL50 + (240.7987 - 173),
         Bhat_KING_FL50 = Bhat_KING_FL50 + (209.8265 - 173 ),
         Bhat_KBSM_FL50 = Bhat_KBSM_FL50 + (261.9135 - 173),
         Bhat_PKLE_FL50 = Bhat_PKLE_FL50 + (210.2195 - 173),
         Bhat_TMPL_FL50 = Bhat_TMPL_FL50 + (240.0100 - 173),
         ) 
write_rds(B_hat_random_df_30000topSNPs_gulf_solstice, file.path(path, "B_hat_random_df_30000topSNPs_gulf_solstice.rds"))
```
what is each site's mean FL50? Just for Gulf plants.

For interpretability of the mash results... may want to scale all these effects so that 0 is, say, the summer solstice? Instead of the mean flowering time at that site for that pop.
```{r}
phe_gulf <- phenotypes$Gulf %>%
  select(PLANT_ID, ends_with("FL50"))

sd(phe_gulf$CLMB_FL50, na.rm = TRUE)
sd(phe_gulf$KBSM_FL50, na.rm = TRUE)
sd(phe_gulf$KING_FL50, na.rm = TRUE)
sd(phe_gulf$PKLE_FL50, na.rm = TRUE)
sd(phe_gulf$TMPL_FL50, na.rm = TRUE)

summary(phe_gulf$CLMB_FL50, na.rm = TRUE)
summary(phe_gulf$KBSM_FL50, na.rm = TRUE)
summary(phe_gulf$KING_FL50, na.rm = TRUE)
summary(phe_gulf$PKLE_FL50, na.rm = TRUE)
summary(phe_gulf$TMPL_FL50, na.rm = TRUE)
summary(phe_gulf)

```
CLMB_FL50 KBSM_FL50 KING_FL50 PKLE_FL50 TMPL_FL50 
240.7987   261.9135  209.8265  210.2195 240.0100

241 +/- 18.6
262 +/- 10.5
210 +/- 34
210 +/- 35
240 +/- 28