### 2020-07-17 R/Plots_manuscript_v0.2.Rmd
Add GR50 and FL50 specific functions of weather for ASReml h2 analysis

---
title: "Analyses v0.2"
author: "Alice MacQueen"
date: "7/17/2020"
output: html_document
---

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
wea_df <- readRDS("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```
```{r}
jday_phe <- wea_df %>%
  group_by(SITE) %>%
  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         FL50 = yday(DOY)) %>%
  select(DOY, DYLN, dyln_change_sec, everything(), -PTT_12C) %>%
  right_join(jday_phe, by = c("SITE", "FL50")) %>%
  select(-(TMAX:CPTT_11C))
jday_phe <- metadata %>%
  rename(Lat_origin = LATITUDE, Long_origin = LONGITUDE) %>%
  select(PLANT_ID, SUBPOP, Lat_origin, Long_origin) %>%
  right_join(jday_phe) %>%
  left_join(sites)

jday_phe$SUBPOP <- factor(jday_phe$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
jday_phe$SITE <- factor(jday_phe$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
jday_phe$manu_site <- factor(jday_phe$manu_site, levels = rev(c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD")))
```

# Overview
Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.2.

# Greenup h2
Greenup as a function of Julian date had extremely low heritability within subpopulations (0.005 +/- 0.003 Gulf; 0.021 +/0 0.011 for Midwest) across all eight common gardens, but high heritability across both subpopulations (h2 = 0.987 +/- 0.006).

<did any environmental cue beat GR50 to estimate greenup?>.

## define environmental cues for greenup
Probably all CGDD related.
```{r}

asreml_fl <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, FL50, GR50, CGDD_12C, CRAIN, CRAIN_5d, DYLN, dyln_change_sec) %>%
  rename(dyln_fl50 = DYLN, cgdd_12c_gr2fl = CGDD_12C, crain_gr2fl = CRAIN) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(between(DOY, as_date(FL50, origin = "2019-01-01") - 6, as_date(FL50, origin = "2019-01-01"))) %>%
  mutate(crain_2d = RAIN_SM + lag(RAIN_SM),
         crain_3d = RAIN_SM + lag(RAIN_SM) + lag(RAIN_SM, n = 2),
         crain_7d = sum(RAIN_SM)) %>%
  filter(DOY == as_date(FL50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX:TMIN_SM), -(GDD_13C_SM:CPTT_13C)) %>%
  rename(crain_1d = RAIN_SM, crain_5d = CRAIN_5d)
saveRDS(asreml_fl, file = "../data/Weather_related_phe_for_asreml_h2_FL50.rds")
asreml_gr <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, GR50) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(DOY <= as_date(GR50, origin = "2019-01-01")) %>%
  mutate(cgdd_12c_jan2gr = sum(GDD_12C_SM),
         cgdd_8c_jan2gr = sum(GDD_8C_SM)
         ) %>%
  select(-(GDD_13C_SM:CPTT_13C), -(TMAX:Longitude), GDD_12C_SM) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 17, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_18d = mean(TMAX_SM),
         tmin_18d = mean(TMIN_SM),
         tave_18d = (tmax_18d + tmin_18d)/2,
         cgdd_12c_18d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 9, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_10d = mean(TMAX_SM),
         tmin_10d = mean(TMIN_SM),
         tave_10d = (tmax_10d + tmin_10d)/2,
         cgdd_12c_10d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 4, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_5d = mean(TMAX_SM),
         tmin_5d = mean(TMIN_SM),
         tave_5d = (tmax_5d + tmin_5d)/2,
         cgdd_12c_5d = mean(GDD_12C_SM))%>%
  filter(DOY == as_date(GR50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX_SM:RAIN_SM), -Day_of_year)
saveRDS(asreml_gr, file = "../data/Weather_related_phe_for_asreml_h2_GR50.rds")
```

## ASReml h2 greenup analysis
```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)
```

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

```

### loop to make heritability table using ASReml mixed models and vpredict.
```{r}
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", IL = "IL", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), nine_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
           "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d",
           "crain_7d")
phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
            "cgdd_12c_5d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)
h2_table <- tibble()

for(p in seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in seq_along(site_v)){
    for(j in seq_along(subpop_v)){
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
    left_join(sites) %>%
    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
    rename(phenotype = eval(phe_v[k])) %>%
    mutate(PLANT_ID = as_factor(PLANT_ID),
           SITE = as_factor(SITE),
           PLOT_GL = as_factor(PLOT_GL)) %>%
    as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(SITE),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
      mutate(site = site_v[i],
             subpop = subpop_v[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  } else {
    asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  }

  if(i == 1 & j == 1 & k == 1){
    h2_table <- h2_est
  } else {
    h2_table <- add_row(h2_table, h2_est)
  }
  tryCatch({
      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
                                                 names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".png")), 
            plot = plot(asr_out))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_", 
                                                  "functions_of_weather.csv")))
}

```

# Make new environmental variables for Li to use?
Can I estimate heritability for the fourway cross?