### 2020-07-18 R/Plots_manuscript_v0.2.Rmd
phenotype df for gwas for G_BLUP GR & FL

---
title: "Analyses v0.2"
author: "Alice MacQueen"
date: "7/17/2020"
output: html_document
---

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
library(switchgrassGWAS)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
wea_df <- readRDS("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```
```{r}
jday_phe <- wea_df %>%
  group_by(SITE) %>%
  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         FL50 = yday(DOY)) %>%
  select(DOY, DYLN, dyln_change_sec, everything(), -PTT_12C) %>%
  right_join(jday_phe, by = c("SITE", "FL50")) %>%
  select(-(TMAX:CPTT_11C))
jday_phe <- metadata %>%
  rename(Lat_origin = LATITUDE, Long_origin = LONGITUDE) %>%
  select(PLANT_ID, SUBPOP, Lat_origin, Long_origin) %>%
  right_join(jday_phe) %>%
  left_join(sites)

jday_phe$SUBPOP <- factor(jday_phe$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
jday_phe$SITE <- factor(jday_phe$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
jday_phe$manu_site <- factor(jday_phe$manu_site, levels = rev(c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD")))
```

# Overview
Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.2.

# Greenup h2
Greenup as a function of Julian date had extremely low heritability within subpopulations (0.005 +/- 0.003 Gulf; 0.021 +/0 0.011 for Midwest) across all eight common gardens, but high heritability across both subpopulations (h2 = 0.987 +/- 0.006).

<did any environmental cue beat GR50 to estimate greenup?>.

## define environmental cues for greenup
Probably all CGDD related.
```{r}

asreml_fl <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, FL50, GR50, CGDD_12C, CRAIN, CRAIN_5d, DYLN, dyln_change_sec) %>%
  rename(dyln_fl50 = DYLN, cgdd_12c_gr2fl = CGDD_12C, crain_gr2fl = CRAIN) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(between(DOY, as_date(FL50, origin = "2019-01-01") - 6, as_date(FL50, origin = "2019-01-01"))) %>%
  mutate(crain_2d = RAIN_SM + lag(RAIN_SM),
         crain_3d = RAIN_SM + lag(RAIN_SM) + lag(RAIN_SM, n = 2),
         crain_7d = sum(RAIN_SM)) %>%
  filter(DOY == as_date(FL50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX:TMIN_SM), -(GDD_13C_SM:CPTT_13C)) %>%
  rename(crain_1d = RAIN_SM, crain_5d = CRAIN_5d)
saveRDS(asreml_fl, file = "../data/Weather_related_phe_for_asreml_h2_FL50.rds")
asreml_gr <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, GR50) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(DOY <= as_date(GR50, origin = "2019-01-01")) %>%
  mutate(cgdd_12c_jan2gr = sum(GDD_12C_SM),
         cgdd_8c_jan2gr = sum(GDD_8C_SM)
         ) %>%
  select(-(GDD_13C_SM:CPTT_13C), -(TMAX:Longitude), GDD_12C_SM) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 17, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_18d = mean(TMAX_SM),
         tmin_18d = mean(TMIN_SM),
         tave_18d = (tmax_18d + tmin_18d)/2,
         cgdd_12c_18d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 9, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_10d = mean(TMAX_SM),
         tmin_10d = mean(TMIN_SM),
         tave_10d = (tmax_10d + tmin_10d)/2,
         cgdd_12c_10d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 4, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_5d = mean(TMAX_SM),
         tmin_5d = mean(TMIN_SM),
         tave_5d = (tmax_5d + tmin_5d)/2,
         cgdd_12c_5d = mean(GDD_12C_SM))%>%
  filter(DOY == as_date(GR50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX_SM:RAIN_SM), -Day_of_year)
saveRDS(asreml_gr, file = "../data/Weather_related_phe_for_asreml_h2_GR50.rds")
```

## ASReml h2 greenup analysis
```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)
```

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

```

### loop to make heritability table using ASReml mixed models and vpredict.
```{r}
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", IL = "IL", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), nine_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
           "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d",
           "crain_7d")
phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
            "cgdd_12c_5d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)
h2_table <- tibble()

for(p in seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in 6:9){
    for(j in seq_along(subpop_v)){
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
    left_join(sites, by = "SITE") %>%
    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
    rename(phenotype = eval(phe_v[k])) %>%
    mutate(PLANT_ID = as_factor(PLANT_ID),
           SITE = as_factor(SITE),
           PLOT_GL = as_factor(PLOT_GL)) %>%
    as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(SITE),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  } else {
    tryCatch({
    asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }

  if(i == 1 & j == 1 & k == 1){
    h2_table <- h2_est
  } else {
    h2_table <- add_row(h2_table, h2_est)
  }
  tryCatch({
      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
                                                 names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".png")), 
            plot = plot(asr_out))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_", 
                                                  "functions_of_weather_v2.csv")))
}

```

## BLUPs for G

We next evaluated genetic effects of greenup as functions of GDD and minimum temperature 10 days prior to greenup, and genetic effects of flowering as functions of daylength at flowering and GDD between greenup and flowering. To do this, we calculated BLUPs for each individual using mixed models with common garden and a kinship matrix as random effects. We then conducted genome-wide association on these BLUPs both within and between genetic subpopulations.
```{r}
site_v <- list(eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl")
phe_gr <- c("GR50", "tmin_10d", "cgdd_12c_10d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)
h2_table <- tibble()

i = 1
j = 1
k = 1
p = 2

for(p in seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in seq_along(site_v)){
    for(j in seq_along(subpop_v)){
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
    left_join(sites, by = "SITE") %>%
    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
    rename(phenotype = eval(phe_v[k])) %>%
    mutate(PLANT_ID = as_factor(PLANT_ID),
           SITE = as_factor(SITE),
           PLOT_GL = as_factor(PLOT_GL)) %>%
    as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(manu_site),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
    blup <- summary(asr_out, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
    write_csv(blup, file.path(outputdir, paste0("blups_", names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".csv")))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  } else {
    tryCatch({
    asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }

  if(i == 1 & j == 1 & k == 1){
    h2_table <- h2_est
  } else {
    h2_table <- add_row(h2_table, h2_est)
  }
  tryCatch({
      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
                                                 names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".png")), 
            plot = plot(asr_out))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

      }
    }
  }

}

```

After this I'll need to select from each csv just the individuals in the subpop that I included phenotypes for, to do GWAS on. Then do GWAS with bigsnpr, and go ahead and look at those results.

# gwas

set up phenotype df for gwas
```{r}
site_v <- list(eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl")
phe_gr <- c("GR50", "tmin_10d", "cgdd_12c_10d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)

i = 1
j = 1
k = 1
p = 2
phe_gwas <- tibble()

for(j in seq_along(subpop_v)){
  for(p in seq_along(phe_dfs)){
    phe_v <- phe_lists[[p]]
    phe <- phe_dfs[[p]]
    for(i in seq_along(site_v)){
      for(k in seq_along(phe_v)){
        phe_raw <- read_csv(file.path(outputdir, paste0("blups_",
                                                        names(site_v)[i], "_",
                                                        names(subpop_v)[j],
                                                        "_", phe_v[k],
                                                        ".csv")))
        names(phe_raw)[2] <- paste0(phe_v[k], "_G_BLUP")
        phe_proc <- phe_raw %>%
          separate(Effect, into = c("factor", "PLANT_ID"), sep = "\\)_") %>%
          filter(!is.na(PLANT_ID)) %>%
          left_join(metadata, by = "PLANT_ID") %>%
          filter(SUBPOP %in% subpop_v[[j]]) %>%
          select(PLANT_ID, 3)
        if(i == 1 & k == 1 & p == 1){
          phe_gwas <- phe_proc
        } else {
          phe_gwas <- phe_gwas %>%
            left_join(phe_proc, by = "PLANT_ID")
        }
      }
    }
  }
write_csv(phe_gwas, path = file.path(outputdir, paste0("G_BLUP_phenotypes_",
                                                       "for_gwas_GR_FL_",
                                                       names(subpop_v)[j],
                                                       ".csv")))
}
```

```{r}
library(bigsnpr)
library(tidyverse)
library(switchgrassGWAS)
library(AnnotationDbi)
txdb <- loadDb(file = file.path("~", "Github", "pvdiv-genome",
                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
inputdir <- file.path(workingdir, "analysis", "heritability")
outputdir <- file.path("~", "Github", "pvdiv-phenology-gxe", "analysis", "gwas", "G_BLUP")
subpop_v <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", Gulf = "Gulf")

inputfiles <- read_delim("~/Github/pvdiv-phenology-gxe/analysis/gwas/inputkinship.txt", delim = " ", col_names = "SNPfiles")

for(i in seq_along(subpop_v)){
  snp <- snp_attach(inputfiles$SNPfiles[i])
  phe_df <- read_rds(file.path(inputdir, paste0("G_BLUP_phenotypes_",
                                                "for_gwas_GR_FL_",
                                                names(subpop_v)[i], ".csv")))
  pvdiv_standard_gwas(snp, df = phe_df, type = "linear", outputdir = outputdir,
                      savegwas = TRUE, saveannos = TRUE, txdb = txdb,
                      minphe = 130)
}
```

# Make new environmental variables for Li to use?

We next evaluated genetic effects of greenup as functions of GDD (12C) and minimum temperature 10 days prior to greenup, and genetic effects of flowering as functions of daylength at flowering and GDD between greenup and flowering.

Can I estimate heritability for the fourway cross?