### 2020-07-22 R/Plots_manuscript_v0.3.Rmd

---
title: "Plots_manuscript_v0.3"
author: "Alice MacQueen"
date: "7/22/2020"
output: html_document
---

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
library(switchgrassGWAS)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```

# Overview
Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.

v0.2 has the ASReml h2 analysis, has BLUPs for G (which as of v0.3 probably needs a GxE term...)