### 2020-07-23 R/Plots_manuscript_v0.3.Rmd
Add manuscript work and GR50 and FL50 distribution correlation and h2 figure

---
title: "Plots_manuscript_v0.3"
author: "Alice MacQueen"
date: "7/22/2020"
output: html_document
---

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
library(switchgrassGWAS)
library(GGally)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
h2fl <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_FL50_related_functions_of_weather_v3.csv")
h2gr <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_GR50_related_functions_of_weather_v2.csv")
```

# Overview
Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.

v0.2 has the ASReml h2 analysis, has BLUPs for G (which as of v0.3 probably needs a GxE term... define some possibilities below.)

Results: h2 at single sites. Then h2 at all eight sites. And so there must be GxE, and there must be a lot of rank-changing GxE, in that we can’t predict ranks across sites.

To show this, make barplots of the different variables and facet them by site. Then, show heritability across all eight sites (in a model with a random effect of site, interacted with some number of eigenvectors from a SVD of the kinship matrix.)

# Heritability Plots

## h2 as functions of GR50 and FL50
```{r}
h2all <- h2fl %>%
  full_join(h2gr)

f1h2 <- h2all %>%
  filter(phe %in% c("FL50", "GR50")) %>%
  mutate(site = case_when(is.na(site) ~ "all",
                          site == "eight_sites" ~ "all",
                          TRUE ~ site),
         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
                                 subpop == "three_subpops" ~ "all",
                                 TRUE ~ subpop),
         phe_name = case_when(phe == "GR50" ~ "50% greenup",
                              phe == "FL50" ~ "50% flowering",
                              TRUE ~ phe))
f1h2$site <- factor(f1h2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "all", "nine_sites"))
f1h2$phe_name <- factor(f1h2$phe_name, levels = c("50% greenup", "50% flowering"))
f1h2$subpop_name <- factor(f1h2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
ggh21 <- f1h2 %>%
  filter(!(site %in% c("nine_sites", "IL")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
  ggplot(aes(x = site, y = Estimate)) +
  geom_bar(aes(fill = site == "all"), stat = "identity") +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  facet_grid(phe_name ~ subpop_name) +
  theme(panel.spacing.x = unit(0.1, 'cm'), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = "none") + 
  ylim(c(0, 1)) +
  xlab("Common Gardens") + 
  ylab(bquote('h'^2)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
  #
save_plot("../manuscript/plots/h2_GR_FL_site_by_subpop.png", plot = ggh21, base_height = 3.5, base_asp = 1.618)

f1h2 %>%
  filter(!(site %in% c("all", "nine_sites", "IL")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
  ungroup() %>%
  group_by(phe) %>%
  summarise(mean = mean(Estimate))
f1h2 %>%
  filter((site %in% c("all")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
  ungroup() %>%
  group_by(phe) %>%
  summarise(mean = mean(Estimate))
```

## h2 flowering weather
```{r}
flh2 <- h2fl %>%
  filter(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")) %>%
  mutate(site = case_when(is.na(site) ~ "all",
                          site == "eight_sites" ~ "all",
                          TRUE ~ site),
         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
                                 subpop == "three_subpops" ~ "all",
                                 TRUE ~ subpop),
         phe_name = case_when(phe == "GR50" ~ "50% greenup",
                              phe == "FL50" ~ "50% flowering",
                              TRUE ~ phe))
flh2$site <- factor(flh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "all", "nine_sites"))
flh2$phe_name <- factor(flh2$phe_name, levels = c("50% greenup", "50% flowering"))
flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))

flh2_p1 <- flh2 %>%
  filter(subpop == "Gulf" & !(site %in% c("nine_sites")) & !is.na(site)) %>%
  ggplot(aes(x = phe, y = Estimate)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  facet_wrap(~ site) +
  geom_hline(data = filter(flh2, subpop == "Gulf" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0, 1))
save_plot("../manuscript/plots/h2_FL_site_weather_functions.png", plot = flh2_p1, base_height = 3.5, base_asp = 1.618)
flh2 %>%
  filter(subpop == "Gulf_and_Midwest" & !(site %in% c("eight_sites", "nine_sites")) & !is.na(site)) %>%
  ggplot(aes(x = phe, y = Estimate)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  facet_wrap(~ site) +
  geom_hline(data = filter(flh2, subpop == "Gulf_and_Midwest" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0,1))

```
## h2 greenup weather
```{r}

h2gr$site <- factor(h2gr$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
flh2 %>%
  filter(subpop == "Gulf" & !(site %in% c("eight_sites", "nine_sites")))

grh2_p1 <- h2gr %>%
  filter(subpop == "Gulf" & !(site %in% c("eight_sites", "nine_sites")) & !is.na(site)) %>%
  ggplot(aes(x = phe, y = Estimate)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  facet_wrap(~ site) +
  geom_hline(data = filter(flh2, subpop == "Gulf" & !(site %in% c("NA", "nine_sites")) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + ylim(c(0, 1))
save_plot("../manuscript/plots/h2_GR_site_weather_functions.png", plot = grh2_p1, base_height = 3.5, base_asp = 1.618)
```
## Phenotype correlations
can you easily look at the correlation between greenup ranks at OK versus NE?
```{r}
phe_gr_df %>%
  filter(SITE %in% c("LINC", "STIL")) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, -PLOT_GL) %>%
  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
  ggplot(aes(x = STIL, y = LINC)) +
  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
  scale_fill_viridis(option = "B",
                     end = 0.9, direction = -1)

phe_gr_df %>%
  filter(SITE %in% c("STIL", "PKLE")) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, -PLOT_GL) %>%
  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
  ggplot(aes(x = STIL, y = Lat_origin)) +
  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
  scale_fill_viridis(option = "B",
                     end = 0.9, direction = -1)

cor_gr <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor_fl <- phe_fl_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  select(PLANT_ID:FL50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "FL50", values_fn = mean)

cor_north <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  filter(Lat_origin > 40) %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor_south <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  filter(Lat_origin < 32) %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor(cor_north[,5:13], use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
cor(cor_south[,5:13], use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)

cor(cor_gr$STIL, cor_gr$LINC, use = "pairwise")
cor(cor_gr[,5:13], use = "pairwise") %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
cor(cor_fl[,5:13], use = "pairwise") %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)

# two different ways. Think I like the second more. Force latitudinal order.
cor_gr %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  cor(use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
```

```{r}
pand <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pane <- cor_fl %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panf <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pana <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = c(0.4, 0.85))
panb <- cor_gr %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panc <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")

pheno_corr <- plot_grid(pana, panb, panc, pand, pane, panf, nrow = 2)
save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.png", plot = pheno_corr, base_height = 5)
```

# --------------
# SVD of K

## Plot eigenvectors to capture predominant patterns in K matrix.
```{r}
k_full <- read_rds("../data/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
```

```{r}
b1 <- svd(k_full)
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
b1$d[2] / sum(b1$d) # 7.1
b1$d[3] / sum(b1$d) # 6.1
b1$d[4] / sum(b1$d) # 1.7
b1$d[5] / sum(b1$d) # 1.3
sum(b1$d[1:5]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
for(g in 1:5){
barplot(b1$v[,g]/k_full[which.max(abs(k_full[,g])),g])
}
order <- enframe(colnames(k_full), name = NULL, value = "PLANT_ID")
ggb1 <- tibble(`First eigenvector (PVE = 11.5%)` = b1$v[,1]/k_full[which.max(abs(k_full[,1])),1],
       `Second eigenvector (PVE = 7.1%)` = b1$v[,2]/k_full[which.max(abs(k_full[,2])),2],
       `Third eigenvector (PVE = 6.1%)` = b1$v[,3]/k_full[which.max(abs(k_full[,3])),3],
       `Fourth eigenvector (PVE = 1.7%)` = b1$v[,4]/k_full[which.max(abs(k_full[,4])),4],
       `Fifth eigenvector (PVE = 1.3%)` = b1$v[,5]/k_full[which.max(abs(k_full[,5])),5],
       `Sixth eigenvector (PVE = 0.9%)` = b1$v[,6]/k_full[which.max(abs(k_full[,6])),6],
       PLANT_ID = order$PLANT_ID) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))

ggb1 <- ggb1 %>%
  arrange(LATITUDE) %>%
  mutate(rank = row_number()) %>%
  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
                           levels = c("First eigenvector (PVE = 11.5%)", 
                                      "Second eigenvector (PVE = 7.1%)", 
                                      "Third eigenvector (PVE = 6.1%)", 
                                      "Fourth eigenvector (PVE = 1.7%)",
                                      "Fifth eigenvector (PVE = 1.3%)",
                                      "Sixth eigenvector (PVE = 0.9%)"
                                      ))
ggb1_plot <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = SUBPOP, fill = SUBPOP), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, end = 0.9, direction = -1) +
  scale_fill_viridis(discrete = TRUE, end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 607, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors.png", base_height = 5, base_asp = 2, plot = ggb1_plot)

ggb1_plot2 <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
                      direction = -1) +
  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 607, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
```

The eigenvectors capture the predominant patterns in the Uk3 covariance matrix.
```
v <- m$fitted_g$Ulist$ED_PCA_1
b1=svd(v)
for(g in 1){
barplot(b1$v[,g]/v[which.max(abs(v[,g])),g])
}

b1$d # the top eigenvector captures the predominant pattern in the ED_PCA_1 covariance matrix. 
```

WAIT. I could also plot this for the svd from the whole SNP set. That might make more sense than doing a svd on the kinship matrix...

# -------------
# ASReml models for eight sites including GxE
```{r}

```

## ASReml h2 greenup analysis
```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)
```

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

```

### loop to make heritability table using ASReml mixed models and vpredict.

## Include just texas sites for greenup and just northern four sites.
```{r}
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", IL = "IL", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), nine_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
           "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d",
           "crain_7d")
phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
            "cgdd_12c_5d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)
h2_table <- tibble()

for(p in 1){#seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in 12:13){#seq_along(site_v)){
    for(j in 1:3){#seq_along(subpop_v)){
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
    left_join(sites, by = "SITE") %>%
    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
    rename(phenotype = eval(phe_v[k])) %>%
    mutate(PLANT_ID = as_factor(PLANT_ID),
           SITE = as_factor(SITE),
           PLOT_GL = as_factor(PLOT_GL)) %>%
    as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(SITE), #:vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  } else {
    tryCatch({
    asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }

  if(i == 1 & j == 1 & k == 1){
    h2_table <- h2_est
  } else {
    h2_table <- add_row(h2_table, h2_est)
  }
  tryCatch({
      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
                                                 names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".png")), 
            plot = plot(asr_out))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_functions_", 
                                                  "of_weather_v4.csv")))
}