### 2020-07-25 R/Plots_manuscript_v0.3.Rmd
Fix phenotypes for 4WCR and plot FL and GR as functions of weather

---
title: "Plots_manuscript_v0.3"
author: "Alice MacQueen"
date: "7/22/2020"
output: html_document
---

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
library(switchgrassGWAS)
library(GGally)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
h2fl <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_FL50_related_functions_of_weather_v3.csv")
h2gr <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_GR50_related_functions_of_weather_v2.csv")
h2gr4 <- read_csv(file.path(workingdir, "analysis", "heritability", "h2_df_GR50_related_functions_of_weather_v4.csv"))
h2gr <- h2gr %>%
  full_join(h2gr4)
```

# Overview
Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.

v0.2 has the ASReml h2 analysis, has BLUPs for G (which as of v0.3 probably needs a GxE term... define some possibilities below.)

Results: h2 at single sites. Then h2 at all eight sites. And so there must be GxE, and there must be a lot of rank-changing GxE, in that we can’t predict ranks across sites.

To show this, make barplots of the different variables and facet them by site. Then, show heritability across all eight sites (in a model with a random effect of site, interacted with some number of eigenvectors from a SVD of the kinship matrix.)

# Heritability Plots

## h2 as functions of GR50 and FL50
```{r}
h2all <- h2fl %>%
  full_join(h2gr)

f1h2 <- h2all %>%
  filter(phe %in% c("FL50", "GR50")) %>%
  mutate(site = case_when(is.na(site) ~ "all",
                          site == "eight_sites" ~ "all",
                          TRUE ~ site),
         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
                                 subpop == "three_subpops" ~ "all",
                                 TRUE ~ subpop),
         phe_name = case_when(phe == "GR50" ~ "50% greenup",
                              phe == "FL50" ~ "50% flowering",
                              TRUE ~ phe))
f1h2$site <- factor(f1h2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "all", "nine_sites"))
f1h2$phe_name <- factor(f1h2$phe_name, levels = c("50% greenup", "50% flowering"))
f1h2$subpop_name <- factor(f1h2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
ggh21 <- f1h2 %>%
  filter(!(site %in% c("nine_sites", "IL")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
  ggplot(aes(x = site, y = Estimate)) +
  geom_bar(aes(fill = site == "all"), stat = "identity") +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  facet_grid(phe_name ~ subpop_name) +
  theme(panel.spacing.x = unit(0.1, 'cm'), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = "none") + 
  ylim(c(0, 1)) +
  xlab("Common Gardens") + 
  ylab(bquote('h'^2)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
  #
save_plot("../manuscript/plots/h2_GR_FL_site_by_subpop.png", plot = ggh21, base_height = 3.5, base_asp = 1.618)

f1h2 %>%
  filter(!(site %in% c("all", "nine_sites", "IL")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
  ungroup() %>%
  group_by(phe) %>%
  summarise(mean = mean(Estimate))
f1h2 %>%
  filter((site %in% c("all")) & !(subpop %in% c("three_subpops", "Atlantic"))) %>%
  ungroup() %>%
  group_by(phe) %>%
  summarise(mean = mean(Estimate))
```

## h2 flowering weather
```{r}
flh2 <- h2fl %>%
  filter(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")) %>%
  mutate(site = case_when(is.na(site) ~ "all",
                          site == "eight_sites" ~ "All",
                          TRUE ~ site),
         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
                                 subpop == "three_subpops" ~ "all",
                                 TRUE ~ subpop),
         phe_name = case_when(phe == "FL50" ~ "50% flowering",
                              phe == "dyln_fl50" ~ "daylength",
                              phe == "dyln_change_sec" ~ "daylength change (s)",
                              phe == "cgdd_12c_gr2fl" ~ "GDD, GR to FL",
                              phe == "crain_gr2fl" ~ "Rainfall, GR to FL",
                              phe == "crain_1d" ~ "Rainfall, 1 day",
                              phe == "crain_2d" ~ "Rainfall, 2d sum",
                              phe == "crain_3d" ~ "Rainfall, 3d sum",
                              phe == "crain_5d" ~ "Rainfall, 5d sum",
                              phe == "crain_7d" ~ "Rainfall, 7d sum",
                              TRUE ~ phe))
flh2$site <- factor(flh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "nine_sites"))
flh2$phe_name <- factor(flh2$phe_name, levels = c("50% flowering", "daylength",
                                                  "daylength change (s)",
                                                  "GDD, GR to FL",
                                                  "Rainfall, GR to FL",
                                                  "Rainfall, 1 day",
                                                  "Rainfall, 2d sum",
                                                  "Rainfall, 3d sum",
                                                  "Rainfall, 5d sum",
                                                  "Rainfall, 7d sum"
                                                  ))
flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD", "All", "Texas", "North")

flh2_p1 <- flh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)

save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_v2.png", plot = flh2_p1, base_height = 7, base_asp = 1.3)

site_plot <- c("All", "Texas", "North")

flh2_p2 <- flh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")

save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_v2.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)
```
## h2 greenup weather
```{r}
h2gr
```

```{r}
unique(h2gr$site)
grh2 <- h2gr %>%
  filter(!(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d", "DYLN"))) %>%
  mutate(site = case_when(is.na(site) ~ "all",
                          site == "eight_sites" ~ "All",
                          site == "tx_sites" ~ "Texas",
                          site == "north_sites" ~ "North",
                          TRUE ~ site),
         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
                                 subpop == "three_subpops" ~ "All",
                                 TRUE ~ subpop),
         phe_name = case_when(phe == "GR50" ~ "50% greenup",
                              phe == "FL50" ~ "50% flowering",
                              phe == "cgdd_12c_jan2gr" ~ "GDD 01-01 to greenup",
                              phe == "cgdd_8c_jan2gr" ~ "GDD 8C 01-01 to greenup",
                              phe == "tmax_18d" ~ "Max T, 18d mean",
                              phe == "tmin_18d" ~ "Min T, 18d mean",
                              phe == "tave_18d" ~ "Mean T, 18d mean",
                              phe == "tmax_10d" ~ "Max T, 10d mean",
                              phe == "tmin_10d" ~ "Min T, 10d mean",
                              phe == "tave_10d" ~ "Mean T, 10d mean",
                              phe == "tmax_5d" ~ "Max T, 5d mean",
                              phe == "tmin_5d" ~ "Min T, 5d mean",
                              phe == "tave_5d" ~ "Mean T, 5d mean",
                              phe == "cgdd_12c_10d" ~ "GDD, 10d",
                              phe == "cgdd_12c_18d" ~ "GDD, 18d",
                              phe == "cgdd_12c_5d" ~ "GDD, 5d",
                              TRUE ~ phe))
grh2$site <- factor(grh2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North", "nine_sites"))
grh2$phe_name <- factor(grh2$phe_name, levels = c("50% greenup", "GDD, 5d",
                                                  "Min T, 5d mean",
                                                  "Mean T, 5d mean",
                                                  "Max T, 5d mean",
                                                  "GDD, 10d",
                                                  "Min T, 10d mean",
                                                  "Mean T, 10d mean",
                                                  "Max T, 10d mean",
                                                  "GDD, 18d",
                                                  "Min T, 18d mean",
                                                  "Mean T, 18d mean",
                                                  "Max T, 18d mean",
                                                  "GDD 8C 01-01 to greenup",
                                                  "GDD 01-01 to greenup"))
grh2$subpop_name <- factor(grh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "All"))

subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
i=3
subpop_v[[i]]
site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")

grh2_p1 <- grh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v2.png", plot = grh2_p1, base_height = 7, base_asp = 1.3)

site_plot <- c( "All", "Texas", "North")

grh2_p2 <- grh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v3.png", plot = grh2_p2, base_height = 5, base_asp = 1.5)

site_plot <- c( "All")

grh2_p3 <- grh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.15)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_v4.png", plot = grh2_p3, base_height = 4, base_asp = 1.5)

gr_fl_p1 <- plot_grid(grh2_p3, flh2_p2, nrow = 2)
```
## Phenotype correlations
can you easily look at the correlation between greenup ranks at OK versus NE?
```{r}
phe_gr_df %>%
  filter(SITE %in% c("LINC", "STIL")) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, -PLOT_GL) %>%
  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
  ggplot(aes(x = STIL, y = LINC)) +
  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
  scale_fill_viridis(option = "B",
                     end = 0.9, direction = -1)

phe_gr_df %>%
  filter(SITE %in% c("STIL", "PKLE")) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, -PLOT_GL) %>%
  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
  ggplot(aes(x = STIL, y = Lat_origin)) +
  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
  scale_fill_viridis(option = "B",
                     end = 0.9, direction = -1)

cor_gr <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor_fl <- phe_fl_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  select(PLANT_ID:FL50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "FL50", values_fn = mean)

cor_north <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  filter(Lat_origin > 40) %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor_south <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  filter(Lat_origin < 32) %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor(cor_north[,5:13], use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
cor(cor_south[,5:13], use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)

cor(cor_gr$STIL, cor_gr$LINC, use = "pairwise")
cor(cor_gr[,5:13], use = "pairwise") %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
cor(cor_fl[,5:13], use = "pairwise") %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)

# two different ways. Think I like the second more. Force latitudinal order.
cor_gr %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  cor(use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
```

```{r}
pand <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pane <- cor_fl %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panf <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pana <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = c(0.4, 0.85))
panb <- cor_gr %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panc <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")

pheno_corr <- plot_grid(pana, panb, panc, pand, pane, panf, nrow = 2)
save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.png", plot = pheno_corr, base_height = 5)
```

# --------------
# SVD of K

## Plot eigenvectors to capture predominant patterns in K matrix.
```{r}
k_full <- read_rds("../data/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
```

```{r}
b1 <- svd(k_full)
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
b1$d[2] / sum(b1$d) # 7.1
b1$d[3] / sum(b1$d) # 6.1
b1$d[4] / sum(b1$d) # 1.7
b1$d[5] / sum(b1$d) # 1.3
sum(b1$d[1:5]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
for(g in 1:5){
barplot(b1$v[,g]/k_full[which.max(abs(k_full[,g])),g])
}
order <- enframe(colnames(k_full), name = NULL, value = "PLANT_ID")
ggb1 <- tibble(`First eigenvector (PVE = 11.5%)` = b1$v[,1]/k_full[which.max(abs(k_full[,1])),1],
       `Second eigenvector (PVE = 7.1%)` = b1$v[,2]/k_full[which.max(abs(k_full[,2])),2],
       `Third eigenvector (PVE = 6.1%)` = b1$v[,3]/k_full[which.max(abs(k_full[,3])),3],
       `Fourth eigenvector (PVE = 1.7%)` = b1$v[,4]/k_full[which.max(abs(k_full[,4])),4],
       `Fifth eigenvector (PVE = 1.3%)` = b1$v[,5]/k_full[which.max(abs(k_full[,5])),5],
       `Sixth eigenvector (PVE = 0.9%)` = b1$v[,6]/k_full[which.max(abs(k_full[,6])),6],
       PLANT_ID = order$PLANT_ID) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))

ggb1 <- ggb1 %>%
  arrange(LATITUDE) %>%
  mutate(rank = row_number()) %>%
  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
                           levels = c("First eigenvector (PVE = 11.5%)", 
                                      "Second eigenvector (PVE = 7.1%)", 
                                      "Third eigenvector (PVE = 6.1%)", 
                                      "Fourth eigenvector (PVE = 1.7%)",
                                      "Fifth eigenvector (PVE = 1.3%)",
                                      "Sixth eigenvector (PVE = 0.9%)"
                                      ))
ggb1_plot <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = SUBPOP, fill = SUBPOP), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, end = 0.9, direction = -1) +
  scale_fill_viridis(discrete = TRUE, end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 607, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors.png", base_height = 5, base_asp = 2, plot = ggb1_plot)

ggb1_plot2 <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
                      direction = -1) +
  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 607, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
```

The eigenvectors capture the predominant patterns in the Uk3 covariance matrix.
```
v <- m$fitted_g$Ulist$ED_PCA_1
b1=svd(v)
for(g in 1){
barplot(b1$v[,g]/v[which.max(abs(v[,g])),g])
}

b1$d # the top eigenvector captures the predominant pattern in the ED_PCA_1 covariance matrix. 
```

WAIT. I could also plot this for the svd from the whole SNP set. That might make more sense than doing a svd on the kinship matrix...

# -------------
# ASReml models for eight sites including GxE
```{r}

```

## ASReml h2 greenup analysis
```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)

source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

```

### loop to make heritability table using ASReml mixed models and vpredict.

## Include just texas sites for greenup and just northern four sites.
e.g.:
for(p in 1){#seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in 12:13){#seq_along(site_v)){ # may need to adjust below for h2_table
    for(j in 1:3){#seq_along(subpop_v)){
      for(k in seq_along(phe_v)){

## Model eight, texas, and north sites with a kinship : site interaction.
Then, find GxE and G proportions of variance.
```{r}
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", IL = "IL", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), nine_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
           "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d",
           "crain_7d")
phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
            "cgdd_12c_5d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)
h2_table <- tibble()

for(p in seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in c(10,12:13)){#seq_along(site_v)){ # may need to adjust below for h2_table
    for(j in 1:3){#seq_along(subpop_v)){
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
    left_join(sites, by = "SITE") %>%
    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
    rename(phenotype = eval(phe_v[k])) %>%
    mutate(PLANT_ID = as_factor(PLANT_ID),
           SITE = as_factor(SITE),
           PLOT_GL = as_factor(PLOT_GL)) %>%
    as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(SITE):vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  } else {
    tryCatch({
    asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }

  if(i == 12 & j == 1 & k == 1){
    h2_table <- h2_est
  } else {
    h2_table <- add_row(h2_table, h2_est)
  }
  tryCatch({
      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
                                                 names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".png")), 
            plot = plot(asr_out))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_functions_", 
                                                  "of_weather_v4.csv")))
}