### 2020-07-29 data-raw/Data_Setup.Rmd

---
title: "Data_Setup"
author: "Alice MacQueen"
date: "6/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# metadata (germplasm)
```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_),
         LATITUDE = abs(LATITUDE),
         LONGITUDE = abs(LONGITUDE))
write_rds(metadata, "../data/metadata.rds")
```

# common gardens
```{r}
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")

sites <- sites %>%
  mutate(manu_site = case_when(State != "TX" ~ State,
                               SITE == "KING" ~ "TX1",
                               SITE == "PKLE" ~ "TX2",
                               SITE == "TMPL" ~ "TX3",
                               SITE == "OVTN" ~ "TX4"))
sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))
write_rds(sites, "../data/sites.rds")
```

# Subpop color coding
```{r}
subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
                                SUBPOP == "Gulf" ~ "#F47F72",    
                                SUBPOP == "Midwest" ~ "#442C83",
                                SUBPOP == "4X" ~ "grey", 
                                SUBPOP == "8X" ~ "#090906"))
subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
write_rds(subpops, path = "../data/subpop_color_coding.rds")
```

### ------
# Make phenotypes for flowering time paper from all the cleaned phenotypes.

Specifically using GR50 & FL50 Julian dates from 2019. And using them to make cGDD between greenup and flowering for another GWAS phenotype.
```{r}
phenotypes <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_cleaned.rds")
ptt_df <- read_rds("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
fourway <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2019_Multiple_Environment.csv")
fwy_prev_raw <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/0-data/Phenology_2016_2018.csv")
```

Also made a year-specific "PLOT_GL" for the 4WCR. Can use this to calculate cGDD by specific plant, when a GR50 and FL50 measurement are linked for the same plant for a F0 or F1, for example.
```{r}
phe_gwas <- phenotypes %>%
  filter(PHE %in% c("EM1", "EM50", "FL1", "FL50", "GR1", "GR50", "GR100")) %>%
  mutate(YEAR = 2019,
         WHERE = "GWAS") %>%
  select(-manu_site)

fwy_prev <- fwy_prev_raw %>%
  separate(ENVI, into = c("SITE", "YEAR"), sep = "_") %>%
  group_by(LINE, SITE, YEAR) %>%
  mutate(PLOT_GL = paste0(str_sub(SITE, end = 1), str_sub(YEAR, start =3, end = 4), "_", row_number())) %>%
  mutate(YEAR = as.numeric(YEAR),
         WHERE = "FWCR") %>%
  pivot_longer(cols = BIOMASS:HT, names_to = "PHE", values_to = "MEAS") %>%
  rename(PLANT_ID = LINE) %>%
  filter(PHE %in% c("FL50", "GR50"))
fwy_2019 <- fourway %>%
  separate(ENVI, into = c("SITE", "YEAR"), sep = "_") %>%
  group_by(LINE, SITE, YEAR) %>%
  mutate(PLOT_GL = paste0(str_sub(SITE, end = 1), str_sub(YEAR, start = 3, 
                                                          end = 4), "_", row_number())) %>%
  mutate(YEAR = as.numeric(YEAR),
         WHERE = "FWCR") %>%
  pivot_longer(cols = FL50:HT, names_to = "PHE", values_to = "MEAS") %>%
  rename(PLANT_ID = LINE) %>%
  filter(PHE %in% c("FL50", "GR50"))

phe_all <- phe_gwas %>%
  full_join(fwy_2019, by = c("SITE", "PLOT_GL", "PLANT_ID", "YEAR", "PHE", "MEAS", "WHERE")) %>%
  full_join(fwy_prev, by = c("SITE", "PLOT_GL", "PLANT_ID", "YEAR", "PHE", "MEAS", "WHERE")) %>%
  filter(!is.na(MEAS))

phe_all %>%
  group_by(WHERE) %>%
  tally() %>% arrange((n))

write_rds(phe_all, path = "../data/Phenology_phenotypes_4wcr_and_gwas.rds")
```

cGDD between GR50 and FL50 for each YEAR, each PLANT_ID (and PLOT_GL)
```{r}
wea_df <- ptt_df %>%
  dplyr::select(DOY:GDD_12C_SM, PTT_12C) %>%
  group_by(SITE) %>%
  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60)

fl_timing <- phe_all %>% 
  filter(YEAR == 2019 & PHE %in% c("GR50", "FL50")) %>%
  pivot_wider(names_from = PHE, values_from = MEAS) %>%
  select(-SUBPOP) %>%
  drop_na()

wea_gr_fl <- wea_df %>%
  left_join(fl_timing)

wea_gr_fl2 <- wea_gr_fl %>%
  group_by(Day_of_year, PLOT_GL, PLANT_ID) %>%
  filter(between(Day_of_year, GR50, FL50)) %>% 
  ungroup()

cgdd_df <- wea_gr_fl2 %>%
  group_by(WHERE, SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
  summarise(CGDD_12C = sum(GDD_12C_SM),
            PTT_12C = sum(PTT_12C),
            CRAIN = sum(RAIN_SM)
            )

write_rds(cgdd_df, path = "../data/Julian_date_and_GDD_phenotypes_4wcr_and_gwas.rds")

wea_gr_fl3 <- wea_gr_fl %>%
  group_by(Day_of_year, PLOT_GL, PLANT_ID) %>%
  filter(between(Day_of_year, FL50-5, FL50)) %>% 
  ungroup()

cumweadf %>%
  group_by(WHERE) %>%
  tally() %>% arrange((n))

crain_df <- wea_gr_fl3 %>%
  group_by(WHERE, SITE, PLANT_ID, PLOT_GL, GR50, FL50) %>%
  summarise(CGDD_12C_5d = sum(GDD_12C_SM),
            PTT_12C_5d = sum(PTT_12C),
            CRAIN_5d = sum(RAIN_SM)
            )
wea_gr_fl4 <- wea_gr_fl %>%
  ungroup() %>%
  group_by(Day_of_year, PLOT_GL, PLANT_ID) %>%
  filter(Day_of_year == FL50) %>% 
  rename(dyln_fl50 = DYLN) %>%
  dplyr::select(-(TMAX:Day_of_year), -(TMAX_SM:PTT_12C))

cumweadf <- cgdd_df %>%
  left_join(crain_df) %>%
  left_join(wea_gr_fl4)

write_rds(cumweadf, path = "../data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```

```{r}
fourway = cumweadf[cumweadf$WHERE == "FWCR",]
df
fourway
```

### -------

# All phenotypes
```{r}
phe_all <- read_rds("../data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
phenology <- read_rds("../data/Phenology_phenotypes_4wcr_and_gwas.rds")
```

```{r}
phenology %>%
  pivot_wider(names_from = PHE, values_from = MEAS) %>%
  left_join(phe_all) %>%
  dplyr::select(-CGDD_12C_5d, -PTT_12C_5d) %>%
  pivot_longer(cols = EM1:CRAIN_5d, names_to = "PHE", values_to = "MEAS") %>%
  filter(!is.na(MEAS)) %>%
  group_by(PHE, WHERE, YEAR) %>%
  tally() %>% arrange(desc(YEAR))

phenology %>%
  pivot_wider(names_from = PHE, values_from = MEAS) %>%
  left_join(phe_all) %>%
  dplyr::select(-CGDD_12C_5d, -PTT_12C_5d) %>%
  pivot_longer(cols = EM1:CRAIN_5d, names_to = "PHE", values_to = "MEAS") %>%
  filter(!is.na(MEAS)) %>%
  write_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
```

# Phenotypes for GWAS
```{r}
phenotypes <- 
  read_rds("../data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")

phe_df <- phenotypes %>%
  filter(WHERE == "GWAS" & !(SITE %in% c("OVTN", "FRMI"))) %>%
  unite(col = "SITE_PHE", SITE, PHE) %>%
  filter(SUBPOP %in% c("Atlantic", "Midwest", "Gulf")) %>%
  select(-PLOT_GL, -WHERE, -YEAR) %>%
  pivot_wider(names_from = SITE_PHE, values_from = MEAS, values_fn = mean) %>%
  select(PLANT_ID, SUBPOP, matches("GDD"), matches("GR50"), matches("FL50"),
         matches("CRAIN"), matches("PTT"), everything())

midwgulf_df <- phe_df %>% filter(SUBPOP %in% c("Gulf", "Midwest")) 
midwgulf_df <- midwgulf_df %>%
  select(names(which(colSums(!is.na(midwgulf_df)) > 100))) %>%
  select(-SUBPOP)
midw_df <- phe_df %>% filter(SUBPOP == "Midwest") 
midw_df <- midw_df %>%
  select(names(which(colSums(!is.na(midw_df)) > 100))) %>%
  select(-SUBPOP)
gulf_df <- phe_df %>% filter(SUBPOP == "Gulf") 
gulf_df <- gulf_df %>%
  select(names(which(colSums(!is.na(gulf_df)) > 100))) %>%
  select(-SUBPOP)
atl_df <- phe_df %>% filter(SUBPOP == "Atlantic") 
atl_df <- atl_df %>%
  select(names(which(colSums(!is.na(atl_df)) > 100))) %>%
  select(-SUBPOP)
midwatl_df <- phe_df %>% filter(SUBPOP %in% c("Atlantic", "Midwest")) 
midwatl_df <- midwatl_df %>%
  select(names(which(colSums(!is.na(midwatl_df)) > 100))) %>%
  select(-SUBPOP)
gulfatl_df <- phe_df %>% filter(SUBPOP %in% c("Gulf", "Atlantic")) 
gulfatl_df <- gulfatl_df %>%
  select(names(which(colSums(!is.na(gulfatl_df)) > 100))) %>%
  select(-SUBPOP)

phe_subsets <- list(Midwest_and_Gulf = midwgulf_df, Midwest = midw_df, Gulf = gulf_df, Atlantic = atl_df, Midwest_and_Atlantic = midwatl_df, Gulf_and_Atlantic = gulfatl_df)
write_rds(phe_subsets, "../data/Phenotype_Subsets_for_Subpop_GWAS.rds")
names(phe_subsets)
```