### 2020-07-30 R/Plots_manuscript_v0.3.Rmd
Finish drafting first Results section v0.3 comparing heritabilities and GxE

---
title: "Plots_manuscript_v0.3"
author: "Alice MacQueen"
date: "7/22/2020"
output: html_document
---

Add any libraries/scripts that need to be loaded for these analyses here.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(lubridate)
library(switchgrassGWAS)
library(GGally)
library(maps)
library(ggmap)
usa <- map_data("usa")
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

# Load data
Add any data that needs to be loaded for these analyses here.
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
h2gr <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_GR50_related_functions_of_weather_v1.csv"))
h2fl <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "h2_df_FL50_related_functions_of_weather_v1.csv"))
h2gr_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_GR50_related_functions_of_weather_v1.csv"))
h2fl_gxe <- read_csv(file.path(workingdir, "analysis", "heritability", "kinship_subpop", "gxe_and_h2_df_FL50_related_functions_of_weather_v1.csv"))

h2gr <- h2gr %>%
  full_join(h2gr_gxe)
h2fl <- h2fl %>%
  full_join(h2fl_gxe)
```
### older files to load
h2fl <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_FL50_related_functions_of_weather_v3.csv")
h2gr <- read_csv("~/Github/pvdiv-phenology-gxe/analysis/heritability/h2_df_GR50_related_functions_of_weather_v2.csv")
h2gr4 <- read_csv(file.path(workingdir, "analysis", "heritability", "h2_df_GR50_related_functions_of_weather_v4.csv"))
h2gr <- h2gr %>%
  full_join(h2gr4)

# Overview
Here I'm putting the code for any new analyses that I'm running to do new writeup for the manuscript v0.3.

v0.2 has the ASReml h2 analysis, has BLUPs for G (which as of v0.3 probably needs a GxE term... define some possibilities below.)

Results: h2 at single sites. Then h2 at all eight sites. And so there must be GxE, and there must be a lot of rank-changing GxE, in that we can’t predict ranks across sites.

To show this, make barplots of the different variables and facet them by site. Then, show heritability across all eight sites (in a model with a random effect of site, interacted with some number of eigenvectors from a SVD of the kinship matrix.)

# Map Fig 1A

```{r}
subpops
library(ggrepel)

metaplot <- metadata %>%
  filter(SUBPOP %in% c("Gulf", "Midwest") & !is.na(LATITUDE)) %>%
  mutate(LONGITUDE = -LONGITUDE)
metaplot$ECOTYPE_NNET <- factor(metaplot$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
siteplot <- sites %>%
  filter(!(SITE %in% c("FRMI", "OVTN")))

ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) +
  geom_jitter(data = metaplot, aes(x = LONGITUDE, y = LATITUDE, 
                                   color = SUBPOP, shape = ECOTYPE_NNET),
              width = 0.45, height = 0.45) + 
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 21, size = 4, stroke = 1.5, fill = "white") +
  geom_point(data = siteplot, aes(x = Longitude, y = Latitude), shape = 4, size = 3, stroke = 1.5) +
  geom_label_repel(data = siteplot, aes(x = Longitude, y = Latitude, label = manu_site), alpha = 0.7) +
  labs(x = "Longitude", y = "Latitude") + 
  theme(legend.position = "right") +
  scale_color_manual(values = c("#F47F72", "#442C83")) +
  scale_shape_manual(values = c(0, 6, 1, 4)) 
  #scale_fill_manual(values = c("#280B54FF","#FAC127FF", "#D44842FF", "white")) +
  #scale_color_manual(values = c(rep("black", 10), "#280B54FF","#FAC127FF", "#D44842FF", "white", "#280B54FF","#FAC127FF", "#D44842FF", "white"), guide = FALSE) +
  #scale_alpha_manual(values = c(0.7, 0.7, 0.7, 0.7), guide = FALSE)
save_plot(filename = "../manuscript/plots/Map_subpops_and_Ecotypes.png", plot = last_plot(), base_height = 3)
?geom_point
```

# -- Figure 1 B
# Phenotype correlations
can you easily look at the correlation between greenup ranks at OK versus NE?
```{r}
phe_gr_df %>%
  filter(SITE %in% c("LINC", "STIL")) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, -PLOT_GL) %>%
  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
  ggplot(aes(x = STIL, y = LINC)) +
  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
  scale_fill_viridis(option = "B",
                     end = 0.9, direction = -1)

phe_gr_df %>%
  filter(SITE %in% c("STIL", "PKLE")) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, -PLOT_GL) %>%
  pivot_wider(names_from = "SITE", values_from = "GR50", values_fn = mean) %>%
  ggplot(aes(x = STIL, y = Lat_origin)) +
  geom_hex() + #aes(color = SUBPOP), height = 1, width = 1) +
  scale_fill_viridis(option = "B",
                     end = 0.9, direction = -1)

cor_gr <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor_fl <- phe_fl_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  select(PLANT_ID:FL50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "FL50", values_fn = mean)

cor_north <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  filter(Lat_origin > 40) %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor_south <- phe_gr_df %>%
  left_join(select(sites, SITE, manu_site)) %>%
  ungroup() %>%
  filter(Lat_origin < 32) %>%
  select(PLANT_ID:GR50, manu_site, -PLOT_GL, -SITE) %>%
  pivot_wider(names_from = "manu_site", values_from = "GR50", values_fn = mean)
cor(cor_north[,5:13], use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
cor(cor_south[,5:13], use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)

cor(cor_gr$STIL, cor_gr$LINC, use = "pairwise")
cor(cor_gr[,5:13], use = "pairwise") %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
cor(cor_fl[,5:13], use = "pairwise") %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)

# two different ways. Think I like the second more. Force latitudinal order.
cor_gr %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  cor(use = "pairwise") %>%
  switchgrassGWAS:::reorder_cormat(.) %>%
  ggcorr(data = NULL, cor_matrix = ., label = TRUE)
```

```{r}
pand <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pane <- cor_fl %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panf <- cor_fl %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
pana <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = c(0.4, 0.85))
panb <- cor_gr %>%
  filter(SUBPOP %in% c("Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")
panc <- cor_gr %>%
  filter(SUBPOP %in% c("Gulf", "Midwest")) %>%
  select(TX1, TX2, TX3, OK, MO, NE, MI, SD) %>%
  ggcorr(method = "pairwise", label = FALSE, layout.exp = 0.2, hjust = 0.95) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  theme(legend.position = "none")

pheno_corr <- plot_grid(pana, panb, panc, pand, pane, panf, nrow = 2)
save_plot(filename = "../manuscript/plots/Phenotypic_correlations_GR50_FL50_2.svg", plot = pheno_corr, base_width = 6.5)
```

# --- Fig 1 C & D
# Heritability Plots

Within common gardens, h2 were typically quite high: 62% on average for greenup date, and 87% for flowering date. However, h2 were not consistently high, notably for greenup at our OK and NE common gardens.
```{r}
h2all <- h2fl %>%
  full_join(h2gr)

h2all %>%
  filter(phe %in% c("FL50", "GR50") & site %in% c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD")) %>%
  group_by(phe) %>%
  summarise(meanh2 = mean(Estimate))

h2all %>%
  filter(phe %in% c("FL50", "GR50") & site %in% c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD")) %>%
  filter(Estimate < 0.2)

h2all %>%
  filter(phe %in% c("FL50", "GR50") & site %in% c("eight_sites")) %>%
  group_by(phe) %>%
  summarise(meanh2 = mean(Estimate))

h2all %>%
  filter(phe %in% c("GR50") & site %in% c("tx_sites", "north_sites")) %>%
  mutate(GandGxE = Estimate + GxE_Estimate) %>%
  select(GandGxE, Estimate:phe, GxE_Estimate) %>%
  group_by(site) %>%
  summarise(meanh2 = mean(Estimate + GxE_Estimate))

h2all %>%
  filter(site %in% c("eight_sites")) %>%
  mutate(GandGxE = Estimate + GxE_Estimate)  %>%
  select(GandGxE, site:phe, Estimate, GxE_Estimate)
```

## h2 as functions of GR50 and FL50
```{r}
h2all <- h2fl %>%
  full_join(h2gr)
unique(h2all$site)
f1h2 <- h2all %>%
  filter(phe %in% c("FL50", "GR50")) %>%
  mutate(site = case_when(is.na(site) ~ "all",
                          site == "eight_sites" ~ "All",
                          site == "tx_sites" ~ "Texas",
                          site == "north_sites" ~ "North",
                          TRUE ~ site),
         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
                                 subpop == "three_subpops" ~ "all",
                                 TRUE ~ subpop),
         phe_name = case_when(phe == "GR50" ~ "50% greenup",
                              phe == "FL50" ~ "50% flowering",
                              TRUE ~ phe))
f1h2$site <- factor(f1h2$site, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All"))#, "Texas", "North"))
f1h2$phe_name <- factor(f1h2$phe_name, levels = c("50% greenup", "50% flowering"))
f1h2$subpop_name <- factor(f1h2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
ggh21 <- f1h2 %>%
  filter(!(site %in% c("nine_sites", "IL")) & !is.na(subpop_name) & !is.na(phe_name) & !is.na(site)) %>%
  ggplot(aes(x = site, y = Estimate)) +
  geom_bar(aes(fill = site %in% c("All", "Texas", "North")), stat = "identity") +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(phe_name ~ subpop_name) +
  theme(panel.spacing.x = unit(0.1, 'cm'), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = "none") + 
  ylim(c(0, 1)) +
  xlab("Common Gardens") + 
  ylab(bquote('h'^2)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
  #
save_plot("../manuscript/plots/h2_GR_FL_site_by_subpop_k_specific.png", plot = ggh21, base_width = 5.5, base_height = 3)
```

## h2 flowering weather
```{r}
flh2 <- h2fl %>%
  filter(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")) %>%
  mutate(site = case_when(is.na(site) ~ "all",
                          site == "eight_sites" ~ "All",
                          site == "tx_sites" ~ "Texas",
                          site == "north_sites" ~ "North",
                          TRUE ~ site),
         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
                                 subpop == "three_subpops" ~ "all",
                                 TRUE ~ subpop),
         phe_name = case_when(phe == "FL50" ~ "50% flowering",
                              phe == "dyln_fl50" ~ "daylength",
                              phe == "dyln_change_sec" ~ "daylength change (s)",
                              phe == "cgdd_12c_gr2fl" ~ "GDD, GR to FL",
                              phe == "crain_gr2fl" ~ "Rainfall, GR to FL",
                              phe == "crain_1d" ~ "Rainfall, 1 day",
                              phe == "crain_2d" ~ "Rainfall, 2d sum",
                              phe == "crain_3d" ~ "Rainfall, 3d sum",
                              phe == "crain_5d" ~ "Rainfall, 5d sum",
                              phe == "crain_7d" ~ "Rainfall, 7d sum",
                              TRUE ~ phe))
flh2$site <- factor(flh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")))
flh2$phe_name <- factor(flh2$phe_name, levels = c("50% flowering", "daylength",
                                                  "daylength change (s)",
                                                  "GDD, GR to FL",
                                                  "Rainfall, GR to FL",
                                                  "Rainfall, 1 day",
                                                  #"Rainfall, 2d sum",
                                                  "Rainfall, 3d sum",
                                                  "Rainfall, 5d sum"#,
                                                  #"Rainfall, 7d sum"
                                                  ))
flh2$subpop_name <- factor(flh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "all"))
site_plot <- rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD", "All", "Texas", "North"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))

flh2_p1 <- flh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)

save_plot("../manuscript/plots/h2_FL_site_weather_functions_single_sites_k_specific.png", plot = flh2_p1, base_height = 7, base_asp = 1.1)

site_plot <- c("All", "Texas", "North")

flh2_p2 <- flh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.8)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")

save_plot("../manuscript/plots/h2_FL_site_weather_functions_site_groups_k_specific.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)

site_plot <- c("All")

flh2_p3 <- flh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.5)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
save_plot("../manuscript/plots/h2_FL_site_weather_functions_All_Sites_k_specific.png", plot = flh2_p3, base_height = 3.5)
```
## gxe flowering all, north, texas
```{r}
site_plot <- c("All", "Texas", "North")

fl_barplot2 <- flh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
  select(site:phe, subpop_name, phe_name, ends_with("SE")) %>%
  pivot_longer(cols = ends_with("SE"), names_to = "Type", values_to = "SE") %>%
  mutate(type_name = case_when(Type == "SE" ~ "G",
                               Type == "GxE_SE" ~ "GxE",
                               Type == "error_SE" ~ "error",
                               Type == "Env_SE" ~ "E")) %>%
  select(-Type)
fl_barplot <- flh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
  select(site:phe, subpop_name, phe_name, ends_with("Estimate")) %>%
  pivot_longer(cols = ends_with("Estimate"), names_to = "Type", values_to = "Estimate") %>%
  mutate(type_name = case_when(Type == "Estimate" ~ "G",
                               Type == "GxE_Estimate" ~ "GxE",
                               Type == "error_Estimate" ~ "error",
                               Type == "Env_Estimate" ~ "E"))
fl_barplot <- fl_barplot %>%
  left_join(fl_barplot2)
fl_barplot$type_name <- factor(fl_barplot$type_name, levels = rev(c("G", "GxE", "E", "error")))

fl_bar_gxe <- fl_barplot %>%
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = type_name)) + 
  geom_errorbar(data = filter(fl_barplot, type_name == "G"), aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  facet_grid(site ~ subpop_name)+
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") +
  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = GxE_Estimate+Estimate), linetype = 2) + xlab("") +
  ylab("Variance Explained")

save_plot("../manuscript/plots/Var_Comp_flowering_site_groups.png", plot = fl_bar_gxe, base_width = 5.5, base_height = 4)

flh2_p4 <- flh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = GxE_Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% flowering")) +
  geom_errorbar(aes(ymin = GxE_Estimate - GxE_SE, ymax = GxE_Estimate + GxE_SE), width = 0.3) +
  geom_linerange(aes(ymin = GxE_Estimate, ymax = GxE_Estimate + GxE_SE)) +
  geom_linerange(aes(ymin = GxE_Estimate - GxE_SE, ymax = GxE_Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(flh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "FL50"), aes(yintercept = GxE_Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")

save_plot("../manuscript/plots/h2_FL_site_weather_functions_site_groups_k_specific.png", plot = flh2_p2, base_height = 4, base_asp = 1.5)

library(scales)
show_col(viridis(4)) # #440154FF   #31688EFF
```

## h2 greenup weather

```{r}
unique(h2gr$site)
grh2 <- h2gr %>%
  filter(!(phe %in% c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d", "DYLN"))) %>%
  mutate(site = case_when(is.na(site) ~ "all",
                          site == "eight_sites" ~ "All",
                          site == "tx_sites" ~ "Texas",
                          site == "north_sites" ~ "North",
                          TRUE ~ site),
         subpop_name = case_when(subpop == "Gulf_and_Midwest" ~ "Gulf and Midwest",
                                 subpop == "three_subpops" ~ "All",
                                 TRUE ~ subpop),
         phe_name = case_when(phe == "GR50" ~ "50% greenup",
                              phe == "FL50" ~ "50% flowering",
                              phe == "cgdd_12c_jan2gr" ~ "GDD 01-01 to GR",
                              phe == "cgdd_8c_jan2gr" ~ "GDD 8C 01-01 to greenup",
                              phe == "tmax_18d" ~ "Max T, 18d mean",
                              phe == "tmin_18d" ~ "Min T, 18d mean",
                              phe == "tave_18d" ~ "Mean T, 18d mean",
                              phe == "tmax_10d" ~ "Max T, 10d mean",
                              phe == "tmin_10d" ~ "Min T, 10d mean",
                              phe == "tave_10d" ~ "Mean T, 10d mean",
                              phe == "tmax_5d" ~ "Max T, 5d mean",
                              phe == "tmin_5d" ~ "Min T, 5d mean",
                              phe == "tave_5d" ~ "Mean T, 5d mean",
                              phe == "cgdd_12c_10d" ~ "GDD, 10d",
                              phe == "cgdd_12c_18d" ~ "GDD, 18d",
                              phe == "cgdd_12c_5d" ~ "GDD, 5d",
                              TRUE ~ phe))
grh2$site <- factor(grh2$site, levels = rev(c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North", "nine_sites")))
grh2$phe_name <- factor(grh2$phe_name, levels = c("50% greenup", "GDD, 5d",
                                                  "Min T, 5d mean",
                                                  #"Mean T, 5d mean",
                                                  #"Max T, 5d mean",
                                                  "GDD, 10d",
                                                  "Min T, 10d mean",
                                                  #"Mean T, 10d mean",
                                                  #"Max T, 10d mean",
                                                  "GDD, 18d",
                                                  "Min T, 18d mean",
                                                  #"Mean T, 18d mean",
                                                  #"Max T, 18d mean",
                                                  #"GDD 8C 01-01 to greenup",
                                                  "GDD 01-01 to GR"))
grh2$subpop_name <- factor(grh2$subpop_name, levels = c("Gulf", "Midwest", "Gulf and Midwest", "Atlantic", "All"))

subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))

site_plot <- c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD", "All", "Texas", "North")

grh2_p1 <- grh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 1.1)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
save_plot("../manuscript/plots/h2_GR_site_weather_functions_single_sites_k_specific.png", plot = grh2_p1, base_height = 7.5, base_asp = 1.1)

site_plot <- c( "All", "Texas", "North")

grh2_p2 <- grh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.6)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1)
save_plot("../manuscript/plots/h2_GR_site_weather_functions_site_groups_k_specific.png", plot = grh2_p2, base_height = 5, base_asp = 1.3)

site_plot <- c( "All")

grh2_p3 <- grh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>% # subpop == subpop_v[[i]] & 
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = phe_name == "50% greenup")) +
  geom_errorbar(aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  geom_linerange(aes(ymin = Estimate, ymax = Estimate + SE)) +
  geom_linerange(aes(ymin = Estimate - SE, ymax = Estimate)) +
  facet_grid(site ~ subpop_name) +
  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = Estimate), linetype = 2) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "none") + ylim(c(0, 0.11)) +
  scale_fill_viridis(discrete = TRUE, end = 0.8, direction = 1) + xlab("")
save_plot("../manuscript/plots/h2_GR_site_weather_functions_All_Sites_k_specific.png", plot = grh2_p3, base_height = 3.5)

gr_fl_p1 <- plot_grid(grh2_p3, flh2_p3, nrow = 2)

save_plot("../manuscript/plots/h2_GR_and_FL_site_weather_functions_All_Sites_k_specific.png", plot = gr_fl_p1, base_height = 4.6)
```
## gxe greenup

```{r}
site_plot <- c("All", "Texas", "North")

gr_barplot2 <- grh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
  select(site:phe, subpop_name, phe_name, ends_with("SE")) %>%
  pivot_longer(cols = ends_with("SE"), names_to = "Type", values_to = "SE") %>%
  mutate(type_name = case_when(Type == "SE" ~ "G",
                               Type == "GxE_SE" ~ "GxE",
                               Type == "error_SE" ~ "error",
                               Type == "Env_SE" ~ "E")) %>%
  select(-Type)
gr_barplot <- grh2 %>%
  filter(subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & !is.na(phe_name)) %>%
  select(site:phe, subpop_name, phe_name, ends_with("Estimate")) %>%
  pivot_longer(cols = ends_with("Estimate"), names_to = "Type", values_to = "Estimate") %>%
  mutate(type_name = case_when(Type == "Estimate" ~ "G",
                               Type == "GxE_Estimate" ~ "GxE",
                               Type == "error_Estimate" ~ "error",
                               Type == "Env_Estimate" ~ "E"))
gr_barplot <- gr_barplot %>%
  left_join(gr_barplot2)
gr_barplot$type_name <- factor(gr_barplot$type_name, levels = rev(c("G", "GxE", "E", "error")))

gr_bar_gxe <- gr_barplot %>%
  ggplot(aes(x = phe_name, y = Estimate)) +
  geom_bar(stat = "identity", aes(fill = type_name)) + 
  geom_errorbar(data = filter(gr_barplot, type_name == "G"), aes(ymin = Estimate - SE, ymax = Estimate + SE), width = 0.3) +
  facet_grid(site ~ subpop_name)+
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme(panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), legend.position = "right") +
  geom_hline(data = filter(grh2, subpop_name %in% c("Gulf and Midwest", subpop_v[[1]], subpop_v[[2]]) & (site %in% site_plot) & !is.na(site) & phe == "GR50"), aes(yintercept = GxE_Estimate+Estimate), linetype = 2)

save_plot("../manuscript/plots/Var_Comp_greenup_site_groups.png", plot = gr_bar_gxe, base_height = 4, base_asp = 1.5)
```

# --------------
# SVD of K

## Plot eigenvectors to capture predominant patterns in K matrix.
```{r}
k_full <- read_rds("../data/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
k_gandm <- read_rds("../data/Kinship_van_Raden_method_363g_Gulf_and_Midwest_subset_maf0.05.rds")
k_midw <- read_rds("../data/Kinship_van_Raden_method_134g_Midwest_subset_maf0.05.rds")
k_gulf <- read_rds("../data/Kinship_van_Raden_method_229g_Gulf_subset_maf0.05.rds")
```

```{r}
b1 <- svd(k_full)
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
b1$d[2] / sum(b1$d) # 7.1
b1$d[3] / sum(b1$d) # 6.1
b1$d[4] / sum(b1$d) # 1.7
b1$d[5] / sum(b1$d) # 1.3
sum(b1$d[1:5]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
for(g in 1:5){
barplot(b1$v[,g]/k_full[which.max(abs(k_full[,g])),g])
}
order <- enframe(colnames(k_full), name = NULL, value = "PLANT_ID")
ggb1 <- tibble(`First eigenvector (PVE = 11.5%)` = b1$v[,1]/k_full[which.max(abs(k_full[,1])),1],
       `Second eigenvector (PVE = 7.1%)` = b1$v[,2]/k_full[which.max(abs(k_full[,2])),2],
       `Third eigenvector (PVE = 6.1%)` = b1$v[,3]/k_full[which.max(abs(k_full[,3])),3],
       `Fourth eigenvector (PVE = 1.7%)` = b1$v[,4]/k_full[which.max(abs(k_full[,4])),4],
       `Fifth eigenvector (PVE = 1.3%)` = b1$v[,5]/k_full[which.max(abs(k_full[,5])),5],
       `Sixth eigenvector (PVE = 0.9%)` = b1$v[,6]/k_full[which.max(abs(k_full[,6])),6],
       PLANT_ID = order$PLANT_ID) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))

ggb1 <- ggb1 %>%
  arrange(LATITUDE) %>%
  mutate(rank = row_number()) %>%
  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
                           levels = c("First eigenvector (PVE = 11.5%)", 
                                      "Second eigenvector (PVE = 7.1%)", 
                                      "Third eigenvector (PVE = 6.1%)", 
                                      "Fourth eigenvector (PVE = 1.7%)",
                                      "Fifth eigenvector (PVE = 1.3%)",
                                      "Sixth eigenvector (PVE = 0.9%)"
                                      ))
ggb1_plot <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = SUBPOP, fill = SUBPOP), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, end = 0.9, direction = -1) +
  scale_fill_viridis(discrete = TRUE, end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 607, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors.png", base_height = 5, base_asp = 2, plot = ggb1_plot)

ggb1_plot2 <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
                      direction = -1) +
  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 607, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
```

## k_gandm

```{r}
b1 <- svd(k_gandm)
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 29.2
b1$d[2] / sum(b1$d) # 18.2
b1$d[3] / sum(b1$d) # 3.7
b1$d[4] / sum(b1$d) # 1.4
b1$d[5] / sum(b1$d) # 0.008
sum(b1$d[1:4]) / sum(b1$d) # First 4 eigenvectors explain 1% or more variance, and 52.5% of total variance.

for(g in 1:5){
barplot(b1$v[,g]/k_gandm[which.max(abs(k_gandm[,g])),g])
}
order <- enframe(colnames(k_gandm), name = NULL, value = "PLANT_ID")
ggb1 <- tibble(`First eigenvector (PVE = 29.2%)` = b1$v[,1]/k_gandm[which.max(abs(k_gandm[,1])),1],
       `Second eigenvector (PVE = 18.2%)` = b1$v[,2]/k_gandm[which.max(abs(k_gandm[,2])),2],
       `Third eigenvector (PVE = 3.7%)` = b1$v[,3]/k_gandm[which.max(abs(k_gandm[,3])),3],
       `Fourth eigenvector (PVE = 1.4%)` = b1$v[,4]/k_gandm[which.max(abs(k_gandm[,4])),4],
       #`Fifth eigenvector (PVE = 0.8%)` = b1$v[,5]/k_gandm[which.max(abs(k_gandm[,5])),5],
       PLANT_ID = order$PLANT_ID) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))

ggb1 <- ggb1 %>%
  arrange(LATITUDE) %>%
  mutate(rank = row_number()) %>%
  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
                           levels = c("First eigenvector (PVE = 29.2%)", 
                                      "Second eigenvector (PVE = 18.2%)", 
                                      "Third eigenvector (PVE = 3.7%)", 
                                      "Fourth eigenvector (PVE = 1.4%)",
                                      "Fifth eigenvector (PVE = 0.8%)"
                                      ))
ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))
ggb1_plot <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = SUBPOP, fill = SUBPOP), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_manual(values = c("#F47F72", "#442C83")) +
  scale_fill_manual(values = c("#F47F72", "#442C83")) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 342, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_and_Midwest_first_four_eigenvectors.png", base_height = 5, base_asp = 2, plot = ggb1_plot)

ggb1_plot2 <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
                      direction = -1) +
  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 342, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_and_Midwest_first_four_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
```

## k_midw

```{r}
b1 <- svd(k_midw)
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 10.2
b1$d[2] / sum(b1$d) # 4.8
b1$d[3] / sum(b1$d) # 2.4
b1$d[4] / sum(b1$d) # 2.0
b1$d[5] / sum(b1$d) # 1.7
b1$d[6] / sum(b1$d) # 1.3
b1$d[7] / sum(b1$d) # 1.3
b1$d[8] / sum(b1$d) # 1.1
b1$d[9] / sum(b1$d) # 1.1
b1$d[10] / sum(b1$d) # 1.1
b1$d[11] / sum(b1$d) # 1.1
b1$d[12] / sum(b1$d) # 1.0
b1$d[13] / sum(b1$d) # 1.0
sum(b1$d[1:13]) / sum(b1$d) # First 13 eigenvectors explain 1% or more variance, and 30% of total variance. Phew! Exhausting.
```

```{r}
k_use <- k_midw
for(g in 1:5){
barplot(b1$v[,g]/k_use[which.max(abs(k_use[,g])),g])
}
order <- enframe(colnames(k_use), name = NULL, value = "PLANT_ID")
ggb1 <- tibble(`First eigenvector (PVE = 10.2%)` = b1$v[,1]/k_use[which.max(abs(k_use[,1])),1],
       `Second eigenvector (PVE = 4.8%)` = b1$v[,2]/k_use[which.max(abs(k_use[,2])),2],
       `Third eigenvector (PVE = 2.4%)` = b1$v[,3]/k_use[which.max(abs(k_use[,3])),3],
       `Fourth eigenvector (PVE = 2.0%)` = b1$v[,4]/k_use[which.max(abs(k_use[,4])),4],
       `Fifth eigenvector (PVE = 1.7%)` = b1$v[,5]/k_use[which.max(abs(k_use[,5])),5],
       `Sixth eigenvector (PVE = 1.3%)` = b1$v[,6]/k_use[which.max(abs(k_use[,6])),6],
       PLANT_ID = order$PLANT_ID) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))

ggb1 <- ggb1 %>%
  arrange(LATITUDE) %>%
  mutate(rank = row_number()) %>%
  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
                           levels = c("First eigenvector (PVE = 10.2%)", 
                                      "Second eigenvector (PVE = 4.8%)", 
                                      "Third eigenvector (PVE = 2.4%)", 
                                      "Fourth eigenvector (PVE = 2.0%)",
                                      "Fifth eigenvector (PVE = 1.7%)",
                                      "Sixth eigenvector (PVE = 1.3%)"
                                      ))
ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))

ggb1_plot2 <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
                      direction = -1) +
  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 133, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Midwest_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
```
## k_gulf

```{r}
b1 <- svd(k_gulf)
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 22.5
b1$d[2] / sum(b1$d) # 9.4
b1$d[3] / sum(b1$d) # 3.1
b1$d[4] / sum(b1$d) # 1.8
b1$d[5] / sum(b1$d) # 1.6
b1$d[6] / sum(b1$d) # 1.3
b1$d[7] / sum(b1$d) # 1.1
b1$d[8] / sum(b1$d) # 1.0
sum(b1$d[1:8]) / sum(b1$d) # First 8 eigenvectors explain 1% or more variance, and 41.8% of total variance. 
```

```{r}
k_use <- k_gulf
for(g in 1:8){
barplot(b1$v[,g]/k_use[which.max(abs(k_use[,g])),g])
}
order <- enframe(colnames(k_use), name = NULL, value = "PLANT_ID")
ggb1 <- tibble(`First eigenvector (PVE = 22.5%)` = b1$v[,1]/k_use[which.max(abs(k_use[,1])),1],
       `Second eigenvector (PVE = 9.4%)` = b1$v[,2]/k_use[which.max(abs(k_use[,2])),2],
       `Third eigenvector (PVE = 3.1%)` = b1$v[,3]/k_use[which.max(abs(k_use[,3])),3],
       `Fourth eigenvector (PVE = 1.8%)` = b1$v[,4]/k_use[which.max(abs(k_use[,4])),4],
       `Fifth eigenvector (PVE = 1.6%)` = b1$v[,5]/k_use[which.max(abs(k_use[,5])),5],
       `Sixth eigenvector (PVE = 1.3%)` = b1$v[,6]/k_use[which.max(abs(k_use[,6])),6],
       PLANT_ID = order$PLANT_ID) %>%
  left_join(metadata, by = "PLANT_ID") %>%
  dplyr::select(PLANT_ID, SUBPOP, ECOTYPE_NNET, LATITUDE, ends_with("%)"))

ggb1 <- ggb1 %>%
  arrange(LATITUDE) %>%
  mutate(rank = row_number()) %>%
  pivot_longer(cols = ends_with("%)"), names_to = "Eigenvector", values_to = "value")
ggb1$Eigenvector <- factor(ggb1$Eigenvector, 
                           levels = c("First eigenvector (PVE = 22.5%)", 
                                      "Second eigenvector (PVE = 9.4%)", 
                                      "Third eigenvector (PVE = 3.1%)", 
                                      "Fourth eigenvector (PVE = 1.8%)",
                                      "Fifth eigenvector (PVE = 1.6%)",
                                      "Sixth eigenvector (PVE = 1.3%)"
                                      ))
ggb1$ECOTYPE_NNET <- factor(ggb1$ECOTYPE_NNET, levels = c("Coastal", "Lowland", "Upland", "Unknown"))

ggb1_plot2 <- ggb1 %>%
  ggplot(aes(x = rank, y = value)) +
  geom_bar(aes(color = ECOTYPE_NNET, fill = ECOTYPE_NNET), stat = "identity") +
  facet_wrap(~ Eigenvector) +
  theme(legend.position = "right", axis.text.x = element_blank(), panel.spacing.x = unit(0.1, 'cm')) +
  scale_color_viridis(discrete = TRUE, option = "B", end = 0.9, 
                      direction = -1) +
  scale_fill_viridis(discrete = TRUE, option = "B", end = 0.9, direction = -1) +
  xlab("Latitude of Origin (Rank Order)") + ylab("") +
  geom_vline(xintercept = 210, linetype = 2)
save_plot("../manuscript/plots/SVD_of_Kinship_matrix_Gulf_first_six_eigenvectors_ecotype.png", base_height = 5, base_asp = 2, plot = ggb1_plot2)
```

The eigenvectors capture the predominant patterns in the Uk3 covariance matrix.
```
v <- m$fitted_g$Ulist$ED_PCA_1
b1=svd(v)
for(g in 1){
barplot(b1$v[,g]/v[which.max(abs(v[,g])),g])
}

b1$d # the top eigenvector captures the predominant pattern in the ED_PCA_1 covariance matrix. 
```

WAIT. I could also plot this for the svd from the whole SNP set. That might make more sense than doing a svd on the kinship matrix...

```{r}
snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/kinship/Pvirgatum_V5_GWAS_134g_Midwest_K_subset_maf0.05.rds")

```

# -------------
# ASReml models for eight sites including GxE
```{r}

```

## ASReml h2 greenup analysis
```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)

source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

```

### loop to make heritability table using ASReml mixed models and vpredict.

## Include just texas sites for greenup and just northern four sites.
e.g.:
for(p in 1){#seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in 12:13){#seq_along(site_v)){ # may need to adjust below for h2_table
    for(j in 1:3){#seq_along(subpop_v)){
      for(k in seq_along(phe_v)){

## Model eight, texas, and north sites with a kinship : site interaction.
Then, find GxE and G proportions of variance.
```{r}
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", IL = "IL", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), nine_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "IL", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"), Atlantic = "Atlantic", three_subpops = c("Atlantic", "Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
           "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d",
           "crain_7d")
phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "cgdd_8c_jan2gr", "tmax_18d", 
            "tmin_18d", "tave_18d", "cgdd_12c_18d", "tmax_10d", "tmin_10d",
            "tave_10d", "cgdd_12c_10d", "tmax_5d", "tmin_5d", "tave_5d",
            "cgdd_12c_5d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)
h2_table <- tibble()

for(p in seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in c(10,12:13)){#seq_along(site_v)){ # may need to adjust below for h2_table
    for(j in 1:3){#seq_along(subpop_v)){
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
    left_join(sites, by = "SITE") %>%
    filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
    rename(phenotype = eval(phe_v[k])) %>%
    mutate(PLANT_ID = as_factor(PLANT_ID),
           SITE = as_factor(SITE),
           PLOT_GL = as_factor(PLOT_GL)) %>%
    as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full) +
                         ~idv(SITE):vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V2/(V1+V2+V3)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  } else {
    tryCatch({
    asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_full),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }

  if(i == 12 & j == 1 & k == 1){
    h2_table <- h2_est
  } else {
    h2_table <- add_row(h2_table, h2_est)
  }
  tryCatch({
      save_plot(filename=file.path(outputdir, paste0("ASReml_performance_",
                                                 names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".png")), 
            plot = plot(asr_out))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_functions_", 
                                                  "of_weather_v4.csv")))
}
```

# ASReml 07-29

## Model eight, texas, and north sites with a kinship : site interaction.
Then, find GxE and G proportions of variance.
## Use Midwest-specific kinship matrix for Midwest-only, and Gulf-specific kinship matrix for Gulf-only comparisons. And use (and visualize eigenvectors of? Midwest and Gulf kinship matrix for the comparisons with both.)
```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)

source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability", "kinship_subpop")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))

kinshipdir <- file.path(workingdir, "data")
k_gandm <- read_rds(file.path(kinshipdir, "Kinship_van_Raden_method_363g_Gulf_and_Midwest_subset_maf0.05.rds"))
k_midw <- read_rds(file.path(kinshipdir,
                "Kinship_van_Raden_method_134g_Midwest_subset_maf0.05.rds"))
k_gulf <- read_rds(file.path(kinshipdir, "Kinship_van_Raden_method_229g_Gulf_subset_maf0.05.rds"))

```

```{r}
site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", MI = "MI", SD = "SD", eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", Gulf_and_Midwest = c("Gulf", "Midwest"))
phe_fl <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl",
           "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")
phe_gr <- c("GR50", "DYLN", "cgdd_12c_jan2gr", "tmin_18d", "cgdd_12c_18d",
            "tmin_10d", "cgdd_12c_10d", "tmin_5d", "cgdd_12c_5d")
phe_dfs <- list(phe_gr_df, phe_fl_df)
phe_lists <- list(phe_gr, phe_fl)
k_list <- list(Gulf = k_gulf, Midwest = k_midw, Gulf_and_Midwest = k_gandm)
h2_table <- tibble()
gxe_table <- tibble()

for(p in seq_along(phe_dfs)){
  phe_v <- phe_lists[[p]]
  phe <- phe_dfs[[p]]

  for(i in seq_along(site_v)){ # may need to adjust below for h2_table
    dir.create(file.path(outputdir, names(site_v)[i]), showWarnings = FALSE)
    for(j in seq_along(subpop_v)){
      k_use <- k_list[[j]]
      for(k in seq_along(phe_v)){
        phe_single <- phe %>%
          left_join(sites, by = "SITE") %>%
          filter(manu_site %in% site_v[[i]] & SUBPOP %in% subpop_v[[j]]) %>%
          rename(phenotype = eval(phe_v[k])) %>%
          mutate(PLANT_ID = as_factor(PLANT_ID),
                 SITE = as_factor(SITE),
                 PLOT_GL = as_factor(PLOT_GL)) %>%
          as.data.frame()

  if(length(site_v[[i]]) > 1){  # need to add site as a random factor if 
    # there is more than one site included in the dataset.
    tryCatch({
     asr_out <- asreml(phenotype ~ 1,
                       random = ~idv(SITE)*vm(PLANT_ID, k_use),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    Env_est <- vpredict(asr_out, E ~ V1/(V1+V2+V3+V4))
    GxE_est <- vpredict(asr_out, GxE_h2 ~ V3/(V1+V2+V3+V4))
    err_est <- vpredict(asr_out, err ~ V4/(V1+V2+V3+V4))
    h2_est <- vpredict(asr_out, G_h2 ~ V2/(V1+V2+V3+V4)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik,
             Env_Estimate = Env_est$Estimate[1],
             Env_SE = Env_est$SE[1],
             GxE_Estimate = GxE_est$Estimate[1],
             GxE_SE = GxE_est$SE[1],
             error_Estimate = err_est$Estimate[1],
             error_SE = err_est$SE[1])

    blup <- summary(asr_out, coef=TRUE)$coef.random %>% 
    as_tibble(rownames = "Effect")
    write_csv(blup, file.path(outputdir, names(site_v)[i], 
                              paste0("blups_", names(site_v)[i], "_",
                                     names(subpop_v)[j], "_", phe_v[k],
                                     ".csv")))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  } else {
    tryCatch({
    asr_out <- asreml(phenotype ~ 1,
                       random = ~vm(PLANT_ID, k_use),
                       residual = ~idv(units),
                       data = phe_single, 
                       workspace = "3gb")

    h2_est <- vpredict(asr_out, h2 ~ V1/(V1+V2)) %>% as_tibble() %>%
      mutate(site = names(site_v)[i],
             subpop = names(subpop_v)[j],
             phe = phe_v[k],
             bic = summary(asr_out)$bic[1],
             loglik = summary(asr_out)$loglik)
    blup <- summary(asr_out, coef=TRUE)$coef.random %>% 
    as_tibble(rownames = "Effect")
    write_csv(blup, file.path(outputdir, names(site_v)[i],
                              paste0("blups_", names(site_v)[i], "_",
                                     names(subpop_v)[j], "_", phe_v[k],
                                     ".csv")))
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }

  if(i == 1 & j == 1 & k == 1){
    h2_table <- h2_est
  } else if(i == 9 & j == 1 & k == 1){
    gxe_table <- h2_est
  } else if(length(site_v[[i]]) > 1){
    gxe_table <- add_row(gxe_table, h2_est)
  } else {
    h2_table <- add_row(h2_table, h2_est)
  }
  tryCatch({
      save_plot(filename=file.path(outputdir, names(site_v)[i],
                                   paste0("ASReml_performance_",
                                                 names(site_v)[i], "_",
                                                 names(subpop_v)[j],
                                                 "_", phe_v[k], ".png")), 
            plot = plot(asr_out))
    }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

      }
    }
  }
  write_csv(h2_table, file.path(outputdir, paste0("h2_df_", phe_v[1],
                                                  "_related_functions_", 
                                                  "of_weather_v1.csv")))
  write_csv(gxe_table, file.path(outputdir, paste0("gxe_and_h2_df_", phe_v[1],
                                                  "_related_functions_", 
                                                  "of_weather_v1.csv")))
}
```

# D2F

If flowering date frequently varies as a function of GDD in switchgrass, this explains observations that moving southern populations northwards delays flowering, and moving northern populations south hastens flowering (Sanderson et al 1996). < Indeed, we observed D2F for Midwest plants was shorter at TX than BRKG, and vice versa?>

```{r}
d2f_lat <- phenotypes %>%
  mutate(D2F = FL50 - GR50) %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  filter(!(SITE %in% c("OVTN", "FRMI")) & WHERE == "GWAS" & SUBPOP %in% c("Gulf", "Midwest")) %>%
  group_by(SUBPOP, SITE) %>%
  summarise(mean_D2F = mean(D2F, na.rm = TRUE)) %>%
  left_join(select(sites, SITE, Latitude, manu_site)) %>%
  arrange(SUBPOP, Latitude) 
d2f_lat %>%
  ggplot(aes(x = Latitude, y = mean_D2F)) +
  geom_point(aes(color = SUBPOP))

summary(lm(mean_D2F ~ Latitude*SUBPOP, data = d2f_lat))
82-55
```

## G_BLUP GWAS

```{r}
gwasdir <- file.path(workingdir, "analysis", "gwas", "G_BLUP")
log10p_gulf <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Gulf.rds"))
log10p_midw <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Midwest.rds"))
log10p_gandm <- readRDS(file.path(gwasdir,                                 "log10p_strong_df_1000topSNPs_GR_FL_Gulf_and_Midwest.rds"))
```

```{r}
log10p_gulf %>%
  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
  ungroup() %>%
  filter(log10p > -log10(0.05/10300000)) %>%
  group_by(PHE) %>%
  tally() %>%
  write_csv(., "../manuscript/tables/Gulf_SNPs_above_Bonferroni_by_trait.csv")
```

```{r}
log10p_midw %>%
  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
  ungroup() %>%
  filter(log10p > -log10(0.05/8800000)) %>%
  group_by(PHE) %>%
  tally() %>%
  write_csv(., "../manuscript/tables/Midwest_SNPs_above_Bonferroni_by_trait.csv")

log10p_gandm %>%
  pivot_longer(cols = -(SNP:POS), names_to = "PHE", values_to = "log10p") %>%
  ungroup() %>%
  filter(log10p > -log10(0.05/12100000)) %>%
  group_by(PHE) %>%
  tally() %>%
  write_csv(., "../manuscript/tables/Gulf_and_Midwest_diversity_panel_SNPs_above_Bonferroni_by_trait.csv")
```

```{r}
gulf_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Gulf.rds"))

gulf_top %>%
  group_by(n_cond_in_top_set) %>%
  tally()

gulf_top_manh2 <- gulf_top %>%
  filter(max_score_log10p > 5.897) %>%
  ggplot(aes(x = POS, y = max_score_log10p)) + 
  geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
  #geom_line(aes(group = Condition)) +
  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
  theme(legend.position = "top",
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
        labs(x = "Chromosome", y = "-log10(p value)") +
        scale_x_continuous(expand = c(0.18, 0.18)) +
  scale_color_viridis(discrete = TRUE, end = 0.97) +
  scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
save_plot("../manuscript/plots/Gulf_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = gulf_top_manh2)

gulf_top_manh <- gulf_top %>%
  filter(max_score_log10p > 5.897) %>%
  ggplot(aes(x = POS, y = max_score_log10p)) + 
  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
  #geom_line(aes(group = Condition)) +
  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
  theme(legend.position = "top",
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
        labs(x = "Chromosome", y = "-log10(p value)") +
        scale_x_continuous(expand = c(0.18, 0.18)) +
  scale_color_viridis(option = "B", end = 0.9) +
  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))

save_plot("../manuscript/plots/Gulf_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = gulf_top_manh)
```

```{r}
midw_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Midwest.rds"))

midw_top %>%
  group_by(n_cond_in_top_set) %>%
  tally()

midw_top_manh2 <- midw_top %>%
  filter(max_score_log10p > 6) %>%
  ggplot(aes(x = POS, y = max_score_log10p)) + 
  geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
  #geom_line(aes(group = Condition)) +
  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
  theme(legend.position = "top",
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
        labs(x = "Chromosome", y = "-log10(p value)") +
        scale_x_continuous(expand = c(0.18, 0.18)) +
  scale_color_viridis(discrete = TRUE, end = 0.97) +
  scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
save_plot("../manuscript/plots/Midwest_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = midw_top_manh2)

midw_top_manh <- midw_top %>%
  filter(max_score_log10p > 5.897) %>%
  ggplot(aes(x = POS, y = max_score_log10p)) + 
  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
  #geom_line(aes(group = Condition)) +
  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
  theme(legend.position = "top",
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
        labs(x = "Chromosome", y = "-log10(p value)") +
        scale_x_continuous(expand = c(0.18, 0.18)) +
  scale_color_viridis(option = "B", end = 0.9) +
  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))

save_plot("../manuscript/plots/Midwest_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = midw_top_manh)
```

```{r}
gandm_top <- readRDS(file.path(gwasdir, "Part-One-Output_Top-Effects-1000-SNPs_GR_FL_Gulf_and_Midwest.rds"))

gandm_top %>%
  group_by(n_cond_in_top_set) %>%
  tally()

gandm_top_manh2 <- gandm_top %>%
  filter(max_score_log10p > 6) %>%
  ggplot(aes(x = POS, y = max_score_log10p)) + 
  geom_point(aes(color = Condition, shape = as_factor(n_cond_in_top_set))) + 
  #geom_line(aes(group = Condition)) +
  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
  theme(legend.position = "top",
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
        labs(x = "Chromosome", y = "-log10(p value)") +
        scale_x_continuous(expand = c(0.18, 0.18)) +
  scale_color_viridis(discrete = TRUE, end = 0.97) +
  scale_shape_manual(values = c(4, 16, 17, 18, 14, 3, 4))
save_plot("../manuscript/plots/Gulf_and_Midwest_G_BLUP_top_500_Manhattan_2.png", base_height = 5, base_asp = 2.5, plot = gandm_top_manh2)

gandm_top_manh <- gandm_top %>%
  filter(max_score_log10p > 5.897) %>%
  ggplot(aes(x = POS, y = max_score_log10p)) + 
  geom_point(aes(color = n_cond_in_top_set, shape = Condition)) + 
  #geom_line(aes(group = Condition)) +
  facet_wrap( ~ CHR, nrow = 1, scales = "free_x", strip.position = "bottom") +
  theme(legend.position = "top",
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        panel.background = element_rect(fill=NA)) +
        labs(x = "Chromosome", y = "-log10(p value)") +
        scale_x_continuous(expand = c(0.18, 0.18)) +
  scale_color_viridis(option = "B", end = 0.9) +
  scale_shape_manual(values = c(15, 16, 17, 18, 14, 3, 4))

save_plot("../manuscript/plots/Gulf_and_Midwest_G_BLUP_top_500_Manhattan.png", base_height = 5, base_asp = 2.5, plot = gandm_top_manh)
```

## Overlap with QTL and between subpops
```{r}

```