### 2020-10-29 R/Manuscript_v0.5_genetic_covariances.Rmd

---
title: "Manuscript_v0.5_genetic_covariances"
author: "Alice MacQueen"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(asreml)
library(cowplot)
library(switchgrassGWAS)
```

# Load data & prep multivariate model data frames.
Add any data that needs to be loaded for these analyses here.
```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

fl50_cov_df <- phe_fl_df %>%
  ungroup() %>%
  left_join(select(sites, SITE, manu_site)) %>%
  select(PLANT_ID, SUBPOP, manu_site, FL50) %>%
  group_by(PLANT_ID, SUBPOP, manu_site) %>%
  mutate(IND = row_number()) %>%
  pivot_wider(names_from = manu_site, values_from = "FL50") %>%
  mutate(PLANT_ID = as.factor(PLANT_ID),
         IND = as.factor(IND)) %>%
  select(-IL) %>%
  select(PLANT_ID:TX2, TX3, OK, MO, NE, MI, SD)
gr50_cov_df <- phe_fl_df %>%
  ungroup() %>%
  left_join(select(sites, SITE, manu_site)) %>%
  select(PLANT_ID, SUBPOP, manu_site, GR50) %>%
  group_by(PLANT_ID, SUBPOP, manu_site) %>%
  mutate(IND = row_number()) %>%
  pivot_wider(names_from = manu_site, values_from = "GR50") %>%
  mutate(PLANT_ID = as.factor(PLANT_ID),
         IND = as.factor(IND)) %>%
  select(-IL) %>%
  select(PLANT_ID:TX2, TX3, OK, MO, NE, MI, SD)
```

# In this document:
For both GR50 and FL50, in Gulf, Midwest, and G&M subpops:
fit 7 (or more) bivariate models; compare h^2 estimates between them for A-H.
Fit trivariate models?
Try to fit 8-way model? Run these models on the server so my laptop isn't overloaded.

## bivariate analyses
```{r}
asr_1 <- asreml(cbind(PKLE, KBSM) ~ trait,
                random = ~us(trait, init = c(1,0.1,1)):vm(PLANT_ID, k_full), #+
                  #us(trait, init = c(1, 0.1, 1)):SUBPOP,
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1)), 
                #na.method(y = "omit", x = "omit"),
                data = fl50_cor_df, 
                workspace = "3gb")
asr_1 <- update.asreml(asr_1)
summary(asr_1)$varcomp

vpredict(asr_1, h2_PKLE ~ V1/(V1+V5))
vpredict(asr_1, h2_KBSM ~ V3/(V3+V7))
vpredict(asr_1, r_G ~ V2/((V1*V3)^0.5))
vpredict(asr_1, r_R ~ V6/((V5*V7)^0.5))

# now try adding a factor to split trait by subpopulation. 
# Nope. I don't think this is a good idea with the kinship matrix there too.
# Maybe try this later if I ever try and partition variance by individuals. Could do for subpops and for individuals/genotypes, also.
asr_2 <- asreml(cbind(PKLE, KBSM) ~ trait,
                random = ~us(trait, init = c(1,0.1,1)):vm(PLANT_ID, k_full) +
                  us(trait, init = c(1, 0.1, 1)):SUBPOP,
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1)), 
                #na.method(y = "omit", x = "omit"),
                data = fl50_cor_df, 
                workspace = "3gb")
asr_2 <- update.asreml(asr_2)
summary(asr_2)$varcomp
vpredict(asr_2, h2_PKLE ~ V4/(V1+V4+V8))
vpredict(asr_2, h2_KBSM ~ V3/(V3+V6+V10))
vpredict(asr_2, r_S ~ V2/((V1*V3)^0.5))
vpredict(asr_2, r_G ~ V5/((V4*V6)^0.5))

# below works way better (with fewer warnings) when initial values are given. Order for init argument is specified below this code box.
asr_3 <- asreml(cbind(PKLE, KBSM, CLMB) ~ trait,
                random = ~us(trait, init = c(1,0.1, 1, 0.1, 0.1, 1)):vm(PLANT_ID, k_full), 
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1)), 
                na.action = na.method(x = "omit"),
                data = fl50_cor_df, 
                workspace = "3gb")
asr_3 <- update.asreml(asr_3)
summary(asr_3)$varcomp

vpredict(asr_3, h2_PKLE ~ V1/(V1+V8)) # 93.80
vpredict(asr_3, h2_KBSM ~ V3/(V3+V10)) # 90.19
vpredict(asr_3, h2_CLMB ~ V6/(V6+V13)) # 70.58
vpredict(asr_3, r_PK ~ V2/((V1*V3)^0.5)) # 41.23
vpredict(asr_3, r_KC ~ V5/((V3*V6)^0.5)) # 76.57
vpredict(asr_3, r_CP ~ V4/((V1*V6)^0.5)) # 52.84
?na.method()
```
order for init: AA, BA, BB, CA, CB, CC
AA BA BB CA CB CC DA DB DC DD EA EB EC ED EE FA FB FC FD FE FF (6int) GG (7int) HH

2 garden model: h2_PKLE	0.9328744	0.009197664	
h2_KBSM	0.9151788	0.01180919	
r_G	0.4239493	0.04072152	
r_R	0.9709351	0.006573728