### 2020-11-04 R/Manuscript_v0.5_genetic_covariances.Rmd

---
title: "Manuscript_v0.5_genetic_covariances"
author: "Alice MacQueen"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(asreml)
library(cowplot)
library(switchgrassGWAS)
```

# Load data & prep multivariate model data frames.
Add any data that needs to be loaded for these analyses here.
```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "heritability")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

fl50_cov_df <- phe_fl_df %>%
  ungroup() %>%
  left_join(select(sites, SITE, manu_site)) %>%
  select(PLANT_ID, SUBPOP, manu_site, FL50) %>%
  group_by(PLANT_ID, SUBPOP, manu_site) %>%
  mutate(IND = row_number()) %>%
  pivot_wider(names_from = manu_site, values_from = "FL50") %>%
  mutate(PLANT_ID = as.factor(PLANT_ID),
         IND = as.factor(IND)) %>%
  select(-IL) %>%
  select(PLANT_ID:TX2, TX3, OK, MO, NE, MI, SD)
gr50_cov_df <- phe_fl_df %>%
  ungroup() %>%
  left_join(select(sites, SITE, manu_site)) %>%
  select(PLANT_ID, SUBPOP, manu_site, GR50) %>%
  group_by(PLANT_ID, SUBPOP, manu_site) %>%
  mutate(IND = row_number()) %>%
  pivot_wider(names_from = manu_site, values_from = "GR50") %>%
  mutate(PLANT_ID = as.factor(PLANT_ID),
         IND = as.factor(IND)) %>%
  select(-IL) %>%
  select(PLANT_ID:TX2, TX3, OK, MO, NE, MI, SD)
```

# In this document:
For both GR50 and FL50, in Gulf, Midwest, and G&M subpops:
fit 7 (or more) bivariate models; compare h^2 estimates between them for A-H.
Fit trivariate models?
Try to fit 8-way model? Run these models on the server so my laptop isn't overloaded.

## bivariate analyses

asr_1: test bivariate analysis.
```{r}
asr_1 <- asreml(cbind(TX2, MI) ~ trait,
                random = ~us(trait, init = c(1,0.1,1)):vm(PLANT_ID, k_full), #+
                  #us(trait, init = c(1, 0.1, 1)):SUBPOP,
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1)), 
                #na.method(y = "omit", x = "omit"),
                data = fl50_cov_df, 
                workspace = "3gb")
asr_1 <- update.asreml(asr_1)
summary(asr_1)$varcomp

vpredict(asr_1, h2_TX2 ~ V1/(V1+V5))
vpredict(asr_1, h2_MO ~ V3/(V3+V7))
vpredict(asr_1, r_G ~ V2/((V1*V3)^0.5))
vpredict(asr_1, r_R ~ V6/((V5*V7)^0.5))

# now try adding a factor to split trait by subpopulation. 
# Nope. I don't think this is a good idea with the kinship matrix there too.
# Maybe try this later if I ever try and partition variance by individuals. Could do for subpops and for individuals/genotypes, also.
asr_2 <- asreml(cbind(PKLE, KBSM) ~ trait,
                random = ~us(trait, init = c(1,0.1,1)):vm(PLANT_ID, k_full) +
                  us(trait, init = c(1, 0.1, 1)):SUBPOP,
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1)), 
                #na.method(y = "omit", x = "omit"),
                data = fl50_cov_df, 
                workspace = "3gb")
asr_2 <- update.asreml(asr_2)
summary(asr_2)$varcomp
vpredict(asr_2, h2_TX2 ~ V4/(V1+V4+V8))
vpredict(asr_2, h2_MI ~ V3/(V3+V6+V10))
vpredict(asr_2, r_S ~ V2/((V1*V3)^0.5))
vpredict(asr_2, r_G ~ V5/((V4*V6)^0.5))
```

asr_3: trivariate analysis.  specifying initial values for variances and covariances.
```{r}
# below works way better (with fewer warnings) when initial values are given. Order for init argument is specified below this code box.
asr_3 <- asreml(cbind(TX2, MI, MO) ~ trait,
                random = ~us(trait, init = c(1,0.1, 1, 0.1, 0.1, 1)):vm(PLANT_ID, k_full), 
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1)), 
                na.action = na.method(x = "omit"),
                data = fl50_cov_df, 
                workspace = "3gb", maxiter = 30)
asr_3 <- update.asreml(asr_3)
summary(asr_3)$varcomp

vpredict(asr_3, h2_TX2 ~ V1/(V1+V8)) # 93.80
vpredict(asr_3, h2_MI ~ V3/(V3+V10)) # 90.19
vpredict(asr_3, h2_MO ~ V6/(V6+V13)) # 70.58
vpredict(asr_3, r_TX2MI ~ V2/((V1*V3)^0.5)) # 41.23
vpredict(asr_3, r_MIMO ~ V5/((V3*V6)^0.5)) # 76.57
vpredict(asr_3, r_TX2MO ~ V4/((V1*V6)^0.5)) # 52.84
?na.method()

asr_4 <- asreml(cbind(TX1, TX2, TX3, OK) ~ trait,
                random = ~us(trait, init = c(1, 0.1, 1, 0.1, 0.1, 1, 0.1, 0.1, 
                                             0.1, 1)):vm(PLANT_ID, k_full), 
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 0.1, 0.1, 
                                                     0.1, 0.1, 0.1, 0.1, 0.1)), 
                na.action = na.method(x = "omit"),
                data = fl50_cov_df, 
                workspace = "3gb", maxiter = 20)
asr_4 <- update.asreml(asr_4)
summary(asr_4)$varcomp

vpredict(asr_4, h2_TX1 ~ V1/(V1+V12)) # 93.80
vpredict(asr_4, h2_TX2 ~ V3/(V3+V14)) # 90.19
vpredict(asr_4, h2_TX3 ~ V6/(V6+V17)) # 70.58
vpredict(asr_4, h2_OK ~ V10/(V10+V21))
vpredict(asr_4, r_TX2MI ~ V2/((V1*V3)^0.5)) # 41.23
vpredict(asr_4, r_MIMO ~ V5/((V3*V6)^0.5)) # 76.57
vpredict(asr_4, r_TX2MO ~ V4/((V1*V6)^0.5)) # 52.84
```
order for init: AA, BA, BB, CA, CB, CC
AA BA BB CA CB CC DA DB DC DD EA EB EC ED EE FA FB FC FD FE FF (6int) GG (7int) HH

2 garden model: h2_PKLE	0.9328744	0.009197664	
h2_KBSM	0.9151788	0.01180919	
r_G	0.4239493	0.04072152	
r_R	0.9709351	0.006573728

# On server:
For both GR50 and FL50, in Gulf, Midwest, and G&M subpops.
-> subset fl50_cov_df and gr50_cov_df to these three individual subsets. Six subsets.
```{r}
subpop_v <- list(Gulf = "Gulf", Midwest = "Midwest", 
                 Gulf_and_Midwest = c("Gulf", "Midwest"))
i=3
phe_df <- list(FL50 = fl50_cov_df, GR50 = gr50_cov_df)
j=2

cov_df_use <- phe_df[[j]] %>%
  filter(SUBPOP %in% subpop_v[[i]])
```

for Midwest, try just gardens with n>100 first.
```{r}
i=2
j=2
asr_3 <- asreml(cbind(TX2, MI, MO) ~ trait,
                random = ~us(trait, init = c(1,0.1, 1, 0.1, 0.1, 1)):vm(PLANT_ID, k_full), 
                residual = ~units:us(trait, init = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1)), 
                na.action = na.method(x = "omit"),
                data = cov_df_use, 
                workspace = "6gb", maxiter = 50)
asr_3 <- update.asreml(asr_3, maxiter = 100)
summary(asr_3)$varcomp

vpredict(asr_3, h2_TX2 ~ V1/(V1+V8)) # 93.80
vpredict(asr_3, h2_MI ~ V3/(V3+V10)) # 90.19
vpredict(asr_3, h2_MO ~ V6/(V6+V13)) # 70.58
vpredict(asr_3, r_TX2MI ~ V2/((V1*V3)^0.5)) # 41.23
vpredict(asr_3, r_MIMO ~ V5/((V3*V6)^0.5)) # 76.57
vpredict(asr_3, r_TX2MO ~ V4/((V1*V6)^0.5)) # 52.84
saveRDS(asr_3, "analysis/genetic-covariances/GR50_asreml_cov_model_3_gardens_Midwest_subpop.rds") 
# for the model running 2020-11-02 that will probably have a loglik around -778, which was run with i=2 and j=2.
```

```{r}

# try all eight gardens first.
asr_8 <- asreml(cbind(TX1, TX2, TX3, OK, MO, NE, MI, SD) ~ trait,
                random = ~us(trait, init = c(1, 0.1, 1, 0.1, 0.1, 1, 
                                             0.1, 0.1, 0.1, 1,
                                             0.1, 0.1, 0.1, 0.1, 1,
                                             0.1, 0.1, 0.1, 0.1, 0.1, 1,
                                             0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 1,
                                             0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 1)):vm(PLANT_ID, k_full), 
                residual = ~units:us(trait, init = rep(0.1, 36)), 
                na.action = na.method(x = "omit"),
                data = cov_df_use, 
                workspace = "6gb", maxiter = 240)
saveRDS(asr_8, "analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_and_Midwest_subpop.rds") 
summary(asr_8)$varcomp
asr_8 <- update.asreml(asr_8)

summary(asr_8)$varcomp

saveRDS(asr_8, "analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_and_Midwest_subpop.rds")

# for the model running 2020-11-03 that will probably have a loglik around ____, which was run with i=3 and j=2.

#saveRDS(asr_8, "analysis/genetic-covariances/FL50_asreml_cov_model_8_gardens_Gulf_subpop.rds") 
# for the model running 2020-11-02 that will probably have a loglik around -3978, which was run with i=1 and j=1.

8+7+6+5+4+3+2+1 # 36 var and covar in eight garden model.
# programmatically, each pair of two gardens, k and k+1. (special case of 8 & 1 also).
```
check model ran and saved ok on the server
```{r}
asr_8 <- readRDS("../analysis/genetic-covariances/GR50_asreml_cov_model_8_gardens_Gulf_subpop.rds")
summary(asr_8)$varcomp
vpredict(asr_8, h2_tx1 ~ V1/(V1+V38))
vpredict(asr_8, h2_tx2 ~ V3/(V3+V40))
vpredict(asr_8, h2_tx3 ~ V6/(V6+V43))
vpredict(asr_8, h2_ok ~ V10/(V10+V47))  # all sig > 0 minus 2SE, ok is the closest to zero... only 3.8%
vpredict(asr_8, h2_mo ~ V15/(V15+V52))
vpredict(asr_8, h2_ne ~ V21/(V21+V58))
vpredict(asr_8, h2_mi ~ V28/(V28+V65))
vpredict(asr_8, h2_sd ~ V36/(V36+V73))

vpredict(asr_8, cov_tx2tx1 ~ V2/((V1*V3)^0.5))  # sig pos

vpredict(asr_8, cov_tx3tx1 ~ V4/((V1*V6)^0.5))   # sig pos
vpredict(asr_8, cov_tx3tx2 ~ V5/((V6*V3)^0.5))   # sig pos

vpredict(asr_8, cov_oktx1 ~ V7/((V1*V10)^0.5)) 
vpredict(asr_8, cov_oktx2 ~ V8/((V3*V10)^0.5)) 
vpredict(asr_8, cov_oktx3 ~ V9/((V6*V10)^0.5))

vpredict(asr_8, cov_motx1 ~ V11/((V1*V15)^0.5)) 
vpredict(asr_8, cov_motx2 ~ V12/((V3*V15)^0.5))  # ** sig neg even plus 2SE
vpredict(asr_8, cov_motx3 ~ V13/((V6*V15)^0.5)) 
vpredict(asr_8, cov_mook ~ V14/((V10*V15)^0.5))

vpredict(asr_8, cov_netx1 ~ V16/((V1*V21)^0.5)) 
vpredict(asr_8, cov_netx2 ~ V17/((V3*V21)^0.5)) 
vpredict(asr_8, cov_netx3 ~ V18/((V6*V21)^0.5)) 
vpredict(asr_8, cov_neok ~ V19/((V10*V21)^0.5)) 
vpredict(asr_8, cov_nemo ~ V20/((V15*V21)^0.5))   # sig pos

vpredict(asr_8, cov_mitx1 ~ V22/((V1*V28)^0.5)) 
vpredict(asr_8, cov_mitx2 ~ V23/((V3*V28)^0.5)) # ** sig neg even plus 2SE
vpredict(asr_8, cov_mitx3 ~ V24/((V6*V28)^0.5)) 
vpredict(asr_8, cov_miok ~ V25/((V10*V28)^0.5)) 
vpredict(asr_8, cov_mimo ~ V26/((V15*V28)^0.5)) # sig pos
vpredict(asr_8, cov_mine ~ V27/((V21*V28)^0.5))

vpredict(asr_8, cov_sdtx1 ~ V29/((V1*V36)^0.5)) # ** sig neg even plus 2SE
vpredict(asr_8, cov_sdtx2 ~ V30/((V3*V36)^0.5)) # ** sig neg even plus 2SE
vpredict(asr_8, cov_sdtx3 ~ V31/((V6*V36)^0.5)) # ** sig neg even plus 2SE
vpredict(asr_8, cov_sdok ~ V32/((V10*V36)^0.5)) 
vpredict(asr_8, cov_sdmo ~ V33/((V15*V36)^0.5)) # sig pos
vpredict(asr_8, cov_sdne ~ V34/((V21*V36)^0.5)) 
vpredict(asr_8, cov_sdmi ~ V35/((V28*V36)^0.5))

#36.80566-(16.49249*2)
```
36.80566-(16.49249*2)

fit 7 (or more) bivariate models; compare h^2 estimates between them for A-H.
-> collect h^2 estimates and SE in tibbles, and compare/plot them.

----------------------
Fit trivariate models?
Try to fit 8-way model? Run these models on the server so my laptop isn't overloaded.