### 2020-11-19 R/Allele_frequencies_and_positions.Rmd
Add mash analysis with Mashattan by covariance matrix and other allele features

---
title: "Allele frequencies and positions"
author: "Alice MacQueen"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(cowplot)
library(maps)
library(ggmap)
library(ggrepel)
library(lubridate)
usa <- map_data("usa")
metadata <- read_rds("../data/metadata.rds")
```

#
```{r}
metadata %>%
  group_by(GENETIC_SUBPOP_KINSHIP) %>%
  tally()
Atlantic <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Atlantic")
Gulf <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Gulf")
Midwest <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Midwest")

metadata %>%
  group_by(ECOTYPE_NNET) %>%
  tally()
Coastal <- metadata %>%
  filter(ECOTYPE_NNET == "Coastal")
Lowland <- metadata %>%
  filter(ECOTYPE_NNET == "Lowland")
Upland <- metadata %>%
  filter(ECOTYPE_NNET == "Upland")
metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP %in% c("Atlantic", "Gulf", "Midwest")) %>%
  group_by(COLLECTION_TYPE) %>%
  tally()

# vcf.gz in SharedUser has PLANT_ID as the library names now.
writeLines(Atlantic$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Atlantic", "_subpop"))
writeLines(Gulf$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Gulf", "_subpop"))
writeLines(Midwest$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Midwest", "_subpop"))
writeLines(Coastal$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Coastal", "_ecotype"))
writeLines(Lowland$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Lowland", "_ecotype"))
writeLines(Upland$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Upland", "_ecotype"))
```

What are things I could color Manhattans by? For mash results?

Ideally, always for alternate alleles (aka not AP13's allele, in reality):

The covariance matrix that the SNP has the highest posterior probability for. Or second highest. Or whatever.

Their allele frequency in each subpopulation.

The average latitudinal and/or longitudinal position in our diversity panel. Or, as a factor, which state or region is this position in?

Which ecotype the allele is predominantly found in? Or is it common everywhere?

```{r}
allele_freq <- readRDS("~/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/Allele_frequencies_site_by_year_100000SNPs.rds")
allele_freq
```

Maybe big_count in bigsnpr could take care of this for me.
```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")

inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles")

metadata <- read_rds("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
```

```{r}
#snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/Pvirgatum_V5_GWAS_229g_Gulf_K_subset_maf0.05.rds")
snp <- snp_attach(inputfiles$SNPfiles[1])

metadata1 <- metadata %>%
  filter(PLANT_ID %in% snp$fam$sample.ID) %>%
  mutate(Atlantic = ifelse(GENETIC_SUBPOP_KINSHIP == "Atlantic", 1, 0),
         Gulf = ifelse(GENETIC_SUBPOP_KINSHIP == "Gulf", 1, 0),
         Midwest = ifelse(GENETIC_SUBPOP_KINSHIP == "Midwest", 1, 0),
         Coastal = ifelse(ECOTYPE_NNET == "Coastal", 1, 0),
         Upland = ifelse(ECOTYPE_NNET == "Upland", 1, 0),
         Lowland = ifelse(ECOTYPE_NNET == "Lowland", 1, 0),
         COLL_YEAR = year(COLL_DATE))
i=1
m1 <- metadata1 %>%
  mutate(SNP = snp$genotypes[,i]) %>%
  summarise(Alt_freq = mean(SNP/2, na.rm = TRUE),
            Alt_Latitude = weighted.mean(LATITUDE, SNP, na.rm = TRUE),
            Alt_Longitude = weighted.mean(-LONGITUDE, SNP, na.rm = TRUE),
            Alt_Elevation = weighted.mean(ELEVATION, SNP, na.rm = TRUE),
            Alt_Year = weighted.mean(COLL_YEAR, SNP, na.rm = TRUE),
            Alt_Coastal_d = weighted.mean(Coastal, SNP, na.rm = TRUE),
            Alt_Coastal_c = weighted.mean(COASTAL_ASSIGNMENT, SNP, 
                                          na.rm = TRUE),
            Alt_Lowland_d = weighted.mean(Lowland, SNP, na.rm = TRUE),
            Alt_Lowland_c = weighted.mean(LOWLAND_ASSIGNMENT, SNP, 
                                          na.rm = TRUE),
            Alt_Upland_d = weighted.mean(Upland, SNP, na.rm = TRUE),
            Alt_Upland_c = weighted.mean(UPLAND_ASSIGNMENT, SNP, na.rm = TRUE),
            Alt_Atlantic_d = weighted.mean(Atlantic, SNP, na.rm = TRUE),
            Alt_Atlantic_c = weighted.mean(ATLANTIC_Q, SNP, na.rm = TRUE),
            Alt_Gulf_d = weighted.mean(Gulf, SNP, na.rm = TRUE),
            Alt_Gulf_c = weighted.mean(GULF_Q, SNP, na.rm = TRUE),
            Alt_Midwest_d = weighted.mean(Midwest, SNP, na.rm = TRUE),
            Alt_Midwest_c = weighted.mean(MIDWEST_DA, SNP, na.rm = TRUE)
            ) %>%
  mutate(Marker = snp$map$marker.ID[i]) %>%
  select(Marker, everything())

for(i in 2:snp$genotypes$ncol){
  m_i <- metadata1 %>%
  mutate(SNP = snp$genotypes[,i]) %>%
  summarise(Alt_freq = mean(SNP/2, na.rm = TRUE),
            Alt_Latitude = weighted.mean(LATITUDE, SNP, na.rm = TRUE),
            Alt_Longitude = weighted.mean(-LONGITUDE, SNP, na.rm = TRUE),
            Alt_Elevation = weighted.mean(ELEVATION, SNP, na.rm = TRUE),
            Alt_Year = weighted.mean(COLL_YEAR, SNP, na.rm = TRUE),
            Alt_Coastal_d = weighted.mean(Coastal, SNP, na.rm = TRUE),
            Alt_Coastal_c = weighted.mean(COASTAL_ASSIGNMENT, SNP, 
                                          na.rm = TRUE),
            Alt_Lowland_d = weighted.mean(Lowland, SNP, na.rm = TRUE),
            Alt_Lowland_c = weighted.mean(LOWLAND_ASSIGNMENT, SNP, 
                                          na.rm = TRUE),
            Alt_Upland_d = weighted.mean(Upland, SNP, na.rm = TRUE),
            Alt_Upland_c = weighted.mean(UPLAND_ASSIGNMENT, SNP, na.rm = TRUE),
            Alt_Atlantic_d = weighted.mean(Atlantic, SNP, na.rm = TRUE),
            Alt_Atlantic_c = weighted.mean(ATLANTIC_Q, SNP, na.rm = TRUE),
            Alt_Gulf_d = weighted.mean(Gulf, SNP, na.rm = TRUE),
            Alt_Gulf_c = weighted.mean(GULF_Q, SNP, na.rm = TRUE),
            Alt_Midwest_d = weighted.mean(Midwest, SNP, na.rm = TRUE),
            Alt_Midwest_c = weighted.mean(MIDWEST_DA, SNP, na.rm = TRUE)
            ) %>%
  mutate(Marker = snp$map$marker.ID[i]) %>%
  select(Marker, everything())
  m1 <- m1 %>% add_row(m_i)
}

write_rds(m1, "~/Github/pvdiv-phenology-gxe/large-files/Metadata_associated_with_Alternate_alleles_363g_Midwest_and_Gulf.rds", compress = "gz")
```
Of the 363 Gulf & Midwest plants:
83.7 have ecotype assignment without the neural network.
67.7% have collection date
93.9% have lat & long; 89% have elevation

# Map this test Lat/Long
```{r}
ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) + 
  geom_point(data = m1, aes(x = SNP_Long, y = SNP_Lat), shape = 13, 
             size = 4, stroke = 1.5) + 
  labs(x = "Longitude", y = "Latitude")
```