### 2020-12-03 R/Analysis_v0.5_snp_allele_frequencies_and_other_info.Rmd
Move R scripts not used in current manuscript and rename those that are

---
title: "Allele frequencies and positions"
author: "Alice MacQueen"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(cowplot)
#library(maps)
#library(ggmap)
#library(ggrepel)
#usa <- map_data("usa")
library(switchgrassGWAS)

workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")

inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles")

metadata <- read_rds("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
```

#
```{r}

metadata %>%
  group_by(GENETIC_SUBPOP_KINSHIP) %>%
  tally()
Atlantic <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Atlantic")
Gulf <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Gulf")
Midwest <- metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP == "Midwest")

metadata %>%
  group_by(ECOTYPE_NNET) %>%
  tally()
Coastal <- metadata %>%
  filter(ECOTYPE_NNET == "Coastal")
Lowland <- metadata %>%
  filter(ECOTYPE_NNET == "Lowland")
Upland <- metadata %>%
  filter(ECOTYPE_NNET == "Upland")
metadata %>%
  filter(GENETIC_SUBPOP_KINSHIP %in% c("Atlantic", "Gulf", "Midwest")) %>%
  group_by(COLLECTION_TYPE) %>%
  tally()

# vcf.gz in SharedUser has PLANT_ID as the library names now.
writeLines(Atlantic$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Atlantic", "_subpop"))
writeLines(Gulf$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Gulf", "_subpop"))
writeLines(Midwest$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Midwest", "_subpop"))
writeLines(Coastal$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Coastal", "_ecotype"))
writeLines(Lowland$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Lowland", "_ecotype"))
writeLines(Upland$PLANT_ID, con = paste0("~/Github/pvdiv-genome/PLANT_ID/", 
                                           "Upland", "_ecotype"))
```

What are things I could color Manhattans by? For mash results?

Ideally, always for alternate alleles (aka not AP13's allele, in reality):

The covariance matrix that the SNP has the highest posterior probability for. Or second highest. Or whatever.

Their allele frequency in each subpopulation.

The average latitudinal and/or longitudinal position in our diversity panel. Or, as a factor, which state or region is this position in?

Which ecotype the allele is predominantly found in? Or is it common everywhere?

If I use vcftools to get allele frequencies, what is the kind of information that I get?
```{r}
allele_freq <- readRDS("~/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/Allele_frequencies_site_by_year_100000SNPs.rds")
allele_freq
```

Maybe big_count in bigsnpr could take care of this for me.

# get_mash_metadata
```{r}
inputdir <- file.path(workingdir, "analysis/gwas/BLUPs/single_sites", "Gulf_and_Midwest")
mash_out <- pvdiv_results_in_folder(inputdir, "Strong")
mash_out
m <- read_rds(file.path(inputdir, mash_out[12]))
markers <- switchgrassGWAS:::get_marker_df(m)
```

```{r}
snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/Pvirgatum_V5_GWAS_229g_Gulf_K_subset_maf0.05.rds")
snp <- snp_attach("~/Github/pvdiv-genome/envGWAS/Pvirgatum_V5_GWAS_678g_33M_envGWAS_maf_0.03.rds")
#snp <- snp_attach(inputfiles$SNPfiles[1])
#
metadata_env <- get_mash_metadata(m, snp, suffix = "678g_envGWAS_m_greenup_2_phe_Gulf_subpop", saveoutput = TRUE)

snp <- snp_attach("~/Github/pvdiv-genome/microbiome/Pvirgatum_V5_GWAS_630g_33M_microbiome.rds")

metadata_630 <- get_mash_metadata(m, snp, suffix = "snp_630g_PKLE_m_greenup_2_phe_Gulf_subpop", saveoutput = TRUE)
metadata_363 <- get_mash_metadata(m, snp, suffix = "snp_363g_m_greenup_2_phe_Gulf_and_Midwest_subpop", saveoutput = TRUE)

```

```{r function to get metadata}

#' @title Get switchgrass metadata for SNPs in a mash object.
#'
#' @description Takes a mash object and a bigsnp object and returns
#'     a dataframe containing summary metadata for each SNP in the mash object.
#'     This information includes details for the alternate allele of each SNP
#'     such as its frequency in the bigsnp object, the average latitude, 
#'     longitude, elevation, and year that the SNP was sampled, and how 
#'     frequently individuals with the SNP have specific subpop and ecotype 
#'     identities in the bigsnp object.
#'     NB: This is VERY slow and best for up to tens of thousands of
#'     SNPs, NOT millions of SNPs.
#'
#' @param m A mash object (outputted by mash).
#' @param snp A bigsnp object 
#' @param metadata 
#' @param saveoutput Logical. Should the output be saved to the path?
#' @param suffix Character. Optional. A unique suffix used to save the files,
#'     instead of the current date & time.
get_mash_metadata <- function(m, snp, metadata, suffix = "", 
                              saveoutput = FALSE){
  markers <- switchgrassGWAS:::get_marker_df(m)
  filtercol <- which(snp$map$marker.ID %in% markers$Marker)
  if(str_sub(suffix, end = 1) %in% c("")){
      suffix <- paste0("_", switchgrassGWAS:::get_date_filename())
    }

  metadata1 <- metadata %>%
    filter(PLANT_ID %in% snp$fam$sample.ID) %>%
    mutate(Atlantic = ifelse(GENETIC_SUBPOP_KINSHIP == "Atlantic", 1, 0),
           Gulf = ifelse(GENETIC_SUBPOP_KINSHIP == "Gulf", 1, 0),
           Midwest = ifelse(GENETIC_SUBPOP_KINSHIP == "Midwest", 1, 0),
           Coastal = ifelse(ECOTYPE_NNET == "Coastal", 1, 0),
           Upland = ifelse(ECOTYPE_NNET == "Upland", 1, 0),
           Lowland = ifelse(ECOTYPE_NNET == "Lowland", 1, 0),
           COLL_YEAR = year(COLL_DATE))
  m1 <- metadata1 %>%
    mutate(SNP = snp$genotypes[,filtercol[1]]) %>%
    summarise(Alt_freq = mean(SNP/2, na.rm = TRUE),
              Alt_Latitude = weighted.mean(LATITUDE, SNP, na.rm = TRUE),
              Alt_Longitude = weighted.mean(-LONGITUDE, SNP, na.rm = TRUE),
              Alt_Elevation = weighted.mean(ELEVATION, SNP, na.rm = TRUE),
              Alt_Year = weighted.mean(COLL_YEAR, SNP, na.rm = TRUE),
              Alt_Coastal_d = weighted.mean(Coastal, SNP, na.rm = TRUE),
              Alt_Coastal_c = weighted.mean(COASTAL_ASSIGNMENT, SNP, 
                                            na.rm = TRUE),
              Alt_Lowland_d = weighted.mean(Lowland, SNP, na.rm = TRUE),
              Alt_Lowland_c = weighted.mean(LOWLAND_ASSIGNMENT, SNP, 
                                            na.rm = TRUE),
              Alt_Upland_d = weighted.mean(Upland, SNP, na.rm = TRUE),
              Alt_Upland_c = weighted.mean(UPLAND_ASSIGNMENT, SNP, na.rm = TRUE),
              Alt_Atlantic_d = weighted.mean(Atlantic, SNP, na.rm = TRUE),
              Alt_Atlantic_c = weighted.mean(ATLANTIC_Q, SNP, na.rm = TRUE),
              Alt_Gulf_d = weighted.mean(Gulf, SNP, na.rm = TRUE),
              Alt_Gulf_c = weighted.mean(GULF_Q, SNP, na.rm = TRUE),
              Alt_Midwest_d = weighted.mean(Midwest, SNP, na.rm = TRUE),
              Alt_Midwest_c = weighted.mean(MIDWEST_DA, SNP, na.rm = TRUE)
              ) %>%
    mutate(Marker = snp$map$marker.ID[1]) %>%
    select(Marker, everything())

  for(i in 2:length(filtercol)){
    m_i <- metadata1 %>%
    mutate(SNP = snp$genotypes[,filtercol[i]]) %>%
    summarise(Alt_freq = mean(SNP/2, na.rm = TRUE),
              Alt_Latitude = weighted.mean(LATITUDE, SNP, na.rm = TRUE),
              Alt_Longitude = weighted.mean(-LONGITUDE, SNP, na.rm = TRUE),
              Alt_Elevation = weighted.mean(ELEVATION, SNP, na.rm = TRUE),
              Alt_Year = weighted.mean(COLL_YEAR, SNP, na.rm = TRUE),
              Alt_Coastal_d = weighted.mean(Coastal, SNP, na.rm = TRUE),
              Alt_Coastal_c = weighted.mean(COASTAL_ASSIGNMENT, SNP, 
                                            na.rm = TRUE),
              Alt_Lowland_d = weighted.mean(Lowland, SNP, na.rm = TRUE),
              Alt_Lowland_c = weighted.mean(LOWLAND_ASSIGNMENT, SNP, 
                                            na.rm = TRUE),
              Alt_Upland_d = weighted.mean(Upland, SNP, na.rm = TRUE),
              Alt_Upland_c = weighted.mean(UPLAND_ASSIGNMENT, SNP, na.rm = TRUE),
              Alt_Atlantic_d = weighted.mean(Atlantic, SNP, na.rm = TRUE),
              Alt_Atlantic_c = weighted.mean(ATLANTIC_Q, SNP, na.rm = TRUE),
              Alt_Gulf_d = weighted.mean(Gulf, SNP, na.rm = TRUE),
              Alt_Gulf_c = weighted.mean(GULF_Q, SNP, na.rm = TRUE),
              Alt_Midwest_d = weighted.mean(Midwest, SNP, na.rm = TRUE),
              Alt_Midwest_c = weighted.mean(MIDWEST_DA, SNP, na.rm = TRUE)
              ) %>%
    mutate(Marker = snp$map$marker.ID[filtercol[i]]) %>%
    select(Marker, everything())
    m1 <- m1 %>% add_row(m_i)
    if(saveoutput == TRUE){
    if(i%%1000 == 0){
    write_rds(m1, paste0("~/Github/pvdiv-phenology-gxe/large-files/Metadata_associated_with_alternate_alleles_", suffix, ".rds"), compress = "gz")
    }
    }
  }
  if(saveoutput == TRUE){
    write_rds(m1, paste0("~/Github/pvdiv-phenology-gxe/large-files/Metadata_associated_with_alternate_alleles_", suffix, "_full.rds"), compress = "gz")
  }
}

((1000/(100))*25)/60
((12100000/(100))*25)/60/60/24 # 35 days to run for all 12M SNPs. Yikes.
((21697/(100))*25)/60/60  # 1.5 hours to run just for the Gulf mash SNP subset.
write_rds(m1, "~/Github/pvdiv-phenology-gxe/large-files/Metadata_associated_with_alternate_alleles_229g_Gulf_subpop_m_greenup_2_phe_full.rds", compress = "gz")

m1
```
Of the 363 Gulf & Midwest plants:
83.7 have ecotype assignment without the neural network.
67.7% have collection date
93.9% have lat & long; 89% have elevation

25 seconds to run 100 SNPs.
# Map this test Lat/Long
```{r}
ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group), 
               fill = "white", color = "black") +
  coord_quickmap(xlim = c(-105, -66), ylim = c(25, 50)) + 
  geom_point(data = m1, aes(x = SNP_Long, y = SNP_Lat), shape = 13, 
             size = 4, stroke = 1.5) + 
  labs(x = "Longitude", y = "Latitude")
```