### 2020-12-03 R/Pre_manuscript/Draft_models_and_stats_2020-06-04.Rmd

---
title: "Draft models and statistics"
author: "Alice MacQueen"
date: "6/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(corpcor)
library(viridis)
```

# ASReml-R analyses on server

```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(corpcor)
```

```{r}
metadata <- readRDS("~/Github/pvdiv-phenotypes/data/metadata.rds")
metadata <- metadata %>%
  mutate(SUBPOP = case_when(!is.na(GENETIC_SUBPOP_KINSHIP) ~ as.character(GENETIC_SUBPOP_KINSHIP),
                            !is.na(PLOIDY) ~ as.character(PLOIDY),
                            TRUE ~ NA_character_),
         LATITUDE = abs(LATITUDE),
         LONGITUDE = abs(LONGITUDE)) 
phe_all <- read_rds("~/Github/pvdiv-phenology-gxe/data/all_phenotype_meas_4wcr_gwas.rds")
sites <- readRDS("~/Github/pvdiv-phenotypes/data/sites.rds")
phe_all$SUBPOP <- factor(phe_all$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
sites <- sites %>%
  mutate(manu_site = case_when(State != "TX" ~ State,
                               SITE == "KING" ~ "TX1",
                               SITE == "PKLE" ~ "TX2",
                               SITE == "TMPL" ~ "TX3",
                               SITE == "OVTN" ~ "TX4"))
sites$manu_site <- factor(sites$manu_site, levels = c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD"))

subpops <- enframe(unique(metadata$SUBPOP), name = NULL, value = "SUBPOP") %>%
  mutate(color_code = case_when(SUBPOP == "Atlantic" ~ "#6E91CB",
                                SUBPOP == "Gulf" ~ "#F47F72",    
                                SUBPOP == "Midwest" ~ "#442C83",
                                SUBPOP == "4X" ~ "grey", 
                                SUBPOP == "8X" ~ "#090906"))
subpops$SUBPOP <- factor(subpops$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))

# gulf: #F47F72, atlantic: #6E91CB, midwest: #442C83
```

```{r}
sites %>% arrange(Latitude)
```

We observed a strong signal of latitude of origin on whether plants grown at Texas common gardens flowered in lengthening or shortening days (generalized linear model on binomial output, p-value, Supp. Figure 1). 
```{r}
lat_photo_cue <- phe_all %>%
  filter(!is.na(LATITUDE) & LATITUDE > 0 & WHERE == "GWAS" & PHE == "FL50" & SITE %in% c("PKLE", "KING", "TMPL")) %>%
  mutate(Photo_cue = case_when(MEAS > 173 ~ 1,
                               MEAS <= 172 ~ 0,
                               TRUE ~ NA_real_))
lat_photo_cue$SUBPOP <- factor(lat_photo_cue$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

lat1 <- glm(Photo_cue ~ LATITUDE:SUBPOP + SUBPOP, data = lat_photo_cue, family = "binomial")

summary(lat1)

#lat2 <- glm(Photo_cue ~ LATITUDE*SUBPOP, data = lat_photo_cue, family = "binomial")
#summary(lat2)  # Think the first model is better. AIC 4610 vs 4604.5. 
#
#

# Create some new data
n2 <- 100
new.df <- data.frame(LATITUDE = 26:45, SUBPOP = rep(c("Midwest", "Gulf", "Atlantic", "4X", "8X"), length(25:45)/5))
new.df <- new.df %>% complete(LATITUDE, SUBPOP)

# Predict the probabilities
# predict(model, new.df) will just give you the logit (z) values
# We want the probabilities so use the "response" type
new.df$probs <- predict(lat1, new.df, "response")

# Draw some random values between 0 and 1
draws <- runif(n2)

# Check if draw value is less than threshold probability
new.df$results <- (draws < probs)

new.df %>%
  filter((SUBPOP == "Midwest" & between(LATITUDE, 30, 46)) | 
           !(SUBPOP %in% c("Midwest", "Gulf")) | 
          (SUBPOP == "Gulf" & between(LATITUDE, 25, 43)) ) %>%
  ggplot(aes(x= SUBPOP, y = LATITUDE)) + 
  geom_tile(aes(fill = probs)) +
  scale_fill_viridis(option = "B")

lat_photo_cue <- lat_photo_cue %>%
  ungroup() %>%
  mutate(predphoto = predict(lat1, ., "response"),
         residphoto = resid(lat1))

color_code <- c("#442C83", "#F47F72", "#6E91CB", "grey", "#090906")

lat_photo_cue %>%
  ggplot(aes(x = LATITUDE, y = predphoto)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_manual(values = color_code) + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) 
lat_photo_cue %>%
  ggplot(aes(x = LATITUDE, y = residphoto)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP))  + 
  facet_wrap(~SUBPOP)

photo_cue_prob <- lat_photo_cue %>%
  group_by(PLANT_ID) %>%
  summarise(photo_prob = mean(predphoto))
```

# Set up weather data for 2019
```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

wea_df <- ptt_df %>%
  dplyr::select(DOY:GDD_12C_SM, PTT_12C)

GR_start <- phe_all %>% 
  filter(WHERE == "GWAS") %>%
  ungroup() %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(PHE == "GR50") %>%
  summarise(GR_start = floor(min(MEAS)))
FL_end <- phe_all %>% 
  filter(WHERE == "GWAS") %>%
  ungroup() %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(PHE == "FL50") %>%
  summarise(FL_end = ceiling(max(MEAS)))

wea_gr_fl <- wea_df %>%
  left_join(GR_start) %>%
  left_join(FL_end) %>%
  group_by(Day_of_year, SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup() # Just the weather data between the minimum GR50 and max FL50.
# wea_df has all the weather, wea_gr_fl has the weather for the subset between
# greenup and flowering.
# 
gdd_12c_cue <- wea_gr_fl %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL, FL_end) %>%
  summarise(CGDD_13C = sum(GDD_13C_SM),
            CGDD_12C = sum(GDD_12C_SM),
            CPTT_12C = sum(PTT_12C)) %>%
  rename(FL50 = FL_end) %>%
  left_join(metadata)

```

# daylength cue

Additionally, we observed no evidence for any specific day length that triggered flowering (model: DYLN at FL50 ~ SUBPOP + PLANT_ID + SUBPOP:LATITUDE does not explain much variance). 
```{r}
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
dyln_cue <- wea_gr_fl %>%
  filter(Day_of_year == FL_end) %>%
  left_join(select(metadata, PLANT_ID, SUBPOP, LATITUDE)) %>%
  mutate(Photo_cue = case_when(FL_end > 173 ~ 1,
                               FL_end <= 172 ~ 0,
                               TRUE ~ NA_real_)) %>%
  filter(Photo_cue == 1) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(PLANT_ID %in% row.names(k_full)) %>%
  mutate(PLANT_ID = as_factor(PLANT_ID),
         SITE = as_factor(SITE)) %>%
  as.data.frame()
dyln_cue$SUBPOP <- factor(dyln_cue$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

asr_dyln <- asreml(DYLN ~ 1, random = 
                     ~idv(SITE) + 
                    ~vm(PLANT_ID, k_full),
                  residual = ~idv(units), data = dyln_cue, 
                  workspace = "3gb")

summary(asr_dyln)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_dyln$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_dyln_fl50_by_site_and_k_full.csv",
            col_names = TRUE)

dyln_cue$predicted <- asr_dyln$linear.predictors
dyln_cue %>% 
  as_tibble() %>%
  group_by(SUBPOP) %>%
  summarise(meanpred = mean(predicted),
            fivepred = quantile(predicted, 0.025),
            ninefivepred = quantile(predicted, 0.975),
            fiveobs = quantile(DYLN, 0.025),
            ninefiveobs = quantile(DYLN, 0.975))

blup <- summary(asr_dyln, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
int <- summary(asr_dyln, coef=TRUE)$coef.fixed %>% as_tibble(rownames = "Effect") %>% rename(std.error = `std error`)

write_csv(int, "manuscript/plots/Effects_dyln_fl50_by_site_and_k_full.csv", col_names = TRUE)
write_csv(blup, "manuscript/plots/Effects_dyln_fl50_by_site_and_k_full.csv", col_names = TRUE, append = TRUE)
```

To identify an additional environmental cue for flowering, we evaluated flowering date as a function of cumulative GDD between greenup and flowering at the northern seven sites (Kiniry et al 2005, Behrman 2013). A model with effects of subpopulation and an interaction between subpopulation and latitude of origin (linear model table) explained 55% of the variation in flowering GDD. The Midwest required the smallest GDD before flowering (585 +/- 12 GDD), and the Gulf required the largest before flowering (1238 +/- 378 GDD), with the other three subpopulations falling in between (708 – 750 +/- 106-123 GDD). 
```{r}
gdd_12c_7site <- filter(gdd_12c_cue, !(SITE %in% c("PKLE", "TMPL", "KING", "OVTN"))) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018), LATITUDE) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  left_join(photo_cue_prob)
```

```{r}

# gdd3 is the best model
gdd_12c_cue$SUBPOP <- factor(gdd_12c_cue$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))
gdd1 <- lm(CGDD_12C ~ SUBPOP+LATITUDE:SUBPOP, data = filter(gdd_12c_cue, !(SITE %in% c("OVTN"))))
summary(gdd1)
gdd9site <- filter(gdd_12c_cue, !(SITE %in% c("OVTN")) & !is.na(LATITUDE) &
                     !is.na(SUBPOP)) %>%
  ungroup() %>%
  mutate(pred9 = predict(gdd1, .),
         resid9 = resid(gdd1))
gdd9site %>% 
  dplyr::select(SUBPOP:resid9, everything()) %>%
  ggplot(aes(x = LATITUDE, y = resid9)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) + 
  facet_wrap(~SUBPOP)
gdd9site %>% 
  dplyr::select(SUBPOP:resid9, everything()) %>%
  ggplot(aes(x = pred9, y = resid9)) + 
  geom_jitter(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) + 
  facet_wrap(~SUBPOP)
```
# ASReml
```{r}
gdd_12c_7site <- filter(gdd_12c_cue, !(SITE %in% c("OVTN"))) %>%
  left_join(select(sites, SITE:Latitude)) %>%
  left_join(photo_cue_prob) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018), LATITUDE) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(!(photo_prob > 0.5 & SITE %in% c("KING", "TMPL", "PKLE", "STIL"))) %>%
    as.data.frame()

gdd_12c_7site$SUBPOP <- factor(gdd_12c_7site$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

asr_gdd <- asreml(CGDD_12C ~ 1, random = ~idv(SUBPOP) +
                    ~id(LATITUDE):idv(SUBPOP),
                  residual = ~idv(units), data = gdd_12c_7site, 
                  workspace = "3gb")

summary(asr_gdd)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_gdd$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_CGDD_12C_Subpop_LatofOrigin_filteredphoto_cue.csv",
            col_names = TRUE)

gdd_12c_7site$predicted <- asr_gdd$linear.predictors
gdd_12c_7site %>% 
  as_tibble() %>%
  group_by(SUBPOP) %>%
  summarise(meanpred = mean(predicted),
            fivepred = quantile(predicted, 0.025),
            ninefivepred = quantile(predicted, 0.975),
            fiveobs = quantile(CGDD_12C, 0.025),
            ninefiveobs = quantile(CGDD_12C, 0.975))

 blup <- summary(asr_gdd, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
int <- summary(asr_gdd, coef=TRUE)$coef.fixed %>% as_tibble(rownames = "Effect") %>% rename(std.error = `std error`)

write_csv(int, "manuscript/plots/Effects_CGDD_12C_Subpop_LatofOrigin_filteredphoto_cue.csv", col_names = TRUE)
write_csv(blup, "manuscript/plots/Effects_CGDD_12C_Subpop_LatofOrigin_filteredphoto_cue.csv", col_names = TRUE, append = TRUE)

gdd_12c_7site %>%
   as_tibble() %>%
   filter(SUBPOP == "Midwest") %>% 
  filter(LATITUDE > quantile(LATITUDE, 0.8)) %>%
   summarise(meanpred = mean(predicted))
gdd_12c_7site %>%
   as_tibble() %>%
   filter(SUBPOP == "Midwest") %>%
  filter(LATITUDE < quantile(LATITUDE, 0.2) & CGDD_12C < 1095) %>%
   summarise(meanpred = mean(predicted))

gdd_12c_7site %>%
   as_tibble() %>%
   filter(SUBPOP == "Gulf") %>% 
  filter(LATITUDE > quantile(LATITUDE, 0.8)) %>%
   summarise(meanpred = mean(predicted))
gdd_12c_7site %>%
   as_tibble() %>%
   filter(SUBPOP == "Gulf") %>%
  filter(LATITUDE < quantile(LATITUDE, 0.2)) %>%
   summarise(meanpred = mean(predicted))

```

# K GDD ASReml
```{r}
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
gdd_12c_h2 <- filter(gdd_12c_cue, !(SITE %in% c("OVTN"))) %>%
  left_join(select(sites, SITE:Latitude)) %>%
  left_join(photo_cue_prob) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018), LATITUDE) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(PLANT_ID %in% row.names(k_full)) %>%
  filter(!(photo_prob > 0.5 & SITE %in% c("KING", "TMPL", "PKLE", "STIL"))) %>%
  ungroup() %>%
  mutate(PLANT_ID = as_factor(PLANT_ID),
         SITE = as_factor(SITE)) %>%
    as.data.frame()

gdd_12c_h2$SUBPOP <- factor(gdd_12c_h2$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

asr_gddh2 <- asreml(CGDD_12C ~ 1, random = ~idv(SITE) + ~vm(PLANT_ID, k_full),
                  residual = ~idv(units), data = gdd_12c_h2, 
                  workspace = "3gb")

summary(asr_gddh2)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_gddh2$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_gdd12c_fl50_by_site_and_k_full.csv",
            col_names = TRUE)
```

# K flowering date ASReml
```{r}
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
fl50_h2 <- filter(gdd_12c_cue, !(SITE %in% c("OVTN"))) %>%
  left_join(select(sites, SITE:Latitude)) %>%
  left_join(photo_cue_prob) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018), LATITUDE) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(PLANT_ID %in% row.names(k_full)) %>%
  filter(!(photo_prob > 0.5 & SITE %in% c("KING", "TMPL", "PKLE", "STIL"))) %>%
  ungroup() %>%
  mutate(PLANT_ID = as_factor(PLANT_ID),
         SITE = as_factor(SITE)) %>%
    as.data.frame()

fl50_h2$SUBPOP <- factor(fl50_h2$SUBPOP, levels = c("Midwest", "Gulf", "Atlantic", "4X", "8X"))

asr_fl50 <- asreml(FL50 ~ 1, random = ~idv(SITE) + ~vm(PLANT_ID, k_full),
                  residual = ~idv(units), data = fl50_h2, 
                  workspace = "3gb")

summary(asr_fl50)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_fl50$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_fl50_photo_insensitive_only_by_site_and_k_full.csv",
            col_names = TRUE)
```

```{r}
gdd2 <- lm(CGDD_12C ~ SUBPOP:LATITUDE + LATITUDE, data = gdd_12c_7site)
summary(gdd2)
gdd3 <- lm(CGDD_12C ~ LATITUDE:SUBPOP + SUBPOP, data = gdd_12c_7site)
summary(gdd3)

gdd4 <- lm(CGDD_12C ~ LATITUDE*photo_prob*SUBPOP, data = gdd_12c_7site)
summary(gdd4)

gdd7site <- gdd_12c_7site %>%
  filter(!is.na(LATITUDE) & !is.na(SUBPOP)) %>%
  ungroup() %>%
  mutate(pred7lat = predict(gdd2, .),
         resid7lat = resid(gdd2),
         pred7int = predict(gdd3,.),
         resid7int = resid(gdd3))

gdd7photo <- gdd_12c_7site %>%
  ungroup() %>%
  filter(!is.na(LATITUDE) & !is.na(SUBPOP) & !is.na(photo_prob)) %>%
  mutate(pred7photo = predict(gdd4, .),
         resid7photo = resid(gdd4))

gdd7site %>%
  filter((SUBPOP == "Midwest" & between(LATITUDE, 30, 46)) | 
           !(SUBPOP %in% c("Midwest", "Gulf")) | 
          (SUBPOP == "Gulf" & between(LATITUDE, 25, 43)) ) %>%
  ggplot(aes(x= SUBPOP, y = LATITUDE)) + 
  geom_jitter(aes(color = pred7int)) +
  scale_color_viridis(option = "B")
gdd7site %>% 
  ggplot(aes(x = LATITUDE, y = resid7int)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) + 
  facet_wrap(~SUBPOP) # int better than lat in terms of residuals

gdd7site %>% 
  ggplot(aes(x = pred7int , y = resid7int)) + 
  geom_point(aes(color = SUBPOP), alpha = 0.4) + 
  scale_color_viridis(discrete = TRUE, option = "B") + 
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth(aes(color = SUBPOP)) + 
  facet_wrap(~SUBPOP) # int better than lat in terms of residuals
```

```{r}

gdd7site %>%
  group_by(SUBPOP) %>%
  summarise(meanpred = mean(pred7int),
            fivepred = quantile(pred7int, 0.05),
            ninefivepred = quantile(pred7int, 0.95),
            fiveobs = quantile(CGDD_12C, 0.05),
            ninefiveobs = quantile(CGDD_12C, 0.95))
```

```{r}

gdddf <- data.frame(LATITUDE = 26:45, SUBPOP = rep(c("Midwest", "Gulf", "Atlantic", "4X", "8X"), length(25:45)/5))
gdddf <- gdddf %>% complete(LATITUDE, SUBPOP)

# Predict the probabilities
gdddf$pred9 <- predict(gdd1, gdddf)
gdddf$pred7int <- predict(gdd3, gdddf)
gdddf$pred7lat <- predict(gdd2, gdddf)
resid(gdd2)

gdddf %>%
  filter((SUBPOP == "Midwest" & between(LATITUDE, 30, 46)) | 
           !(SUBPOP %in% c("Midwest", "Gulf")) | 
          (SUBPOP == "Gulf" & between(LATITUDE, 25, 43)) ) %>%
  ggplot(aes(x= SUBPOP, y = LATITUDE)) + 
  geom_tile(aes(fill = pred7int)) +
  scale_fill_viridis(option = "B")

gdddf %>%
  mutate(diff = pred7int-pred7lat) %>%
  ggplot(aes(x= SUBPOP, y = LATITUDE)) + 
  geom_tile(aes(fill = diff)) +
  scale_fill_viridis(option = "B")
```