### 2020-12-03 R/Pre_manuscript/Draft_models_and_stats_2020-06-19.Rmd

---
title: "Draft_models_and_stats_2020-06-19"
author: "Alice MacQueen"
date: "6/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(cowplot)
library(corpcor)
library(viridis)
```

# ASReml-R analyses on server

```{r}
library(tidyverse)
library(asreml)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
library(corpcor)
```

# Load data

```{r}
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "manuscript", "plots")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe <- read_rds(file.path(workingdir, "data",
                          "Phenology_CGDD_PTT_and_rainfall_phenotypes.rds"))
```

Unfortunately, no parents flowered exclusively during shortening days at the three Texas sites; WBC flowered during shortening days in the majority years in the Texas common gardens.
```{r}
phe %>%
  filter(WHERE == "FWCR" & PLANT_ID %in% c("AP13", "WBC", "DAC", "VS16", "VxW", "AxD")) %>%
  filter(PHE == "FL50" & SITE %in% c("PKLE", "TMPL", "KING", "OVTN", "STIL")) %>%
  group_by(PLANT_ID, SITE, MEAS >= 173) %>%
  tally() 
```
AxD KING 0 / 41
PKLE 0 / 39
TMPL 0/39

VxW 
PKLE 2 / 43
TMPL: 21 / 41
KING: 0 / 43

AP13:
TMPL: 56 / 66 >= 173
PKLE: 52 / 67 
KING: 0 / 40

WBC: 
TMPL: 40/40
PKLE: 39 / 46
KING: 10 / 18

DAC: 0% of 15
VS16: 3.6% of 56

```{r}
56/66
52/67
0
(56+52)/(66+67+40)

100
39/46
10/18
(40+39+10)/(40+46+18)
2/56
23/(43+41+43)
41+39+39
```

# daylength cue

Additionally, we observed no evidence for any specific day length that triggered flowering (model: DYLN at FL50 ~ SUBPOP + PLANT_ID + SUBPOP:LATITUDE does not explain much variance). 
```{r}
ptt_df <- readRDS(file = "~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")

wea_df <- ptt_df %>%
  dplyr::select(DOY:GDD_12C_SM, PTT_12C)

GR_start <- phe %>% 
  filter(WHERE == "GWAS") %>%
  ungroup() %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(PHE == "GR50") %>%
  summarise(GR_start = floor(min(MEAS)))
FL_end <- phe %>% 
  filter(WHERE == "GWAS") %>%
  ungroup() %>%
  group_by(SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(PHE == "FL50") %>%
  summarise(FL_end = ceiling(max(MEAS)))

wea_gr_fl <- wea_df %>%
  left_join(GR_start) %>%
  left_join(FL_end) %>%
  group_by(Day_of_year, SITE, YEAR, PLANT_ID, PLOT_GL) %>%
  filter(between(Day_of_year, GR_start, FL_end)) %>% 
  ungroup() #
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

wea_gr_fl %>%
  filter(Day_of_year == FL_end & FL_end == 175)
```

```{r}
dyln_cue <- wea_gr_fl %>%
  filter(Day_of_year == FL_end) %>%
  left_join(select(metadata, PLANT_ID, SUBPOP, LATITUDE)) %>%
  mutate(Photo_cue = case_when(FL_end >= 174 ~ 1,
                               FL_end <= 173 ~ 0,
                               TRUE ~ NA_real_)) %>%
  filter(Photo_cue == 1) %>%
  filter(!is.na(SUBPOP) & !is.na(LATITUDE)) %>%
  filter(PLANT_ID %in% row.names(k_full)) %>%
  arrange(SITE) %>%
  mutate(PLANT_ID = as_factor(PLANT_ID),
         SITE = as_factor(SITE),
         PLOT_GL = as_factor(PLOT_GL)) %>%
  as.data.frame()
dyln_cue$SUBPOP <- factor(dyln_cue$SUBPOP, levels = c("Midwest", "Gulf",
                                                      "Atlantic", "4X", "8X"))

asr_dyln <- asreml(DYLN ~ 1, random = 
                    ~vm(PLANT_ID, k_full) + ~idv(SITE),
                  residual = ~dsum(~units|SITE), data = dyln_cue, 
                  workspace = "3gb")
summary(asr_dyln)$varcomp
summary(asr_dyln)$bic
vpredict(asr_dyln, h2 ~ V2/(V1+V2+V3+V4+V5+V6+V7+V8+V9+V10+V11))
save_plot(filename=file.path(outputdir, "dyln_asreml_1.png"), 
          plot = plot(asr_dyln))

asr_dsum <- asreml(DYLN ~ 1,
                   random = ~vm(PLANT_ID, k_full) +
                     ~idv(SITE),
                   residual = ~idv(units),
                   data = dyln_cue, 
                   workspace = "3gb")
summary(asr_dsum)$varcomp
summary(asr_dsum)$bic
vpredict(asr_dsum, h2 ~ V2/(V1+V2+V3))
save_plot(filename=file.path(outputdir, "dyln_asreml_dsum.png"), 
          plot = plot(asr_dsum))

summary(asr_dyln)$bic    # first model is better
summary(asr_dsum)$bic

asr_corv <- asreml(DYLN ~ SITE,
                   random = ~vm(PLANT_ID, k_full),
                   residual = ~id(units):corv(SITE),
                   data = dyln_cue,
                   workspace = "3gb"
                   )

model.corv <- asreml(fixed = RESP ~ TIME,
                     residual = ~id(RAT):corv(TIME),
                     data = LUNG)

summary(asr_dyln)$varcomp %>%
  as_tibble(rownames = "Effect") %>%
  mutate(`%Var` = (component*100)/sum(component),
         `VarSE` = (std.error*100)/sum(component),
         lowVar = `%Var` - VarSE,
         hiVar = `%Var` + VarSE,
         loglik = asr_dyln$loglik) %>%
  write_csv(., "manuscript/plots/Variance_components_dyln_fl50_by_site_and_k_full.csv",
            col_names = TRUE)

dyln_cue$predicted <- asr_dyln$linear.predictors
dyln_cue %>% 
  as_tibble() %>%
  group_by(SUBPOP) %>%
  summarise(meanpred = mean(predicted),
            fivepred = quantile(predicted, 0.025),
            ninefivepred = quantile(predicted, 0.975),
            fiveobs = quantile(DYLN, 0.025),
            ninefiveobs = quantile(DYLN, 0.975))

blup <- summary(asr_dyln, coef=TRUE)$coef.random %>% as_tibble(rownames = "Effect")
int <- summary(asr_dyln, coef=TRUE)$coef.fixed %>% as_tibble(rownames = "Effect") %>% rename(std.error = `std error`)

write_csv(int, "manuscript/plots/Effects_dyln_fl50_by_site_and_k_full.csv", col_names = TRUE)
write_csv(blup, "manuscript/plots/Effects_dyln_fl50_by_site_and_k_full.csv", col_names = TRUE, append = TRUE)
```