### 2020-12-03 R/Pre_manuscript/Flowering_time_environmental_index_analysis.Rmd

---
title: "Environmental Index"
author: "Alice MacQueen"
date: "3/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(viridis)
library(corrplot)
library(zoo)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r}
weather <- read_rds("~/Github/pvdiv-phenotypes/data/weather_10_sites_2019.rds") %>%
  dplyr::select(DOY:DYLN) %>%
  group_by(SITE) %>%
  mutate(TMIN = ifelse(TMIN < -50, NA, TMIN),
         TMAX_SM = ifelse(is.na(TMAX),
                          (na.locf(TMAX) + na.locf(TMAX, fromLast = TRUE))/2,
                          TMAX),
         TMIN_SM = ifelse(is.na(TMIN),
                          (na.locf(TMIN) + na.locf(TMIN, fromLast = TRUE))/2,
                          TMIN),
         RAIN_SM = ifelse(is.na(RAIN), 0, RAIN))

metadata <- read_rds("~/Github/pvdiv-phenotypes/data/metadata.rds")
phe_all <- read_rds("~/Github/pvdiv-phenotypes/data/Phenotypes_long_non_Blackwell_replants_pre_and_post_2019_replant.rds")
sites <- read_rds("~/Github/pvdiv-phenotypes/data/sites.rds")
fourway_sites <- c("BRKG", "CLMB", "KBSM", "KING", "LINC", "PKLE", "STIL")
metadata$ECOTYPE <- factor(metadata$ECOTYPE, levels = c("Upland", "Coastal", "Lowland", "Unknown"))
metadata$GENETIC_SUBPOP_50PER <- factor(metadata$GENETIC_SUBPOP_50PER, levels = c("Midwest", "Atlantic", "Gulf", "Admixed", "Unknown"))
```

# My analysis based on Li's code (below)

From my pvdiv-phenotypes repository, I already had generated a data frame of weather data for all 10 GWAS sites with daylength included.

Actually, get and load smoothed weather data from pvdiv-phenotypes.

Make D2F variable which is FL50 - GR50.
```{r}
gwas_d2f <- phe_all %>%
  filter(SITE %in% fourway_sites) %>%
  filter(PHE %in% c("GR50", "FL50")) %>%
  pivot_wider(names_from = PHE, values_from = MEAS) %>%
  mutate(D2F = FL50 - GR50) %>%
  filter(!is.na(D2F))

weather <- weather %>%
  filter(SITE %in% fourway_sites)
```
fourway_seven set should include these 7 site_year combinations:
BRKG_2019
CLMB_2019
KBSM_2019
KING_2019
LINC_2019
PKLE_2019
STIL_2019

# Make GDD & PTT using a range of base temperatures
```{r}

Tbase <- seq(5, 13)
Topt <- seq(24,33)

GR_start <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(GR_start = round(mean(GR50)))
FL_end <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(FL_end = round(mean(FL50)))

GDD_PT <- NULL

for (j in Tbase){
  for (k in Topt){

    tmp <- weather %>% 
      left_join(GR_start, by = "SITE") %>%
      left_join(FL_end, by = "SITE") %>%
      group_by(SITE, Day_of_year) %>%
      filter(between(Day_of_year, GR_start, FL_end)) %>% 
      ungroup %>%
      mutate(TMEAN = (TMAX_SM + TMIN_SM) / 2,
             GDD = ifelse(TMEAN >= j, TMEAN - j, 0),
             GDD = ifelse(TMEAN >= k, k - j, GDD),
             PT = GDD * DYLN) %>%
      group_by(SITE) %>%
      summarize_at(vars(GDD, PT), funs(sum))
    colnames(tmp)[2:3] = c(paste0('GDD_', j, "_", k), paste0('PT_', j, "_", k))
    GDD_PT = c(GDD_PT, tmp)
    }
}

  GDD_PT <- GDD_PT %>% 
    as.data.frame()

  GDD_PT <- GDD_PT %>%
    select(SITE, starts_with("GDD"), starts_with("PT"))

  gwas_avegdd <- gwas_d2f %>%
    left_join(GDD_PT, by = "SITE") %>%
    ungroup()

###correlation

M = cor(gwas_avegdd %>% select(D2F, starts_with('GDD')) %>% filter(complete.cases(.)))

# corrplot::corrplot.mixed(M)

(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))

M[1,best1]

M = cor(gwas_avegdd %>% select(D2F, starts_with('PT')) %>% filter(complete.cases(.)))

# corrplot::corrplot.mixed(M)

(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))

M[1,best1]
```

# Make GDD & PTT using a range of base temperatures

And starting every individual's GDD and PTT from its own GR50 time
```{r}

Tbase <- seq(5, 13)
Topt <- seq(24,33)

GR_start <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(FL_end = ceiling(max(FL50)))

GDD_PT <- NULL

tmp1 <- weather %>% 
      left_join(GR_start, by = "SITE") %>%
      left_join(FL_end, by = "SITE") %>%
      group_by(SITE, Day_of_year) %>%
      filter(between(Day_of_year, GR_start, FL_end)) %>% 
      ungroup %>%
  left_join(gwas_d2f)

for (j in Tbase){
  for (k in Topt){

    tmp <- tmp1 %>%
      group_by(SITE, PLANT_ID, Day_of_year) %>%
      filter(between(Day_of_year, floor(GR50), ceiling(FL50))) %>%
      mutate(TMEAN = (TMAX_SM + TMIN_SM) / 2,
             GDD = ifelse(TMEAN >= j, TMEAN - j, 0),
             GDD = ifelse(TMEAN >= k, k - j, GDD),
             PT = GDD * DYLN) %>%
      ungroup %>%
      group_by(SITE, PLANT_ID) %>%
      summarize_at(vars(GDD, PT), funs(sum))
    colnames(tmp)[3:4] = c(paste0('GDD_', j, "_", k), paste0('PT_', j, "_", k))
    GDD_PT = c(GDD_PT, tmp)
    }
}

  GDD_PT <- GDD_PT %>% 
    as.data.frame()

  GDD_PT <- GDD_PT %>%
    select(SITE, PLANT_ID, starts_with("GDD"), starts_with("PT"))

  gwas_avegdd <- gwas_d2f %>%
    left_join(GDD_PT, by = c("PLANT_ID", "SITE")) %>%
    ungroup()

###correlation

M = cor(gwas_avegdd %>% select(D2F, starts_with('GDD')) %>% filter(complete.cases(.)))

# corrplot::corrplot.mixed(M)

(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))

M[1,best1]

M = cor(gwas_avegdd %>% select(D2F, starts_with('PT')) %>% filter(complete.cases(.)))

# corrplot::corrplot.mixed(M[1:15,1:15])

(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))

M[1,best1]

ggplot(data = M, aes(x = D2F, y = PT_8_33)) + 
  geom_point()
M2 <- enframe(M[1,], name = "PT", value = "corr") %>%
  filter(PT != "D2F") %>%
  separate(col = PT, into = c("Var", "Tbase", "Topt")) %>%
  dplyr::select(-Var) %>%
  pivot_wider(names_from = Topt, values_from = corr, names_prefix = "Topt_") %>%
  mutate(Tbase = paste0("Tbase_", Tbase))
M3 <- M2 %>%
  column_to_rownames(var = "Tbase") %>%
  as.matrix(.)
corrplot::corrplot.mixed(M3)
```
```{r}
write_rds(gwas_avegdd, path = "../analysis/environmental-index/GDD_and_PTT_GWAS.rds")
```

# Do analysis just using daylength, not temperature

And starting every individual's daylength received from its own GR50 time
```{r}

Tbase <- seq(5, 13)
Topt <- seq(24,33)

GR_start <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- gwas_d2f %>% 
  group_by(SITE) %>%
  summarise(FL_end = ceiling(max(FL50)))

GDD_PT <- NULL

tmp1 <- weather %>% 
      left_join(GR_start, by = "SITE") %>%
      left_join(FL_end, by = "SITE") %>%
      group_by(SITE, Day_of_year) %>%
      filter(between(Day_of_year, GR_start, FL_end)) %>% 
      ungroup %>%
  left_join(gwas_d2f)

    tmp <- tmp1 %>%
      group_by(SITE, PLANT_ID, Day_of_year) %>%
      filter(between(Day_of_year, floor(GR50), ceiling(FL50))) %>%
      ungroup %>%
      group_by(SITE, PLANT_ID) %>%
      summarize_at(vars(DYLN), funs(sum))
    GDD_PT = c(GDD_PT, tmp)

  GDD_PT <- GDD_PT %>% 
    as.data.frame()

  GDD_PT <- GDD_PT %>%
    select(SITE, PLANT_ID, starts_with("DYLN"), starts_with("PT"))

  gwas_avegdd <- gwas_d2f %>%
    left_join(GDD_PT, by = c("PLANT_ID", "SITE")) %>%
    ungroup()

###correlation

M = cor(gwas_avegdd %>% select(D2F, starts_with('DYLN')) %>% filter(complete.cases(.)))

 corrplot::corrplot.mixed(M)

(best1 <- which(M[1,] == max(M[1,2:ncol(M)])))

M[1,best1]

gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(PLANT_ID %in% c("AP13", "J001.B", "J005.B", "J006.B", "J016.B", "J024.A", "J026.B", "J350.C", "J499.B", "J517.B")) %>%
  ggplot(aes(x = D2F, y = DYLN)) + 
  geom_point(aes(color = PLANT_ID)) + 
  geom_smooth(aes(color = PLANT_ID), se = FALSE, method = lm) +
  scale_color_viridis(discrete = TRUE)

gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown", "Midwest"))) %>%
  filter(!(ECOTYPE %in% c("Upland", "Unknown"))) %>%
  ggplot(aes(x = D2F, y = DYLN.8)) + 
  geom_point(aes(color = GENETIC_SUBPOP_50PER)) + 
  geom_smooth(aes(color = GENETIC_SUBPOP_50PER), se = FALSE, method = lm) +
  scale_color_viridis(discrete = TRUE) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")

gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown", "Midwest"))) %>%
  filter(!(ECOTYPE %in% c("Upland", "Unknown"))) %>%
  ggplot(aes(x = D2F, y = DYLN.8)) + 
  geom_point(aes(color = ECOTYPE)) + 
  geom_smooth(aes(color = ECOTYPE), se = FALSE, method = lm) +
  scale_color_viridis(discrete = TRUE) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")
```

```{r}
gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown"))) %>%
  filter(!(ECOTYPE %in% c("Unknown"))) %>%
  ggplot(aes(x = D2F, y = DYLN)) + 
  geom_point(aes(color = GENETIC_SUBPOP_50PER)) + 
  geom_smooth(aes(color = GENETIC_SUBPOP_50PER), se = TRUE, method = "gam",
              formula = y ~ poly(x,2)) +
  scale_color_viridis(discrete = TRUE, end = 0.8) + ylab("One site-year of cumulative daylength") + xlab("Days to flowering (FL50 - GR50)")

save_plot(filename = "../analysis/environmental-index/Days_to_flowering_by_Genetic_subpop.png", plot = last_plot())

gwas_avegdd %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(!(GENETIC_SUBPOP_50PER %in% c("Admixed", "Unknown"))) %>%
  filter(!(ECOTYPE %in% c("Unknown"))) %>%
  ggplot(aes(x = D2F, y = DYLN)) + 
  geom_point(aes(color = ECOTYPE)) + 
  geom_smooth(aes(color = ECOTYPE), se = TRUE, method = "gam", 
              formula = y ~ poly(x,2)) +
  scale_color_viridis(discrete = TRUE, end = 0.8) + 
  ylab("One site-year of cumulative daylength") + 
  xlab("Days to flowering (FL50 - GR50)") + 
  theme(legend.position = "right")

save_plot(filename = "../analysis/environmental-index/Days_to_flowering_by_Ecotype.png", plot = last_plot())

```

# Li's code

```{r}
rm(list=ls())

dir = "~/Desktop/Phenology_4way/"

setwd(paste0(dir,"0-data/"))

library(tidyverse)

###1. combine weather data together

##function to calculate daylength

day.length.hrs <- function(day, latitude) {

  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))

  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))

  return(daylength.hrs)

}

##read into the latitude of the 7 sites to calculate the day length

Lat = read.csv('4WCR_GWAS_PlantingLocations.csv')

wthfiles = list.files(pattern='[A-Z]_2019.csv')

wth = lapply(wthfiles, function(x){

  tmp = read.csv(x)

  SITE = strsplit(x,'_',fixed = T)[[1]][1]

  lat = Lat %>% filter(Site.Code==SITE)

  tmp = tmp %>% dplyr::select(MONTH, DAY,DOY, TMAX, TMIN, RAIN)%>%

    mutate(DL = round(day.length.hrs(DOY, lat$Latitude),2)) %>% mutate(SITE =SITE, YEAR='2019')

})

wth_2019 = do.call(rbind, wth)

write.csv(wth_2019,'WeatherFile_2019_7_Sites.csv',row.names = F)

###2. read in GR50, FL50 phenology data for reation norm plot

FL50 = read.csv('FL50_7Sites.csv')

FL50 = FL50 %>% separate(ENVI, c('SITE','YEAR'))%>% mutate(D2F = FL50-GR50)

Tbase = seq(8, 13)

Topt = seq(30, 35)

FL50_Tbase = NULL

for (i in 1:nrow(FL50)){

  s = FL50$SITE[i]

  gr = floor(FL50$GR50[i])

  fl = floor(FL50$FL50[i])

  GDD_PT = NULL

  for (j in Tbase){

    for (k in Topt){

      tmp = wth_2019 %>% filter(SITE==s)%>% filter(DOY>= gr & DOY <= fl) %>%

            mutate(TMEAN=(TMAX+TMIN)/2)%>%  mutate(GDD =ifelse(TMEAN>=j, TMEAN-j, 0))%>%

            mutate(GDD =ifelse(TMEAN>=k, k-j, GDD) )%>% mutate(PT = GDD*DL)%>%

            summarize_at(vars(GDD,PT),funs(sum))

            colnames(tmp) = c(paste0('GDD_',j, k), paste0('PT_',j, k))

            GDD_PT = c(GDD_PT,tmp)

            }

  }

    GDD_PT = GDD_PT %>% as.data.frame()

    FL50_Tbase = rbind(FL50_Tbase, GDD_PT)

}

FL50 = FL50 %>% bind_cols(FL50_Tbase)

###correlation

M = cor(FL50 %>% select(D2F, matches('PT')) %>% filter(complete.cases(.)))

corrplot::corrplot.mixed(M)
```