### 2020-12-03 R/Pre_manuscript/Overlap_with_4way_QTL.Rmd

---
title: "FL50_4way_overlap"
author: "Alice MacQueen"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r}
mash <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/FL50/Strong_Effects29475SNPs.rds")
qtl <- read_csv("QTLsWithFlankMarkers_full_multiple_sites.csv")
load("~/Github/pvdiv-phenology-gxe/data/sites.rda")
```

```{r}
df <- get_significant_results(m=mash) %>%
  enframe(name = "SNP")
df1 <- switchgrassGWAS:::get_marker_df(mash)
df1 <- df1 %>%
  mutate(log10bf = switchgrassGWAS:::get_log10bf(m=mash),
         sigcond = switchgrassGWAS:::get_n_significant_conditions(mash)
         )
```

```{r}
switchgrassGWAS:::get_log10bf(m=mash)
switchgrassGWAS::mash_plot_effects(m=mash, i = 26711)

df1 %>%
  filter(log10bf > 5)

panela <- df1 %>%
  filter(grepl("Chr09K_5", Marker) & value > 26500) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B") # 26711
save_plot(filename = "Chr09K_FL50_mash_result_zoom.png", plot = last_plot(), base_asp = 1.5)

p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
save_plot(filename = "Chr09K_FL50_mash_result_zoom_2panel.png", plot = p3, base_asp = 3)
```
#Chr05N
```{r}
panela <- df1 %>%
  filter(grepl("Chr05N_1", Marker) & value > 15717) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.9) 
panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 15864)
panelb$ggobject 
p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
save_plot(filename = "Chr05N_FL50_mash_result_zoom.png", plot = p3, base_asp = 3)

```
Chr05N_15045423	15847	4.688166	8	  affects flowering at the 3 major sites
Chr05N_15048576	15849	4.101891	8	
Chr05N_15951722	15863	4.048425	7	  affects flowering at KBSM & CLMB
Chr05N_16084715	15864	4.226715	8	  affects flowering at the 3 major sites

```{r}
panela <- df1 %>%
  filter(grepl("Chr04K_1", Marker) & value > 11158) %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.9) 
panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = 11493)
panelb$ggobject 
p3 <- plot_grid(panela, panelb$ggobject, rel_widths = c(2,1))
save_plot(filename = "Chr04K_FL50_mash_result_zoom.png", plot = p3, base_asp = 3)
df1 %>%
  filter(grepl("Chr04K_1", Marker) & value > 11158 & log10bf > 4)
```
Chr04K_14442056	11332	4.954623	7	
Chr04K_14919129	11382	4.263826	8	qtl
Chr04K_14920267	11384	4.655388	8	
Chr04K_14955235	11388	4.693233	7	
Chr04K_16627327	11460	4.240948	7	
Chr04K_16687695	11493	4.405048	7	
Chr04K_18870690	11610	5.308748	6

# Li's QTL CSV

```{r}
qtl <- qtl %>%
  separate(flank_hi, into = c("Chr_hi", "_hi", "Position_hi"), sep = c(6,7), remove = FALSE) %>%
  separate(flank_lo, into = c("Chr_lo", "_lo", "Position_lo"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position_hi = as.numeric(Position_hi)*1000000,
         Position_lo = as.numeric(Position_lo)*1000000)
qtl
```

```{r}

# load("~/Github/pvdiv-phenology-gxe/data/sites.rda")

levelsSN = rev(c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
levelsNS = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING")
```

```{r}
i=1
width = 200000

for(i in 1:nrow(qtl)){
df_a <- df1 %>%
  separate(Marker, into = c("Chr", "_", "Position"), sep = c(6,7), remove = FALSE) %>%
  mutate(Position = as.numeric(Position)) %>%
  filter(grepl(qtl$chr[i], Chr) & between(Position, qtl$Position_lo[i]-width, qtl$Position_hi[i]+width)) 
panel_a <- df_a %>%
  ggplot(aes(x=Position, y = log10bf)) +
  geom_point(aes(color = sigcond)) +
  scale_color_viridis(option = "B", end = 0.9)  + 
  labs(x = "Position", y = "log10(Bayes Factor)") +
  geom_label(data = df_a[df_a$log10bf==max(df_a$log10bf),], mapping = aes(x = Position, y = log10bf, label = Marker), nudge_y = -0.4)

panelb <- switchgrassGWAS::mash_plot_effects(m=mash, i = df_a[df_a$log10bf==max(df_a$log10bf),]$value)
df_b <- panelb$effect_df %>%
  mutate(Site = str_sub(value, start =11, end = 14)) %>%
  left_join(sites) 
sites$Site <- factor(sites$Site, levels = levelsSN)
df_b$Site <- factor(df_b$Site, levels = levelsSN)
panel_b <- df_b %>%
  ggplot(aes(x = Site)) +
    geom_point(mapping = aes(y = mn)) +
    geom_errorbar(mapping = aes(ymin = mn - se,
                                ymax = mn + se, width = 0.3)) +
    geom_hline(yintercept = 0, lty = 2) +
    labs(x = "Conditions", y = "Effect Size") + coord_flip()

p3 <- plot_grid(panel_a, panel_b, rel_widths = c(2,1))
save_plot(filename = paste(qtl$Chr_lo[i], qtl$Position_lo[i], "to", qtl$Position_hi[i], "FL50_mash_result_zoom.png", sep = "_"), plot = p3, base_asp = 3)
}
```