### 2020-12-03 R/Pre_manuscript/Plots_PAG_2020.Rmd

---
title: "Plots_PAG_2020"
author: "Alice MacQueen"
date: "1/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(XLConnect)
library(rlang)
library(switchgrassGWAS)
library(cowplot)
library(viridis)
library(bigsnpr)
library(scales)
library(lubridate)

source("../../../Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
theme_set(theme_poster)

load("../data/sites.rda")
load("../data/metadata.rda")
load("../data/Taxa.rda")
load("../data/phenotypes_2019.rda")

usa <- map_data("state")
```

## Set colors

John's scheme
dark blue = MW
light blue = NEC
yellow = SEC
orange = GC
red = TX

blue = UP
orange = SEC, GC
red = TX

Plot: GR50, EM50, FL50, T_GR_FL as facets on x
10 sites as facets on y, in latitudinal order
colored by ecotype

```{r}
phenotypes_dates <- phenotypes_wide_2019 %>%
  dplyr::select(PLANT_ID, ends_with("GR50"), ends_with("EM50"), ends_with("FL50")) %>%
  pivot_longer(cols = BRKG_GR50:TMPL_FL50, names_to = "SITE_PHE", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  separate(SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-`_`)

phenotypes_dates$SITE <- factor(phenotypes_dates$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
phenotypes_dates$PHE <- factor(phenotypes_dates$PHE, levels = c("GR50", "EM50", "FL50"))

phenotypes_values <- phenotypes_wide_2019 %>%
  dplyr::select(PLANT_ID, ends_with("TC_EOS"), ends_with("HT_PAN_EOS"), ends_with("T_GR_FL")) %>%
  pivot_longer(cols = BRKG_TC_EOS:TMPL_T_GR_FL, names_to = "SITE_PHE", values_to = "Value") %>%
  filter(!is.na(Value)) %>%
  separate(SITE_PHE, into = c("SITE", "_", "PHE"), sep = c(4,5)) %>%
  dplyr::select(-`_`)

phenotypes_values$SITE <- factor(phenotypes_values$SITE, levels = c("BRKG", "KBSM", "FRMI", "LINC", "CLMB", "STIL", "OVTN", "TMPL", "PKLE", "KING"))
phenotypes_values$PHE <- factor(phenotypes_values$PHE, levels = c("T_GR_FL", "TC_EOS", "HT_PAN_EOS"))

```

```{r}
 metadata <- metadata %>%
   filter(PLANT_ID != "AP13")

metadata <- add_row(metadata, PLANT_ID = "AP13", LIBRARY = "REF", LIB_CLIMATE = "Y", LIB_PHENO = "Y", LIB_ACC = "Y", JL_genetic_subpop = "TX", JL_detail_genetic = "TX", PLOIDY = "4X", Ecotype = "Texas", Genetic_subpop = "TX")

metadata$Genetic_subpop <- factor(metadata$Genetic_subpop, levels = c("TX", "GC,TX", "GC", "SEC", "NEC,SEC", "NEC", "MW", "Admixed", "Unknown"))
metadata$Ecotype <- factor(metadata$Ecotype, levels = c("Texas", "Coastal", "Upland"))
```

```{r}

phenotypes_dates %>%
  mutate(Days = as_date(Value)) %>%
  filter(PHE != "EM50") %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
  ggplot(aes(x = Days)) + 
  geom_histogram(aes(fill = Ecotype)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#CC0000", "#FF9900", "#000077")) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) + 
  scale_y_continuous(breaks = c(20,60,100)) + 
  xlab("Date") + 
  ylab("Number of Individuals") +
  geom_vline(xintercept = c(120), linetype = 2) +
  geom_vline(xintercept = c(181), linetype = 3, color = "grey20") +
  geom_vline(xintercept = c(243), linetype = 4, color = "grey80")

save_plot("Phenology_Data_by_Site_Ecotype_noAP13_2020-01-07.png", plot = last_plot(), base_height = 8, base_asp = 1.618/2)

phenotypes_values %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(PHE != "T_GR_FL") %>%
  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
  ggplot(aes(x = Value)) + 
  geom_histogram(aes(fill = Ecotype)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#CC0000", "#FF9900", "#000077")) +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) + 
  scale_y_continuous(breaks = c(20,60,100)) +
  xlab("Value") + 
  ylab("Number of Individuals") +
  geom_vline(xintercept = c(100), linetype = 2) +
  geom_vline(xintercept = c(300), linetype = 3, color = "grey20")

save_plot("Phenotype_Data_by_Site_Ecotype_noAP13_2020-01-07.png", plot = last_plot(), base_height = 8, base_asp = 1.8/2)

phenotypes_values %>%
  left_join(metadata, by = "PLANT_ID") %>%
  filter(PHE == "T_GR_FL") %>%
  filter(Genetic_subpop != "Unknown" & PLANT_ID != "AP13") %>%
  ggplot(aes(x = Value)) + 
  geom_histogram(aes(fill = Ecotype)) + 
  facet_grid(SITE~PHE, scales = "free") +
  scale_fill_manual(values = c("#CC0000", "#FF9900", "#000077")) +
  theme(legend.position = "right", 
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 14)) + 
  scale_y_continuous(breaks = c(20,60,100)) +
  xlab("Value") + 
  ylab("Number of Individuals") +
  geom_vline(xintercept = c(100), linetype = 2) +
  geom_vline(xintercept = c(300), linetype = 3, color = "grey20")

save_plot("T_GR_FL_Data_by_Site_Ecotype_noAP13_2020-01-07.png", plot = last_plot(), base_height = 8, base_asp = 1.8/3)
```

```{r}
mashhattan <- mash_plot_manhattan_by_condition(m = mash_output)
mhtn <- mashhattan$ggmanobject + scale_color_viridis(option = "B",  end = 0.8) + scale_fill_viridis(option = "B", end = 0.8)

save_plot("Manhattan_test.png", plot = last_plot(), base_height = 4, base_asp = 2.5)

pairwise_plot <- mash_plot_pairwise_sharing(effectRDS = "Pairwise_sharing_Strong_Effects_21349SNPs.rds", reorder = FALSE)
save_plot("Pairwise_test_unordered.svg", plot = pairwise_plot$gg_corr, base_height = 2.5, base_asp = 1.618)

```

```{r}
pairwise_plot <- mash_plot_pairwise_sharing(effectRDS = "T_GR_FL/Pairwise_sharing_sign_Strong_Effects_26989SNPs.rds", reorder = TRUE)
save_plot("T_GR_FL/Pairwise_test_T_GR_FL.svg", plot = pairwise_plot$gg_corr, base_height = 2.5, base_asp = 1.618)

anno_dfs <- readRDS("T_GR_FL/Annotation_tables_Time_greenup_to_flowering_date_mash_7_sites_9_PCs.rds")
anno_dfs$top10SNPs_within20000bp
```

# Logistic winter survival

```{r}
annoswk <- readRDS("~/Github/pvdiv-climate-gwas/results-upland-vs-lowland/Logistic_winter_survival/Annotation_tables_logistic_winter_survival_subset_9_PCs.rds")
annosldly_wk <- readRDS("~/Github/pvdiv-climate-gwas/results-upland-vs-lowland/Logistic_winter_survival/Logistic_winter_survival_Lindley_sigZones_with_allele_frequencies.rds")
annoswk$top1000SNPs_within0bp %>%
  mutate(Posbin = ceiling(region_end/10000))
annosldly <- annosldly_wk %>% arrange(chr, end) %>%
  mutate(Posbin = ceiling(end/10000))
annosuncor <- annoswk$top1000SNPs_within100000bp %>%
  mutate(Posbin = ceiling(region_end/10000))
annosldly %>%
  left_join(annosuncor, by = c("chr" = "CHR", "Posbin")) %>%
  dplyr::select(chr, Posbin, everything()) %>%
  group_by(chr, Posbin) %>% tally()

pvdiv_table_topsnps(df = annosldly_wk, type = "table")

annosuncor %>%
  filter(CHR == "Chr03K")
```

```{r}
annos_winterkill <- annosldly %>%
  left_join(annosuncor, by = c("chr" = "CHR", "Posbin")) %>%
  dplyr::select(chr, Posbin, everything()) %>%
  arrange(desc(peak))

write_csv(annos_winterkill, path = "Annotations_Lindley_Winter_Killed.csv")
```

```{r}
winterkill_df <- annos_winterkill %>%
  dplyr::select(chr, beg, end) %>%
  dplyr::rename(CHR = chr, start = beg)  %>%
      mutate(start = as.integer(.data$start),
             end = as.integer(.data$end),
             POS = as.integer(.data$end))

library(VariantAnnotation)
save(winterkill_df, file = "winterkill_df.rda")

anno_info <-
  read_delim(file = file.path("~",  "Github", "pvdiv-phenology-gxe", 
                              "large-files",
                              "Pvirgatum_516_v5.1.annotation_info.txt"),
             col_names = TRUE, delim = "\t")
gene_anno_info <- tbl_df(anno_info) %>%
  distinct(locusName, .keep_all = TRUE)
txdb <- loadDb(file = file.path("~",  "Github", "pvdiv-phenology-gxe", 
                              "large-files",
                                "Pvirgatum_516_v5.1.gene.txdb.sqlite"))

 target <- with(winterkill_df, GRanges(seqnames = Rle(CHR),
                                      ranges = IRanges(start = start,
                                                       end = end,
                                                       names = NULL),
                                      strand = Rle(strand("*"))))

        ### Find genes that overlap input SNPs with VariantAnnotation
        loc <-
          VariantAnnotation::locateVariants(target, txdb,
                                            VariantAnnotation::AllVariants(promoter =
                                                                             VariantAnnotation::PromoterVariants(upstream = 2000L,
                                                        downstream = 2000L),
                                     intergenic =
                                       VariantAnnotation::IntergenicVariants(upstream = 20000L,
                                                          downstream = 20000L,
                                                          idType = "gene")))

test_df <- pvdiv_table_topsnps(df = winterkill_df, type = "table", anno_info = gene_anno_info, txdb = txdb)

load("Winterkill_lindley_annotation.rda")
library(tidyverse)
winterkill_anno_lindley <- test_df[[1]] %>% # CHR POS
  left_join(annosldly, by = c("CHR" = "chr", "POS" = "end"))
write_csv(winterkill_anno_lindley, "Winterkill_lindley_annotation_df.csv")
```