### 2020-12-03 R/Pre_manuscript/Visualize_FL50_33M_Results.Rmd

---
title: "FL50_33M_Results"
author: "Alice MacQueen"
date: "2/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(switchgrassGWAS)
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
```

##

```{r}
annos$top10SNPs_within0bp
```

```{r}
m <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/FL50/gwas_33M/Strong_Effects28384SNPs.rds")
switchgrassGWAS::mash_plot_pairwise_sharing(effectRDS = "Pairwise_sharing_Strong_Effects_28384SNPs.rds")

mash_plot_manhattan_by_condition(m = m)
mash_plot_sig_by_condition(m=m)
mash_plot_effects(m=m, n=1)
```

# Overlaps with fourway set!