### 2020-12-03 R/Analysis_v0.5_weather_covariance_Uhyp_flowering.Rmd
Add analysis with weather phenotypes as covariance U_hyp

---
title: "Weather_covariance_matrices_flowering"
author: "Alice MacQueen"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(mashr)
library(switchgrassGWAS)
library(rlist)
```
# TRIAL script
# -----------------------------
NB: Trial using covariance matrices derived from weather-related environmental cues as U_hyp in mash.

# Load data

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
workingdir <- file.path("~", "Github", "pvdiv-phenology-gxe")
outputdir <- file.path(workingdir, "analysis", "gwas", "BLUPs", "single_sites")
metadata <- read_rds(file.path(workingdir, "data", "metadata.rds"))
sites <- read_rds(file.path(workingdir, "data", "sites.rds"))
subpops <- read_rds(file.path(workingdir, "data", "subpop_color_coding.rds"))
phe_gr_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_GR50.rds"))
phe_fl_df <- read_rds(file.path(workingdir, "data",
                                "Weather_related_phe_for_asreml_h2_FL50.rds"))
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")

site_v <- list(TX1 = "TX1", TX2 = "TX2", TX3 = "TX3", OK = "OK", MO = "MO",
               NE = "NE", MI = "MI", SD = "SD")#, eight_sites = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"), tx_sites = c("TX1", "TX2", "TX3"), north_sites = c("MO", "NE", "MI", "SD"))
subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
                 Gulf = "229g_10.3M")
subpop_v2 <- list(Gulf_and_Midwest = c("Gulf", "Midwest"), Midwest = "Midwest", 
                 Gulf = "Gulf")
phe_single <- c("FL50", "dyln_fl50", "cgdd_12c_gr2fl", "crain_gr2fl",
               "crain_1d", "GR50", "cgdd_12c_10d")
phe_mult <- list(flowering_5 = c("FL50", "dyln_fl50", "cgdd_12c_gr2fl",
                                 "crain_gr2fl", "crain_1d"), 
                 greenup_2 = c("GR50", "cgdd_12c_10d"))
inputfiles <- read_delim(file.path(workingdir, "analysis", "gwas",
                                   "inputkinship.txt"),  delim = " ", 
                                   col_names = "SNPfiles")
```
# 1. Weather U_hyp

Make weather corr/covar matrices

1. Make dataframe of each env variable across all 8 sites
2. Compute correlations with pairwise complete observations
  2a. Compare correlation matrices - how dissimilar are they?
  2b. Compute correlations for Gulf, Midwest, & G&M subsets & compare
3. Experiment with changing within-site covariance to coefficient of variation for each site. Wouldn't expect these to be 1, really.

Four phenotypes to do this for, and three pop subsets.

### plot covariance matrices to look at their differences
```{r}
pvdiv_plot_cov <- function(covdf){
  U1 <- cor_phe
  if(isSymmetric(U1)){
        for(i in 1:nrow(U1)){
          for(j in 1:ncol(U1)){
            if(i < j){
              U1[i, j] <- NA
            }
          }
        }
  }
  U1 <- as_tibble(U1, rownames = "rowU", .name_repair = "unique") %>%
          pivot_longer(cols = -.data$rowU, names_to = "colU",
                       values_to = "covar") %>%
          filter(!is.na(.data$covar))
  U1$colU <- factor(U1$colU, levels = colnames(cor_phe))
  U1$rowU <- factor(U1$rowU, levels = colnames(cor_phe))

  U1_covar <- U1 %>%
    ggplot(aes(x = .data$rowU, y = .data$colU)) +
    switchgrassGWAS::theme_oeco +
    geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
    scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                    "white", "#27AD81FF", "#5DC863FF",
                                    "#FDE725FF"), limits = c(-1,1)) +
    geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
    theme(legend.position = c(0.2, 0.9),
          axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    xlab("") + ylab("") 
  return(U1_covar)
}
```

```{r}
U_hyp <- list()
ggcov <- list()
phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")

for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

  cov_df <- phe_fl_df %>%
    ungroup() %>%
    left_join(select(sites, SITE, manu_site), by = "SITE") %>%
    filter(SUBPOP %in% subpop_v2[[k]]) %>%
    select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
    group_by(PLANT_ID, SUBPOP, manu_site) %>%
    mutate(IND = row_number()) %>%
    pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
    mutate(PLANT_ID = as.factor(PLANT_ID),
           IND = as.factor(IND)) %>%
    select(-IL) %>%
    select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3)

  cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
  diag(cor_phe) <- cov_df %>%
    ungroup() %>%
    summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),
              MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
              NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
              OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
              SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
              TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
              TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
              TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
    as_vector()

  #ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
  #ggcov <- list.prepend(ggcov, ggcov1)
  U_hyp <- list.prepend(U_hyp, cor_phe)
  names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
  }
}

ggcov
```

# 2. make mashr inputs for flowering 50 phenotype

FL50 for GWAS with 90 or more individuals. 8 gardens for Gulf & G&M. 3 gardens (TX2, MO, MI) for Midwest.

## Set up mash inputs

### mash G&M
```{r}
i=1 # site
j=1 # phe
k=3 # subpop: G&M, M, Gulf
for(k in c(3)){
phe_suffix <- paste("FL50", subpop_v[k], 
                    sep = "*.*")
snp <- snp_attach(inputfiles$SNPfiles[k])
gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       phe_suffix))
phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
                           snp = snp, gwas_rds = gwas_rds, 
                           phenotypes = phenotypes, numSNPs = numSNPs, 
                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
                           suffix = paste0("FL50_", names(subpop_v)[k]))
}
```
Getting a lot of this: "error: chol(): decomposition failed"

## Run mash Gulf
```{r}
for(k in c(1)){
  phe_suffix <- paste("FL50", subpop_v[k], 
                      sep = "*.*")
  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                        pattern = paste0("GWAS_datatable*.*",
                                                         phe_suffix))
  phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
      str_replace(., "_$", "")
  phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
  numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
  m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
                             U_hyp = U_hyp,
                             suffix = paste0("FL50_", names(subpop_v)[k]),
                             saveoutput = TRUE)
  }
```

## 2b. Midwest NB
Some univariate GWAS for the Midwest had population structure issues. Mostly for greenup - even GR50 at MO was quite bad. Drop MO for GR50. Otherwise they look ok for FL50 and GR50.

This is the error I get if I try and run mash on all eight sites with just the Midwest subpopulation as input:
Error in mashr::mash_set_data(Bhat_random, Shat_random) :
  Both Bhat and Shat are zero (or near zero) for some input data. Please check your input. If it is expected please set Shat to a positive number to avoid numerical issues; or lower the threshold to call zeros using zero_check_tol (currently set to 2.22044604925031e-16). To replace zero elements you can use `zero_Bhat_Shat_reset`.

What's happening with Midwest input?  NE & OK inputs are all zeros.
```{r}
mwdf <- readRDS(file.path(outputdir,
                          "B_hat_strong_df_19000topSNPs_FL50_Midwest.rds"))

summary(mwdf) # NE, OK random Bhats are all 0's. Shat random are all vv small.

k=2
phe_suffix <- paste("FL50", subpop_v[k], 
                    sep = "*.*")
snp <- snp_attach(inputfiles$SNPfiles[k])
gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                      pattern = paste0("GWAS_datatable*.*",
                                                       phe_suffix))
gwas_rds <- gwas_rds[c(1:2,5:8)]

phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
    str_replace(., "_$", "")
phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
out1phe <- pvdiv_bigsnp2mashr(path = file.path(outputdir), 
                           snp = snp, gwas_rds = gwas_rds, 
                           phenotypes = phenotypes, numSNPs = numSNPs, 
                           clump = TRUE, scaled = TRUE, saveoutput = TRUE,
                           suffix = paste0("FL50_", names(subpop_v)[k]))

U_hyp <- list()
ggcov <- list()
phe_hyp <- c("FL50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_2d", "crain_3d", "crain_5d", "crain_7d")

for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

  cov_df <- phe_fl_df %>%
    ungroup() %>%
    left_join(select(sites, SITE, manu_site), by = "SITE") %>%
    filter(SUBPOP %in% subpop_v2[[k]]) %>%
    select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
    group_by(PLANT_ID, SUBPOP, manu_site) %>%
    mutate(IND = row_number()) %>%
    pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
    mutate(PLANT_ID = as.factor(PLANT_ID),
           IND = as.factor(IND)) %>%
    select(-IL) %>%
    select(PLANT_ID, SUBPOP, IND, MI, MO, SD, TX1, TX2, TX3)

  cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
  diag(cor_phe) <- cov_df %>%
    ungroup() %>%
    summarise(MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),
              MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
              SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
              TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
              TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
              TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
    as_vector()

  #ggcov1 <- pvdiv_plot_cov(cor_phe) + ggtitle(label = paste0(phe_hyp[i], "_", names(subpop_v2)[k]))
  #ggcov <- list.prepend(ggcov, ggcov1)
  U_hyp <- list.prepend(U_hyp, cor_phe)
  names(U_hyp)[1] <- paste0(phe_hyp[i], "_", names(subpop_v2)[k])
  }
}

for(k in c(2)){
  phe_suffix <- paste("FL50", subpop_v[k], 
                      sep = "*.*")
  gwas_rds <- pvdiv_results_in_folder(path = file.path(outputdir), 
                                        pattern = paste0("GWAS_datatable*.*",
                                                         phe_suffix))
  phenotypes_1 <- str_sub(gwas_rds, start = 16, end = -42) %>%
      str_replace(., "_$", "")
  phenotypes <- paste0(names(subpop_v)[k], "_", phenotypes_1)
  numSNPs <- round(ceiling(1200000/length(phenotypes)^2)/1000)*1000
  m_out <- mash_standard_run(path = outputdir, numSNPs = numSNPs, 
                             U_hyp = U_hyp,
                             suffix = paste0("FL50_", names(subpop_v)[k]),
                             saveoutput = TRUE)
  }
```

# -----------------------------

# 3. Analyse mash output