### 2021-01-26 R/Analysis_v0.6_QTL_GxE_and_mash_comparison.Rmd
Finish mash Results section writeup

---
title: "Analysis_v0.7_QTL_GxE"
author: "Alice MacQueen"
date: "1/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TRIAL script
# -------------------------
NB: Trial script to compare the QTL to the mash results using env covar matrices.

## 3a. Log-likelihoods
```{r}
m_gulf_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_GR50_Gulf_Uhyp.rds")
m_gulf_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf/Strong_Effects_19000_SNPs_FL50_Gulf_Uhyp.rds")

m_midw_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_24000_SNPs_GR50_Midwest_Uhyp.rds")
m_midw_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Midwest/Strong_Effects_33000_SNPs_FL50_Midwest_Uhyp.rds")

m_gam_gr_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_GR50_Gulf_and_Midwest_Uhyp.rds")
m_gam_fl_hyp <- readRDS("~/Github/pvdiv-phenology-gxe/analysis/mash/Gulf_and_Midwest/Strong_Effects_19000_SNPs_FL50_Gulf_and_Midwest_Uhyp.rds")

```

## QTL
```{r}
qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")

qtldf <- qtldf %>%
  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
  arrange(CHR, Pos_Mb_lo)

qtlplot <- qtldf %>%
  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
                              lodcolumn == "dyln_fl50" ~ "Daylength",
                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
  filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
  mutate(POS_lo = Pos_Mb_lo * 1000000,
         POS_hi = Pos_Mb_hi * 1000000)

qtlplot$phe <- factor(qtlplot$phe, 
                      levels = c("50% flowering", "Rainfall, GR to FL",
                                 "GDD, GR to FL", "Daylength",
                                          "daylength change (s)",
                                          "Rainfall, 1 day",
                                                  #"Rainfall, 2d sum",
                                                  "Rainfall, 3d sum",
                                                  "Rainfall, 5d sum"#,
                                                  #"Rainfall, 7d sum"
                                                  ))
```

```{r}
get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

  pw_df <- as_tibble(m$posterior_weights)
  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
    bind_cols(pw_df)

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    left_join(log10bf_df, by = "Marker") %>%
    arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

pw_gg <- get_post_weights(m_gulf_gr_hyp) %>%
  mutate(subpop = "Gulf",
         phenotype = "greenup")
pw_gf <- get_post_weights(m_gulf_fl_hyp) %>%
  mutate(subpop = "Gulf",
         phenotype = "flowering")

pw_mg <- get_post_weights(m_midw_gr_hyp) %>%
  mutate(subpop = "Midwest",
         phenotype = "greenup")
pw_mf <- get_post_weights(m_midw_fl_hyp) %>%
  mutate(subpop = "Midwest",
         phenotype = "flowering")

pw_gamg <- get_post_weights(m_gam_gr_hyp) %>%
  mutate(subpop = "Both",
         phenotype = "greenup")
pw_gamf <- get_post_weights(m_gam_fl_hyp) %>%
  mutate(subpop = "Both",
         phenotype = "flowering")
```

```{r}
pw_3f <- pw_gf %>%
  full_join(pw_mf) %>%
  full_join(pw_gamf) %>%
  arrange(Chr, Pos) %>%
  filter(log10BayesFactor > 2) %>%
  select(Chr:Pos, log10BayesFactor, subpop, phenotype, everything())

qtlplot

pw_3f %>%
  filter(Chr %in% qtlplot$CHR[1] & between(Pos, qtlplot$POS_lo[1], qtlplot$POS_hi[1])) %>%
  group_by(subpop) %>% tally()
```