### 2021-01-27 R/Analysis_v0.6_QTL_interval_plotting.Rmd

---
title: "QTL_interval_plotting"
author: "Alice MacQueen"
date: "1/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(qtl)
library(qtl2)
library(qtl2convert)
library(tidyverse)
```

#

```{r prep QTL_Eff df}
###QTL Effect plot
###load the original cross and reduce the markers from 4700 to 1185
load('../phenology-fourway/5-QTLmappingForAlice/sexAveraged.cross.rda', verbose = T)
map  =  lapply(pull.map(cross), function(x) x[1,])
# reduced_markers  =  unlist(
#   lapply(
#     reduce_markers(
#       lapply(pull.map(cross), function(x)
#         x[1,]), min_distance = 1), names))
#
# mars2drop  =  markernames(cross)[!markernames(cross) %in% reduced_markers]
# cross  =  drop.markers(cross, mars2drop)
cross = calc.genoprob(cross, step=1, error.prob = 1e-3,map.function = "kosambi",stepwidth = "max")
######### Converting probalibity from qtl1 to qtl2
probs  =  probs_qtl_to_qtl2(cross)$probs
##calculate the kinship matrix for single site
kinship_loco_Single = calc_kinship(probs,use_allele_probs = F,type = "loco")

###extract the genotype ID from the original cross file, so we can align the phenology file with genotype ID
ID <- data.frame(ID=cross$pheno$id)
ID <- ID %>% mutate_at(vars('ID'), list(as.character)) %>% mutate_at(vars('ID'), list(as.numeric))

######get the phenotype at each site
phenology = read.csv('../phenology-fourway/5-QTLmappingForAlice/FL50_phenotypes_Alice.csv', header = TRUE)
colnames(phenology)[1:2] = c('ID','Covariate')
str(phenology) ###make sure your phenology traits are numeric, if not, change them into numeric
phenos <- phenology %>% mutate_at(vars('ID'), list(as.character)) %>%
  mutate_at(vars('ID'), list(as.numeric))%>% group_by(Covariate,ID ) %>%
  summarize_all(mean, na.rm=T) %>% ungroup()%>% as.data.frame()
```

```{r setup QTL effects for all traits}
###QTL effect for each trait
EffectMatrix = vector('list', length=0)
t=0
for (envi in unique(phenos$Covariate)){
    t=t+1
    pheno = phenos %>% filter(Covariate==envi) %>% dplyr::select(-Covariate) %>%
      select(ID, FL50) %>% right_join(ID) %>%  column_to_rownames('ID') %>% as.matrix()

    ##ran effect scan for each chromosome
    eff_out = vector('list',length = 0)
    se_out = vector('list',length = 0)
    add_contrasts=cbind(mu=c(1,1,1,1), add1=c(-1, 1, -1, 1),add2=c(-1, -1, 1, 1), dom=c(-1, 1, 1, -1))

    z = 0
    for (chr in chrnames(cross)){
      print(c(chr, envi,t))
      z= z+1
      eff = scan1coef(probs[,chr],pheno,kinship = kinship_loco_Single[chr], se=T, contrasts = add_contrasts)
      se = attr(eff, 'SE')
      eff_out[[z]] =  eff %>% as.data.frame() %>% rownames_to_column('MARKER')
      se_out[[z]] =  se %>% as.data.frame() %>% rownames_to_column('MARKER')
    }
    eff_out = bind_rows(eff_out)
    se_out = bind_rows(se_out)

    se_out2 = se_out %>% dplyr::select(-mu) %>% gather('CROSS','SE',-MARKER)
    eff_out2 = eff_out %>% dplyr::select(-mu) %>% gather('CROSS','EFF',-MARKER) %>% mutate(SE=se_out2$SE, Covariate=envi)

    EffectMatrix[[t]] = eff_out2
  }

EffectMatrix2 = bind_rows(EffectMatrix)

###unlist nmap and merge with Effect file
nmap = data.frame(pos = unlist(map))
nnmap = nmap %>% rownames_to_column('MARKER') %>% rowwise() %>%
  mutate(chr = str_sub(MARKER, 1,2), MARKER=str_sub(MARKER, 4,nchar(MARKER)))%>%
  ungroup() %>% mutate(pos=round(pos,4))

flank_marker_full = read.csv('../phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv')
flank_marker_full = flank_marker_full %>% mutate_at(vars(chr),list(as.character))

for (i in 1:nrow(flank_marker_full)){
  tmp = nnmap %>% filter(chr==flank_marker_full$chr[i] & pos==round(flank_marker_full$pos[i],4))
  flank_marker_full$MARKER[i] = tmp$MARKER
}

QTL_full_model = flank_marker_full %>% dplyr::rename(trait=lodcolumn)

```

```{r ggplots of QTL effects}
QTLdf <- QTL_full_model %>% left_join(EffectMatrix2)%>%
  mutate(CROSS = str_replace(CROSS,'add2','C x D')) %>%
  mutate(CROSS = str_replace(CROSS,'add1','A x B')) %>%
  mutate(Garden = case_when(Covariate == "BRKG" ~ "SD",
                            Covariate == "CLMB" ~ "MO",
                            Covariate == "KBSM" ~ "MI",
                            Covariate == "KING" ~ "TX1",
                            Covariate == "LINC" ~ "NE",
                            Covariate == "PKLE" ~ "TX2",
                            Covariate == "STIL" ~ "OK",
                            TRUE ~ "Other"),
         Cross = case_when(CROSS == "A x B" ~ "AP13 x DAC",
                           CROSS == "C x D" ~ "VS16 x WBC",
                           CROSS == "dom" ~ "Dominance",
                           TRUE ~ "Other")) %>%
  separate(MARKER, into = c("Chr", "Pos"), sep = "_", convert = TRUE, 
           remove = FALSE) %>%
  mutate(Pos_plot = paste0(Chr, "\n", round(Pos, 1), " Mb"))

QTLdf$Covariate = factor(QTLdf$Covariate, levels=c('KING','PKLE','STIL','CLMB',
                                                       'LINC','KBSM','BRKG'))
QTLdf$Garden = factor(QTLdf$Garden, levels = c("TX1", "TX2", "OK", "MO", 
                                                   "NE", "MI", "SD"))

QTLdf %>% 
  filter(Cross != "Dominance") %>% # trait == 'FL50' &
  ggplot(aes(x = Garden, y = EFF, group = Cross)) +
  geom_point(size = 2) + facet_grid(Cross + trait ~ Pos_plot) +
  geom_hline(yintercept = 0, linetype=2) + 
  geom_errorbar(aes(ymin=EFF-SE, ymax=EFF+SE), size=0.6, width=0.5) +
  xlab("") +
  ylab('QTL Effect') + 
  coord_flip() + 
  theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        text=element_text(size=12, face="bold"),
        axis.title=element_text(size=12,face="bold"),
        axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 1))
# shared QTL for more than one trait differ in LOD score, but not meaningfully in effect patterns across gardens. Useful to know.

QTLdf %>% 
  filter(trait == 'CGDD_12C') 
QTLdf %>% 
  filter(trait == 'dyln_fl50')
QTLdf %>% 
  filter(trait == 'dyln_change_sec') 
```