### 2021-06-29 R/UT_Dataverse_files.Rmd
Prep supplement and stable data replication links for SNP data

---
title: "UT Dataverse"
author: "Alice MacQueen"
date: 2021-06-29
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(bigsnpr)
```

# Make bed files

Make bed/bim/fam SNP files for Midwest, Gulf, and Both genetic subpopulations.
I've used '.rds' and '.bk' files for the analyses in this manuscript, but these
are not common genome file formats, while bed formatted files are. So uploading
files in this format to a permanent DOI will help reproducibility for this
dataset.

I only need to do this for the 'Both' subpopulations; the Midwest and Gulf
subpopulation SNP sets were generated for the genome paper.

```{r}
snp <- snp_attach("~/Github/pvdiv-genome/genetic_subpops/kinship/Pvirgatum_V5_GWAS_363g_Midwest_and_Gulf_subset_maf0.05.rds")
```

```{r}
snp_writeBed(snp, bedfile = "~/Github/pvdiv-genome/Dataverse/Pvirgatum_V5_GWAS_363g_Midwest_and_Gulf_subset_maf0.05.bed")
```