### 2022-01-09 R/Analysis_v1.1_weather_2020.Rmd

---
title: "Analysis_v1.1_weather"
author: "Alice MacQueen"
date: "1/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(zoo)
```

#

```{r}
source("~/Github/pv-weather/R/get_functions.R")
sites <- read_rds(here("data", "sites.rds"))
metadata <- read_rds(here("data", "metadata.rds"))
mesonet_stations <- read_csv("~/Github/pv-weather/data-raw/All_Mesonet_Stations.csv")

get_station_info <- function(df, decdegrees = 0.1, loc_key = "Location_code"){
  loc_key <- sym(loc_key)
  locs <- list()
  Location_finder <- df %>%
    group_by(!! loc_key) %>%
    summarize(Lat1 = mean(Latitude, na.rm = TRUE) - decdegrees,
              Long1 = mean(Longitude, na.rm = TRUE) - decdegrees,
              Lat2 = Lat1 + decdegrees*2,
              Long2 = Long1 + decdegrees*2)
  for(i in 1:nrow(df)){
    locs[[i]] <- mesonet_stations %>%
      filter(between(lon, Location_finder$Long1[i], Location_finder$Long2[i]) &
               between(lat, Location_finder$Lat1[i], Location_finder$Lat2[i])) %>%
      arrange(begints)
  }
  station_info <- do.call(rbind.data.frame, locs)
  return(station_info)
}

Station1 <- metadata %>%
  filter(PLANT_ID == "J001.A") %>%
  rename(Latitude = LATITUDE, Longitude = LONGITUDE) %>%
  mutate(Longitude = -Longitude)
get_station_info(Station1, loc_key = "PLANT_ID")
Station1$station <- "BKV"
J001_2018 <- get_mesonet_weather(station = Station1$station, year1 = "2018", 
                                 month1 = "06", day1 = "01", year2 = "2021", 
                                 month2 = "11", day2 = "01", 
                                 tz = "America/New_York")
```

SD stid: SMCS2 (or MDYS2)
MO stid: MO027
IL stid: WSCI2 (don't need this)
MI stid: KBSM4 (or GLLM4)
NE stid: MPHN1

```{r}
meso_url <- "https://mesonet.agron.iastate.edu/cgi-bin/request/asos.py/"
library(httr)
library(data.table)
library(lubridate)
#' @import httr
#' @import data.table
#' @import dplyr
#' @import lubridate
#'
#' @export
get_mesonet_weather <- function(station, year1, month1, day1, year2, month2, day2, tz){
  MESO_WEA <- GET(url = meso_url,
                  query = list(station = station,
                               data = "all",
                               year1 = year1,
                               month1 = month1,
                               day1 = day1,
                               year2 = year2,
                               month2 = month2,
                               day2 = day2,
                               tz = tz,
                               format = "comma")) %>%
    content() %>%
    read_csv(skip = 5, na = "M") %>%
    mutate(Working_date = as_date(valid)) %>%
    group_by(Working_date) %>%
    summarise(tmin = (min(as.numeric(tmpf), na.rm = TRUE)-32)*(5/9), # convert F to C for all temperatures
              tmax = (max(as.numeric(tmpf), na.rm = TRUE)-32)*(5/9),
              RHmean = mean(as.numeric(relh), na.rm = TRUE),
              dewp = (mean(as.numeric(dwpf), na.rm = TRUE)-32)*(5/9),
              wndk = mean(as.numeric(sknt), na.rm = TRUE),
              prcp = (sum(as.numeric(p01i), na.rm = TRUE)*254), # convert inches to 0.1x mm
              Tdrnl = mean(tmax - tmin),
              Tmean = (mean(as.numeric(tmpf), na.rm = TRUE)-32)*(5/9)) %>%
    mutate_if(is.double, funs(na_if(., NaN)))
  DT <- data.table(MESO_WEA)
  invisible(lapply(names(DT),function(.name) set(DT,
                                                 which(is.infinite(DT[[.name]])),
                                                 j = .name,value =NA)))
  invisible(lapply(names(DT),function(.name) set(DT, which(is.nan(DT[[.name]])),
                                                 j = .name,value =NA)))
  return(DT)
}

get_mesonet_weather(station = "WSCI2", year = "2019", month1 = "01", day1 = "01", year2 = "2021", month2 = "10", day2 = "31", tz = "America/Chicago")
```
Only the ASOS stations seem to work. Still, nothing is farther than 0.18 decimal degrees away from the field site. Given that the techs haven't added weather data for the site proper, that's just going to have to be good enough.
```{r}
sites[1,]
get_station_info(sites[1,], decdegrees = 0.15, loc_key = "manu_site")
Station1$station <- "BKX"
test_SD <- get_mesonet_weather(station = Station1$station, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[2,]
get_station_info(sites[2,], decdegrees = 0.1, loc_key = "manu_site")
Station2 <- "COU"
test_MO <- get_mesonet_weather(station = Station2, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[3,]
get_station_info(sites[3,], decdegrees = 0.1, loc_key = "manu_site")
Station3 <- "DPA"
test_IL <- get_mesonet_weather(station = Station3, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[4,]
get_station_info(sites[4,], decdegrees = 0.15, loc_key = "manu_site")
Station4 <- "BTL"
test_MI <- get_mesonet_weather(station = Station4, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[5,]
get_station_info(sites[5,], decdegrees = 0.1, loc_key = "manu_site")
Station5 <- "NQI"
test_TX1 <- get_mesonet_weather(station = Station5, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[6,]
get_station_info(sites[6,], decdegrees = 0.18, loc_key = "manu_site")
Station6 <- "AHQ"
test_NE <- get_mesonet_weather(station = Station6, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[7,]
get_station_info(sites[7,], decdegrees = 0.18, loc_key = "manu_site")
Station7 <- "RFI"
test_TX4 <- get_mesonet_weather(station = Station7, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[8,]
get_station_info(sites[8,], decdegrees = 0.1, loc_key = "manu_site")
Station8 <- "ATT"
test_TX2 <- get_mesonet_weather(station = Station8, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[9,]
get_station_info(sites[9,], decdegrees = 0.18, loc_key = "manu_site")
Station9 <- "SWO"
test_OK <- get_mesonet_weather(station = Station9, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

sites[10,]
get_station_info(sites[10,], decdegrees = 0.15, loc_key = "manu_site")
Station10 <- "TPL"
test_TX3 <- get_mesonet_weather(station = Station10, year1 = "2018", 
                                 month1 = "01", day1 = "01", year2 = "2021", 
                                 month2 = "12", day2 = "31", 
                                 tz = "America/Chicago")

```

```{r}
wea_MI <- test_MI %>%
  mutate(manu_site = "MI") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_SD <- test_SD %>%
  mutate(manu_site = "SD") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_MO <- test_MO %>%
  mutate(manu_site = "MO") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_TX1 <- test_TX1 %>%
  mutate(manu_site = "TX1") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_TX2 <- test_TX2 %>%
  mutate(manu_site = "TX2") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_TX3 <- test_TX3 %>%
  mutate(manu_site = "TX3") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_OK <- test_OK %>%
  mutate(manu_site = "OK") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_NE <- test_NE %>%
  mutate(manu_site = "NE") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_TX4 <- test_TX4 %>%
  mutate(manu_site = "TX4") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))
wea_IL <- test_IL %>%
  mutate(manu_site = "IL") %>%
  left_join(select(sites, manu_site, Latitude, Longitude, Elevation))

wea_ten_sites <- wea_MI %>%
  full_join(wea_SD) %>%
  full_join(wea_MO) %>%
  full_join(wea_TX1) %>%
  full_join(wea_TX2) %>%
  full_join(wea_TX3) %>%
  full_join(wea_OK) %>%
  full_join(wea_NE) %>%
  full_join(wea_TX4) %>%
  full_join(wea_IL)

##function to calculate daylength

day.length.hrs <- function(day, latitude) {

  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))

  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))

  return(daylength.hrs)

}

```

```{r}
wea_ten_sites

phenotypes <- readRDS("~/Github/pvdiv-phenotypes/data/GWAS/Phenotypes_gwas_all_subpops_2018_to_2021.rds")

phe_ten_sites <- phenotypes %>%
  select(PLANT_ID, matches("GR50"), matches("FL50")) %>%
  pivot_longer(cols = -PLANT_ID, names_to = "SITE_YEAR_PHE", values_to = "JDAY") %>%
  separate(SITE_YEAR_PHE, into = c("SITE", "Year", "PHE"), sep = "_", convert = TRUE) %>%
  filter(!is.na(JDAY)) %>%
  left_join(select(sites, SITE, manu_site))

phe_ten_sites %>%
  write_rds(file = here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
wea_ten_sites %>%
  write_rds(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))

```

Can make most weather variables before I join the weather dataframe to the phenotype dataframe. Can make most of these with mutate and the lag function, even if it's a bit ugly looking in terms of code.

Greenup
#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
#             "tave_5d", "tave_10d", "tave_18d")
```{r}
colSums(is.na(wea_ten_sites))

wea_ten_sites <- wea_ten_sites %>%
  mutate(JDAY = yday(Working_date),
         DYLN = day.length.hrs(day = JDAY, latitude = Latitude),
         tmax_sm = ifelse(is.na(tmax),
                          (na.locf(tmax) + na.locf(tmax, fromLast = TRUE))/2,
                          tmax),
         tmin_sm = ifelse(is.na(tmin),
                          (na.locf(tmin) + na.locf(tmin, fromLast = TRUE))/2,
                          tmin),
         tmean_sm = ifelse(is.na(Tmean),
                          (na.locf(Tmean) + na.locf(Tmean, fromLast = TRUE))/2,
                          Tmean),
         GDD_12Cbase = ifelse((tmax_sm + tmin_sm)/2 >= 12.0,
                              (tmax_sm + tmin_sm)/2 - 10,
                              0),
         Year = year(Working_date)) %>%
  arrange(Latitude, Year, JDAY) %>%
  group_by(manu_site) %>%
  mutate(crain_2d = prcp + lag(prcp, default = 0),
         crain_3d = prcp + lag(prcp, default = 0) + lag(prcp, default = 0, 
                                                        n = 2),
         crain_5d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4),
         dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         dyln_change_sec_7d_prior = (lag(DYLN, n = 6) - 
                                       lag(DYLN, n = 7))*60*60,
         dyln_change_sec_14d_prior = (lag(DYLN, n = 13) - 
                                       lag(DYLN, n = 14))*60*60,
         dyln_7d_prior = lag(DYLN, n = 6),
         dyln_14d_prior = lag(DYLN, n = 13),
         cgdd_12c_5d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4))/5,
         tave_5d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4))/5,
         cgdd_12c_10d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9))/10,
         tave_10d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9))/10,
         cgdd_12c_20d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13) +
           lag(GDD_12Cbase, default = 0, n = 14) +
           lag(GDD_12Cbase, default = 0, n = 15) +
           lag(GDD_12Cbase, default = 0, n = 16) +
           lag(GDD_12Cbase, default = 0, n = 17) +
           lag(GDD_12Cbase, default = 0, n = 18) +
           lag(GDD_12Cbase, default = 0, n = 19))/20,
         tave_20d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13) +
           lag(tmean_sm, default = 0, n = 14) +
           lag(tmean_sm, default = 0, n = 15) +
           lag(tmean_sm, default = 0, n = 16) +
           lag(tmean_sm, default = 0, n = 17) +
           lag(tmean_sm, default = 0, n = 18) +
           lag(tmean_sm, default = 0, n = 19))/20)

```
Flowering
phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

If I join the current wea_ten_sites to the greenup and flowering df, that should leave me with variables I want just for the dates & greenup & flowering.
Including some new daylength ones I made which will be 1-2 weeks before flowering.

The greenup to flowering variables will need to be made in a separate step. 
These are cgdd_12c_gr2fl and crain_gr2fl. This cumulative sum is different for each year, site, and plant id.
```{r}

Tbase <- 12

phe_gr_fl <- phe_ten_sites %>%
  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
                              Year == 2019 ~ JDAY + 365,
                              Year == 2020 ~ JDAY + (365*2),
                              Year == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  select(-JDAY) %>%
  pivot_wider(names_from = PHE, values_from = JDAY_EXP)

GR_start <- phe_gr_fl %>% 
  group_by(manu_site, Year) %>%
  filter(!is.na(GR50)) %>%
  summarise(GR_start = floor(min(GR50)))
FL_end <- phe_gr_fl %>% 
  group_by(manu_site, Year) %>%
  filter(!is.na(FL50)) %>%
  summarise(FL_end = ceiling(max(FL50)))

tmp1 <- wea_ten_sites %>%
  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
                              Year == 2019 ~ JDAY + 365,
                              Year == 2020 ~ JDAY + 365 + 365,
                              Year == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  left_join(phe_gr_fl)

tmp <- tmp1 %>%
  group_by(manu_site, PLANT_ID, Year, JDAY_EXP) %>%
  filter(between(JDAY_EXP, floor(GR50), ceiling(FL50))) %>%
  ungroup %>%
  group_by(manu_site, Year, PLANT_ID) %>%
  summarize_at(vars(GDD_12Cbase, prcp), funs(sum))

tmp %>%
  filter(manu_site == "TX1")
tmp
```

```{r}
phe_ten_sites %>%
  left_join(wea_ten_sites)
```