### 2023-08-02 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
revision: More covariance matrices to add to greedy mash algorithm

---
title: "Analysis v1.2 mash"
author: "Alice MacQueen"
date: "2023-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
library(rlist)
library(cowplot)
library(here)
library(zoo)
```

# Installation reminders for any reinstalls
On debian: `apt install libgsl-dev` to install the GNU Scientific Library, zB.

#

Sam used a greedy mash algorithm to find the best set of covariance matrices to include, using 19K randomly associated variants.

Now use the best set of covariance matrices on 19K strongly associated variants.

Step 0. Set up.
  0a. Look at phenotype distribution & heritability; run GWAS to get univariate SNP effect estimates & standard errors. Already done.
  0b. Make hypothesis covariance matrices based on weather data from 2018 - 2020. For cumulative weather variables, use a range of days, 1 - 20, say. Come back and add more if 20 is chosen by greedy mash.
Step 1. Run greedy mash to find best set of covariance matrices that explain SNP effect variation across environments, using 19K 'randomly associated variants'. Go back to 0b. if high number of day covar matrices tend to be chosen by greedy mash.
Step 2. Use covariances from (1) to run mash on 19K 'strongly associated variants'. 
3. Look at loadings on covariance matrices and meanings of covariance matrices as the writeup (Figure 2, 3; Table 1). Look at amount of antagonistic pleiotropy vs same sign significant effects.
4. Repeat the QTL analysis... with Li.

How much of a lift is it to re-run the greedy mash algorithm? If we, or if reviewers are thinking about this in a Jiaming Yu sense, I think the next logical question to ask is 'Why these covariance matrices'? The overly honest answer for the current set is that there were computational limits to running mash with so many covariance matrices, so I chose a few arbitrary time points for each weather variable we thought might be a environmental driver. But now we have introduced a method to choose between highly correlated covariance matrices, perhaps we could do better. 
2.

# Deprecated analysis
### Code from Plots_v0.2_manuscript_figures.Rmd to make weather variables

Actually, code from Analysis_v1.1_weather_2020.Rmd might be more suitable, if I am remaking these anyway.

```{r}
wea_df <- readRDS("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```

```{r}
jday_phe <- wea_df %>%
  group_by(SITE) %>%
  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         FL50 = yday(DOY)) %>%
  select(DOY, DYLN, dyln_change_sec, everything()) %>%
  select(-PTT_12C) %>%
  right_join(jday_phe, by = c("SITE", "FL50")) %>%
  select(-(TMAX:CPTT_11C))
jday_phe <- metadata %>%
  rename(Lat_origin = LATITUDE, Long_origin = LONGITUDE) %>%
  select(PLANT_ID, SUBPOP, Lat_origin, Long_origin) %>%
  right_join(jday_phe) %>%
  left_join(sites)

jday_phe$SUBPOP <- factor(jday_phe$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
jday_phe$SITE <- factor(jday_phe$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
jday_phe$manu_site <- factor(jday_phe$manu_site, levels = rev(c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD")))
```

## define environmental cues for greenup
Probably all CGDD related.
```{r}

asreml_fl <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, FL50, GR50, CGDD_12C, CRAIN, CRAIN_5d, DYLN, dyln_change_sec) %>%
  rename(dyln_fl50 = DYLN, cgdd_12c_gr2fl = CGDD_12C, crain_gr2fl = CRAIN) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(between(DOY, as_date(FL50, origin = "2019-01-01") - 6, as_date(FL50, origin = "2019-01-01"))) %>%
  mutate(crain_2d = RAIN_SM + lag(RAIN_SM),
         crain_3d = RAIN_SM + lag(RAIN_SM) + lag(RAIN_SM, n = 2),
         crain_7d = sum(RAIN_SM)) %>%
  filter(DOY == as_date(FL50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX:TMIN_SM), -(GDD_13C_SM:CPTT_13C)) %>%
  rename(crain_1d = RAIN_SM, crain_5d = CRAIN_5d)
saveRDS(asreml_fl, file = "../data/Weather_related_phe_for_asreml_h2_FL50.rds")
asreml_gr <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, GR50) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(DOY <= as_date(GR50, origin = "2019-01-01")) %>%
  mutate(cgdd_12c_jan2gr = sum(GDD_12C_SM),
         cgdd_8c_jan2gr = sum(GDD_8C_SM)
         ) %>%
  select(-(GDD_13C_SM:CPTT_13C), -(TMAX:Longitude), GDD_12C_SM) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 17, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_18d = mean(TMAX_SM),
         tmin_18d = mean(TMIN_SM),
         tave_18d = (tmax_18d + tmin_18d)/2,
         cgdd_12c_18d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 9, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_10d = mean(TMAX_SM),
         tmin_10d = mean(TMIN_SM),
         tave_10d = (tmax_10d + tmin_10d)/2,
         cgdd_12c_10d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 4, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_5d = mean(TMAX_SM),
         tmin_5d = mean(TMIN_SM),
         tave_5d = (tmax_5d + tmin_5d)/2,
         cgdd_12c_5d = mean(GDD_12C_SM))%>%
  filter(DOY == as_date(GR50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX_SM:RAIN_SM), -Day_of_year)
saveRDS(asreml_gr, file = "../data/Weather_related_phe_for_asreml_h2_GR50.rds")
```

##

# ----
## load phe & wea data
```{r}
phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
wea_ten_sites <- readRDS(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))

##function to calculate daylength

day.length.hrs <- function(day, latitude) {

  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))

  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))

  return(daylength.hrs)

}

phe_2019 <- jday_phe %>%
  select(-(CGDD_12C:dyln_change_sec)) %>%
  pivot_longer(cols = GR50:FL50, names_to = "PHE", values_to = "JDAY")
```

```{r}

wea_ten_sites <- wea_ten_sites %>%
  mutate(JDAY = yday(Working_date),
         DYLN = day.length.hrs(day = JDAY, latitude = Latitude),
         tmax_sm = ifelse(is.na(tmax),
                          (na.locf(tmax) + na.locf(tmax, fromLast = TRUE))/2,
                          tmax),
         tmin_sm = ifelse(is.na(tmin),
                          (na.locf(tmin) + na.locf(tmin, fromLast = TRUE))/2,
                          tmin),
         tmean_sm = ifelse(is.na(Tmean),
                          (na.locf(Tmean) + na.locf(Tmean, fromLast = TRUE))/2,
                          Tmean),
         GDD_12Cbase = ifelse((tmax_sm + tmin_sm)/2 >= 12.0,
                              (tmax_sm + tmin_sm)/2 - 10,
                              0),
         Year = year(Working_date)) %>%
  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
                              Year == 2019 ~ JDAY + 365,
                              Year == 2020 ~ JDAY + (365*2),
                              Year == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  arrange(Latitude, Year, JDAY_EXP) %>%
  select(-JDAY) %>%
  group_by(manu_site) %>%
  mutate(crain_2d = prcp + lag(prcp, default = 0),
         crain_3d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2),
         crain_4d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + 
           lag(prcp, default = 0, n = 3),
         crain_5d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4),
         crain_6d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4) +
           lag(prcp, default = 0, n = 5),
         crain_7d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4) +
           lag(prcp, default = 0, n = 5) +
           lag(prcp, default = 0, n = 6),
         dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         dyln_change_sec_2d_prior = (lag(DYLN, n = 1) - 
                                       lag(DYLN, n = 2))*60*60,
         dyln_change_sec_3d_prior = (lag(DYLN, n = 2) - 
                                       lag(DYLN, n = 3))*60*60,
         dyln_change_sec_4d_prior = (lag(DYLN, n = 3) - 
                                       lag(DYLN, n = 4))*60*60,
         dyln_change_sec_5d_prior = (lag(DYLN, n = 4) - 
                                       lag(DYLN, n = 5))*60*60,
         dyln_change_sec_6d_prior = (lag(DYLN, n = 5) - 
                                       lag(DYLN, n = 6))*60*60,
         dyln_change_sec_7d_prior = (lag(DYLN, n = 6) - 
                                       lag(DYLN, n = 7))*60*60,
         dyln_change_sec_14d_prior = (lag(DYLN, n = 13) - 
                                       lag(DYLN, n = 14))*60*60,
         dyln_2d_prior = lag(DYLN, n = 1),
         dyln_3d_prior = lag(DYLN, n = 2),
         dyln_4d_prior = lag(DYLN, n = 3),
         dyln_5d_prior = lag(DYLN, n = 4),
         dyln_6d_prior = lag(DYLN, n = 5),
         dyln_7d_prior = lag(DYLN, n = 6),
         dyln_14d_prior = lag(DYLN, n = 13),
         cgdd_12c_2d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0))/2,
         tave_2d = (tmean_sm + lag(tmean_sm, default = 0))/2,
         cgdd_12c_3d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2))/3,
         tave_3d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2))/3,
         cgdd_12c_4d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3))/4,
         tave_4d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3))/4,
         cgdd_12c_5d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4))/5,
         tave_5d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4))/5,
         cgdd_12c_6d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5))/6,
         tave_6d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5))/6,
         cgdd_12c_7d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6))/7,
         tave_7d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6))/7,
         cgdd_12c_14d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13))/14,
         tave_14d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13))/14,
         cgdd_12c_21d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13) +
           lag(GDD_12Cbase, default = 0, n = 14) +
           lag(GDD_12Cbase, default = 0, n = 15) +
           lag(GDD_12Cbase, default = 0, n = 16) +
           lag(GDD_12Cbase, default = 0, n = 17) +
           lag(GDD_12Cbase, default = 0, n = 18) +
           lag(GDD_12Cbase, default = 0, n = 19) +
           lag(GDD_12Cbase, default = 0, n = 20))/21,
         tave_21d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13) +
           lag(tmean_sm, default = 0, n = 14) +
           lag(tmean_sm, default = 0, n = 15) +
           lag(tmean_sm, default = 0, n = 16) +
           lag(tmean_sm, default = 0, n = 17) +
           lag(tmean_sm, default = 0, n = 18) +
           lag(tmean_sm, default = 0, n = 19) +
           lag(tmean_sm, default = 0, n = 20))/21,
         cgdd_12c_28d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13) +
           lag(GDD_12Cbase, default = 0, n = 14) +
           lag(GDD_12Cbase, default = 0, n = 15) +
           lag(GDD_12Cbase, default = 0, n = 16) +
           lag(GDD_12Cbase, default = 0, n = 17) +
           lag(GDD_12Cbase, default = 0, n = 18) +
           lag(GDD_12Cbase, default = 0, n = 19) +
           lag(GDD_12Cbase, default = 0, n = 20) +
           lag(GDD_12Cbase, default = 0, n = 21) +
           lag(GDD_12Cbase, default = 0, n = 22) +
           lag(GDD_12Cbase, default = 0, n = 23) +
           lag(GDD_12Cbase, default = 0, n = 24) +
           lag(GDD_12Cbase, default = 0, n = 25) +
           lag(GDD_12Cbase, default = 0, n = 26) +
           lag(GDD_12Cbase, default = 0, n = 27))/28,
         tave_28d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13) +
           lag(tmean_sm, default = 0, n = 14) +
           lag(tmean_sm, default = 0, n = 15) +
           lag(tmean_sm, default = 0, n = 16) +
           lag(tmean_sm, default = 0, n = 17) +
           lag(tmean_sm, default = 0, n = 18) +
           lag(tmean_sm, default = 0, n = 19) +
           lag(tmean_sm, default = 0, n = 20) +
           lag(tmean_sm, default = 0, n = 21) +
           lag(tmean_sm, default = 0, n = 22) +
           lag(tmean_sm, default = 0, n = 23) +
           lag(tmean_sm, default = 0, n = 24) +
           lag(tmean_sm, default = 0, n = 25) +
           lag(tmean_sm, default = 0, n = 26) +
           lag(tmean_sm, default = 0, n = 27))/28,
         )

```

```{r}
wea_ten_sites
```

# Define phenotypes
```{r}
Tbase <- 12

phe_gr_fl <- phe_2019 %>%
  mutate(JDAY_EXP = case_when(YEAR == 2018 ~ JDAY,
                              YEAR == 2019 ~ JDAY + 365,
                              YEAR == 2020 ~ JDAY + (365*2),
                              YEAR == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  select(-JDAY) %>%
  pivot_wider(names_from = PHE, values_from = JDAY_EXP) %>%
  left_join(select(sites, SITE, manu_site)) %>%
  mutate(Region = case_when(manu_site %in% c("SD", "MO", "IL", "MI", "NE") ~ "North",
                            manu_site %in% c("TX1", "TX2", "TX3", "TX4") ~ "Texas",
                                             TRUE ~ "OK"))

tmp2 <- wea_ten_sites %>%
  select(-Year)

gr_ten_sites <- phe_gr_fl %>%
  ungroup() %>%
  mutate(GR50 = floor(GR50)) %>%
  filter(!is.na(GR50)) %>%
  left_join(tmp2, by = c("manu_site", "GR50" = "JDAY_EXP")) %>%
  select(-SITE, -FL50, -(Working_date:wndk), -Tdrnl, -(Latitude:Elevation), -(tmax_sm:tmean_sm), -DLYN) %>%
  rename(crain_1d = prcp, dyln_gr50 = DYLN, cgdd_12c_1d = GDD_12Cbase, tave_1d = Tmean)

fl_ten_sites <- phe_gr_fl %>%
  ungroup() %>%
  filter(!is.na(FL50)) %>%
  mutate(FL50 = floor(FL50),
         GR50 = floor(GR50)) %>%
  left_join(tmp2, by = c("manu_site", "FL50" = "JDAY_EXP")) %>%
  select(-SITE, -(Working_date:wndk), -Tdrnl, -(Latitude:Elevation), -(tmax_sm:tmean_sm), -DLYN) %>%
  rename(crain_1d = prcp, dyln_gr50 = DYLN, cgdd_12c_1d = GDD_12Cbase, tave_1d = Tmean)
# 11 September flowering dates in 2021 at Pickle have NA's for weather values. Decided to let this be.

```

# Covar csv for greedy mash algorithm

Make & save hypothesis matrices in a hypothesis_matrices directory in data. 
Then hopefully it's not much of a lift for Sam to run the greedy mash algorithm again. Otherwise I will figure out how to run it.

phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
             "tave_5d", "tave_10d", "tave_18d")
phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

```{r}
gr_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_FL50_2019.rds"))
```

### load weather phenotypes
```{r}
gr_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_FL50_2019.rds"))
```

```{r}
phe_hyp <- colnames(fl_ten_sites)[c(5,9:51)]
colnames(gr_ten_sites)[8:50]
```

```{r}
metadata <- read_rds(here("data", "metadata.rds"))
subpop_v2 <- list(all = c("Gulf", "Midwest"), midwest = "Midwest",
                  gulf = "Gulf")
year_v = c(2019, 2020, 2021)

fl_ten_sites <- fl_ten_sites %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  rename(Year = YEAR)

# change based on phenotype ---------------
#phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", 
#             "dyln_change_sec_7d_prior", "dyln_change_sec_14d_prior", 
#             "dyln_7d_prior", "dyln_14d_prior", "cgdd_12c_gr2fl", 
#             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

phe_hyp <- colnames(fl_ten_sites)[c(5,9:51)]
i=1
k=1

U_hyp_2019 <- list()
for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

    cov_df <- fl_ten_sites %>%
      ungroup() %>%
      filter(SUBPOP %in% subpop_v2[[k]]) %>%
      filter(Year %in% 2019) %>%
      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
      group_by(PLANT_ID, SUBPOP, manu_site) %>%
      mutate(IND = row_number()) %>%
      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
      mutate(PLANT_ID = as.factor(PLANT_ID),
             IND = as.factor(IND)) %>%
      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, SD, TX1, TX2, TX3) #####  IL,

    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      summarise(#IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
      as_vector()

    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
    names(U_hyp_2019)[1] <- paste0("flowering.", names(subpop_v2)[k], ".", phe_hyp[i], ".U.mat")
  }
}

U_hyp_2019
length(U_hyp_2019)
for (i in 1:length(U_hyp_2019)) {
  d_out <- as.data.frame(U_hyp_2019[[i]], optional = FALSE) %>%
    rownames_to_column(var = " ")
  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")), )
}
```

```{r}
gr_ten_sites <- gr_ten_sites %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  rename(Year = YEAR)

# change based on phenotype ---------------
#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_20d", "cgdd_12c_5d", 
#             "tave_5d", "tave_10d", "tave_20d")
phe_hyp <- colnames(gr_ten_sites)[8:50]
i=1
k=1

U_hyp_2019 <- list()
for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

    cov_df <- gr_ten_sites %>%
      ungroup() %>%
      filter(SUBPOP %in% subpop_v2[[k]]) %>%
      filter(Year %in% 2019) %>%
      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
      group_by(PLANT_ID, SUBPOP, manu_site) %>%
      mutate(IND = row_number()) %>%
      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
      mutate(PLANT_ID = as.factor(PLANT_ID),
             IND = as.factor(IND)) %>%
      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, SD, TX1, TX2, TX3) #####  IL,

    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      summarise(#IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
      as_vector()

    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
    names(U_hyp_2019)[1] <- paste0("greenup.", names(subpop_v2)[k], ".", phe_hyp[i], ".U.mat")
  }
}
for (i in 1:length(U_hyp_2019)) {
  d_out <- as.data.frame(U_hyp_2019[[i]], optional = FALSE) %>%
    rownames_to_column(var = " ")
  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")), )
}
```

Visualize correlation matrices
```{r}

i=7
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+1]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+2]],
                                            reorder = FALSE) 
names(U_hyp_2019)[[i+2]]
```