### 2023-09-11 R/Analysis_v1.2_mash_using_greedy_covar.Rmd
analysis: re-run mash using only greedy algorithm covariance matrices

---
title: "Analysis v1.2 mash"
author: "Alice MacQueen"
date: "2023-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bigsnpr)
library(mashr)
library(switchgrassGWAS)
library(rlist)
library(cowplot)
library(here)
library(zoo)
```

# Installation reminders for any reinstalls
On debian: `apt install libgsl-dev` to install the GNU Scientific Library, zB.

For mashr: Sometimes install.packages("RcppGSL") is also needed.
#

Sam used a greedy mash algorithm to find the best set of covariance matrices to include, using 19K randomly associated variants.

Now use the best set of covariance matrices on 19K strongly associated variants.

Step 0. Set up.
  0a. Look at phenotype distribution & heritability; run GWAS to get univariate SNP effect estimates & standard errors. Already done.
  0b. Make hypothesis covariance matrices based on weather data from 2018 - 2020. For cumulative weather variables, use a range of days, 1 - 20, say. Come back and add more if 20 is chosen by greedy mash.
Step 1. Run greedy mash to find best set of covariance matrices that explain SNP effect variation across environments, using 19K 'randomly associated variants'. Go back to 0b. if high number of day covar matrices tend to be chosen by greedy mash.
Step 2. Use covariances from (1) to run mash on 19K 'strongly associated variants'. 
3. Look at loadings on covariance matrices and meanings of covariance matrices as the writeup (Figure 2, 3; Table 1). Look at amount of antagonistic pleiotropy vs same sign significant effects.
4. Repeat the QTL analysis... with Li.

How much of a lift is it to re-run the greedy mash algorithm? If we, or if reviewers are thinking about this in a Jiaming Yu sense, I think the next logical question to ask is 'Why these covariance matrices'? The overly honest answer for the current set is that there were computational limits to running mash with so many covariance matrices, so I chose a few arbitrary time points for each weather variable we thought might be a environmental driver. But now we have introduced a method to choose between highly correlated covariance matrices, perhaps we could do better. 
2.

# Deprecated analysis
### Code from Plots_v0.2_manuscript_figures.Rmd to make weather variables

Actually, code from Analysis_v1.1_weather_2020.Rmd might be more suitable, if I am remaking these anyway.

```{r}
wea_df <- readRDS("~/Github/pvdiv-phenology-gxe/data/Weather_with_PTT_2019.rds")
metadata <- readRDS("~/Github/pvdiv-phenology-gxe/data/metadata.rds")
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
phenotypes <- readRDS("~/Github/pvdiv-phenology-gxe/data/Phenology_CGDD_PTT_and_rainfall_phenotypes.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
```

```{r}
jday_phe <- wea_df %>%
  group_by(SITE) %>%
  mutate(dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         FL50 = yday(DOY)) %>%
  select(DOY, DYLN, dyln_change_sec, everything()) %>%
  select(-PTT_12C) %>%
  right_join(jday_phe, by = c("SITE", "FL50")) %>%
  select(-(TMAX:CPTT_11C))
jday_phe <- metadata %>%
  rename(Lat_origin = LATITUDE, Long_origin = LONGITUDE) %>%
  select(PLANT_ID, SUBPOP, Lat_origin, Long_origin) %>%
  right_join(jday_phe) %>%
  left_join(sites)

jday_phe$SUBPOP <- factor(jday_phe$SUBPOP, levels = c("4X", "Atlantic", "Gulf", "Midwest", "8X"))
jday_phe$SITE <- factor(jday_phe$SITE, levels = c("KING", "PKLE", "TMPL", "OVTN", "STIL", "CLMB", "LINC", "FRMI", "KBSM", "BRKG"))
jday_phe$manu_site <- factor(jday_phe$manu_site, levels = rev(c("TX1", "TX2", "TX3", "TX4", "OK", "MO", "NE", "IL", "MI", "SD")))
```

## define environmental cues for greenup
Probably all CGDD related.
```{r}

asreml_fl <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, FL50, GR50, CGDD_12C, CRAIN, CRAIN_5d, DYLN, dyln_change_sec) %>%
  rename(dyln_fl50 = DYLN, cgdd_12c_gr2fl = CGDD_12C, crain_gr2fl = CRAIN) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(between(DOY, as_date(FL50, origin = "2019-01-01") - 6, as_date(FL50, origin = "2019-01-01"))) %>%
  mutate(crain_2d = RAIN_SM + lag(RAIN_SM),
         crain_3d = RAIN_SM + lag(RAIN_SM) + lag(RAIN_SM, n = 2),
         crain_7d = sum(RAIN_SM)) %>%
  filter(DOY == as_date(FL50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX:TMIN_SM), -(GDD_13C_SM:CPTT_13C)) %>%
  rename(crain_1d = RAIN_SM, crain_5d = CRAIN_5d)
saveRDS(asreml_fl, file = "../data/Weather_related_phe_for_asreml_h2_FL50.rds")
asreml_gr <- jday_phe %>%
  select(PLANT_ID:Long_origin, PLOT_GL, SITE, GR50) %>%
  left_join(wea_df) %>%
  group_by(SITE, PLOT_GL) %>%
  filter(DOY <= as_date(GR50, origin = "2019-01-01")) %>%
  mutate(cgdd_12c_jan2gr = sum(GDD_12C_SM),
         cgdd_8c_jan2gr = sum(GDD_8C_SM)
         ) %>%
  select(-(GDD_13C_SM:CPTT_13C), -(TMAX:Longitude), GDD_12C_SM) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 17, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_18d = mean(TMAX_SM),
         tmin_18d = mean(TMIN_SM),
         tave_18d = (tmax_18d + tmin_18d)/2,
         cgdd_12c_18d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 9, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_10d = mean(TMAX_SM),
         tmin_10d = mean(TMIN_SM),
         tave_10d = (tmax_10d + tmin_10d)/2,
         cgdd_12c_10d = mean(GDD_12C_SM)) %>%
  filter(between(DOY, as_date(GR50, origin = "2019-01-01") - 4, as_date(GR50, origin = "2019-01-01"))) %>%
  mutate(tmax_5d = mean(TMAX_SM),
         tmin_5d = mean(TMIN_SM),
         tave_5d = (tmax_5d + tmin_5d)/2,
         cgdd_12c_5d = mean(GDD_12C_SM))%>%
  filter(DOY == as_date(GR50, origin = "2019-01-01")) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf", "Atlantic") & !(SITE %in% c("OVTN"))) %>%
  select(-(TMAX_SM:RAIN_SM), -Day_of_year)
saveRDS(asreml_gr, file = "../data/Weather_related_phe_for_asreml_h2_GR50.rds")
```

##

# ----
## load phe & wea data
```{r}
phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
wea_ten_sites <- readRDS(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))

##function to calculate daylength

day.length.hrs <- function(day, latitude) {

  daylength.coeff <- asin(0.39795*cos(0.2163108 + 2*atan(0.9671396*tan(0.00860*(day - 186)))))

  daylength.hrs <- 24 - (24/pi)*acos((sin(0.8333*pi/180) + sin(latitude*pi/180)*sin(daylength.coeff))/(cos(latitude*pi/180)*cos(daylength.coeff)))

  return(daylength.hrs)

}

phe_2019 <- jday_phe %>%
  select(-(CGDD_12C:dyln_change_sec)) %>%
  pivot_longer(cols = GR50:FL50, names_to = "PHE", values_to = "JDAY")
```

```{r}

wea_ten_sites <- wea_ten_sites %>%
  mutate(JDAY = yday(Working_date),
         DYLN = day.length.hrs(day = JDAY, latitude = Latitude),
         tmax_sm = ifelse(is.na(tmax),
                          (na.locf(tmax) + na.locf(tmax, fromLast = TRUE))/2,
                          tmax),
         tmin_sm = ifelse(is.na(tmin),
                          (na.locf(tmin) + na.locf(tmin, fromLast = TRUE))/2,
                          tmin),
         tmean_sm = ifelse(is.na(Tmean),
                          (na.locf(Tmean) + na.locf(Tmean, fromLast = TRUE))/2,
                          Tmean),
         GDD_12Cbase = ifelse((tmax_sm + tmin_sm)/2 >= 12.0,
                              (tmax_sm + tmin_sm)/2 - 10,
                              0),
         Year = year(Working_date)) %>%
  mutate(JDAY_EXP = case_when(Year == 2018 ~ JDAY,
                              Year == 2019 ~ JDAY + 365,
                              Year == 2020 ~ JDAY + (365*2),
                              Year == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  arrange(Latitude, Year, JDAY_EXP) %>%
  select(-JDAY) %>%
  group_by(manu_site) %>%
  mutate(crain_2d = prcp + lag(prcp, default = 0),
         crain_3d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2),
         crain_4d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + 
           lag(prcp, default = 0, n = 3),
         crain_5d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4),
         crain_6d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4) +
           lag(prcp, default = 0, n = 5),
         crain_7d = prcp + lag(prcp, default = 0) + 
           lag(prcp, default = 0, n = 2) + lag(prcp, default = 0, n = 3) +
           lag(prcp, default = 0, n = 4) +
           lag(prcp, default = 0, n = 5) +
           lag(prcp, default = 0, n = 6),
         dyln_change_sec = (DYLN - lag(DYLN))*60*60,
         dyln_change_sec_2d_prior = (lag(DYLN, n = 1) - 
                                       lag(DYLN, n = 2))*60*60,
         dyln_change_sec_3d_prior = (lag(DYLN, n = 2) - 
                                       lag(DYLN, n = 3))*60*60,
         dyln_change_sec_4d_prior = (lag(DYLN, n = 3) - 
                                       lag(DYLN, n = 4))*60*60,
         dyln_change_sec_5d_prior = (lag(DYLN, n = 4) - 
                                       lag(DYLN, n = 5))*60*60,
         dyln_change_sec_6d_prior = (lag(DYLN, n = 5) - 
                                       lag(DYLN, n = 6))*60*60,
         dyln_change_sec_7d_prior = (lag(DYLN, n = 6) - 
                                       lag(DYLN, n = 7))*60*60,
         dyln_change_sec_14d_prior = (lag(DYLN, n = 13) - 
                                       lag(DYLN, n = 14))*60*60,
         dyln_2d_prior = lag(DYLN, n = 1),
         dyln_3d_prior = lag(DYLN, n = 2),
         dyln_4d_prior = lag(DYLN, n = 3),
         dyln_5d_prior = lag(DYLN, n = 4),
         dyln_6d_prior = lag(DYLN, n = 5),
         dyln_7d_prior = lag(DYLN, n = 6),
         dyln_14d_prior = lag(DYLN, n = 13),
         cgdd_12c_2d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0))/2,
         tave_2d = (tmean_sm + lag(tmean_sm, default = 0))/2,
         cgdd_12c_3d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2))/3,
         tave_3d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2))/3,
         cgdd_12c_4d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3))/4,
         tave_4d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3))/4,
         cgdd_12c_5d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4))/5,
         tave_5d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4))/5,
         cgdd_12c_6d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5))/6,
         tave_6d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5))/6,
         cgdd_12c_7d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6))/7,
         tave_7d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6))/7,
         cgdd_12c_14d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13))/14,
         tave_14d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13))/14,
         cgdd_12c_21d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13) +
           lag(GDD_12Cbase, default = 0, n = 14) +
           lag(GDD_12Cbase, default = 0, n = 15) +
           lag(GDD_12Cbase, default = 0, n = 16) +
           lag(GDD_12Cbase, default = 0, n = 17) +
           lag(GDD_12Cbase, default = 0, n = 18) +
           lag(GDD_12Cbase, default = 0, n = 19) +
           lag(GDD_12Cbase, default = 0, n = 20))/21,
         tave_21d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13) +
           lag(tmean_sm, default = 0, n = 14) +
           lag(tmean_sm, default = 0, n = 15) +
           lag(tmean_sm, default = 0, n = 16) +
           lag(tmean_sm, default = 0, n = 17) +
           lag(tmean_sm, default = 0, n = 18) +
           lag(tmean_sm, default = 0, n = 19) +
           lag(tmean_sm, default = 0, n = 20))/21,
         cgdd_12c_28d = (GDD_12Cbase + lag(GDD_12Cbase, default = 0) + 
           lag(GDD_12Cbase, default = 0, n = 2) + 
           lag(GDD_12Cbase, default = 0, n = 3) +
           lag(GDD_12Cbase, default = 0, n = 4) +
           lag(GDD_12Cbase, default = 0, n = 5) +
           lag(GDD_12Cbase, default = 0, n = 6) +
           lag(GDD_12Cbase, default = 0, n = 7) +
           lag(GDD_12Cbase, default = 0, n = 8) +
           lag(GDD_12Cbase, default = 0, n = 9) +
           lag(GDD_12Cbase, default = 0, n = 10) +
           lag(GDD_12Cbase, default = 0, n = 11) +
           lag(GDD_12Cbase, default = 0, n = 12) +
           lag(GDD_12Cbase, default = 0, n = 13) +
           lag(GDD_12Cbase, default = 0, n = 14) +
           lag(GDD_12Cbase, default = 0, n = 15) +
           lag(GDD_12Cbase, default = 0, n = 16) +
           lag(GDD_12Cbase, default = 0, n = 17) +
           lag(GDD_12Cbase, default = 0, n = 18) +
           lag(GDD_12Cbase, default = 0, n = 19) +
           lag(GDD_12Cbase, default = 0, n = 20) +
           lag(GDD_12Cbase, default = 0, n = 21) +
           lag(GDD_12Cbase, default = 0, n = 22) +
           lag(GDD_12Cbase, default = 0, n = 23) +
           lag(GDD_12Cbase, default = 0, n = 24) +
           lag(GDD_12Cbase, default = 0, n = 25) +
           lag(GDD_12Cbase, default = 0, n = 26) +
           lag(GDD_12Cbase, default = 0, n = 27))/28,
         tave_28d = (tmean_sm + lag(tmean_sm, default = 0) + 
           lag(tmean_sm, default = 0, n = 2) + 
           lag(tmean_sm, default = 0, n = 3) +
           lag(tmean_sm, default = 0, n = 4) +
           lag(tmean_sm, default = 0, n = 5) +
           lag(tmean_sm, default = 0, n = 6) +
           lag(tmean_sm, default = 0, n = 7) +
           lag(tmean_sm, default = 0, n = 8) +
           lag(tmean_sm, default = 0, n = 9) +
           lag(tmean_sm, default = 0, n = 10) +
           lag(tmean_sm, default = 0, n = 11) +
           lag(tmean_sm, default = 0, n = 12) +
           lag(tmean_sm, default = 0, n = 13) +
           lag(tmean_sm, default = 0, n = 14) +
           lag(tmean_sm, default = 0, n = 15) +
           lag(tmean_sm, default = 0, n = 16) +
           lag(tmean_sm, default = 0, n = 17) +
           lag(tmean_sm, default = 0, n = 18) +
           lag(tmean_sm, default = 0, n = 19) +
           lag(tmean_sm, default = 0, n = 20) +
           lag(tmean_sm, default = 0, n = 21) +
           lag(tmean_sm, default = 0, n = 22) +
           lag(tmean_sm, default = 0, n = 23) +
           lag(tmean_sm, default = 0, n = 24) +
           lag(tmean_sm, default = 0, n = 25) +
           lag(tmean_sm, default = 0, n = 26) +
           lag(tmean_sm, default = 0, n = 27))/28,
         )

```

```{r}
wea_ten_sites
```

# Define phenotypes
```{r}
Tbase <- 12

phe_gr_fl <- phe_2019 %>%
  mutate(JDAY_EXP = case_when(YEAR == 2018 ~ JDAY,
                              YEAR == 2019 ~ JDAY + 365,
                              YEAR == 2020 ~ JDAY + (365*2),
                              YEAR == 2021 ~ JDAY + (365*3),
                              TRUE ~ NA_real_)) %>%
  select(-JDAY) %>%
  pivot_wider(names_from = PHE, values_from = JDAY_EXP) %>%
  left_join(select(sites, SITE, manu_site)) %>%
  mutate(Region = case_when(manu_site %in% c("SD", "MO", "IL", "MI", "NE") ~ "North",
                            manu_site %in% c("TX1", "TX2", "TX3", "TX4") ~ "Texas",
                                             TRUE ~ "OK"))

tmp2 <- wea_ten_sites %>%
  select(-Year)

gr_ten_sites <- phe_gr_fl %>%
  ungroup() %>%
  mutate(GR50 = floor(GR50)) %>%
  filter(!is.na(GR50)) %>%
  left_join(tmp2, by = c("manu_site", "GR50" = "JDAY_EXP")) %>%
  select(-SITE, -FL50, -(Working_date:wndk), -Tdrnl, -(Latitude:Elevation), -(tmax_sm:tmean_sm), -DLYN) %>%
  rename(crain_1d = prcp, dyln_gr50 = DYLN, cgdd_12c_1d = GDD_12Cbase, tave_1d = Tmean)

fl_ten_sites <- phe_gr_fl %>%
  ungroup() %>%
  filter(!is.na(FL50)) %>%
  mutate(FL50 = floor(FL50),
         GR50 = floor(GR50)) %>%
  left_join(tmp2, by = c("manu_site", "FL50" = "JDAY_EXP")) %>%
  select(-SITE, -(Working_date:wndk), -Tdrnl, -(Latitude:Elevation), -(tmax_sm:tmean_sm), -DLYN) %>%
  rename(crain_1d = prcp, dyln_gr50 = DYLN, cgdd_12c_1d = GDD_12Cbase, tave_1d = Tmean)
# 11 September flowering dates in 2021 at Pickle have NA's for weather values. Decided to let this be.

```

# Covar csv for greedy mash algorithm

Make & save hypothesis matrices in a hypothesis_matrices directory in data. 
Then hopefully it's not much of a lift for Sam to run the greedy mash algorithm again. Otherwise I will figure out how to run it.

phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
             "tave_5d", "tave_10d", "tave_18d")
phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

```{r}
gr_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_FL50_2019.rds"))
```

### load weather phenotypes
```{r}
gr_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_FL50_2019.rds"))
```

```{r}
phe_hyp <- colnames(fl_ten_sites)[c(5,9:51)]
colnames(gr_ten_sites)[8:50]
```

```{r}
metadata <- read_rds(here("data", "metadata.rds"))
subpop_v2 <- list(all = c("Gulf", "Midwest"), midwest = "Midwest",
                  gulf = "Gulf")
year_v = c(2019, 2020, 2021)

fl_ten_sites <- fl_ten_sites %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  rename(Year = YEAR)

# change based on phenotype ---------------
#phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", 
#             "dyln_change_sec_7d_prior", "dyln_change_sec_14d_prior", 
#             "dyln_7d_prior", "dyln_14d_prior", "cgdd_12c_gr2fl", 
#             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

phe_hyp <- colnames(fl_ten_sites)[c(5,9:51)]
i=1
k=1

U_hyp_2019 <- list()
for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

    cov_df <- fl_ten_sites %>%
      ungroup() %>%
      filter(SUBPOP %in% subpop_v2[[k]]) %>%
      filter(Year %in% 2019) %>%
      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
      group_by(PLANT_ID, SUBPOP, manu_site) %>%
      mutate(IND = row_number()) %>%
      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
      mutate(PLANT_ID = as.factor(PLANT_ID),
             IND = as.factor(IND)) %>%
      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3) #####  IL,

    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      summarise(#IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
                OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
      as_vector()

    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
    names(U_hyp_2019)[1] <- paste0("flowering.", names(subpop_v2)[k], ".", phe_hyp[i], ".U.mat")
  }
}

U_hyp_2019
length(U_hyp_2019)
for (i in 1:length(U_hyp_2019)) {
  d_out <- as.data.frame(U_hyp_2019[[i]], optional = FALSE) %>%
    rownames_to_column(var = " ")
  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")))
}
```

```{r}
gr_ten_sites <- gr_ten_sites %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  rename(Year = YEAR)

# change based on phenotype ---------------
#phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_20d", "cgdd_12c_5d", 
#             "tave_5d", "tave_10d", "tave_20d")
phe_hyp <- colnames(gr_ten_sites)[8:50]
i=1
k=1

U_hyp_2019 <- list()
for(i in seq_along(phe_hyp)){
  for(k in seq_along(subpop_v2)){

    cov_df <- gr_ten_sites %>%
      ungroup() %>%
      filter(SUBPOP %in% subpop_v2[[k]]) %>%
      filter(Year %in% 2019) %>%
      select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
      group_by(PLANT_ID, SUBPOP, manu_site) %>%
      mutate(IND = row_number()) %>%
      pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
      mutate(PLANT_ID = as.factor(PLANT_ID),
             IND = as.factor(IND)) %>%
      select(PLANT_ID, SUBPOP, IND, MI, MO, NE, OK, SD, TX1, TX2, TX3) #####  IL,

    cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      summarise(#IL = sqrt(var(IL, na.rm = TRUE))/mean(IL, na.rm = TRUE),  #####
                MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),  #####
                MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
                NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
                OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
                SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE),
                TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
                TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
                TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE)) %>%
      as_vector()

    U_hyp_2019 <- list.prepend(U_hyp_2019, cor_phe)
    names(U_hyp_2019)[1] <- paste0("greenup.", names(subpop_v2)[k], ".", phe_hyp[i], ".U.mat")
  }
}
for (i in 1:length(U_hyp_2019)) {
  d_out <- as.data.frame(U_hyp_2019[[i]], optional = FALSE) %>%
    rownames_to_column(var = " ")
  write_csv(d_out, file = here("data", "greedy_more_cov", "hypothesis_matrices", paste0(names(U_hyp_2019)[i], ".csv")), )
}
```

Visualize correlation matrices
```{r}

i=7
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+1]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+2]],
                                            reorder = FALSE) 
names(U_hyp_2019)[[i+2]]
```

# -----------------------------
# Run mash

```{r}
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
library(rlist)
library(here)

source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
mashdir <- here("analysis", "mash_greedy_algorithm")

#phe_gr_df <- read_rds(here("data",
#                           "Weather_related_phe_for_asreml_h2_GR50.rds"))
# all greenup related phenotypes, including greenup as functions of weather variables.
#phe_fl_df <- read_rds(here("data",
#                           "Weather_related_phe_for_asreml_h2_FL50.rds"))
# all flowering related phenotypes, including functions of weather variables.

# Files with the six suffixes I have/want mash results for:
numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
              "GR50_Gulf", "GR50_Midwest", "FL50_Midwest")
suffix_vli <- c("Gulf_and_Midwest", "Gulf", "Gulf_and_Midwest",
              "Gulf", "Midwest", "Midwest")
subpop_v <- list(Gulf_and_Midwest = "363g_12.1M", Midwest = "134g_8.8M", 
                 Gulf = "229g_10.3M")

pop_sc <- c("all", "gulf", "all", "gulf", "midwest", "midwest")
phe_sc <- c("flowering", "flowering", "greenup", "greenup", "greenup",
            "flowering")

Upath <- here("analysis", "mash_greedy_algorithm", "mash_switchgrass_models", 
                "mash", "all_matrices", "matrices")
stopcondpath <- here("analysis", "mash_greedy_algorithm",
                     "mash_switchgrass_models", "mash", 
                     "all_matrices", "likelihood_paths")

for (i in 1:6) {  # change this for different scripts
  # B_hat_strong_df_19000topSNPs_FL50_Gulf_and_Midwest.rds
  # B_hat_strong_df_19000topSNPs_FL50_Gulf.rds
  # B_hat_strong_df_19000topSNPs_GR50_Gulf_and_Midwest.rds
  # B_hat_strong_df_19000topSNPs_GR50_Gulf.rds
  # B_hat_strong_df_24000topSNPs_GR50_Midwest.rds
  # B_hat_strong_df_33000topSNPs_FL50_Midwest.rds
  ## mash Gulf greenup
  numSNPs <- numSNPs_v[i]
  suffix <- suffix_v[i]
  # use this one for the likelihood paths
  # Read in likelihood paths, then use that to read in matrices, then use those to run mash.

  U_greedy <- 
    read_csv(file = file.path(stopcondpath, 
                              paste0(phe_sc[i], ".", pop_sc[i], 
                                     ".random.stop.condition.csv"))) %>%
    filter(matrix != "0.0")
  ## Read in Uhyp

  U_list <- list()
  for (k in 1:nrow(U_greedy)) {
    U_single <- read_csv(file = file.path(Upath, U_greedy$matrix[k]), show_col_types = FALSE)
    # if conditions for Midwest
    if (i == 5) {
    # MI, MO, NE, OK, TX1, TX2, TX3 used for GR50  
      U_single <- U_single %>%
        filter(`...1` %in% c("MI", "MO", "NE", "OK", "TX1", "TX2", "TX3")) %>%
        select(`...1`, .data$MI, .data$MO, .data$NE, .data$OK, .data$TX1,
               .data$TX2, .data$TX3)
    } else if (i == 6) {
    # MI, MO, SD, TX1, TX2, TX3 used for FL50  
      U_single <- U_single %>%
        filter(`...1` %in% c("MI", "MO", "SD", "TX1", "TX2", "TX3")) %>%
        select(`...1`, .data$MI, .data$MO, .data$SD, .data$TX1, .data$TX2,
               .data$TX3)
    }
    U_single <- U_single %>%
      select(-`...1`)
    U_single <- as.matrix(unname(U_single))

    # format cov matrices correctly for mash
    U_list <- list.append(U_list, U_single)
  }
  names(U_list) <- str_remove(U_greedy$matrix, ".csv")

  lipath <- here("analysis", "mash", suffix_vli[i], "inputs")

  list_input <- switchgrassGWAS:::load_mash_df(path = lipath, 
                                               numSNPs = numSNPs,
                                               suffix = suffix)
  # Just run mash itself as the greedy choice of cov matrices changes mash run
  # enough that mash_standard_run will not perform correctly (eg it always adds
  # all canonical covariance matrices).
  Bhat_strong <- as.matrix(list_input$B_hat_strong)
  Shat_strong <- as.matrix(list_input$S_hat_strong)
  Bhat_random <- as.matrix(list_input$B_hat_random)
  Shat_random <- as.matrix(list_input$S_hat_random)

  data_r <- mashr::mash_set_data(Bhat_random, Shat_random)
  Vhat <- mashr::estimate_null_correlation_simple(data = data_r)

  data_strong <- mashr::mash_set_data(Bhat_strong, Shat_strong, V = Vhat)
  data_random <- mashr::mash_set_data(Bhat_random, Shat_random, V = Vhat)

  # Run mash on the random dataset using the random data w/ correlation structure
  message(paste0("Fit mash to the random tests using both data-driven and ",
                 "canonical covariances."))
  # U_c <- mashr::cov_canonical(data_random)
  m = mashr::mash(data_random, Ulist = U_list, outputlevel = 1)

  # Run mash on the strong dataset (or all data) using
  # the previous results from the random data
  message(paste0("Compute posterior matrices for the strong effects",
                 " using the mash fit from the
                   random tests."))
  m2 = mashr::mash(data_strong, g = get_fitted_g(m), fixg = TRUE)

  saveRDS(m2, file = file.path(mashdir, suffix_vli[i], 
                                   paste0("Strong_mash_posterior_",
                                          "est_w_random_greedy_Ulist", 
                                          numSNPs, "_SNPs_",
                                          suffix, ".rds")))

  print(paste0(suffix, " mash outputs: "))
  print("Log likelihood with specified covariance matrices: ")
  print(get_loglik(m2), digits = 10)
  print("How many significant markers?")
  print(length(get_significant_results(m2)))
  switchgrassGWAS::mash_plot_covar(m2)
  switchgrassGWAS::mash_plot_manhattan_by_condition(m2)
  switchgrassGWAS::mash_plot_pairwise_sharing(m2)
  gxe <- switchgrassGWAS::get_GxE(m2)
  saveRDS(gxe, file = file.path(mashdir, suffix_vli[i], 
                                   paste0("GxE_default_",
                                          "est_w_random_greedy_Ulist",
                                          numSNPs, "_SNPs_",
                                          suffix, "_greedy_Ulist.rds")))
}
```