### 2024-05-17 R/SI_Appendix_Datasets.Rmd
revision2: Add new SI Datasets 1-6 to the SI Appendix

---
title: "SI_Appendix_Datasets"
author: "Alice MacQueen"
date: 2021-06-24
output: html_document
---

Edited 2024-05-17 for revision2.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(cowplot)
library(switchgrassGWAS)
library(ashr)
```

#

Make all these datasets:

    Dataset 1. SNP-associated effects and standard errors for green-up estimated using the Gulf genetic subpopulation.
    Dataset 2. SNP-associated effects and standard errors for green-up estimated using the Midwest genetic subpopulation.
    Dataset 3. SNP-associated effects and standard errors for green-up estimated using Both genetic subpopulations.
    Dataset 4. SNP-associated effects and standard errors for flowering estimated using the Gulf genetic subpopulation.
    Dataset 5. SNP-associated effects and standard errors for flowering estimated using the Midwest genetic subpopulation.
    Dataset 6. SNP-associated effects and standard errors for flowering estimated using Both genetic subpopulations.
    Dataset 7. Genes within QTL that have functionally validated roles in flowering in rice.
    Dataset 8. Overlap between quantitative trait loci for flowering measured using an outbred pseudo-F2 mapping population, and significant mash effect estimates (with a log10Bayes Factor > 2, or in the 1% tail of most significant mash results) found using three overlapping diversity panels. Genes within 20kb with a functionally validated role in flowering in rice are  indicated.

# Dataset 1-6: mash objects

```{r}
mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")

numSNPs_v <- c(19000, 19000, 19000, 19000, 24000, 33000)
suffix_v <- c("FL50_Gulf_and_Midwest", "FL50_Gulf", "GR50_Gulf_and_Midwest",
              "GR50_Gulf", "GR50_Midwest", "FL50_Midwest")
```

```{r}
m_gulf_gr_hyp <- readRDS(file.path(mashdir,  
                                   paste0("Strong_mash_posterior_",
                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
                                          numSNPs_v[4], "_SNPs_",
                                          suffix_v[4], ".rds")))
m_gulf_fl_hyp <- readRDS(file.path(mashdir,  
                                   paste0("Strong_mash_posterior_",
                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
                                          numSNPs_v[2], "_SNPs_",
                                          suffix_v[2], ".rds")))

m_midw_gr_hyp <- readRDS(file.path(mashdir,  
                                   paste0("Strong_mash_posterior_",
                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
                                          numSNPs_v[5], "_SNPs_",
                                          suffix_v[5], ".rds")))
m_midw_fl_hyp <- readRDS(file.path(mashdir,  
                                   paste0("Strong_mash_posterior_",
                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
                                          numSNPs_v[6], "_SNPs_",
                                          suffix_v[6], ".rds")))

m_gam_gr_hyp <- readRDS(file.path(mashdir,  
                                   paste0("Strong_mash_posterior_",
                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
                                          numSNPs_v[3], "_SNPs_",
                                          suffix_v[3], ".rds")))
m_gam_fl_hyp <- readRDS(file.path(mashdir,  
                                   paste0("Strong_mash_posterior_",
                                          "est_w_ahm_random_greedy_h2_diag_Ulist", 
                                          numSNPs_v[1], "_SNPs_",
                                          suffix_v[1], ".rds")))
```

```{r}
m <- m_gulf_gr_hyp
dataset_creation_function <- function(m, filename){

  all_effects <- switchgrassGWAS:::get_marker_df(m)
  BF <- switchgrassGWAS:::get_log10bf(m) |> 
    as.vector()
  all_effects$log10BF <- BF

  Means <- get_pm(m)
  StdErr <- get_psd(m)
  lfsr <- get_lfsr(m)
  # increase clarity of column names for Supplement data
  colnames(StdErr) <- colnames(StdErr) %>%
    str_replace(., "Stand_Bhat", "Effect_StandardError_") %>%
    str_replace(., "Gulf_and_Midwest", "Both") %>%
    str_replace(., "GR50", "GreenupDate") %>%
    str_replace(., "FL50", "FloweringDate")
  colnames(Means) <- colnames(Means) %>%
    str_replace(., "Stand_Bhat", "Effect_Mean_") %>%
    str_replace(., "Gulf_and_Midwest", "Both") %>%
    str_replace(., "GR50", "GreenupDate") %>%
    str_replace(., "FL50", "FloweringDate")
  colnames(lfsr) <- colnames(lfsr) %>%
    str_replace(., "Stand_Bhat", "Effect_lfsr_") %>%
    str_replace(., "Gulf_and_Midwest", "Both") %>%
    str_replace(., "GR50", "GreenupDate") %>%
    str_replace(., "FL50", "FloweringDate")

  all_effects <- cbind(all_effects, Means, StdErr, lfsr) %>%
    select(starts_with("Marker"), log10BF, ends_with("TX1"), ends_with("TX2"),
           ends_with("TX3"), ends_with("OK"), ends_with("MO"), ends_with("NE"),
           ends_with("MI"), ends_with("SD"))

  all_effects |> 
    write_csv(filename)

}

dataset_creation_function(m_gulf_gr_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_1_Gulf_vegetative_growth_date.csv"))
dataset_creation_function(m_gulf_fl_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_4_Gulf_reproductive_growth_date.csv"))

dataset_creation_function(m_gam_gr_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_3_Both_subpopulations_vegetative_growth_date.csv"))
dataset_creation_function(m_gam_fl_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_6_Both_subpopulations_reproductive_growth_date.csv"))

dataset_creation_function(m_midw_fl_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_5_Midwest_reproductive_growth_date.csv"))
dataset_creation_function(m_midw_gr_hyp, filename = here("manuscript", "revision2", "SI_Appendix", "Dataset_2_Midwest_vegetative_growth_date.csv"))
```

# Dataset 7

# Dataset 8

## QTL
```{r}
qtldf <- read_csv("~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv")

qtldf <- qtldf %>%
  separate(flank_lo, into = c("CHR", "Pos_Mb_lo"), sep = "_", convert = TRUE) %>%
  separate(flank_hi, into = c("flank", "Pos_Mb_hi"), sep = "_", convert = TRUE) %>%
  mutate(Pos_Mb_lo = round(Pos_Mb_lo, digits = 2),
         Pos_Mb_hi = round(Pos_Mb_hi, digits = 2)) %>%
  select(CHR, Pos_Mb_lo, Pos_Mb_hi, lodcolumn, lod, everything()) %>%
  arrange(CHR, Pos_Mb_lo)

qtlplot <- qtldf %>%
  mutate(phe = case_when(lodcolumn == "FL50" ~ "50% flowering",
                              lodcolumn == "dyln_fl50" ~ "Daylength",
                              lodcolumn == "dyln_change_sec" ~ "daylength change (s)",
                              lodcolumn == "CGDD_12C" ~ "GDD, GR to FL",
                              lodcolumn == "crain_gr2fl" ~ "Rainfall, GR to FL")) %>%
  #filter(phe %in% c("Daylength", "GDD, GR to FL")) %>%
  mutate(POS_lo = Pos_Mb_lo * 1000000,
         POS_hi = Pos_Mb_hi * 1000000)

qtlplot$phe <- factor(qtlplot$phe, 
                      levels = c("50% flowering", "Rainfall, GR to FL",
                                 "GDD, GR to FL", "Daylength",
                                          "daylength change (s)",
                                          "Rainfall, 1 day",
                                                  #"Rainfall, 2d sum",
                                                  "Rainfall, 3d sum",
                                                  "Rainfall, 5d sum"#,
                                                  #"Rainfall, 7d sum"
                                                  ))
```

```{r}
m <- m_gulf_fl_hyp
qtl_df <- qtlplot
subpop = "Gulf"
phe = "flowering"

qtl_mash_overlap_function <- function(m, subpop, phe){

  mash_df <- switchgrassGWAS:::get_marker_df(m) %>%
    mutate(log10BF = switchgrassGWAS:::get_log10bf(m)) %>%
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, 
                    sep = "_", extra = "merge", convert = TRUE) %>%
    dplyr::arrange(.data$Chr, .data$Pos) %>%
    select(-value)

  for(i in 1:nrow(qtl_df)){  # from 1 to 16
    ## In block & > 2
    qtl_sub <- mash_df %>%
      filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
                                               qtl_df$POS_hi[i])) %>% 
      filter(log10BF > 2)

    if(i == 1){
    outputdf <- tibble(Chr = qtl_df$CHR[i], QTL_Pos_lo = qtl_df$POS_lo[i], 
                       QTL_Pos_hi = qtl_df$POS_hi[i], 
                       QTL_phenotype = qtl_df$lodcolumn[i], 
                       QTL_LOD = qtl_df$lod[i], mash_Pos = qtl_sub$Pos, 
                       mash_log10BF = qtl_sub$log10BF, mash_subpop = subpop, 
                       mash_phenotype = phe)
    } else {
      outputdf1 <- tibble(Chr = qtl_df$CHR[i], QTL_Pos_lo = qtl_df$POS_lo[i], 
                          QTL_Pos_hi = qtl_df$POS_hi[i], 
                          QTL_phenotype = qtl_df$lodcolumn[i], 
                          QTL_LOD = qtl_df$lod[i], mash_Pos = qtl_sub$Pos, 
                          mash_log10BF = qtl_sub$log10BF, mash_subpop = subpop, 
                          mash_phenotype = phe)
      outputdf <- outputdf %>%
        full_join(outputdf1)
    }
  }
  return(outputdf)
}

overlap1 <- qtl_mash_overlap_function(m_gulf_fl_hyp, subpop = "Gulf", 
                                      phe = "flowering") 
overlap2 <- qtl_mash_overlap_function(m_midw_fl_hyp, subpop = "Midwest", 
                                      phe = "flowering") 
overlap3 <- qtl_mash_overlap_function(m_gam_fl_hyp, subpop = "Both", 
                                      phe = "flowering")

dataset_8 <- overlap1 %>%
  full_join(overlap2) %>%
  full_join(overlap3) %>%
  arrange(Chr, mash_Pos)

write_csv(dataset_8, file = "~/Github/pvdiv-phenology-gxe/manuscript/Coauthors/PNAS/Dataset_8_QTL_mash_overlap.csv")
```

Is there an enrichment of SNPs in the top 1% (5%? 10%?) of the distribution for each mash result in the QTL regions?
1% is probably best out of these three arbitrary cutoffs.
```{r}
m <- m_gulf_fl_hyp
qtl_df <- qtlplot

get_QTL_enrichment <- function(m, qtl_df, quantile = 0.95){
  pw_df_all <- get_post_weights(m)

  BFthresh <- quantile(pw_df_all$log10BayesFactor, quantile)
  pw_df <- pw_df_all %>%
  filter(log10BayesFactor > BFthresh) %>%
  select(Chr:Pos, log10BayesFactor, everything())

  for(i in 1:nrow(qtl_df)){  # from 1 to 16
  ## In block & > 2
  QTLandBF <- pw_df %>%
    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
                                             qtl_df$POS_hi[i])) %>% 
    tally()
  ## BF>2
  inBF <- pw_df %>%
    tally()
  ## BF not > 2 (all minus BF > 2, all is below)
  allwrtBF <- pw_df_all %>%
    tally()
  ## All in block
  inQTL <- pw_df_all %>%
    filter(Chr %in% qtl_df$CHR[i] & between(Pos, qtl_df$POS_lo[i],
                                             qtl_df$POS_hi[i])) %>%
    tally()
  notinBF <- (allwrtBF$n[1]-inBF$n[1])
  pval <- phyper(QTLandBF$n[1], inBF$n[1], allwrtBF$n[1]-inBF$n[1], 
                 inQTL$n[1], lower.tail = FALSE)

  if(i == 1){
  outputdf <- tibble(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], Pos_hi = qtl_df$POS_hi[i], inQTLandBF = QTLandBF$n[1], inBF = inBF$n[1], notinBF = notinBF, inQTL = inQTL$n[1], pvalue = pval)
  } else {
    outputdf <- outputdf %>%
      add_row(Chr = qtl_df$CHR[i], Pos_lo = qtl_df$POS_lo[i], Pos_hi = qtl_df$POS_hi[i],inQTLandBF = QTLandBF$n[1], inBF = inBF$n[1], notinBF = notinBF, inQTL = inQTL$n[1], pvalue = pval)
    }
  }
  return(outputdf)
}

get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      dplyr::mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

  pw_df <- as_tibble(m$posterior_weights)
  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
    bind_cols(pw_df)

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    dplyr::left_join(log10bf_df, by = "Marker") %>%
    dplyr::arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

qtlplot <- qtlplot %>%
  arrange(desc(lod)) %>%
  mutate(pos = as.integer(pos)) %>%
  group_by(CHR, pos) %>%
  slice(1) %>%
  arrange(CHR, pos)
enrich_gf <- get_QTL_enrichment(m_gulf_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Gulf", phenotype = "flowering")
enrich_mf <- get_QTL_enrichment(m_midw_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Midwest", phenotype = "flowering")
enrich_gamf <- get_QTL_enrichment(m_gam_fl_hyp, qtl_df = qtlplot, 0.99) %>%
  mutate(subpop = "Both", phenotype = "flowering")

enrich_df <- enrich_gf %>%
  full_join(enrich_mf) %>%
  full_join(enrich_gamf) %>%
  arrange(Chr, Pos_lo)
enrich_df <- enrich_df %>%
  left_join(qtlplot, by = c("Chr" = "CHR", "Pos_lo" = "POS_lo", "Pos_hi" = "POS_hi")) %>%
  select(Chr, Pos_lo, Pos_hi, inQTLandBF:subpop, lod, lodcolumn, pos) %>%
  unique() %>%
  arrange(Chr, Pos_lo, lodcolumn)

enrich_df %>%
  filter(pvalue < 0.05)

t.test(enrich_df[enrich_df$pvalue < 0.05,]$lod, enrich_df[enrich_df$pvalue >= 0.05,]$lod) 
qtlplot %>% group_by(CHR, POS_lo) %>% tally() # 21-24 unique QTL by Pos lo/hi. 23 by QTL maximum. 12 of 24 have enriched mash hits.

```

16 QTL. In all cases there was at least one SNP in the QTL interval that was in the 1% tail of SNPs in one of the three flowering mash runs.
We also looked for enrichment of SNPs in the 1% tail of significance in the mash distribution within each QTL interval. (in random genomic intervals of the same size do we see a similar level of enrichment?)
Half of these QTL had a significant enrichment of significant results for mash compared to expectations given the number of SNPs in the QTL interval.

2K: 2 QTL (lod 14 & 17), 1 enriched (in Midwest) (lod 17: CGDD_12C)
2N: 4 QTL (3 non-overlapping) (lod 19, 39, 17, 15), 2 enriched (lod 39 dylnfl50 in Midwest, in Both & 15 CGDD_12C in all three)
4K: 2 QTL, 2 enriched (lod 40 dyln in Midwest & 33 dyln in Gulf)
5N: 4 QTL (3 non-overlapping) (lod 55, 25, 17, 33), 3 enriched (2 non-overlapping) (lod 25 dyln in Midwest, 17 CGDD in Gulf, in Midwest, 33 dyln in Gulf, in Both)
7K: 2 QTL, 0 en (lod 14, 13)
8N: 1 QTL, 0 en (lod 15)
9N: 1 QTL, 0 en (lod 18)