### 2024-05-17 manuscript/revision2/revision2.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  9pt,
  twocolumn,
  twoside]{pnas-new}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{luacolor}
  \usepackage[soul]{lua-ul}
\else
  \usepackage{soul}

\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% TODO: Add custom LaTeX header directives here
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Diverse Genotype-by-Weather Interactions in Switchgrass},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel A. Smith; Jason Bonnette; Arvid R. Boe; Phillip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative rainfall, genetic variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\templatetype{pnasresearcharticle}

\title{Diverse Genotype-by-Weather Interactions in Switchgrass}

\newcommand{\equalcont}{1}

\newcommand{\correspond}{2}

\author[a%
,\equalcont%
,\correspond%
]{Alice H. MacQueen}
\author[a%
,\equalcont%
%
]{Li Zhang}
\author[a%
,\equalcont%
%
]{Samuel A. Smith}
\author[a%
%
%
]{Jason Bonnette}
\author[b%
%
%
]{Arvid R. Boe}
\author[c%
%
%
]{Phillip A. Fay}
\author[d%
%
%
]{Felix B. Fritschi}
\author[e%
%
%
]{David B. Lowry}
\author[f%
%
%
]{Robert B. Mitchell}
\author[g%
%
%
]{Francis M. Rouquette Jr}
\author[h%
%
%
]{Yanqi Wu}
\author[a%
%
%
]{Arbel Harpak}
\author[a%
%
,\correspond%
]{Thomas E. Juenger}

\affil[a]{University of Texas at Austin, Department of Integrative
Biology, Austin, 78712}
\affil[b]{South Dakota State University, Department of
Agronomy, Brookings, 57006}
\affil[c]{USDA-ARS, Grassland, Soil and Water Research
Laboratory, Temple, 76502}
\affil[d]{University of Missouri, Division of Plant
Sciences, Columbia, 65211}
\affil[e]{Michigan State University, Department of Plant Biology, East
Lansing, 48824}
\affil[f]{USDA-ARS, Wheat, Sorghum, and Forage Research
Unit, Lincoln, 68583}
\affil[g]{Texas A\&M University, Texas A\&M AgriLife Research and
Extension Center, Overton, 75684}
\affil[h]{Oklahoma State University, Department of Plant and Soil
Sciences, Stillwater, 74078}

\leadauthor{MacQueen}

\significancestatement{The timing of plant seasonal development
(phenology) has major impacts on fitness because of the steep fitness
cost of plant-environment mismatches. We infer the mixture of ways that
genetic effects on phenological traits depend on plant environment,
focusing on how effects covary with weather prior to the phenological
event (GxWeather). Most effects do not have GxWeather, but a minority
do. GxWeather is population-specific. The majority of effects on the
timing of vegetative growth in the Gulf population have GxWeather: these
effects differ in sign across environments and covary with a photoperiod
cue two weeks prior. A minority of effects on flowering date in the Gulf
and Midwest populations had GxWeather, covarying with a cumulative
rainfall and a photoperiod cue, respectively.}

\authorcontributions{T.E.J. designed research. D.B.L. contributed plant
material and resources. J.B., D.B.L., and T.E.J. designed and executed
field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
and T.E.J. hosted field experiments. A.H.M., L.Z., and S.A.S. conducted
statistical and computational analyses. The manuscript was written by
A.H.M. with contributions from all authors.}

\authordeclaration{The authors declare no conflicts of interest.}

\equalauthors{\textsuperscript{\equalcont} A.H.M. contributed equally to
this work with L.Z. and S.A.S.}

\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: alicem@uchicago.edu}
\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: tjuenger@utexas.edu}

\keywords{allele-by-environment effect variation | antagonistic
pleiotropy | photoperiod | cumulative rainfall | genetic variation}

\begin{document}
\maketitle

\begin{abstract}
The timing of vegetative and reproductive growth in plants
(``phenological timings'') depends on genetic effects, environmental
(e.g., weather) cues, and possibly their interaction. Quantifying these
GxWeather interactions can aid prediction and manipulation of
phenological timings. Here, we map GxWeather effects on phenological
timings in two highly divergent switchgrass (\emph{Panicum virgatum})
populations using repeated plantings of cloned individuals from these
populations at eight sites across the central United States. Some
genetic effects covary with interpretable weather-based cues. For
example, in the Gulf population, 65\% of genetic effects on the timing
of vegetative growth covary with daylength 14 days prior to green-up
date, and 33\% of genetic effects on the timing of flowering covary with
cumulative rainfall in the seven days prior to flowering. However, most
variation in genetic effects across sites is site-specific and can not
be attributed to variation in weather variables alone. We demonstrate
that we can identify genetic variation with GxWeather and assign these
loci to specific weather-based cues or other patterns. Breeding for
particular alleles at the GxWeather loci we identified could change
flowering responsiveness in a photoperiod and rainfall-specific way.
More broadly, our approach proposes a refined, powerful characterization
of genotype-by-environment interactions in any species phenotyped in
multiple environments.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
\firstpage[17]{2}

Plant phenological timings are major components of plant fitness
affected by multiple external environmental cues (e.g.~degree of winter
chilling, day length, temperature, and water availability) that signal
existing or upcoming growing conditions (1--3). Genetic responses to
environmental cues determine the speed, timing, and energy apportioned
to vegetative and reproductive growth and shape both the individual's
lifespan and its lifetime production of viable seed. Day length (or
photoperiod) is one of the most predictable environmental cues, and
genetic sensitivity to photoperiod protects plants from potentially
fatal consequences of phenological responses to temperature cues at the
``wrong'' time of year. However, the usefulness of specific
environmental cues depends on both features of the environment, such as
cue predictability and relevance, and the species' adaptive strategies
(4). Species with wide natural distributions can have multiple distinct
environmentally cued phenological responses: for example, populations of
sunflower (\emph{Helianthus annuus}) exhibit day-neutral, facultative
short day, and facultative long-day flowering responses, which vary with
their environments (5, 6). Distinct genetic responses in different
environments are known as genotype-by-environment interactions, or GxE.

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (22). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or sign-changing GxE, at the level of
individual loci (23--26), previous empirical work has found limited
evidence of antagonistic pleiotropy (12, 27). However, this work has
been limited by a known statistical bias that reduced detection of
genetic effects that differ in sign (27--29). Thus, despite substantial
interest in the frequencies of various forms of GxE, the prevalence of
antagonistic pleiotropy relative to other forms of GxE remains unknown.

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year (30).
However, as part of its wide environmental adaptation across the eastern
half of North America, its photoperiodicity has been predicted to differ
by plant latitude of origin (31, 32). We previously found divergent
Midwest and Gulf genetic subpopulations of switchgrass with distinct
sets of environmental adaptations, in that both populations had distinct
genetic variation associated with each of two fitness proxies, biomass
and overwinter survival (33). The Midwest genetic subpopulation is
primarily composed of individuals from the well-studied upland
switchgrass ecotype (34, 35), while the Gulf subpopulation has
individuals from the well-studied lowland ecotype and the phenotypically
intermediate coastal ecotype (33).

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology across gardens to
many patterns of weather covariance at these gardens. To do this, we
phenotype a diversity panel of hundreds of switchgrass genotypes from
the Midwest and Gulf subpopulations for the timing of vegetative and
reproductive development at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
We define multiple ways phenological timings might covary with weather
(Table 1) and additional ways phenological timings might vary by site
(SI Appendix, Table S1, Section SX), then jointly re-estimate genetic
effects on these timings at all eight sites using the set of these
covariance matrices that significantly improved the modeled
log-likelihood when included (SI Appendix, Section SX) (36). We use the
Bayesian framework \emph{mash} (multivariate adaptive shrinkage)
developed by (36), to refine effect size estimates from genome-wide
association (GWAS) conducted on individuals at each of the eight sites.
\emph{mash} allows us to identify and specify multiple covariance
structures among genetic effect estimates across sites, including
structures that represent covariance in weather variables of interest.
Importantly, this method circumvents statistical biases in detecting
genetic effects with the same or opposite signs (37). To confirm our
genetic mapping of GxWeather, we compare our posterior genetic effect
estimates from \emph{mash} to mapping results from an outbred pseudo-F2
cross grown at the same sites. Our analyses allow us to describe the
weather cues and genetic variation affecting phenology in two divergent
natural populations of switchgrass.

\section{Results}\label{results}

Genotypes from the Gulf and Midwest subpopulations had distinct
phenological timings and distinct patterns of phenological correlations
across our eight common garden sites (Figure~\ref{fig-map}). At the
three Texas common gardens (hereafter `Texas' gardens), located within
the natural range of the Gulf subpopulation, Gulf vegetative growth
(hereafter `green-up') occurred before Midwestern green-up, and Gulf
reproductive growth (hereafter `flowering') occurred after Midwestern
flowering (Figure~\ref{fig-map} A). At the four northernmost common
gardens (hereafter `North' gardens), located within the natural range of
the Midwest subpopulation, both Gulf green-up and flowering occurred
after Midwest green-up and flowering. At the Oklahoma common garden,
located near the natural range limits of both the Gulf and the Midwest
subpopulations, Gulf and Midwest green-up occurred over the same period,
and Gulf flowering occurred after Midwestern flowering
(Figure~\ref{fig-map} A). These patterns led to strong negative
phenotypic correlations for the timing of the start of vegetative growth
between the North and Texas gardens, particularly in the Gulf and across
all individuals from the Gulf and Midwest subpopulations (hereafter,
`Both' subpopulations), and contributed to positive phenotypic
correlations for the timing of flowering that had larger magnitudes at
more northern gardens (Figure~\ref{fig-map} B).

Narrow-sense heritabilities (h\textsuperscript{2}) suggested that
rank-changing GxE for these phenotypes was present across the common
gardens (Figure~\ref{fig-map} C). h\textsuperscript{2} were typically
high at individual gardens: 59\% on average for green-up date, and 87\%
for flowering date. However, h\textsuperscript{2} were variable across
gardens, and green-up dates were uncorrelated (r\textsuperscript{2}
\textless{} 0.2) or negatively correlated between pairs of gardens
(Figure~\ref{fig-map} B). These negative and small correlations
undoubtedly contributed to the low h\textsuperscript{2} values for
green-up and flowering date when estimated jointly at all eight gardens:
h\textsuperscript{2} was 0.8\% for green-up and 23.2\% for flowering
date.

\subsubsection{\texorpdfstring{Inference of GxWeather effects for
vegetative and reproductive timing using
\emph{mash}}{Inference of GxWeather effects for vegetative and reproductive timing using mash}}\label{inference-of-gxweather-effects-for-vegetative-and-reproductive-timing-using-mash}

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which can flexibly capture many modes of covariance across gardens
(Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We specified three
qualitative categories of covariance structures across gardens:
``canonical'' covariance, with simple patterns of effect size covariance
introduced in the initial \emph{mash} manuscript; ``data-driven''
covariances derived from common patterns of SNP effects observed in the
data, and ``GxWeather'' covariance, estimated from the covariance of
empirical weather patterns at each garden at specific times before the
phenological event (Table 1; SI Appendix, Section S2). For example, the
more similar that the amount of rainfall prior to flowering is for
genotypes A and B between gardens in MI and MO, the nearer these
covariances will be to one. If rainfall amounts are uncorrelated,
covariance will be near zero; if there is rank changing (e.g.~high
rainfall before flowering of A at MO but not MI, and high rainfall
before flowering for B at MI but not MO), this covariance will be near
negative one. We specified multiple covariance structures for the
canonical, data-driven, and GxWeather categories, each of which
represented a specific pattern of covariance in genetic effect size
estimates that could be included in the model. The phenotypic
correlations for the timing of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the mash model.

The GxWeather covariance structures allow hypothesis testing of specific
weather variables as cues for the start of vegetative and reproductive
growth. Say that a SNP in FLC controls flowering in a
photoperiod-dependent manner. In that case, the joint estimate of
effects for that SNP could have a high mixture proportion, or mass, on a
covariance matrix created using a photoperiod-based environmental cue,
such as day length at some interval prior to flowering. In our data, we
would infer that the effect of that SNP on flowering was caused by a
response to the environmental cue used to construct the GxWeather
covariance structure with the largest mass. Overall, loadings of genetic
effects on these GxWeather matrices in the model, along with the
posterior effect size estimates, provide information on genome-wide
patterns of SNP-environment interaction.

We used a greedy algorithm to select covariance matrices from each
model's set that significantly improved the model likelihood. Five
GxWeather covariance structures were selected, zero to two per
subpopulation: two for green-up date, and three for flowering date. Of
these five, three had mass on them in mash models of the strong effects
(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
covariances between sites between Texas and North gardens
(Figure~\ref{fig-covar} A), while one had all positive or near-zero
covariances. SNP-associated phenotypic effects covaried with different
weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf green-up fell on a
covariance matrix constructed using the covariance of day length 14 days
prior to the date of green-up. The covariance matrix for this weather
cue was visually similar to the observed pattern of phenotypic
correlation for the timing of vegetative growth in the Gulf
subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash
models of the timing of vegetative growth in the Midwest subpopulation
and Both subpopulations did not include this GxWeather covariance; the
Midwest had no weight on any GxWeather covariance, while Both
subpopulations had non-zero weights on two additional GxWeather
covariance types, average temperature one day prior to green-up, and the
day length change in seconds in the day prior to green-up
(Figure~\ref{fig-covar} B). The average temperature covariance matrix
had negative covariances between Texas and North gardens, though not as
strong as the negative phenotypic correlations seen in Both
subpopulations (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Only
the Gulf subpopulation and Both subpopulations had mass on any GxWeather
covariance matrices (Figure~\ref{fig-covar} C) for vegetative growth.

For flowering date, distinct GxWeather covariance structures captured
covariance in effect sizes for SNPs in the Gulf and Midwest
subpopulations. 33\% of SNP effects on flowering in the Gulf
subpopulation covaried with cumulative rainfall in the seven days prior
to flowering (Figure~\ref{fig-covar} D). 22.6\% of SNP effects on
flowering in the Midwest subpopulation covaried with day length change
in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
negative covariances between Texas and North gardens. Neither of these
covariance matrices significantly improved the log-likelihood of the
\emph{mash} model of Both subpopulations, and no GxWeather covariance
matrices had non-zero mass in models of Both subpopulations
(Figure~\ref{fig-covar} E). In five of the six mash models of strong
effects, the GxWeather covariance matrices captured a minority of the
posterior weights of the strong effects (Figure~\ref{fig-covar} C,E); in
these five models, the majority of this mass was on various canonical
covariance matrices. These matrices included simple heterozygosity, with
intermediate, positive covariances between all gardens, and
garden-specific effects present only at one garden.

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. If the overall
pattern of phenotypic expression differs between a pair of gardens, this
difference may be comprised of SNP effects that differ in sign between
these gardens (effects with antagonistic pleiotropy), SNP effects that
differ in magnitude (effects that are large in one garden or region and
smaller in others), and SNP effects that are indistinguishable in the
two gardens (Figure~\ref{fig-effects} A). We used the local false sign
rate (lfsr), an analogue of the lfdr that establishes confidence in the
effect sign, not the effect's difference from zero, to determine
significance. We required lfsr significance (p \textless{} 0.05) in both
gardens to include effects. This means that our tests for antagonistic
pleiotropy, or a sign change between conditions, carry an equal
statistical burden to those for effects with the same sign.

For green-up date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens
(Figure~\ref{fig-effects} B; SI Appendix, Fig SX). 78.7\% of pairwise
comparisons between North and Texas gardens had a difference in sign,
while only 28.6\% and 0.2\% of North-North or Texas-Texas comparisons
had a difference in sign, respectively (Figure~\ref{fig-effects} B). The
majority of pairwise effects for greenup for the Midwest
(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
same sign, and effects most often differed in magnitude between gardens
within and between the regions (Figure~\ref{fig-effects} B).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Figure~\ref{fig-effects} B). More
effects differed in magnitude between the Texas and North regions than
within these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects}
B), while the majority of effects had no GxE. The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign (Figure~\ref{fig-effects}
B). Most differences in sign were between TX1, the southernmost garden,
and all other gardens (SI Appendix, Fig SX). Similarly, more effects
that differed in magnitude included gardens in the Texas region
(52.3-55.9\%), and most effect pairs in the North region were not
distinguishable (91.5\%).

\subsubsection{Confirmation of effects on phenology using an independent
mapping
population}\label{confirmation-of-effects-on-phenology-using-an-independent-mapping-population}

We sought additional experimental support for our joint re-estimates of
SNP effects using an independent pseudo-F2 mapping population created
from Gulf \& Midwest individuals and grown at the same sites
(Figure~\ref{fig-qtl} A,B). We conducted quantitative trait loci (QTL)
mapping of flowering as functions of five environmental cues that we
also used as covariance matrices in mash, and identified eight QTL for
flowering date, eight QTL for flowering as a function of day length
change two days prior, one QTL for the start of vegetative growth, and
two QTL for vegetative growth as a function of daylength change one day
prior, all of which showed QTL by environment interactions (SI Appendix,
Fig. S3). All QTL for flowering overlapped one or more homologs from
rice or \emph{A. thaliana} with functionally validated roles in
flowering (SI Appendix, Dataset 7). All flowering and green-up QTL
intervals contained at least one SNP significant in at least one mash
run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
tail of significance, whichever was stricter (SI Appendix, Dataset 8).
We also looked for enrichments of mash SNPs in the 1\% tail of
significance (the `mash 1\% tail') within each QTL interval. At the 5\%
level, three QTL had enrichments of SNPs in the mash 1\% tail. Overall,
there were five significant enrichments (p \textless{} 0.05,
hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
Thus, we were able to experimentally support some of our re-estimates of
SNP effects with a QTL mapping experiment using a separate mapping
population.

\section{Discussion}\label{discussion}

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered. Here, we demonstrate that we can associate multiple
patterns of GxWeather with specific genomic regions using a switchgrass
diversity panel grown at eight common gardens. We can assign genetic
effects to both GxWeather patterns with interpretable weather-based
cues, and to agnostic, site-based patterns. We use this approach to
study GxWeather for the timings of vegetative and reproductive
development in the deeply genetically diverged Gulf and Midwest
subpopulations of switchgrass.

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (Figure~\ref{fig-effects} B). This
result supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (23--26), and is
the first experimental work using GWAS across common gardens to find
antagonistic pleiotropy to be common in small genomic regions (12, 38,
39).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related variation, in that flowering
timing covaries with a day length change signal two days before
flowering occurs. In contrast, the Gulf subpopulation does not have
variation in flowering based on a photoperiod cue. Instead, the Gulf
subpopulation has variation in flowering that covaries with the rainfall
that occurs in the week prior to flowering. Three genomic regions
affecting flowering that we re-estimated across all eight sites were
also supported by QTL from an independent mapping population at these
sites (Figure~\ref{fig-qtl} B).

Identifying the environmental cues that are predictive of, or even
correlated with, plant phenotypic responses remains a major challenge to
studies interrogating gene action across many natural environments. The
GxWeather photoperiod and cumulative rainfall cues we identify here are
functions of the genotypes measured and capture only a minority of SNP
effects on flowering. We could only assign SNP effects to a GxWeather
covariance structures in four of the six phenotype \& genetic
subpopulations we modeled. It's likely we did not include some GxWeather
important to these phenological cues - for example, overwintering
parameters that might cause variation in the start of vegetative growth.
More generally, it is difficult to predict the time scales over which
individuals may integrate environmental cues, particularly in perennial
species which may integrate these cues over longer time scales. If this
integration time itself varies between individuals, the covariance
structures we modeled cannot reflect this, though these structures would
likely be highly correlated with GxWeather structures we did include.
Our approach offers an opportunity to specify multiple environmental
cues and compete them to explain patterns of genetic effects, allowing
us to detect how important these cues are genome-wide, and how strongly
each cue influences each SNP. This is a key development to further
improve our understanding of genetic variation in GxE.

\section{Figures and Tables}\label{figures-and-tables}

\clearpage
\onecolumn
\begin{table}[t!]
\centering
\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices. Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices \\
\midrule
1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
\bottomrule
\end{tabular}

\end{table}
\clearpage
\twocolumn

\begin{figure}

\centering{

\includegraphics{images/Fig1_200dpi.png}

}

\caption{\label{fig-map}Characterization of the timing of vegetative
(green-up) and reproductive (flowering) growth in the switchgrass
diversity panel. (a) Map and trait histograms of green-up and flowering
dates across two genetically distinct switchgrass subpopulations and
eight common gardens. Purple represents individuals from the Midwest
genetic subpopulation, and pink individuals from the Gulf subpopulation;
map positions represent the original collection locations for the
genotypes, and shapes represent the ecotype of the genotype. Vertical
dashed lines indicate the summer solstice. Common gardens are arranged
in latitudinal order. (b) Phenotypic correlations between clonal
replicates planted at eight common gardens, within and between two
genetic subpopulations. (c) Narrow sense heritability of green-up and
flowering within single common gardens (purple) and across all eight
common gardens (green), within and between two genetic subpopulations.
(d) Flow diagram of the methods applied to the green-up and flowering
dates to jointly estimate SNP effects across all sites. Mash was fit to
SNP effect data and used to find covariance matrices that improved the
mash model likelihood using a large set of randomly selected, relatively
unlinked SNP effects; this model was applied to a ``strong'' set of SNP
effects with large effect sizes in the univariate GWAS.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png}

}

\caption{\label{fig-covar}Example hypothesis-driven covariance matrices
specified in mash and the posterior weights placed on all covariance
matrices. (a) Left column: Five example hypothesized covariance matrices
specified for the green-up date or flowering date phenotype; these
matrices were created from environment-specific correlations across
eight common gardens, and are described in Table 1. Common gardens are
arranged in latitudinal order within the matrices. Right column: Five
example canonical covariance matrices. Canonical matrices (purple) have
simple interpretations, such as equal effects across all common gardens,
or effects specific to a single common garden. (b,d) Total posterior
weight placed on each covariance matrix type specified for (b) green-up
date and (d) flowering date mash models, within and between two genetic
subpopulations. Hypothesized covariance matrices (green). Covariance
matrices included in mash that had zero posterior weight in all three
mash runs on the genetic subpopulations, such as the identity matrix,
are not shown. (c,e) Total posterior weight placed on covariance
matrices that were hypothesized or canonical, for the (c) green-up date
phenotype and (e) flowering date phenotype.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{images/Figure_3_GxE.png}

}

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated effects in eight common gardens, for effects with
lfsr \textless{} 0.05. a) Examples of effect patterns at three pairs of
sites with three types of GxE. Sign: Effects that differ in sign at
these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude: Effects
identical in sign (p \textless{} 0.05, lfsr) that differ in magnitude by
a factor of \textgreater0.4. None: Effects not distinguishable by
magnitude nor sign of the effect. b) The fraction of effects with each
GxE type for the start of vegetative growth (Greenup Date) and
reproductive growth (Flowering Date), within and between two genetic
subpopulations. Common gardens are grouped by the larger region they
came from: North gardens are within the natural range of the Midwest
subpopulation, while Texas gardens are within the natural range of the
Gulf subpopulation.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{images/Figure_4_QTL_Overlaps.png}

}

\caption{\label{fig-qtl}Overlaps of QTL from an outbred pseudo-F2 cross
and with jointly re-estimated SNP effects in the 1\% tail of
significance from a diversity panel. Dotted lines indicate
permutation-based significance thresholds for each weather-related
function. Stars indicate QTL with significant enrichment for SNPs in the
1\% mash tail; G, M, and B indicate which subpopulation had enrichment:
G - Gulf subpopulation, M -Midwest subpopulation, B - both
subpopulations. Rug plots show genomic locations of SNPs in the 1\% mash
tail for flowering date for each subpopulation. a) QTL mapping for the
start of vegetative growth (Greenup Date), and three weather-related
functions of greenup date. b) QTL mapping for the start of reproductive
growth (Flowering Date), and two weather-related functions of flowering
date.}

\end{figure}%

\section{Materials and Methods}\label{materials-and-methods}

Whenever possible, plant material will be shared upon request. Source
data and code to replicate these analyses are available at:
https://github.com/Alice-MacQueen/pvdiv-phenology-gxe.git. SNP data to
replicate these analyses are available from the UT dataverse at
https://doi.org/link.

\subsubsection{Genotype-by-environment effects on green-up and flowering
as functions of weather-based
cues}\label{genotype-by-environment-effects-on-green-up-and-flowering-as-functions-of-weather-based-cues}

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (35, 39, 40). We scored the
start of vegetative growth, or green-up date, as the day of the year
when 50\% of the tiller area of the crown of the plant cut the previous
year had green growth. The start of reproductive growth, or flowering
date, was the day of the year when 50\% of the plant tillers had
panicles undergoing anthesis. We scored green-up and flowering as day of
the year, then linked these dates to multiple weather-based
environmental factors measured daily at each common garden (SI Appendix,
Section S1, Table S1).

The formation and resequencing of the diversity panel has been described
previously (33). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (41), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both'). Analyses to determine narrow-sense
heritability (h\textsuperscript{2}) for green-up and flowering were done
using linear mixed models and followed (33). Details on these models can
be found in (SI Appendix, Section S3,S4).

\subsubsection{Mapping major patterns of genotype-by-environment effects
on green-up and
flowering}\label{mapping-major-patterns-of-genotype-by-environment-effects-on-green-up-and-flowering}

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(36). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (40). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for the two traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

We generated hypothesis-based covariance matrices derived from
correlations in environmental cues in the green-up or flowering date
windows for the three populations (Table 1; SI Appendix, Table S1,
Section S1). These covariance matrices represent correlations between
identical genotypes drawn from a specific population at pairs of common
gardens; covariances near one mean that the population has a strong,
positive linear relationship in individual responses at that pair of
gardens, while covariances near zero mean that there is no relationship
within the population for individual responses at that pair of gardens.
Mash SNP effects will undergo strong shrinkage towards one another in
the first case, and little shrinkage in the second case. Mash also
generates data-driven covariance matrices corresponding to major
patterns of SNP effects present in the data. We generated six
data-driven matrices per mash run, five produced by singular value
decomposition (SVD) of an overall matrix.

Last, we characterized the overall patterns of antagonistic pleiotropy
in the set of SNPs where there was pairwise significance of effects at
pairs of gardens. To do this, we used the `get\_GxE' function of the
switchgrassGWAS R package. First, this determines the set of SNPs with
evidence of significant effects in both conditions for all pairs of
conditions using local false sign rates (lfsr) as the significance
criteria. Second, to determine antagonistic pleiotropy, this function
determines if effects significant in both conditions are of opposite
sign.

Using the lfsr rather than the local false discovery rate (lfdr) is a
critical change in our ability to detect antagonistic pleiotropy. The
lfdr, like other measures of FDR, focuses on if we have enough evidence
to reject the null hypothesis that an effect j is 0, or that there is a
significant effect. Previous studies of antagonistic pleiotropy (e.g.
(39)) have used the lfdr or equivalent statistical tests to detect
antagonistic pleiotropy. These tests were conservative, in that they
required two non-zero effects of different signs, while tests for
differential sensitivity required only one non-zero effect. This
previous work recognized that this testing bias could lead to
undercounting occurrences of antagonistic pleiotropy (27, 29), and
sought to reduce it by permutation (28). However, using the lfsr to test
for antagonistic pleiotropy does not undercount occurrences of
antagonistic pleiotropy, as this statistic answers a fundamentally
different question. For each effect j, the lfsr\textsubscript{j} is
defined as the probability that we make an error in the sign of effect j
if we were forced to declare the effect positive or negative (37). Thus,
rather than asking ``Are these two effects different?'' - as we
reasonably expect two effects to be, even if this difference cannot be
measured - the local false sign rate answers a more meaningful question:
Can we be confident in the sign of this effect?

In addition, the get\_GxE function also sets an arbitrary threshold to
count an effect as changing in magnitude between environments, commonly
known as differential sensitivity or a change in amplitude of the
effect. For differential sensitivity, this function determines if
effects significant in both conditions are of the same sign and of a
magnitude (not tested for significance) that differs by a factor of 0.4
or more. The remaining effects that are significant in both conditions
have the same effect sign and similar effect magnitudes and we denote
these effects as having no GxE. The distinction between effects with
different magnitudes is arbitrary but useful to fully characterize how
effects vary across environments to ultimately influence phenotypes. Our
use of the lfsr to determine significance and our specification that SNP
effects must be significant in both conditions to be included means that
our tests for antagonistic pleiotropy carry an equal statistical burden
to those measuring differential sensitivity and effects without GxE.

\subsubsection{Confirmation of genotype-by-environment effects using an
independent mapping
population}\label{confirmation-of-genotype-by-environment-effects-using-an-independent-mapping-population}

To confirm candidate genomic regions and patterns of allelic effects
found in the diversity panel, we analyzed flowering in an outbred
pseudo-F2 cross between four individuals, two Midwest and two Gulf
individuals. The formation of this mapping population has been described
previously (35); additional details on QTL mapping can be found in SI
Appendix, Section S6. To be directly comparable to the diversity panel
data, only 2019 phenology data from the pseudo-F2 cross from the same
eight common garden sites were used.

\section{References}\label{references}

\bibsplit[2]

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-bauerle_photoperiodic_2012}
\CSLLeftMargin{1. }%
\CSLRightInline{W. L. Bauerle, \emph{et al.},
\href{https://doi.org/10.1073/pnas.1119131109}{Photoperiodic regulation
of the seasonal pattern of photosynthetic capacity and the implications
for carbon cycling}. \emph{Proceedings of the National Academy of
Sciences} \textbf{109}, 8612--8617 (2012).}

\bibitem[\citeproctext]{ref-andres2012genetic}
\CSLLeftMargin{2. }%
\CSLRightInline{F. Andrés, G. Coupland, The genetic basis of flowering
responses to seasonal cues. \emph{Nature Reviews Genetics} \textbf{13},
627--639 (2012).}

\bibitem[\citeproctext]{ref-korner2010phenology}
\CSLLeftMargin{3. }%
\CSLRightInline{C. Körner, D. Basler, Phenology under global warming.
\emph{Science} \textbf{327}, 1461--1462 (2010).}

\bibitem[\citeproctext]{ref-botero2015evolutionary}
\CSLLeftMargin{4. }%
\CSLRightInline{C. A. Botero, F. J. Weissing, J. Wright, D. R.
Rubenstein, Evolutionary tipping points in the capacity to adapt to
environmental change. \emph{Proceedings of the National Academy of
Sciences} \textbf{112}, 184--189 (2015).}

\bibitem[\citeproctext]{ref-blackman2013interacting}
\CSLLeftMargin{5. }%
\CSLRightInline{B. K. Blackman, Interacting duplications, fluctuating
selection, and convergence: The complex dynamics of flowering time
evolution during sunflower domestication. \emph{Journal of experimental
botany} \textbf{64}, 421--431 (2013).}

\bibitem[\citeproctext]{ref-henry2014transitions}
\CSLLeftMargin{6. }%
\CSLRightInline{L. P. Henry, R. H. Watson, B. K. Blackman, Transitions
in photoperiodic flowering are common and involve few loci in wild
sunflowers (helianthus; asteraceae). \emph{American Journal of Botany}
\textbf{101}, 1748--1758 (2014).}

\bibitem[\citeproctext]{ref-aagren2017adaptive}
\CSLLeftMargin{7. }%
\CSLRightInline{J. Ågren, C. G. Oakley, S. Lundemo, D. W. Schemske,
Adaptive divergence in flowering time among natural populations of
arabidopsis thaliana: Estimates of selection and QTL mapping.
\emph{Evolution} \textbf{71}, 550--564 (2017).}

\bibitem[\citeproctext]{ref-brachi2010linkage}
\CSLLeftMargin{8. }%
\CSLRightInline{B. Brachi, \emph{et al.}, Linkage and association
mapping of arabidopsis thaliana flowering time in nature. \emph{PLoS
genetics} \textbf{6}, e1000940 (2010).}

\bibitem[\citeproctext]{ref-dittmar2014flowering}
\CSLLeftMargin{9. }%
\CSLRightInline{E. L. Dittmar, C. G. Oakley, J. Ågren, D. W. Schemske,
Flowering time QTL in natural populations of arabidopsis thaliana and
implications for their adaptive value. \emph{Molecular ecology}
\textbf{23}, 4291--4303 (2014).}

\bibitem[\citeproctext]{ref-li2018genomic}
\CSLLeftMargin{10. }%
\CSLRightInline{X. Li, T. Guo, Q. Mu, X. Li, J. Yu, Genomic and
environmental determinants and their interplay underlying phenotypic
plasticity. \emph{Proceedings of the National Academy of Sciences}
\textbf{115}, 6679--6684 (2018).}

\bibitem[\citeproctext]{ref-romero2017study}
\CSLLeftMargin{11. }%
\CSLRightInline{J. A. Romero Navarro, \emph{et al.}, A study of allelic
diversity underlying flowering-time adaptation in maize landraces.
\emph{Nature genetics} \textbf{49}, 476--480 (2017).}

\bibitem[\citeproctext]{ref-wadgymar2017identifying}
\CSLLeftMargin{12. }%
\CSLRightInline{S. M. Wadgymar, \emph{et al.}, Identifying targets and
agents of selection: Innovative methods to evaluate the processes that
contribute to local adaptation. \emph{Methods in Ecology and Evolution}
\textbf{8}, 738--749 (2017).}

\bibitem[\citeproctext]{ref-blumel2015flowering}
\CSLLeftMargin{13. }%
\CSLRightInline{M. Blümel, N. Dally, C. Jung, Flowering time regulation
in crops---what did we learn from arabidopsis? \emph{Current opinion in
biotechnology} \textbf{32}, 121--129 (2015).}

\bibitem[\citeproctext]{ref-jung2009flowering}
\CSLLeftMargin{14. }%
\CSLRightInline{C. Jung, A. E. Müller, Flowering time control and
applications in plant breeding. \emph{Trends in plant science}
\textbf{14}, 563--573 (2009).}

\bibitem[\citeproctext]{ref-turner2005pseudo}
\CSLLeftMargin{15. }%
\CSLRightInline{A. Turner, J. Beales, S. Faure, R. P. Dunford, D. A.
Laurie, The pseudo-response regulator ppd-H1 provides adaptation to
photoperiod in barley. \emph{Science} \textbf{310}, 1031--1034 (2005).}

\bibitem[\citeproctext]{ref-faure2012mutation}
\CSLLeftMargin{16. }%
\CSLRightInline{S. Faure, \emph{et al.}, Mutation at the circadian clock
gene EARLY MATURITY 8 adapts domesticated barley (hordeum vulgare) to
short growing seasons. \emph{Proceedings of the National Academy of
Sciences} \textbf{109}, 8328--8333 (2012).}

\bibitem[\citeproctext]{ref-hung2012zmcct}
\CSLLeftMargin{17. }%
\CSLRightInline{H.-Y. Hung, \emph{et al.}, ZmCCT and the genetic basis
of day-length adaptation underlying the postdomestication spread of
maize. \emph{Proceedings of the National Academy of Sciences}
\textbf{109}, E1913--E1921 (2012).}

\bibitem[\citeproctext]{ref-zakhrabekova2012induced}
\CSLLeftMargin{18. }%
\CSLRightInline{S. Zakhrabekova, \emph{et al.}, Induced mutations in
circadian clock regulator mat-a facilitated short-season adaptation and
range extension in cultivated barley. \emph{Proceedings of the National
Academy of Sciences} \textbf{109}, 4326--4331 (2012).}

\bibitem[\citeproctext]{ref-yang2013oself3}
\CSLLeftMargin{19. }%
\CSLRightInline{Y. Yang, Q. Peng, G.-X. Chen, X.-H. Li, C.-Y. Wu, OsELF3
is involved in circadian clock regulation for promoting flowering under
long-day conditions in rice. \emph{Molecular Plant} \textbf{6}, 202--215
(2013).}

\bibitem[\citeproctext]{ref-pin2012multifaceted}
\CSLLeftMargin{20. }%
\CSLRightInline{P. Pin, O. Nilsson, The multifaceted roles of FLOWERING
LOCUS t in plant development. \emph{Plant, cell \& environment}
\textbf{35}, 1742--1755 (2012).}

\bibitem[\citeproctext]{ref-weller2019parallel}
\CSLLeftMargin{21. }%
\CSLRightInline{J. L. Weller, \emph{et al.}, Parallel origins of
photoperiod adaptation following dual domestications of common bean.
\emph{Journal of Experimental Botany} \textbf{70}, 1209--1219 (2019).}

\bibitem[\citeproctext]{ref-Weine2023.06.21.545998}
\CSLLeftMargin{22. }%
\CSLRightInline{E. Weine, S. P. Smith, R. K. Knowlton, A. Harpak,
Tradeoffs in modeling context dependency in complex trait genetics.
\emph{bioRxiv} (2024)
https:/doi.org/\href{https://doi.org/10.1101/2023.06.21.545998}{10.1101/2023.06.21.545998}.}

\bibitem[\citeproctext]{ref-levene1953genetic}
\CSLLeftMargin{23. }%
\CSLRightInline{H. Levene, Genetic equilibrium when more than one
ecological niche is available. \emph{The American Naturalist}
\textbf{87}, 331--333 (1953).}

\bibitem[\citeproctext]{ref-felsenstein1976theoretical}
\CSLLeftMargin{24. }%
\CSLRightInline{J. Felsenstein, The theoretical population genetics of
variable selection and migration. \emph{Annual review of genetics}
\textbf{10}, 253--280 (1976).}

\bibitem[\citeproctext]{ref-kawecki2004conceptual}
\CSLLeftMargin{25. }%
\CSLRightInline{T. J. Kawecki, D. Ebert, Conceptual issues in local
adaptation. \emph{Ecology letters} \textbf{7}, 1225--1241 (2004).}

\bibitem[\citeproctext]{ref-hedrick1986genetic}
\CSLLeftMargin{26. }%
\CSLRightInline{P. W. Hedrick, Genetic polymorphism in heterogeneous
environments: A decade later. \emph{Annual review of ecology and
systematics} \textbf{17}, 535--566 (1986).}

\bibitem[\citeproctext]{ref-des2013genotype}
\CSLLeftMargin{27. }%
\CSLRightInline{D. L. Des Marais, K. M. Hernandez, T. E. Juenger,
Genotype-by-environment interaction and plasticity: Exploring genomic
responses of plants to the abiotic environment. \emph{Annual Review of
Ecology, Evolution, and Systematics} \textbf{44}, 5--29 (2013).}

\bibitem[\citeproctext]{ref-anderson2013genetic}
\CSLLeftMargin{28. }%
\CSLRightInline{J. T. Anderson, C.-R. Lee, C. A. Rushworth, R. I.
Colautti, T. Mitchell-Olds, Genetic trade-offs and conditional
neutrality contribute to local adaptation. \emph{Molecular ecology}
\textbf{22}, 699--708 (2013).}

\bibitem[\citeproctext]{ref-anderson2011evolutionary}
\CSLLeftMargin{29. }%
\CSLRightInline{J. T. Anderson, J. H. Willis, T. Mitchell-Olds,
Evolutionary genetics of plant adaptation. \emph{Trends in Genetics}
\textbf{27}, 258--266 (2011).}

\bibitem[\citeproctext]{ref-mitchell1997predicting}
\CSLLeftMargin{30. }%
\CSLRightInline{R. B. Mitchell, K. J. Moore, L. E. Moser, J. O. Fritz,
D. D. Redfearn, Predicting developmental morphology in switchgrass and
big bluestem. \emph{Agronomy Journal} \textbf{89}, 827--832 (1997).}

\bibitem[\citeproctext]{ref-parrish2005biology}
\CSLLeftMargin{31. }%
\CSLRightInline{D. J. Parrish, J. H. Fike, The biology and agronomy of
switchgrass for biofuels. \emph{BPTS} \textbf{24}, 423--459 (2005).}

\bibitem[\citeproctext]{ref-casler2004latitudinal}
\CSLLeftMargin{32. }%
\CSLRightInline{M. Casler, K. P. Vogel, C. Taliaferro, R. Wynia,
Latitudinal adaptation of switchgrass populations. \emph{Crop Science}
\textbf{44}, 293--303 (2004).}

\bibitem[\citeproctext]{ref-lovell2021genomic}
\CSLLeftMargin{33. }%
\CSLRightInline{J. T. Lovell, \emph{et al.}, Genomic mechanisms of
climate adaptation in polyploid bioenergy switchgrass. \emph{Nature}
\textbf{590}, 438--444 (2021).}

\bibitem[\citeproctext]{ref-porter1966analysis}
\CSLLeftMargin{34. }%
\CSLRightInline{C. L. Porter Jr, An analysis of variation between upland
and lowland switchgrass, panicum virgatum l., in central oklahoma.
\emph{Ecology} \textbf{47}, 980--992 (1966).}

\bibitem[\citeproctext]{ref-milano2016genetic}
\CSLLeftMargin{35. }%
\CSLRightInline{E. R. Milano, D. B. Lowry, T. E. Juenger, The genetic
basis of upland/lowland ecotype divergence in switchgrass (panicum
virgatum). \emph{G3: Genes, Genomes, Genetics} \textbf{6}, 3561--3570
(2016).}

\bibitem[\citeproctext]{ref-urbut2019flexible}
\CSLLeftMargin{36. }%
\CSLRightInline{S. M. Urbut, G. Wang, P. Carbonetto, M. Stephens,
Flexible statistical methods for estimating and testing effects in
genomic studies with multiple conditions. \emph{Nature genetics}
\textbf{51}, 187--195 (2019).}

\bibitem[\citeproctext]{ref-10.1093ux2fbiostatisticsux2fkxw041}
\CSLLeftMargin{37. }%
\CSLRightInline{M. Stephens,
\href{https://doi.org/10.1093/biostatistics/kxw041}{{False discovery
rates: a new deal}}. \emph{Biostatistics} \textbf{18}, 275--294 (2016).}

\bibitem[\citeproctext]{ref-savolainen2013ecological}
\CSLLeftMargin{38. }%
\CSLRightInline{O. Savolainen, M. Lascoux, J. Merilä, Ecological
genomics of local adaptation. \emph{Nature Reviews Genetics}
\textbf{14}, 807--820 (2013).}

\bibitem[\citeproctext]{ref-lowry2019qtl}
\CSLLeftMargin{39. }%
\CSLRightInline{D. B. Lowry, \emph{et al.}, QTL\(\times\) environment
interactions underlie adaptive divergence in switchgrass across a large
latitudinal gradient. \emph{Proceedings of the National Academy of
Sciences} \textbf{116}, 12933--12941 (2019).}

\bibitem[\citeproctext]{ref-lovell2021}
\CSLLeftMargin{40. }%
\CSLRightInline{J. T. Lovell, \emph{et al.},
\href{https://doi.org/10.1038/s41586-020-03127-1}{Genomic mechanisms of
climate adaptation in polyploid bioenergy switchgrass}. \emph{Nature},
1--7 (2021).}

\bibitem[\citeproctext]{ref-kortefarlow2013}
\CSLLeftMargin{41. }%
\CSLRightInline{A. Korte, A. Farlow, The advantages and limitations of
trait analysis with GWAS: A review. \emph{Plant methods} \textbf{9}, 29
(2013).}

\end{CSLReferences}

\acknow{We thank the Brackenridge Field laboratory, the Ladybird Johnson
Wildflower Center, and the Juenger laboratory for support with plant
care and propagation. This material is based upon work supported in part
by the Great Lakes Bioenergy Research Center, U.S. Department of Energy,
Office of Science, Office of Biological and Environmental Research under
Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of
Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, National
Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and
the Long-term Ecological Research Program (DEB 1832042) at the Kellogg
Biological Station.}

\showacknow{} % Display the acknowledgments section

\end{document}