### 2024-08-18 manuscript/revision2/revision_biorxiv.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
\documentclass[
  9pt,
  twocolumn,
  twoside]{simple-article}% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{flushend.sty}{\usepackage{flushend}}{}

\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{authblk}
\usepackage{orcidlink}
\definecolor{mypink}{RGB}{219, 48, 122}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{simple-article}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs

\hypersetup{
  pdftitle={Diverse Genotype-by-Weather Interactions in Switchgrass},
  pdfauthor={Alice H. MacQueen; Li Zhang; Samuel P. Smith; Jason Bonnette; Arvid R. Boe; Philip A. Fay; Felix B. Fritschi; David B. Lowry; Robert B. Mitchell; Francis M. Rouquette Jr; Yanqi Wu; Arbel Harpak; Thomas E. Juenger},
  colorlinks=true,
  linkcolor={Blue},
  filecolor={Blue},
  citecolor={Blue},
  urlcolor={Maroon},
  pdfcreator={LaTeX via pandoc}}

\title{Diverse Genotype-by-Weather Interactions in Switchgrass}

\newif\ifnoAffil\noAffilfalse
\newif\ifoneAffil\oneAffilfalse

\newcommand{\setNumAffil}[1]{%
    \ifnum#1=0
        \noAffiltrue
    \else
        \ifnum#1=1
            \oneAffiltrue
        \fi
    \fi
}

\setNumAffil{8}

\ifnoAffil
%% No affiliation given....
\author{Alice H. MacQueen, Li Zhang, Samuel P. Smith, Jason
Bonnette, Arvid R. Boe, Philip A. Fay, Felix B. Fritschi, David B.
Lowry, Robert B. Mitchell, Francis M. Rouquette Jr, Yanqi Wu, Arbel
Harpak, Thomas E. Juenger}
\else
\ifoneAffil
%% Single affiliation given...
\author{Alice H. MacQueen}
\author{Li Zhang}
\author{Samuel P. Smith}
\author{Jason Bonnette}
\author{Arvid R. Boe}
\author{Philip A. Fay}
\author{Felix B. Fritschi}
\author{David B. Lowry}
\author{Robert B. Mitchell}
\author{Francis M. Rouquette Jr}
\author{Yanqi Wu}
\author{Arbel Harpak}
\author{Thomas E. Juenger}
\affil{University of Texas at Austin, Austin}
\else
%% Multiple affiliations given
\author[1]{Alice H. MacQueen}
\author[1]{Li Zhang}
\author[1]{Samuel P. Smith}
\author[1]{Jason Bonnette}
\author[2]{Arvid R. Boe}
\author[3]{Philip A. Fay}
\author[4]{Felix B. Fritschi}
\author[5]{David B. Lowry}
\author[6]{Robert B. Mitchell}
\author[7]{Francis M. Rouquette Jr}
\author[8]{Yanqi Wu}
\author[1]{Arbel Harpak}
\author[1]{Thomas E. Juenger}
\affil[1]{University of Texas at Austin, Austin}
\affil[2]{South Dakota State University, Brookings}
\affil[3]{USDA-ARS, Temple}
\affil[4]{University of Missouri, Columbia}
\affil[5]{Michigan State University, East Lansing}
\affil[6]{USDA-ARS, Lincoln}
\affil[7]{Texas A\&M University, Overton}
\affil[8]{Oklahoma State University, Stillwater}
\fi
\fi

\date{}

\fancyhead[C]{Alice H. MacQueen $\bullet$ Diverse Genotype-by-Weather
Interactions in Switchgrass}

\begin{document}

\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent{}The timing of vegetative and reproductive growth in plants
(``phenological timings'') depend on genetic effects (G), environmental
(e.g., weather) cues, and their interaction. Here, we measure
phenological timings in two highly divergent switchgrass (\emph{Panicum
virgatum}) subpopulations using repeated plantings of cloned individuals
at eight sites across the central United States. The timing of
vegetative growth for the two subpopulations reversed between their two
natural ranges and had strong negative correlations between these
regions; in contrast, the timing of flowering was positively~correlated
between gardens. We expect that these phenotypic correlations consist of
polygenic effects on phenology which have distinct patterns of GxE
segregating at different mapped loci. Thus, we infer the mixture of ways
genetic effects impact phenological timings, such as across common
gardens (GxE) or with weather cues (GxWeather). We demonstrate that we
can identify genetic variation with GxWeather and assign genetic loci to
specific weather-based cues or other patterns. For example, in the Gulf
subpopulation, 65\% of genetic effects on the timing of vegetative
growth covary with daylength 14 days prior to green-up date, and 33\% of
genetic effects on the timing of flowering covary with cumulative
rainfall in the week prior to flowering. However, most variation in
genetic effects cannot be attributed to variation in weather variables.
Selective breeding for particular alleles at GxWeather loci could alter
flowering responsiveness in a photoperiod or rainfall-specific way. More
broadly, our approach refines the characterization of
genotype-by-environment interactions and can be implemented in any
species phenotyped in multiple environments.
\end{abstract}
}

\maketitle

\section{Significance Statement}\label{significance-statement}

The timing of plant seasonal development (phenology) has major impacts
on fitness because of the steep fitness cost of plant-environment
mismatches. We infer the mixture of ways that genetic effects on
phenological traits depend on plant environment, focusing on how effects
covary with weather prior to the phenological event (GxWeather). Most
effects do not have GxWeather, but a minority do. GxWeather is
population-specific. The majority of effects on the timing of vegetative
growth in the Gulf population have GxWeather: these effects differ in
sign across environments and covary with a photoperiod cue two weeks
prior. A minority of effects on flowering date in the Gulf and Midwest
populations had GxWeather, covarying with a cumulative rainfall and a
photoperiod cue, respectively.

\section{Introduction}\label{introduction}

Plant phenological traits are important components of plant fitness that
are affected by multiple external environmental cues (e.g.~degree of
winter chilling, day length, temperature, soil fertility, and water
availability), which signal existing or upcoming growing conditions
\citep{bauerle_photoperiodic_2012, andres2012genetic, korner2010phenology, maple2024}.
Genetic responses to environmental cues determine the speed, timing, and
energy apportioned to vegetative and reproductive growth and shape plant
physiological responses, lifespan, and lifetime production of viable
seed. Day length (or photoperiod) is one of the most predictable
environmental cues, and genetic sensitivity to photoperiod protects
plants from potentially fatal consequences of phenological responses to
temperature cues at the ``wrong'' time of year. However, the utility of
specific environmental cues depends on both features of the environment,
such as cue predictability and relevance, and the species' adaptive
strategies \citet{botero2015evolutionary}. Species with wide natural
distributions can have multiple distinct environmentally cued
phenological responses. For example, populations of sunflower
(\emph{Helianthus annuus}) exhibit day-neutral, facultative short day,
and facultative long-day flowering responses, which vary with their
environments \citep{blackman2013interacting, henry2014transitions}.
These distinct genetic responses in different environments are known as
genotype-by-environment interactions, or GxE.

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research
\citep{blackman2013interacting, henry2014transitions, aagren2017adaptive, brachi2010linkage, dittmar2014flowering, li2018genomic, romero2017study, bustos-korts2019},
a key output of selection driving adaptation to local environments
\citep{unterseer2016, andres2012genetic, wadgymar2017identifying, blumel2015flowering, fournier-level2016},
and a selection target for crop improvement to adapt crops to local or
future environments \citet{jung2009flowering}. Changing flowering
responsiveness to photoperiod cues has allowed geographic range
expansion and increased yields in several cereal species
\citep{unterseer2016, turner2005pseudo, faure2012mutation, hung2012zmcct, zakhrabekova2012induced, yang2013oself3}
and other crops \citep{pin2012multifaceted, weller2019parallel}. Recent
statistical advances in studying phenological GxE have involved
determining critical environmental indices before the phenological event
occurs, such as photothermal time within a critical growth window
\citet{li2018genomic}. However, most studies of flowering GxE focus on
finding a single, best fitting form of genotype-environment covariance,
despite the key expectation that different genetic subpopulations, and
even different genomic regions, have likely evolved distinct patterns of
GxE \citep{Weine2023.06.21.545998}. Additionally, despite theoretical
predictions that local adaptation should involve trade-offs caused by
antagonistic pleiotropy, or alleles with effects with opposing fitness
outcomes
\citep{levene1953genetic, felsenstein1976theoretical, kawecki2004conceptual, hedrick1986genetic},
previous empirical work has found limited evidence of trade-offs caused
by this form of GxE
\citep{des2013genotype, wadgymar2017identifying, millet2016}. However,
this work has been limited by a known statistical bias that reduced
detection of genetic effects that differ in sign
\citep{des2013genotype, anderson2013genetic, anderson2011evolutionary}.
Thus, despite substantial interest in the frequencies of various forms
of GxE, the frequency of sign-changing GxE relative to other forms of
GxE remains unknown.

Previous research suggests that switchgrass phenological timings should
have GxWeather and that these timings could differ by genetic
subpopulation. Switchgrass is considered a short-day plant with
reproductive development strongly linked to day of the year
\citet{mitchell1997predicting}. However, as part of its wide
environmental adaptation across the eastern half of North America, its
photoperiodicity has been predicted to differ by plant latitude of
origin \citep{parrish2005biology, casler2004latitudinal}. We previously
found that divergent Midwest and Gulf genetic subpopulations of
switchgrass have distinct sets of environmental adaptations associated
with each of two fitness proxies, biomass and overwinter survival
\citet{lovell2021}. The Midwest genetic subpopulation is primarily
composed of individuals from the well-studied upland switchgrass ecotype
\citep{porter1966analysis, milano2016genetic}, while the Gulf
subpopulation has individuals from the well-studied lowland ecotype and
the phenotypically intermediate coastal ecotype \citet{lovell2021}.

Here, we test for GxWeather for the timing of two phenological traits by
loading patterns of genetic effects on phenology at eight common gardens
onto many patterns of weather covariance at these gardens. To do this,
we phenotyped a diversity panel of hundreds of switchgrass genotypes
from the Midwest and Gulf subpopulations for the timing of vegetative
and reproductive development. We did this at eight common garden
locations spanning 17 degrees of latitude: these gardens covered the
majority of the latitudinal and climatic range of switchgrass and
captured the most comprehensive picture to date of the environmental
variation this species encounters. We defined multiple ways phenological
traits might covary with weather (Table 1) and additional ways
phenological traits might vary by garden (SI Appendix, Section S1), then
jointly re-estimated genetic effects on these timings at all eight
common gardens using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included (SI
Appendix, Section S2-S5) \citet{urbut2019flexible}. We used the Bayesian
framework \emph{mash} (multivariate adaptive shrinkage) developed by
\citet{urbut2019flexible}, to refine effect size estimates from
genome-wide association (GWAS). \emph{Mash} allowed us to identify and
specify multiple covariance structures among genetic effect estimates
across sites, including structures that represent covariance in weather
variables of interest. Importantly, this method circumvented statistical
biases in detecting genetic effects with the same or opposite signs
\citet{10.1093/biostatistics/kxw041}. To confirm our genetic mapping of
GxWeather, we compared genomic locations of the significant posterior
effect estimates from \emph{mash} to mapping results from an outbred
mapping population grown at the same sites. Our analyses allowed us to
describe the weather cues and types of GxE affecting phenology in two
divergent natural populations of switchgrass across the species'
latitudinal range.

\section{Results}\label{results}

Genotypes from the Gulf and Midwest subpopulations had distinct
phenological trait timings and distinct patterns of phenological trait
correlations across our eight common garden sites
(Figure~\ref{fig-map}). At the three Texas common gardens (hereafter
`Texas' gardens), located within the natural range of the Gulf
subpopulation, the onset of vegetative growth, or `green-up date', for
Gulf genotypes occurred before Midwestern green-up, and the onset of
reproductive growth, or `flowering date', for Gulf genotypes occurred
after Midwestern flowering (Figure~\ref{fig-map} A). At the four
northernmost common gardens (hereafter `North' gardens), located within
the natural range of the Midwest subpopulation, both green-up and
flowering of Gulf genotypes occurred after that of Midwest genotypes. At
the Oklahoma common garden, located near the natural range limits of
both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up
occurred over the same period, and Gulf genotype flowering occurred
after Midwestern flowering (Figure~\ref{fig-map} A). These patterns led
to strong negative phenotypic correlations for the onset of vegetative
growth between the North and Texas gardens, particularly in the Gulf and
in the population containing all individuals from the Gulf and Midwest
subpopulations (hereafter, `Both' subpopulations) and contributed to
positive phenotypic correlations for the onset of flowering that had
larger magnitudes at more northern gardens (Figure~\ref{fig-map} B).

Narrow-sense heritabilities (h\textsuperscript{2}) suggested that
rank-changing GxE for these phenotypes was present across the common
gardens (SI Appendix, Figure S1). h\textsuperscript{2} were typically
high at individual gardens: 59\% on average for green-up date, and 87\%
for flowering date. However, h\textsuperscript{2} was variable across
gardens, and green-up dates were uncorrelated (r\textsuperscript{2}
\textless{} 0.2) or negatively correlated between pairs of gardens
(Figure~\ref{fig-map} B). These negative and small correlations
undoubtedly contributed to the low h\textsuperscript{2} values for
green-up and flowering date when estimated jointly at all eight gardens:
h\textsuperscript{2} was 0.8\% for green-up and 23.2\% for flowering
date (SI Appendix, Figure S1).

\subsubsection{Inference of genome-wide patterns of GxE and GxWeather
for vegetative and reproductive
timing}\label{inference-of-genome-wide-patterns-of-gxe-and-gxweather-for-vegetative-and-reproductive-timing}

We expected patterns of GxE and GxWeather to vary at the locus,
subpopulation, and trait level. Thus, we were interested in modeling the
different ways polygenic effects might covary across our eight common
gardens\citep{weine2024}. We jointly re-estimated the genetic effects of
a subset of SNPs across all eight common gardens using \emph{mash}
models, which can flexibly capture many modes of covariance across
gardens (SI Appendix, Sections S1-S5, Datasets 1-6). To obtain genetic
effects, we used subsets of effects from garden- and subpopulation GWAS
(SI Appendix, Section S2-S3; Table S1). To obtain covariance structures,
we specified both covariance structures introduced in the initial
\emph{mash} manuscript (e.g., garden-specific effects; equal effects at
all gardens, Figure~\ref{fig-covar} A), and GxWeather covariance
matrices estimated from the covariance of empirical weather patterns at
each garden at specific times before the phenological event (Table 1; SI
Appendix, Section S1). Then, we used a greedy algorithm to select
covariance matrices from each model's set that significantly improved
the model likelihood (SI Appendix, Section S4; Table S2). Finally, we
fit \emph{mash} models using the selected set of covariance matrices on
a subset of relatively unlinked (r\textsuperscript{2} \textless{} 0.2),
`strong' genetic effects with low \emph{p}-values (SI Appendix, Section
S5).

The loadings of genetic effects onto the GxE and GxWeather covariance
matrices included in each model provide information on genome-wide
patterns of SNP-environment interaction. For example, the phenotypic
correlations for the onset of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the \emph{mash} model. In this case, the joint estimate of effects
for many SNPs could have high mixture proportions, or mass, on a
covariance matrix that captures the negative correlation between Texas
(TX1, TX2, and TX3) and North (MO, NE, MI, and SD) gardens. In our data,
daylength 14 days prior to green-up date was negatively correlated
between Texas and North gardens (Figure~\ref{fig-covar} A); if many SNP
effects have mass on this GxWeather matrix, we would infer that the
effects of those SNPs have that form of GxWeather covariation.

Nine GxWeather covariance structures were selected by the greedy
algorithm, one to two per subpopulation: six for green-up date, and
three for flowering date (SI Appendix, Table S2). Of these nine, six had
mass (\textgreater0.1\%) on them in \emph{mash} models of the strong
effects (Figure~\ref{fig-covar} B,D). Five of these six matrices had
negative covariances between gardens in the Texas and North regions
(Figure~\ref{fig-covar} A), while one had all positive or near-zero
covariances.

For green-up date, SNP-associated phenotypic effects covaried with
different weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the \emph{mash} model of Gulf green-up fell on a
covariance matrix constructed using the covariance of day length 14 days
prior to the date of green-up. The covariance matrix for this weather
cue was visually similar to the observed pattern of phenotypic
correlation for the timing of vegetative growth in the Gulf
subpopulation (Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A).
\emph{Mash} models of the timing of vegetative growth in the Midwest
subpopulation and Both subpopulations did not include this GxWeather
covariance. The Midwest had 0.15\% weight on a GxWeather covariance
matrix of cumulative GDD for the 28d prior to green-up, a GxWeather
matrix that was visually similar to the Midwest's observed pattern of
phenotypic correlation (Figure~\ref{fig-map} B). Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in the
day prior to green-up (Figure~\ref{fig-covar} B). The average
temperature covariance matrix had negative covariances between Texas and
North gardens, though not as strong as the negative phenotypic
correlations observed in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). The Gulf subpopulation and Both
subpopulations had substantially more mass on GxWeather covariance
matrices than the Midwest population (Figure~\ref{fig-covar} C) for the
onset of vegetative growth.

For flowering date, distinct GxWeather covariance structures captured
covariance in effect sizes for SNPs in the Gulf and Midwest
subpopulations. 33\% of SNP effects on flowering in the Gulf
subpopulation covaried with cumulative rainfall in the seven days prior
to flowering (Figure~\ref{fig-covar} D). 22.6\% of SNP effects on
flowering in the Midwest subpopulation covaried with day length change
in the two days prior to flowering (Figure~\ref{fig-covar} D), which had
negative covariances between Texas and North gardens. Neither of these
covariance matrices significantly improved the log-likelihood of the
\emph{mash} model of Both subpopulations, and no GxWeather covariance
matrices had non-zero mass in models of Both subpopulations
(Figure~\ref{fig-covar} E). In five of the six \emph{mash} models of
strong effects, the GxWeather covariance matrices captured a minority of
the posterior weights of the strong effects (Figure~\ref{fig-covar}
C,E); in these five models, the majority of this mass was on various
canonical covariance matrices. These matrices included simple
heterozygosity, with intermediate, positive covariances between all
gardens, and single effect matrices with garden-specific effects.
Re-estimation of genetic effects with \emph{mash} allowed us to identify
and quantify the fraction of loci exhibiting GxWeather patterns; a
minority of GxE was GxWeather covariance, and the majority of GxE
covariance in these models is driven by other, unknown drivers of GxE.

\subsubsection{Frequency of rank-changing GxE in significant SNP
effects}\label{frequency-of-rank-changing-gxe-in-significant-snp-effects}

We expected to observe common rank-changing GxE at the level of
individual loci as this is a key theoretical prediction of local
adaptation
\citep{felsenstein1976theoretical, levene1953genetic, kawecki2004conceptual, hedrick1986genetic}.
Previous empirical work has found limited evidence of trade-offs caused
by antagonistic pleiotropy; however, this work had a known statistical
bias reducing the detection of effects that differed in sign
\citep{wadgymar2017identifying, des2013genotype, millet2016}. To
determine the frequency of rank-changing GxE, we used the local false
sign rate (lfsr), an analogue of the local false discovery rate that
establishes confidence in the effect sign, not the effect's difference
from zero, to determine significance. We required lfsr significance (p
\textless{} 0.05) in both gardens to include effects. This means that
our tests for a sign change between gardens carry an equal statistical
burden to those for effects with the same sign. We separated kinds of
effects at the level of individual loci into SNP effects that differ in
sign between gardens (effects with rank-changing GxE between gardens),
SNP effects that differ in magnitude (effects that are large in one
garden or region and smaller in others), and SNP effects that are
indistinguishable in two gardens (similarly large or small in
comparisons between gardens) (Figure~\ref{fig-effects} A).

For green-up date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited rank-changing GxE, or a difference in sign
between pairs of Texas and North gardens (Figure~\ref{fig-effects} B; SI
Appendix, Fig S2). 78.7\% of pairwise comparisons between North and
Texas gardens had a difference in sign, while only 28.6\% and 0.2\% of
North-North or Texas-Texas comparisons had a difference in sign,
respectively (Figure~\ref{fig-effects} B). The majority of pairwise
effects for greenup for the Midwest (\textgreater73\%; 68-342 effects)
were indistinguishable, or had no GxE. The majority of pairwise effects
for Both subpopulations (\textgreater85\%) were the same sign, and
effects most often differed in magnitude between gardens within and
between the regions; however, these differences in effect sign and
magnitude were mostly between the MO site and other sites
(Figure~\ref{fig-effects} B; SI Appendix, Fig S2).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited rank-changing GxE, or a difference in effect sign
within or between regions (Figure~\ref{fig-effects} B). More effects
differed in magnitude between the Texas and North regions than within
these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} B),
while the majority of effects had no GxE. The Midwest population had
relatively few significant effects for flowering (0-174 per pairwise
comparison), but a large proportion of these differed in sign between
Texas and North regions (42.7\%) or within the North region (65.4\%).
Finally, in Both subpopulations, less than 20\% of pairwise effects
differed in sign (Figure~\ref{fig-effects} B). Most differences in sign
were between TX1, the southernmost garden, and all other gardens (SI
Appendix, Figure S3). Similarly, more effects that differed in magnitude
included gardens in the Texas region (52.3-55.9\%), and most effect
pairs in the North region were not distinguishable (91.5\%).

\subsubsection{Confirmation of effects on phenology using an independent
mapping
population}\label{confirmation-of-effects-on-phenology-using-an-independent-mapping-population}

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Figure S4). All QTL for flowering overlapped
with one or more homologs from rice or \emph{A. thaliana} with
functionally validated roles in flowering (SI Appendix, Dataset 7); the
QTL on Chr04K overlapped a gene recently functionally validated for
flowering in switchgrass \citep{choi2023}. All flowering and green-up
QTL intervals contained at least one SNP significant in at least one
\emph{mash} run at a log10-transformed Bayes Factor \textgreater{} 2, or
in the 1\% tail of significance, whichever was stricter (SI Appendix,
Dataset 8). We also looked for enrichments of \emph{mash} SNPs in the
1\% tail of significance (the `\emph{mash} 1\% tail') within each QTL
interval. At the 5\% level, three QTL had enrichments of SNPs in the
\emph{mash} 1\% tail. Overall, there were five significant enrichments
(p \textless{} 0.05, hypergeometric test) of SNPs in the \emph{mash} 1\%
tail in the QTL intervals. Thus, we were able to experimentally support
the genomic windows of some re-estimates of significant SNP effects with
a QTL mapping experiment using a separate mapping population.

\section{Discussion}\label{discussion}

As the climate and the natural environment change, it is increasingly
critical to understand how patterns of plant-environment interactions
will change in response. To do this, we must understand the current
patterns of trait covariation across environments, the genetic
underpinnings of these patterns, and the cases where this covariation
can be altered through selection at individual loci. Here, we
demonstrate that we can associate multiple patterns of GxWeather with
specific genomic regions using a switchgrass diversity panel grown at
eight common gardens. We assigned genetic effects to both GxWeather
patterns with interpretable weather-based cues, and to unmeasured,
site-based patterns. We used this approach to study GxWeather for the
timings of vegetative and reproductive development in the deeply
genetically diverged Gulf and Midwest subpopulations of switchgrass.

Our analysis of the timing of vegetative and reproductive growth
revealed many alleles with rank-changing GxE, or sign changes in their
effects between the Texas and North gardens (Figure~\ref{fig-effects}
B). The Gulf \& Both subpopulations had rank-changing GxE for green-up
date, while the Midwest subpopulation had rank-changing GxE for
flowering date. As phenological timings are major components of plant
fitness, this result supports theoretical models that local adaptation
should involve trade-offs due to antagonistic pleiotropy at the level of
individual loci
\citep{kawecki2004conceptual, felsenstein1976theoretical, levene1953genetic, hedrick1986genetic}.
Experimental designs in local adaptation research now often use more
than two field sites and a wide range of genetic variation; \emph{mash}
and the local false sign rate facilitate the analysis of these
experimental designs. We use \emph{mash} to determine that rank-changing
GxE for phenological traits is common in small genomic regions
\citep{wadgymar2017identifying, savolainen2013ecological, lowry2019qtl}.

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related genetic variation, in that
flowering timing covaries with a day length change signal two days
before flowering occurs. In contrast, the Gulf subpopulation does not
have genetic variation in flowering that covaries with a photoperiod
cue. Instead, the Gulf subpopulation has genetic variation in flowering
that covaries with the rainfall that occurs in the week prior to
flowering. Three genomic regions affecting flowering that we
re-estimated across all eight sites were also supported by QTL from an
independent mapping population at these sites (Figure~\ref{fig-qtl} B).
Models combining both subpopulations showed less signal, perhaps due to
the distinctness of cues in both subpopulations or to confounding caused
by population structure.

Identifying the environmental cues that are predictive of, or even
correlated with, plant phenotypic responses remains a major challenge to
studies interrogating gene action across many natural environments. The
GxWeather photoperiod and cumulative rainfall cues we identify here are
functions of the genotypes measured and capture only a minority of SNP
effects on flowering. We could only assign SNP effects to a GxWeather
covariance structures in four of the six phenotype \& genetic
subpopulations we modeled. It's likely we did not include some GxWeather
important to these phenological cues - for example, overwintering
parameters that might cause variation in the start of vegetative growth.
More generally, it is difficult to predict the time scales over which
individuals may integrate environmental cues, particularly in perennial
species which may integrate these cues over longer time scales. If this
integration time itself varies between individuals, the covariance
structures we modeled cannot reflect this, though these structures would
likely be highly correlated with GxWeather structures we did include.
Our approach offers an opportunity to specify multiple environmental
cues and compete them to explain patterns of genetic effects, allowing
us to detect how important these cues are genome-wide, and how strongly
each cue influences each SNP. This is a key development to further
improve our understanding of genetic variation in GxE.

\section{Author Contributions}\label{author-contributions}

T.E.J. designed research. D.B.L. contributed plant material and
resources. J.B., D.B.L., and T.E.J. designed and executed field
experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W., and
T.E.J. hosted field experiments. A.H.M., L.Z., and S.P.S. conducted
statistical and computational analyses with input from T.E.J. and A.H.
The manuscript was written by A.H.M. with contributions from all
authors.

\section{Conflict of Interest}\label{conflict-of-interest}

The authors declare no conflicts of interest.

\section{Equal Contributor}\label{equal-contributor}

A.H.M. contributed equally to this work with L.Z. and S.P.S.

\section{Acknowledgements}\label{acknowledgements}

We thank the Brackenridge Field laboratory, the Ladybird Johnson
Wildflower Center, and the Juenger laboratory for support with plant
care and propagation. The authors acknowledge the Texas Advanced
Computing Center (TACC) at The University of Texas at Austin for
providing HPC storage resources that have contributed to the research
results reported within this paper.This material is based upon work
supported in part by the Great Lakes Bioenergy Research Center, U.S.
Department of Energy, Office of Science, Office of Biological and
Environmental Research under Award Numbers DE-SC0018409 and
DE-FC02-07ER64494, the US Department of Energy Awards DE-SC0021126 and
DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, NIH grant R35GM151108 and
a Pew Scholarship to A.H., National Science Foundation PGRP Awards
IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological
Research Program (DEB 1832042) at the Kellogg Biological Station.

\section{Materials and Methods}\label{materials-and-methods}

Whenever possible, plant material will be shared upon request. Source
data and code to replicate these analyses are available at:
https://github.com/Alice-MacQueen/pvdiv-phenology-gxe.git. SNP data to
replicate these analyses are available from the UT dataverse at
\url{https://doi.org/10.18738/T8/A604BU}.

\subsection{Scoring onset of vegetative and reproductive development in
two mapping
panels}\label{scoring-onset-of-vegetative-and-reproductive-development-in-two-mapping-panels}

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations
\citep{lovell2021, milano2016genetic, lowry2019qtl}. We scored the onset
of vegetative growth, or green-up date, as the day of the year when 50\%
of the tiller area of the crown of the plant cut the previous year had
green growth. The onset of reproductive growth, or flowering date, was
the day of the year when 50\% of the plant tillers had panicles
undergoing anthesis.

The formation and resequencing of the diversity panel has been described
previously \citet{lovell2021}. The diversity panel contained 134
sequenced, clonally propagated individuals from the Midwest genetic
subpopulation, and 229 from the Gulf genetic subpopulation. To allow for
the possibility that different subpopulations had different strengths of
connection between our phenotypes and genotypes \citep{kortefarlow2013},
we conducted three sets of genetic analyses: on Gulf and Midwest
genotypes separately, and on both subpopulations together (`Both'
subpopulations). Analyses to determine narrow-sense heritability
(h\textsuperscript{2}) for green-up and flowering were done using linear
mixed models and followed \citet{lovell2021} (SI Appendix, Section S6).

To confirm candidate genomic regions found using \emph{mash} on the
diversity panel, we analyzed flowering in an outbred pseudo-F2 cross
between four individuals, two Midwest and two Gulf individuals. The
formation of this mapping population has been described previously
\citep{milano2016genetic}; additional details on QTL mapping can be
found in SI Appendix, Section S7. To be directly comparable to the
diversity panel data, only 2019 phenology data from the pseudo-F2 cross
from the same eight common garden sites were used.

\subsection{Joint re-estimation of SNP effects to assess the frequency
of rank-changing GxE and assignment of genome-wide patterns of GxE and
GxWeather}\label{joint-re-estimation-of-snp-effects-to-assess-the-frequency-of-rank-changing-gxe-and-assignment-of-genome-wide-patterns-of-gxe-and-gxweather}

We were interested in specifying genetic models for trait variation that
allowed more than one form of GxE, as we reasoned that different loci
should display different forms of GxE (e.g., changes in effect magnitude
vs sign). In addition, we were interested in an unbiased estimation of
the frequency of rank-changing GxE relative to other forms of GxE, as
the presence of rank-changing GxE at the level of individual loci is a
key theoretical prediction of local adaptation. \emph{Mash} allowed us
to both specify multiple forms of GxE and GxWeather and conduct unbiased
statistical tests for when SNP effects changed sign between common
gardens.

To use \emph{mash} on our diversity panel, we had to specify both a
relatively uncorrelated set of covariance matrices, which in our case
defined types of GxE and GxWeather between gardens, and we had to
specify subsets of SNP effect estimates and standard errors for our
traits at each common garden. To specify a set of covariance matrices,
we first defined many covariance matrices, including GxWeather matrices
that represented the correlation in weather cues between gardens before
the phenological event (SI Appendix, Section S1), then implemented a
model selection approach that used a greedy algorithm to evaluate if the
log likelihood of the \emph{mash} model was significantly improved as
additional covariance matrices were included (SI Appendix, Section S4).
To specify subsets of SNP effect estimates and standard errors to use
for both the greedy algorithm and with the optimal set of covariance
matrices, we first calculated best linear unbiased predictors (BLUPs)
for each phenological trait in each genetic subpopulation and each
common garden (SI Appendix, Section S2). Next, we determined effect
estimates for 8.8 to 12.3 million SNPs per subpopulation by conducting
garden-specific GWAS on these BLUPs using \(k\) vectors of singular
values to correct for population structure, where \(k\) was the smallest
integer value that made the genomic control coefficient closest to 1 (SI
Appendix, Section S2). Singular values were computed using singular
value decomposition of the matrix of all SNPs, with iterative SNP
pruning and removal of regions in long-range linkage disequilibrium
\citep{privé2017}. Third, to make the \emph{mash} models computationally
feasible, we extracted two subsets of SNP effect estimates and standard
errors from our GWAS effect estimates: (i) effects from a subset of
``strong'' tests corresponding to stronger effects on our traits; (ii)
results from a random subset of all tests to correspond to an unbiased
representation of all effects (SI Appendix, Section S3). We used the
subset of random effects in our greedy \emph{mash} algorithm (SI
Appendix, Section S4) and used the subset of strong effects in our
\emph{mash} models with the optimal set of covariance matrices (SI
Appendix, Section S5).

The loadings of genetic effects onto the multiple covariance structures
specified in our \emph{mash} models provided information on genome-wide
patterns of SNP-environment interaction. In addition, the the GxWeather
covariance structures allowed hypothesis testing of specific weather
variables as cues for the start of vegetative and reproductive growth.
Say that our diversity panel contains a SNP in the gene CONSTANS (CO), a
well-known flowering time regulator, and that only one of the alleles
affects the promotion of flowering in a photoperiod-dependent manner. In
that case, the joint estimate of effects for that SNP could have a high
mixture proportion, or mass, on a covariance matrix created using a
photoperiod-based environmental cue, such as day length at some interval
prior to flowering. In our data, we would infer that the effect of that
SNP on flowering was cued by the weather variable used to create the
GxWeather covariance structure.

Our joint re-estimation of SNP effects also allowed us to characterize
the overall patterns of GxE in the set of SNPs where there was pairwise
significance of effects at pairs of gardens. To do this, we used the
`get\_GxE' function of the switchgrassGWAS R package. First, this
function determines the set of SNPs with evidence of significant effects
in both conditions for all pairs of conditions using local false sign
rates (lfsr) as the significance criteria. Then, this function
determines if effects significant in both conditions are of opposite
sign.

Using the lfsr rather than the local false discovery rate (local FDR) is
a critical change in our ability to detect alleles directly contributing
to rank-changing GxE between environments. The local FDR, like other
measures of FDR, focuses on if we have enough evidence to reject the
null hypothesis that an effect j is 0, or that there is a significant
effect. Previous studies of antagonistic pleiotropy (e.g.
\citep{lowry2019qtl}) have used the local FDR or equivalent statistical
tests to detect antagonistic pleiotropy. These tests were conservative,
in that they required two non-zero effects of different signs, while
tests for differential sensitivity required only one non-zero effect.
This previous work recognized that this testing bias could lead to
undercounting occurrences of antagonistic pleiotropy
\citep{anderson2011evolutionary, des2013genotype}, and sought to reduce
it by permutation \citep{anderson2013genetic}. However, using the lfsr
to test for allelic effects that differ in sign does not undercount
these occurrences, as this statistic answers a fundamentally different
question. For each effect \(j\), the \(lfsr_j\) is defined as the
probability that we make an error in the sign of effect \(j\) if we were
forced to declare the effect positive or negative
\citet{10.1093/biostatistics/kxw041}. Thus, rather than asking ``Are
these two effects different?'' - as we reasonably expect two effects to
be, even if this difference cannot be measured - the local false sign
rate answers a more meaningful question: Can we be confident in the sign
of this effect?

In addition, the get\_GxE function also sets an arbitrary threshold to
count an effect as changing in magnitude between environments, commonly
known as differential sensitivity or a change in amplitude of the
effect. For differential sensitivity, this function determines if
effects significant in both conditions are of the same sign and of a
magnitude (not tested for significance) that differs by a factor of 0.4
or more. The remaining effects that are significant in both conditions
have the same effect sign and similar effect magnitudes and we denote
these effects as having no GxE. The distinction between effects with
different magnitudes is arbitrary but useful to fully characterize how
effects vary across environments to ultimately influence phenotypes. Our
use of the lfsr to determine significance and our specification that SNP
effects must be significant in both conditions to be included means that
our tests for alleles with rank-changing GxE carry an equal statistical
burden to those measuring differential sensitivity and effects without
GxE.

\section{Figures and Tables}\label{figures-and-tables}

\clearpage
\onecolumn
\begin{table}[t!]
\centering
\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices. Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices \\
\midrule
1. cumulative growing degree days at 12C in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
\bottomrule
\end{tabular}

\end{table}
\clearpage
\twocolumn

\begin{figure}

\centering{

\includegraphics{images/Fig1_AB.png}

}

\caption{\label{fig-map}Characterization of the timing of the onset of
vegetative (green-up) and reproductive (flowering) growth in the
switchgrass diversity panel. (a) Map and trait histograms of green-up
and flowering dates across two genetically distinct switchgrass
subpopulations and eight common gardens. Purple represents individuals
from the Midwest genetic subpopulation, and pink individuals from the
Gulf subpopulation; map positions represent the original collection
locations for the genotypes, and shapes represent the ecotype of the
genotype. Histogram vertical dashed lines indicate the summer solstice.
Common gardens are arranged in latitudinal order. (b) Phenotypic
correlations between clonal replicates planted at eight common gardens,
within and between two genetic subpopulations.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{images/Figure_2_six_sites_Midwest.png}

}

\caption{\label{fig-covar}Example Canonical and GxWeather covariance
matrices specified in \emph{mash} and the posterior weights placed on
all covariance matrices. (a) Common gardens are arranged in latitudinal
order within the matrices. Top row: Four example canonical covariance
matrices. Canonical matrices (purple) have simple interpretations, such
as equal effects across all common gardens, or effects specific to a
single common garden. Bottom row: Five example GxWeather covariance
matrices specified for the green-up date or flowering date phenotype;
these matrices were created from environment-specific correlations
across eight common gardens, and are described in Table 1. (b,d) Total
posterior weight placed on each covariance matrix type specified for (b)
green-up date and (d) flowering date \emph{mash} models, within and
between two genetic subpopulations. Covariance matrices included in
\emph{mash} that had zero posterior weight in all three \emph{mash} runs
on the genetic subpopulations, such as the identity matrix, are not
shown. (c,e) Total posterior weight placed on covariance matrices that
were GxWeather or Canonical, for the (c) green-up date phenotype and (e)
flowering date phenotype.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{images/Figure_3_AH_comments.png}

}

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated SNP effects in eight common gardens, for effects
with lfsr \textless{} 0.05 at both gardens for each pair of gardens
contrasted. a) Examples of effect patterns at three pairs of sites with
three types of GxE. All effects are for an alternate allele, with the
reference allele effect defined to be zero at both gardens and
represented by the dashed vertical line. Sign: Effects that differ in
sign at these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude:
Effects identical in sign (p \textless{} 0.05, lfsr) that differ in
magnitude by a factor of \textgreater0.4. Not Distinguishable: Effects
not distinguishable by magnitude nor sign of the effect, with no
measurable GxE. Confidence intervals are illustrative that the effect
estimate does not overlap zero. b) The fraction of effects with each GxE
type for the onset of vegetative growth (green-up date) and reproductive
growth (flowering date), within and between two genetic subpopulations.
Common gardens are grouped by the larger region they came from: North
gardens are within the natural range of the Midwest subpopulation, and
include MO, NE, MI, and SD, while Texas gardens are within the natural
range of the Gulf subpopulation, and include TX1, TX2, and TX3.}

\end{figure}%

\begin{figure}

\centering{

\includegraphics{images/Figure_4_QTL_Overlaps_6_Midwest_Gardens.png}

}

\caption{\label{fig-qtl}Overlaps of QTL from an outbred pseudo-F2 cross
and with jointly re-estimated SNP effects in the 1\% tail of
significance from a diversity panel. Dotted lines indicate
permutation-based significance thresholds for each weather-related
function. Stars indicate QTL with significant enrichment for SNPs in the
1\% \emph{mash} tail; G, M, and B indicate which subpopulation had
enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both
subpopulations. Rug plots show genomic locations of SNPs in the 1\%
\emph{mash} tail for flowering date for each subpopulation: Both
subpopulations are above the plot panel, the Gulf subpopulation is above
the x-axis, and the Midwest subpopulation is below the x-axis. a) QTL
mapping for the onset of vegetative growth (green-up date), and three
weather-related functions of green-up date. b) QTL mapping for the onset
of reproductive growth (flowering date), and two weather-related
functions of flowering date.}

\end{figure}%

\hfill\break
~

\section{References}\label{references}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\end{document}